---
title: "RNA seq Analysis"
output: html_document
date: "2023-02-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

```{r}
library(AnnotationDbi)
library(org.Mm.eg.db)
library(stringi)
library(DESeq2)
library(dplyr)
library(mixOmics)
library(sva)
library(edgeR)
library(IHW)
library(biomaRt)
```

## Loading the data

```{r,include=TRUE, results='hide'}
myFactors <- read.csv("MERGED_gfcv&LIMA_metadata.csv",
                         stringsAsFactors = T,
                         colClasses = "factor" ,
                         row.names = 1)

myCounts <- read.csv("MERGED_gfcv&LIMA_raw_counts.csv",
                         stringsAsFactors = F,
                         colClasses = c("character",rep("numeric",74)),
                         row.names = 1)
```

### Model data frames and metadata
Create 2 different data frames for each model, with counts and metadata
```{r}
#counts GFCV
myGFCVcounts <- dplyr::select(myCounts, -contains("X"))

myGFCVcounts <- as.data.frame(t(myGFCVcounts))

#metadata GFCV
myGFCVfactors <- dplyr::slice(myFactors,37:74)

#write.csv("gfcv_metadata.csv")
#counts LIMA
myLIMAcounts <- dplyr::select(myCounts, contains("X"))

cln <- colnames(myLIMAcounts)

cln <- gsub("X","",cln)

colnames(myLIMAcounts) <- cln

myLIMAcounts <- as.data.frame(t(myLIMAcounts))

#metadata LIMA
myLIMAfactors <- dplyr::slice(myFactors,1:36)
```

### Length function
Create a function to calculate trasncript length from GTF file (Sonia Tarazona)
```{r}
gtf2length <-  function (mydir, myfile) {
  
  mygtf = read.delim(file.path(mydir, myfile), header = FALSE, as.is = TRUE)
  
  colnames(mygtf) = c("entity", "start", "end", "gene", "transcript")
  
  mygtf = mygtf[which(mygtf[,"entity"] == "exon"),-1]
  
  mylength = abs(mygtf[,"end"]-mygtf[,"start"]) + 1
  
  trans.length = aggregate(mylength, 
                           by = list("gene" = mygtf[,"gene"],
                                     "transcript" = mygtf[,"transcript"]), sum)
  
  gene.length = aggregate(trans.length[,"x"],
                          by = list("gene" = trans.length[,"gene"]), median)
  
  colnames(gene.length)[2] = "length"

  gene.length
  
}
```

### Transcript length calculation
Calculate transcript length from GTF file
```{r}
mydir <- "./"
myfile <- "mmuGTF.txt"

lengthGTF <- data.frame(gtf2length(mydir,myfile))

lengthGTF <- data.frame(ensembl_gene_id=lengthGTF$gene,
                        lengthGTF=lengthGTF$length)

lengthGTF$ensembl_gene_id <- stri_sub(lengthGTF$ensembl_gene_id, 1,18)
```


### TPM calculation
Calculate TPM
```{r}
tpm <- function(counts, lengths) {
  rate <- counts / lengths
  rate / sum(rate) * 1e6
}

myGFCVcounts_t <- as.data.frame(t(myGFCVcounts))

myGFCVcounts_t$ensembl_gene_id <- rownames(myGFCVcounts_t)

lengthGTF <- dplyr::inner_join(lengthGTF,
                               myGFCVcounts_t,
                               by="ensembl_gene_id")

lengthGTF <- dplyr::select(lengthGTF,
                           ensembl_gene_id,
                           lengthGTF)

myGFCVcounts_t <- dplyr::select(myGFCVcounts_t,
                                -ensembl_gene_id)

rownames(lengthGTF) <- lengthGTF$ensembl_gene_id

lengthGTF <- dplyr::select(lengthGTF, lengthGTF)

dim(lengthGTF)
dim(myGFCVcounts_t)

gfcvTPM <- data.frame(ensembl_gene_id=rownames(myGFCVcounts_t),NA)

for (i in 1:ncol(myGFCVcounts_t)){ 
  sample <- myGFCVcounts_t[,i,drop=T]
  TPM <- tpm(myGFCVcounts_t[,i],lengthGTF)
  gfcvTPM[,i] <- cbind(data.frame(TPM))
  colnames(gfcvTPM) <- colnames(myGFCVcounts_t[,i])
}

colnames(gfcvTPM) <- colnames(myGFCVcounts_t)

rownames(gfcvTPM) <- rownames(myGFCVcounts_t)
```


###Filtering 80% and TPM > 0.1
Filtering out genes with TPM > 0.1, output in counts 
```{r}

# filtering 80% from TPM
dim(gfcvTPM)

gfcvTPMall_t <- as.data.frame(t(gfcvTPM))

dim(gfcvTPMall_t)

myGFCVfactors$groups_m <- factor(myGFCVfactors$groups_m, 
                                 levels = c("CVF","CVM", 
                                            "GFF", "GFM"))

groups <- levels(myGFCVfactors$groups_m)
cutoff <- 0.2
thresholds <- round(table(myGFCVfactors$groups_m) * cutoff)
selected_cols <- lapply(groups, function(i){
  gr <- rownames(myGFCVfactors[myGFCVfactors$groups_m == i,])
  ceros <- gfcvTPMall_t[gr,] == 0
  selc_gr <- which(colSums(ceros) < thresholds[i])
  return(selc_gr)
})
selected_index <- unique(unlist(selected_cols))
myGFCVcountsTPM_80 <- gfcvTPMall_t[,selected_index]

dim(myGFCVcountsTPM_80)
myGFCVcountsTPM_80 <- as.data.frame(t(myGFCVcountsTPM_80))
dim(myGFCVcountsTPM_80)

# Filtering TPM>0.1 after the 80% filter

dim(myGFCVcountsTPM_80)
myGFCVcountsTPM_80 <- myGFCVcountsTPM_80 %>% dplyr::mutate(aver = rowSums(.)/38)

myGFCVcountsTPM_80ord <- dplyr::arrange(myGFCVcountsTPM_80, aver)

dim(myGFCVcountsTPM_80ord)
  
myGFCVcountsTPM_80ord_f <- dplyr::filter(myGFCVcountsTPM_80ord, aver > 0.1)

dim(myGFCVcountsTPM_80ord_f) # same results, the order does not matter

myGFCVcountsTPM_80ord_f <- dplyr::select(myGFCVcountsTPM_80ord_f, -aver)

dim(myGFCVcountsTPM_80ord_f)

```

###80% and 0.1 filtered, raw counts
Only keep genes that appear in 80% of the samples in at least one group. 
```{r}
myGFCVcountsTPM_80_f_id <- data.frame(ensembl_gene_id=rownames(myGFCVcountsTPM_80ord_f))


dim(myGFCVcounts_t)
myGFCVcounts_t$ensembl_gene_id <- rownames(myGFCVcounts_t)



# keep raw counts for genes with 80% filter
myGFCVcounts80 <- dplyr::inner_join(myGFCVcountsTPM_80_f_id,
                                    myGFCVcounts_t,
                                    by="ensembl_gene_id")

dim(myGFCVcounts80)

rownames(myGFCVcounts80) <- myGFCVcounts80$ensembl_gene_id

myGFCVcounts80 <- dplyr::select(myGFCVcounts80, -ensembl_gene_id)


#Eliminate genes with variance 0
dim(myGFCVcounts80)
gene_var <- data.frame(gene_var=apply(myGFCVcounts80, 1, var))
dim(gene_var)
gene_var_id <- dplyr::filter(gene_var,gene_var>0) # Michael Love against that
dim(gene_var_id)



##################################################
##################################################
##### Matrix to use for DESeq2 and downstream# ###
                dim(myGFCVcounts80)############# #
##################################################
##################################################

#write_xlsx(myGFCVcounts80, "myGFCVcounts80.xlsx")
write.csv(myGFCVcounts80, file = "GFCV_transcriptomics_filtered.csv", 
          row.names = TRUE)

dim(myGFCVcounts80)

myGFCVcounts80_t <- as.data.frame(t(myGFCVcounts80))

dim(myGFCVcounts80_t)


p_rna_raw <-mixOmics::pca(myGFCVcounts80_t, ncomp = 6, center = T, scale = T,
         max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
         V = NULL, multilevel = NULL)


pdf("GFCV_liv_PCAmixomics_raw80_jun2022.pdf")

plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$gfcv, legend = T, title= "rna tpm liv by gfcv")

plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$groups_m, legend = T, title= "rna tpm liv by group",ellipse = T, ellipse.level = 0.99)

plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$sex, legend = T, title= "rna tpm liv by sex",ellipse = T, ellipse.level = 0.99)

plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$batch,comp = c(1,2), legend = T, title= "rna tpm liv by batch",ellipse = T, ellipse.level = 0.99)
plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$batch,comp = c(1,3), legend = T, title= "rna tpm liv by batch",ellipse = T, ellipse.level = 0.99)
plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$batch,comp = c(2,3), legend = T, title= "rna tpm liv by batch",ellipse = T, ellipse.level = 0.99)

dev.off()

metapei_id <- data.frame(ensembl_gene_id=rownames(myGFCVcounts80))


save(metapei_id, file="GFCV_metaepi1_20k_ensembl_gene_id_80filter.Rdata")

#write_xlsx(metapei_id, "GFCV_metaepi1_20k_ensembl_gene_id_80filter.xlsx")

```

###DESeq2 and VST transformation
```{r}
myGFCVfactors$sex_m <- factor(myGFCVfactors$sex_m , ordered = F)

myGFCVfactors_gs <- dplyr::select(myGFCVfactors, gfcv, sex_m)



ddsGFCV80 <- DESeqDataSetFromMatrix(countData = myGFCVcounts80,
                                      colData = myGFCVfactors_gs,
                                      design= ~ gfcv + sex_m + gfcv:sex_m)


# ddsGFCV80 <- DESeq(ddsGFCV80, test = "Wald")
vsdGFCV80 <- vst(ddsGFCV80, blind = FALSE)

#plotPCA with vst() transformed data

pdf("GFCV_liv_PCAdeseq2_vst_jun2022.pdf")
plotPCA(vsdGFCV80, intgroup=c("gfcv", "sex_m"))
dev.off()


# Extract transformed matrix

vsdGFCV80_mat <- assay(vsdGFCV80)

dim(vsdGFCV80_mat)

class(vsdGFCV80_mat)

vsdGFCV80_df <- as.data.frame(vsdGFCV80_mat)

# filter vst transformed data var=0
dim(vsdGFCV80_mat)
gene_vst_var <- data.frame(gene_var=apply(vsdGFCV80_mat, 1, var))
dim(gene_vst_var)
gene_var_vst_id <- dplyr::filter(gene_vst_var,gene_var>0)
dim(gene_var_vst_id)

df <- bind_rows(
  as_tibble(assay(vsdGFCV80)[, 1:2]) %>% mutate(transformation = "vst"))
  
colnames(df)[1:2] <- c("x", "y")  


ggplot(df, aes(x = x, y = y)) + geom_hex(bins = 80) +
  coord_fixed(xlim = c(0,25), ylim =  c(0,25))
dim(df)

```

### ComBat batch correction
There is a mild batch effect that we already saw in other omics and we opted to correct this batch effect with the same algorithm (ComBat)
```{r}
# Combat on 80% and 0.1 filtered data
dim(myGFCVcounts80)
class(myGFCVcounts80)

myGFCVcounts80_m <- as.matrix(myGFCVcounts80)


myGFCVcounts80_CB <-  ComBat(myGFCVcounts80_m, myGFCVfactors$batch)


sum(table(myGFCVcounts80_CB[myGFCVcounts80_CB<0]))


myGFCVcounts80_CB[myGFCVcounts80_CB<0] <- 0


sum(table(myGFCVcounts80_CB[myGFCVcounts80_CB<0]))


myGFCVcounts80_CB <- as.data.frame(round(myGFCVcounts80_CB, digits = 0))


dim(myGFCVcounts80_CB)

pdf("GFCV_liv_boxplot_80&01_raw&vst_jun2022.pdf")
boxplot(log10(myGFCVcounts80_CB), las=2, main= "80 and 01 raw combat corrected")
boxplot(log10(myGFCVcounts80), las=2, main= "80 and 01 raw not combat corrected")


boxplot(vsdGFCV80_mat, las=2, main= "80 and 01 vst not combat corrected")
dev.off()

myGFCVcounts80_CB <- as.data.frame(t(myGFCVcounts80_CB))


dim(myGFCVcounts80_CB)
dim(myGFCVcounts80_t)


p_rnaCB <-mixOmics::pca(myGFCVcounts80_CB, ncomp = 6, center = T, scale = T,
         max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
         V = NULL, multilevel = NULL)

p_rna <-mixOmics::pca(myGFCVcounts80_t, ncomp = 6, center = T, scale = T,
         max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
         V = NULL, multilevel = NULL)


pdf("GFCV_liv_PCAmixom_rna_CB&noCB_feb2022_jun2022.pdf")

plotIndiv(p_rnaCB, ind.names = T, group = myGFCVfactors$gfcv, legend = T, title= "rna liv CB by gfcv")

plotIndiv(p_rna, ind.names = T, group = myGFCVfactors$gfcv, legend = T, title= "rna liv not CB by gfcv")


plotIndiv(p_rnaCB, ind.names = T, group = myGFCVfactors$groups_m, legend = T, title= "rna liv CB by group")

plotIndiv(p_rna, ind.names = T, group = myGFCVfactors$groups_m, legend = T, title= "rna liv not CB by group")


plotIndiv(p_rnaCB, ind.names = T, group = myGFCVfactors$sex, legend = T, title= "rna liv CB by sex",ellipse = T, ellipse.level = 0.99)

plotIndiv(p_rna, ind.names = T, group = myGFCVfactors$sex, legend = T, title= "rna liv not CB by sex",ellipse = T, ellipse.level = 0.99)


plotIndiv(p_rnaCB, ind.names = T, group = myGFCVfactors$batch, legend = T,comp = c(1,2), title= "rna liv CB by batch",ellipse = T, ellipse.level = 0.99)

plotIndiv(p_rna, ind.names = T, group = myGFCVfactors$batch, legend = T,comp = c(1,2), title= "rna liv not CB by batch",ellipse = T, ellipse.level = 0.99)


plotIndiv(p_rnaCB, ind.names = T, group = myGFCVfactors$batch, legend = T,comp = c(1,3), title= "rna liv CB by batch",ellipse = T, ellipse.level = 0.99)

plotIndiv(p_rna, ind.names = T, group = myGFCVfactors$batch, legend = T,comp = c(1,3), title= "rna liv not CB by batch",ellipse = T, ellipse.level = 0.99)


dev.off()

#Eliminate genes with variance 0
myGFCVcounts80_CB <- as.data.frame(t(myGFCVcounts80_CB))

gene_var <- data.frame(gene_var=apply(myGFCVcounts80_CB, 1, var))
dim(gene_var)
gene_var_id <- dplyr::filter(gene_var,gene_var>0) # Michael Love against eliminating low variance genes for DESeq2
dim(gene_var_id)


```

### TMM & VST normalization
Using myGFCVcounts80_CB (no TPM filtering) corrected data
```{r}
# TMM transformation

#myGFCVcounts80_CB

dim(myGFCVcounts80)
dim(myGFCVfactors)

myGFCVfactors_tmm <- data.frame(sample=rownames(myGFCVfactors),
                                group=stri_sub(rownames(myGFCVfactors),1,2),
                                gfcv=stri_sub(rownames(myGFCVfactors),1,1),
                                sex=stri_sub(rownames(myGFCVfactors),2,2),
                                batch=stri_sub(rownames(myGFCVfactors),4,4),
                                lib.size=colSums(myGFCVcounts80_CB))

dim(myGFCVfactors_tmm)



#DGEList object 80%filter
y <- DGEList(counts=myGFCVcounts80,
             samples = myGFCVfactors_tmm)

y <- calcNormFactors(y,method = "TMM",lib.size=y$samples$lib.size)


cpms <- cpm(y, log=FALSE)

logcpms <- cpm(y, log=TRUE)

#MDS plot
pdf("GFCV_plotMDS_cpms_edgeR_jun2022.pdf")
plotMDS(cpms, labels=y$samples$sample,main="CPMs")

plotMDS(logcpms, labels=y$samples$sample,main="logCPMs")
dev.off()



#Eliminate genes with variance 0
gene_var <- data.frame(gene_var=apply(cpms, 1, var))
dim(gene_var)
gene_var_id <- dplyr::filter(gene_var,gene_var>0) # Michael Love against that for input to DESeq2
dim(gene_var_id)


#PCA tmm counts

dim(cpms)
cpms_t <- as.data.frame(t(cpms))



p_cpms <-mixOmics::pca(cpms_t, ncomp = 6, center = TRUE, scale = TRUE,
         max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
         V = NULL, multilevel = NULL)


pdf("GFCV_liv_PCAmixom_rna_tmm_80_jun2022.pdf")
plotIndiv(p_cpms, ind.names = T, group = myGFCVfactors$gfcv, legend = T, title= "rna tmm liv by gfcv, size norm")


plotIndiv(p_cpms, ind.names = T, group = myGFCVfactors$groups_m, legend = T, title= "rna tmm liv by group, size norm")

plotIndiv(p_cpms, ind.names = T, group = myGFCVfactors$sex, legend = T, title= "rna tmm liv by sex, size norm",ellipse = T, ellipse.level = 0.99)


plotIndiv(p_cpms, ind.names = T, group = myGFCVfactors$batch, legend = T, title= "rna tmm liv by batch, size norm",ellipse = T, ellipse.level = 0.99)


dev.off()

dim(cpms_t)
dim(cpms)


cpms_for_hist_t <- as.data.frame(cpms_t)

cpms_for_hist <- as.data.frame(cpms)


hist(log10(cpms_for_hist$CF21), breaks = 50)

cpms_for_hist_sc <- as.data.frame(t(base::scale(t(cpms_for_hist))))

cpms_for_hist_sc <- as.data.frame(cpms_for_hist_sc)

hist(cpms_for_hist_sc$CF22, breaks = 50, xlim = c(-2,2))


hist(log10(cpms_for_hist_t$ENSMUSG00000004043), breaks = 5)

cpms_for_hist_sc_t <- as.data.frame(t(base::scale(t(cpms_for_hist_t))))

cpms_for_hist_sc_t <- as.data.frame(cpms_for_hist_sc_t)

hist(cpms_for_hist_sc_t$ENSMUSG00000004043, breaks = 5)


pdf("RNAseq_histo_tmm_80_jun2022.pdf")
hist(cpms_for_hist_sc$CF21, breaks = 50)
hist(cpms_for_hist_sc_t$ENSMUSG00000004043, breaks = 5)
dev.off()


# VST transformation for MOFA (leght size + vst)
dim(myGFCVcounts80)
dim(myGFCVfactors)

ddsGFCV80 <- DESeqDataSetFromMatrix(countData = myGFCVcounts80,
                                       colData = myGFCVfactors,
                                       design= ~ gfcv + sex_m +
                                         gfcv:sex_m)


ddsGFCV80 <- DESeq(ddsGFCV80, test = "LRT", reduced = ~1)



ddsGFCV80_vst <- vst(ddsGFCV80,blind = FALSE)

myGFCVcounts80_vst <- as.data.frame(assay(ddsGFCV80_vst))

hist(as.numeric(c(myGFCVcounts80_vst[1,])))
hist(myGFCVcounts80_vst$CF11)

dim(myGFCVcounts80_vst)

myGFCVcounts80_vst_t <- as.data.frame(t(myGFCVcounts80_vst))
dim(myGFCVcounts80_vst_t)


p_vst <-mixOmics::pca(myGFCVcounts80_vst_t, ncomp = 6, center = TRUE, scale = F,
         max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
         V = NULL, multilevel = NULL)


pdf("GFCV_liv_PCAmixom_rna_80_vst_jun2022.pdf")
plotIndiv(p_vst, ind.names = T, group = myGFCVfactors$gfcv, legend = T, title= "rna vst liv by gfcv, size norm")


plotIndiv(p_vst, ind.names = T, group = myGFCVfactors$groups_m, legend = T, title= "rna vst liv by group, size norm")

plotIndiv(p_vst, ind.names = T, group = myGFCVfactors$sex, legend = T, title= "rna vst liv by sex, size norm",ellipse = T, ellipse.level = 0.99)


plotIndiv(p_vst, ind.names = T, group = myGFCVfactors$batch, legend = T, title= "rna vst liv by batch, size norm",ellipse = T, ellipse.level = 0.99)


dev.off()

```


## GFCV DESeq2
Run DESeq with LRT

### DESeq

#### Simplified model parametrization

```{r}
gfcv_80_dt <- read.csv("./GFCV_transcriptomics_filtered.csv", row.names = 1)
myFactors <- read.csv("MERGED_gfcv&LIMA_metadata.csv",
                         stringsAsFactors = T,
                         colClasses = "factor" ,
                         row.names = 1)

myGFCVfactors <- dplyr::slice(myFactors,37:74)

write.csv(myGFCVfactors, file = "RNASeq_gfcv_metadata.csv")

gfcv_DESeq <- DESeqDataSetFromMatrix(countData = gfcv_80_dt,
                                      colData = myGFCVfactors,
                                      design= ~ batch + gfcv + sex_m + gfcv:sex_m)
gfcv_DESeq$gfcv <- relevel(gfcv_DESeq$gfcv, ref = "C")
gfcv_DESeq$sex_m <- relevel(gfcv_DESeq$sex_m, ref = "M")

gfcv_DESeq <- DESeq(gfcv_DESeq, test = "Wald")

ctrs_vec <- resultsNames(gfcv_DESeq)
ctrs_vec
gfcv_DESeq_results <- lapply(ctrs_vec[5], function(ctrs){results(gfcv_DESeq, name = ctrs, filterFun = ihw)})
sapply(gfcv_DESeq_results, summary)
names(gfcv_DESeq_results) <- ctrs_vec[5]

saveRDS(gfcv_DESeq_results, "RNASeq_gfcv_siplified_model_results_list.rds")
```

#### Complete model parametrization

```{r}
gfcv_80_dt <- read.csv("./GFCV_transcriptomics_filtered.csv", row.names = 1)
myGFCVfactors <- read.csv("./RNASeq_gfcv_metadata.csv")
gfcv_DESeq <- DESeqDataSetFromMatrix(countData = gfcv_80_dt,
                                      colData = myGFCVfactors,
                                      design= ~ 0 + groups_m + batch)

gfcv_DESeq <- DESeq(gfcv_DESeq, test = "Wald")

ctrs_vec <- resultsNames(gfcv_DESeq)
ctrs_vec_list <- list(c("groups_m", "GFM", "CVM"),c("groups_m", "GFF", "CVF"), 
                      c("groups_m", "CVF", "CVM"), c("groups_m", "GFF", "GFM"),
                      c("groups_m", "GFM", "CVF"), c("groups_m", "GFF", "CVM"))

gfcv_DESeq_results_complete <- lapply(ctrs_vec_list, function(ctrs){
    results(gfcv_DESeq, contrast = ctrs, filterFun = ihw)
})
lapply(gfcv_DESeq_results_complete, summary)

names(gfcv_DESeq_results_complete) <- c("GFM_CVM", "GFF_CVF", "CVF_CVM", 
                                        "GFF_GFM", "GFM_CVF", "GFF_CVM")

saveRDS(gfcv_DESeq_results_complete, "RNASeq_gfcv_complete_model_ihw_results_list.rds")
```

#### Simplified model parametrization without interecept

```{r}
gfcv_80_dt <- read.csv("./GFCV_transcriptomics_filtered.csv", row.names = 1)
myGFCVfactors <- read.csv("./RNASeq_gfcv_metadata.csv")

gfcv_DESeq <- DESeqDataSetFromMatrix(countData = gfcv_80_dt,
                                      colData = myGFCVfactors,
                                      design= ~ batch + gfcv + sex_m)
gfcv_DESeq$gfcv <- relevel(gfcv_DESeq$gfcv, ref = "C")
gfcv_DESeq$sex_m <- relevel(gfcv_DESeq$sex_m, ref = "M")

gfcv_DESeq <- DESeq(gfcv_DESeq, test = "Wald")

ctrs_vec <- resultsNames(gfcv_DESeq)
ctrs_vec
gfcv_DESeq_results <- lapply(ctrs_vec[3], function(ctrs){results(gfcv_DESeq, name = ctrs, filterFun = ihw)})
sapply(gfcv_DESeq_results, summary)
names(gfcv_DESeq_results) <- ctrs_vec[3]

saveRDS(gfcv_DESeq_results, "RNASeq_gfcv_no_interaction_model_results_list.rds")
```
