---
title: "Integrative Correlation Analysis"
output:
    html_document:
        toc: true
        toc_depth: 4
date: "2023-02-01"
params:
    corr_threshold: 0.9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(knitr.duplicate.label = "allow")
```

## Required packages

```{r}
library(DESeq2)
library(sva)
library(mixOmics)
library(dplyr)
library(vegan)
library(AnnotationDbi)
library(org.Mm.eg.db)
library(clusterProfiler)
library(enrichplot)
library(compositions)
library(MultiAssayExperiment)
library(knitr)
library(DT)
library(RColorBrewer)
source("../../Utility Functions/utilities.R")
```

## Loading the data

### RNAseq

```{r}
rna_seq_results <- readRDS("../RNA_seq/RNASeq_gfcv_siplified_model_results_list.rds")
rna_seq_full_results <- readRDS("../RNA_seq/RNASeq_gfcv_complete_model_ihw_results_list.rds")
rna_seq_no_int <- readRDS("../RNA_seq/RNASeq_gfcv_no_interaction_model_results_list.rds")

rna_seq_full_results <- c(rna_seq_results, rna_seq_no_int, rna_seq_full_results)

sig_genes <- lapply(rna_seq_full_results, function(ctrs){
    rownames(subset(ctrs ,!is.na(padj) & padj <= 0.1))
})

annotated_results <- lapply(rna_seq_full_results, annotate_data, org.Mm.eg.db)

library(VennDiagram)
venn.diagram(x = sig_genes[1:3], filename = "overlap_sig_genes.png")


rna_seq_data <- read.csv("../RNA_seq/GFCV_transcriptomics_filtered.csv", row.names = 1)
rna_seq_metadata <- read.csv("../RNA_seq/RNASeq_gfcv_metadata.csv", row.names = 1, colClasses = "factor")


gfcv_DESeq <- DESeqDataSetFromMatrix(countData = rna_seq_data,
                                      colData = rna_seq_metadata,
                                      design= ~ batch + gfcv + sex_m + gfcv:sex_m)

rna_seq_data_proc <- vst(gfcv_DESeq, blind = FALSE)
#rna_seq_data_proc <- as.data.frame(assay(rna_seq_data_proc))

colnames(colData(rna_seq_data_proc)) <- c("groups_m", "batch", "sex", "group", "samplename", "sizeFactor")

pca_dt <- pca(X = t(as.data.frame(assay(rna_seq_data_proc))), scale = FALSE)
plotIndiv(pca_dt, group = rna_seq_metadata$groups_m)
```

### DNAepi

```{r}
dna_epi_data <- read.csv("../DNA_met/DNAmet_DMR_data.csv", row.names = 1)

dna_epi_metadata <- read.csv("../DNA_met/DNAmet_metadata.csv", row.names = 1, colClasses = "factor")
dna_epi_data_num <- dna_epi_data[,22:61]

transformToMvalues <- function(Beta) {
  Beta[Beta == 1] <- 0.99
  Beta[Beta == 0] <- 0.01
  M <- log2(Beta / (1 - Beta))
  return(M)
}
dna_epi_data_proc <- transformToMvalues(dna_epi_data_num)

dna_epi_data_proc <- dna_epi_data_proc[,colnames(dna_epi_data_proc) %in% colnames(rna_seq_data_proc)]
dna_epi_metadata_proc <- dna_epi_metadata[dna_epi_metadata$samplename %in% rna_seq_metadata$samplename,]

dna_epi_data_proc <- SummarizedExperiment(assays = dna_epi_data_proc, colData = dna_epi_metadata_proc)
```

### Untargeted metabolomics

#### Data re-organization to match epigenomics/transcriptomics and merging of different results

```{r}
metabo_dirs <- list.dirs("../Untargeted_Metabolomics/Hermes_Results/", recursive = FALSE)

metabo_metadata <- lapply(metabo_dirs, function(col_dir){
    read.csv(paste0(col_dir, "/filtered_metadata.csv"))
})
names(metabo_metadata) <- basename(metabo_dirs)

metabo_results <- lapply(metabo_dirs, function(col_dir){
    readRDS(paste0(col_dir, "/contrasts_results_anova_list.rds"))
})
names(metabo_results) <- basename(metabo_dirs)

metabo_results_merged <- lapply(seq_along(metabo_results[[1]][[1]]), function(ctrs_id){
    ctrs_results <- lapply(names(metabo_results), function(metabo_exp_id){
        metabo_ctrs <- metabo_results[[metabo_exp_id]][[1]][[ctrs_id]]
        rownames(metabo_ctrs) <- paste0(metabo_exp_id, "_", rownames(metabo_ctrs))
        metabo_ctrs
    })
    do.call(rbind, ctrs_results)
})
names(metabo_results_merged) <- names(metabo_results[[1]][[1]])

qcrs_corrected_data_reordered <- lapply(seq_along(metabo_results), function(exp_data_id){
    exp_data <- metabo_results[[exp_data_id]]$qcrsc_corrected_data
    rownames(exp_data) <- paste0(names(metabo_results)[exp_data_id], "_", rownames(exp_data))
    
    colnames(exp_data) <- metabo_metadata[[exp_data_id]]$samplename
    
    exp_data <- exp_data[,match(dna_epi_metadata$samplename,
                                colnames(exp_data))]
    return(exp_data)
})
qcrs_corrected_data_merged <- do.call(rbind, qcrs_corrected_data_reordered)

saveRDS(list("Merged_Metadata" = dna_epi_metadata,
             "Merged_Data" = qcrs_corrected_data_merged),
        "../Untargeted_Metabolomics/merged_gfcv_results.rds")

metabo_metadata <- dna_epi_metadata
metabo_metadata_proc <- dna_epi_metadata_proc

metabo_data <- impute::impute.knn(data = qcrs_corrected_data_merged)$data

metabo_data <- DESeq2::DESeqDataSetFromMatrix(round(metabo_data),
                                              colData = metabo_metadata,
                                              design = ~ batch + sex + gfcv + sex:gfcv)
metabo_data <- estimateSizeFactors(metabo_data)
metabo_data <- estimateDispersions(metabo_data)
metabo_data <- assay(varianceStabilizingTransformation(metabo_data, blind=FALSE))
metabo_data <- ComBat(metabo_data, batch = metabo_metadata$batch)

plotIndiv(pca(t(metabo_data), scale = TRUE), group = metabo_metadata$type_gfcv)

metabo_data_proc <- metabo_data[,colnames(metabo_data) %in% metabo_metadata_proc$samplename]

all(colnames(metabo_data_proc) == metabo_metadata_proc$samplename)

metabo_data_proc <- SummarizedExperiment(metabo_data_proc, colData = metabo_metadata_proc)
```

### Targeted metabolomics

```{r}
targeted_metabo_data <- readRDS("../epi_metabolomics/targeted_metabolomics_data.rds")

sample_names <- lapply(targeted_metabo_data, function(organ_data){
    colData(organ_data)$samplename
})
intersect_samples <- Reduce(intersect, sample_names)

targeted_metabo_data_proc <- lapply(names(targeted_metabo_data), function(organ_data_id){
    organ_data <- targeted_metabo_data[[organ_data_id]]
    rownames(organ_data) <- paste0(organ_data_id, "_", rownames(organ_data))
    organ_data[, organ_data$samplename %in% intersect_samples]
})

targeted_metabo_data_proc <- Reduce(combineRows, targeted_metabo_data_proc)

assay(targeted_metabo_data_proc[grep("liver_scfa", rownames(targeted_metabo_data_proc)),]) <- assay(targeted_metabo_data_proc[grep("liver_scfa", rownames(targeted_metabo_data_proc)),])*1000

targeted_metabo_data_temp <- impute::impute.knn(data = as.matrix(assay(targeted_metabo_data_proc)))$data

targeted_metabo_data_temp <- DESeq2::DESeqDataSetFromMatrix(countData = round(targeted_metabo_data_temp),
                                              colData = colData(targeted_metabo_data_proc),
                                              design = ~ batch + sex + group + sex:group)
targeted_metabo_data_temp <- estimateSizeFactors(targeted_metabo_data_temp)
targeted_metabo_data_temp <- estimateDispersions(targeted_metabo_data_temp)
targeted_metabo_data_temp <- assay(varianceStabilizingTransformation(targeted_metabo_data_temp, blind=FALSE))
targeted_metabo_data_temp <- ComBat(targeted_metabo_data_temp, batch = targeted_metabo_data_proc$batch)
assay(targeted_metabo_data_proc) <- as.data.frame(targeted_metabo_data_temp)

pca_dt <- pca(t(assay(targeted_metabo_data_proc)))
plotIndiv(pca_dt, group = paste(targeted_metabo_data_proc$group, targeted_metabo_data_proc$sex))
```

```{r}
rownames(targeted_metabo_data_proc) <- paste0("targeted_", rownames(targeted_metabo_data_proc))

samples_metabo <- intersect(colnames(targeted_metabo_data_proc), colnames(metabo_data_proc))

full_metabo_data <- lapply(list("Targeted_Data" = targeted_metabo_data_proc, "Untargeted_Data" = metabo_data_proc), function(metabo_data){
    metabo_data[,samples_metabo]
})
full_metabo_data$Untargeted_Data$samplename <- as.character(full_metabo_data$Untargeted_Data$samplename)

full_metabo_data <- c(full_metabo_data[[1]], full_metabo_data[[2]])
full_metabo_data$gfcv <- full_metabo_data$group
```


```{r}
targeted_metabo_res <- readRDS("../epi_metabolomics/targeted_metabolomics_results.rds")

targeted_metabo_res <- lapply(seq_along(targeted_metabo_res[[1]]), function(ctrs_id){
    ctrs_results <- lapply(names(targeted_metabo_res), function(metabo_exp_id){
        metabo_ctrs <- targeted_metabo_res[[metabo_exp_id]][[ctrs_id]]
        rownames(metabo_ctrs) <- paste0("targeted_", rownames(metabo_ctrs))
        metabo_ctrs
    })
    do.call(rbind, ctrs_results)
})
names(targeted_metabo_res) <- names(metabo_results[[1]][[1]])

metabo_results_merged <- lapply(seq_along(targeted_metabo_res), function(ctrs_id){
    rbind(metabo_results_merged[[ctrs_id]], targeted_metabo_res[[ctrs_id]])
})
names(metabo_results_merged) <- names(targeted_metabo_res)
```

### 16Smeta

```{r}
x16S_met_data_list <- readRDS("../16S_meta/16S_data_list.rds")
x16S_met_metadata_list <- readRDS("../16S_meta/16S_metadata_list.rds")


x16S_met_data_list_proc <- lapply(seq_along(x16S_met_data_list), function(organ_id){
    organ <- x16S_met_data_list[[organ_id]] 
    organ_metadata <- x16S_met_metadata_list[[organ_id]]
    lapply(organ, function(category){
        category <- ComBat(category, batch = organ_metadata$batch)
        category <- clr(t(category))
        t(category)        
    })
})
names(x16S_met_data_list_proc) <- names(x16S_met_data_list)

x16S_met_data_list_proc <- lapply(names(x16S_met_data_list_proc), function(organ_id){
    organ_data <- x16S_met_data_list_proc[[organ_id]]
    organ_proc <- lapply(names(organ_data), function(hierarchy_id){
        hierarchy_data <- organ_data[[hierarchy_id]]
        rownames(hierarchy_data) <- paste0(organ_id, "_", hierarchy_id, 
                                        "_", rownames(hierarchy_data))  
        hierarchy_data
    })
    do.call(rbind, organ_proc)
})

intersect_samples <- Reduce(intersect, lapply(x16S_met_data_list_proc, colnames))

x16S_met_data_list_proc <- lapply(seq_along(x16S_met_data_list_proc), function(organ_id){
    SummarizedExperiment(x16S_met_data_list_proc[[organ_id]], colData = x16S_met_metadata_list[[organ_id]])
})

x16S_met_data_list_proc <- lapply(x16S_met_data_list_proc, function(organ_data){
    organ_data[,intersect_samples]
})

x16S_met_data_list_proc <- Reduce(combineRows, x16S_met_data_list_proc)
colData(x16S_met_data_list_proc)$samplename <- rownames(colData(x16S_met_data_list_proc))

x16S_met_data_list_proc$gfcv <- x16S_met_data_list_proc$group
```

### Metallomics data

```{r}
metall_data_cec <- read.csv("../Metallomics/GFCV_cec_metalo_4_stats.csv", row.names = 1)
metall_data_liv <- read.csv("../Metallomics/GFCV_liv_metalo_4_stats.csv", row.names = 1)

intersect_samples <- intersect(rownames(metall_data_cec), rownames(metall_data_liv))

metall_data_cec <- SummarizedExperiment(assays = t(metall_data_cec[,5:ncol(metall_data_cec)]),
                                        colData = metall_data_cec[,1:4])
metall_data_liv <- SummarizedExperiment(assays = t(metall_data_liv[,5:ncol(metall_data_liv)]),
                                        colData = metall_data_liv[,1:4])

rownames(metall_data_cec) <- paste0("cecum_", rownames(metall_data_cec))
rownames(metall_data_liv) <- paste0("liver_", rownames(metall_data_liv))

metall_data_cec <- metall_data_cec[,intersect_samples]
metall_data_liv <- metall_data_liv[,intersect_samples]

metall_data_merged <- rbind(metall_data_cec, metall_data_liv)

metall_data_merged_temp <- DESeq2::DESeqDataSetFromMatrix(round(assay(metall_data_merged)),
                                              colData = colData(metall_data_merged),
                                              design = ~ batch + sex + gfcv + sex:gfcv)
metall_data_merged_temp <- estimateSizeFactors(metall_data_merged_temp)
metall_data_merged_temp <- estimateDispersions(metall_data_merged_temp)
metall_data_merged_temp <- assay(varianceStabilizingTransformation(metall_data_merged_temp, blind=FALSE))
metall_data_merged_temp <- ComBat(metall_data_merged_temp, batch = metall_data_merged$batch)
assay(metall_data_merged) <- as.data.frame(metall_data_merged_temp)

metall_data_merged$samplename <- colnames(metall_data_merged)
```

## Combining all datasets in a multi-assay experiment

```{r}
multi_exp <- MultiAssayExperiment(experiments = list("RNASeq" = rna_seq_data_proc, 
                                      "DNAEpi" = dna_epi_data_proc, 
                                      "Metabolomics" = full_metabo_data, 
                                      "x16S_Metagenomics" = x16S_met_data_list_proc,
                                      "Metallomics" = metall_data_merged))
```

```{r}
saveRDS(multi_exp, file = "./Merged_gnotobiomics.rds")
```


## Pearson Correlations 

```{r}
cor_threshold <- params$corr_threshold
```


### Interaction Genes that only vary in GFM-CVM

We search the genes/metabolites that only change in male by the effect of microbiota.

```{r}
#keep_genes <- intersect(sig_genes$gfcvG.sex_mF, sig_genes$GFM_CVM)
keep_genes <- setdiff(intersect(sig_genes$gfcvG.sex_mF,
                                sig_genes$GFM_CVM),
                    sig_genes$GFF_CVF)

sig_metabolites <- lapply(metabo_results_merged, function(ctrs_data){
    rownames(subset(ctrs_data, !is.na(adj.p.value) & adj.p.value <= 0.1))
})
keep_metabo <- setdiff(intersect(sig_metabolites$Interaction, sig_metabolites$GFM_CVM), sig_metabolites$GFF_CVF)
#keep_metabo <- intersect(sig_metabolites$Interaction, sig_metabolites$GFM_CVM)
```

There are `r length(keep_genes)` significant genes and `r length(keep_metabo)` significant metabolites.

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("RNASeq", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M"]}))

rna_seq_metabo_m_gfcv_data <- multi_exp[c(keep_genes, keep_metabo), 
                                intersect_samples, 
                                c("RNASeq", "Metabolomics")]

all(colnames(assay(experiments(rna_seq_metabo_m_gfcv_data)$RNASeq)) == 
colnames(assay(experiments(rna_seq_metabo_m_gfcv_data)$Metabolomics)))
```

#### Correlations between RNASeq-Metabolites

```{r}
cor_res <- cor(t(assay(experiments(rna_seq_metabo_m_gfcv_data)$RNASeq)), 
               t(assay(experiments(rna_seq_metabo_m_gfcv_data)$Metabolomics)))
cor_gene_met <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)
cor_gene_met <- data.frame("gene" = rownames(cor_res)[cor_gene_met[,1]], "feature" = colnames(cor_res)[cor_gene_met[,2]])
```

##### Results Annotation

```{r}
mapped_genes <- mapIds(x = org.Mm.eg.db, keys = cor_gene_met$gene, column = c("GENENAME"), keytype = "ENSEMBL")

putative_metabolites <- sapply(cor_gene_met$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})

res_annotated_gene_met <- cbind(cor_gene_met, data.frame("Gene_Name" = mapped_genes, 
                                            "Putative_Metabolites" = putative_metabolites))
```

#### Correlations between metabolites and microorganisms

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("x16S_Metagenomics", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M" & exp_data$gfcv == "C"]}))

metabo_meta_m_gfcv_data <- multi_exp[c(keep_metabo, rownames(experiments(multi_exp)$x16S_Metagenomics)), 
                                intersect_samples, 
                                c("x16S_Metagenomics", "Metabolomics")]

all(colnames(assay(experiments(metabo_meta_m_gfcv_data)$x16S_Metagenomics)) == 
colnames(assay(experiments(metabo_meta_m_gfcv_data)$Metabolomics)))

cor_res <- cor(t(assay(experiments(metabo_meta_m_gfcv_data)$Metabolomics)),
               t(as.matrix(assay(experiments(metabo_meta_m_gfcv_data)$x16S_Metagenomics))))

metabo_org_cor <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)

cor_metag_metab <- data.frame("feature" = rownames(cor_res)[metabo_org_cor[,1]], 
                              "microorganism" = colnames(cor_res)[metabo_org_cor[,2]])
```

##### Results Annotation

```{r}
putative_metabolites <- sapply(cor_metag_metab$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})
res_annotated_met_metag <- cbind(cor_metag_metab, putative_metabolites)
```

#### Merged results data

```{r}
rna_met_meta <- left_join(x = res_annotated_gene_met, y = res_annotated_met_metag, "feature", multiple = "all")

rna_met_meta_complete_male_genes <- rna_met_meta[apply(rna_met_meta, 1, function(x){!any(is.na(x))}),]
```

### Interaction Genes that only vary in GFF-CVF

We search the genes/metabolites that change only in females by the effect of the microbiota.

```{r}
keep_genes <- setdiff(intersect(sig_genes$gfcvG.sex_mF,
                                sig_genes$GFF_CVF), sig_genes$GFM_CVM)

sig_metabolites <- lapply(metabo_results_merged, function(ctrs_data){
    rownames(subset(ctrs_data, !is.na(adj.p.value) & adj.p.value <= 0.1))
})
keep_metabo <- setdiff(intersect(sig_metabolites$GFF_CVF, sig_metabolites$Interaction), 
                       sig_metabolites$GFM_CVM)
```

There are `r length(keep_genes)` significant genes and `r length(keep_metabo)` significant metabolites.

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("RNASeq", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "F"]}))

rna_seq_metabo_f_gfcv_data <- multi_exp[c(keep_genes, keep_metabo), 
                                intersect_samples, 
                                c("RNASeq", "Metabolomics")]

all(colnames(assay(experiments(rna_seq_metabo_f_gfcv_data)$RNASeq)) == 
colnames(assay(experiments(rna_seq_metabo_f_gfcv_data)$Metabolomics)))
```

#### Correlations between RNASeq-Metabolites

```{r}
cor_res <- cor(t(assay(experiments(rna_seq_metabo_f_gfcv_data)$RNASeq)), 
               t(assay(experiments(rna_seq_metabo_f_gfcv_data)$Metabolomics)))
cor_gene_met <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)
cor_gene_met <- data.frame("gene" = rownames(cor_res)[cor_gene_met[,1]], "feature" = colnames(cor_res)[cor_gene_met[,2]])

```

##### Results Annotation

```{r}
mapped_genes <- mapIds(x = org.Mm.eg.db, keys = cor_gene_met$gene, column = c("GENENAME"), keytype = "ENSEMBL")
# dmr_information <- lapply(cor_gene_dmr$DMR, function(dmr_region){
#     dna_epi_data[rownames(dna_epi_data) == dmr_region,]
# })
# dmr_information <- do.call(rbind, dmr_information)
putative_metabolites <- sapply(cor_gene_met$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})

res_annotated_gene_met <- cbind(cor_gene_met, data.frame("Gene_Name" = mapped_genes, 
                                            "Putative_Metabolites" = putative_metabolites))
```

#### Correlations between metabolites and microorganisms

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("x16S_Metagenomics", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "F" & exp_data$gfcv == "C"]}))

metabo_meta_f_gfcv_data <- multi_exp[c(keep_metabo, rownames(experiments(multi_exp)$x16S_Metagenomics)), 
                                intersect_samples, 
                                c("x16S_Metagenomics", "Metabolomics")]

all(colnames(assay(experiments(metabo_meta_f_gfcv_data)$x16S_Metagenomics)) == 
colnames(assay(experiments(metabo_meta_f_gfcv_data)$Metabolomics)))

cor_res <- cor(t(assay(experiments(metabo_meta_f_gfcv_data)$Metabolomics)),
               t(as.matrix(assay(experiments(metabo_meta_f_gfcv_data)$x16S_Metagenomics))))

metabo_org_cor <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)

cor_metag_metab <- data.frame("feature" = rownames(cor_res)[metabo_org_cor[,1]], 
                              "microorganism" = colnames(cor_res)[metabo_org_cor[,2]])
```

##### Results Annotation

```{r}

putative_metabolites <- sapply(cor_metag_metab$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})
res_annotated_met_metag <- cbind(cor_metag_metab, putative_metabolites)
```

#### Merged results data

```{r}
rna_met_meta <- left_join(x = res_annotated_gene_met, y = res_annotated_met_metag, "feature", multiple = "all")

rna_met_meta_complete_female_genes <- rna_met_meta[apply(rna_met_meta, 1, function(x){!any(is.na(x))}),]
```

### Genes that are sex independent

We search the genes that vary by the effect of the microbiota and are sex-independent.

```{r}
keep_genes <- setdiff(sig_genes$gfcv_G_vs_C, 
                    sig_genes$gfcvG.sex_mF)

sig_metabolites <- lapply(metabo_results_merged, function(ctrs_data){
    rownames(subset(ctrs_data, !is.na(adj.p.value) & adj.p.value <= 0.1))
})
keep_metabo <- setdiff(sig_metabolites$GF_CV_Full, sig_metabolites$Interaction)
```

There are `r length(keep_genes)` significant genes and `r length(keep_metabo)` significant metabolites.

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("RNASeq", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M"]}))

rna_seq_metabo_m_gfcv_data <- multi_exp[c(keep_genes, keep_metabo), 
                                intersect_samples, 
                                c("RNASeq", "Metabolomics")]

all(colnames(assay(experiments(rna_seq_metabo_m_gfcv_data)$RNASeq)) == 
colnames(assay(experiments(rna_seq_metabo_m_gfcv_data)$Metabolomics)))
```

#### Correlations between RNASeq-Metabolites

```{r}
cor_res <- cor(t(assay(experiments(rna_seq_metabo_m_gfcv_data)$RNASeq)), 
               t(assay(experiments(rna_seq_metabo_m_gfcv_data)$Metabolomics)))
cor_gene_met <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)
cor_gene_met <- data.frame("gene" = rownames(cor_res)[cor_gene_met[,1]], "feature" = colnames(cor_res)[cor_gene_met[,2]])
```

##### Results Annotation

```{r}
mapped_genes <- mapIds(x = org.Mm.eg.db, keys = cor_gene_met$gene, column = c("GENENAME"), keytype = "ENSEMBL")
# dmr_information <- lapply(cor_gene_dmr$DMR, function(dmr_region){
#     dna_epi_data[rownames(dna_epi_data) == dmr_region,]
# })
# dmr_information <- do.call(rbind, dmr_information)
putative_metabolites <- sapply(cor_gene_met$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})

res_annotated_gene_met <- cbind(cor_gene_met, data.frame("Gene_Name" = mapped_genes, 
                                            "Putative_Metabolites" = putative_metabolites))
```

#### Correlations between metabolites and microorganisms

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("x16S_Metagenomics", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M" & exp_data$gfcv == "C"]}))

metabo_meta_m_gfcv_data <- multi_exp[c(keep_metabo, rownames(experiments(multi_exp)$x16S_Metagenomics)), 
                                intersect_samples, 
                                c("x16S_Metagenomics", "Metabolomics")]

all(colnames(assay(experiments(metabo_meta_m_gfcv_data)$x16S_Metagenomics)) == 
colnames(assay(experiments(metabo_meta_m_gfcv_data)$Metabolomics)))

cor_res <- cor(t(assay(experiments(metabo_meta_m_gfcv_data)$Metabolomics)),
               t(as.matrix(assay(experiments(metabo_meta_m_gfcv_data)$x16S_Metagenomics))))

metabo_org_cor <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)

cor_metag_metab <- data.frame("feature" = rownames(cor_res)[metabo_org_cor[,1]], 
                              "microorganism" = colnames(cor_res)[metabo_org_cor[,2]])
```

##### Results Annotation

```{r}

putative_metabolites <- sapply(cor_metag_metab$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})
res_annotated_met_metag <- cbind(cor_metag_metab, putative_metabolites)
```

#### Merged results data

```{r}
rna_met_meta <- left_join(x = res_annotated_gene_met, y = res_annotated_met_metag, "feature", multiple = "all")

rna_met_meta_complete_all_genes <- rna_met_meta[apply(rna_met_meta, 1, function(x){!any(is.na(x))}),]
```

### Genes that vary in male/females and the interaction is significant

We search the genes that vary in male/females and they change differently.

```{r}
keep_genes <- intersect(intersect(sig_genes$GFM_CVM, sig_genes$GFF_GFM), 
                    sig_genes$gfcvG.sex_mF)

sig_metabolites <- lapply(metabo_results_merged, function(ctrs_data){
    rownames(subset(ctrs_data, !is.na(adj.p.value) & adj.p.value <= 0.1))
})
keep_metabo <- intersect(intersect(sig_metabolites$GFF_CVF, sig_metabolites$GFM_CVM), sig_metabolites$Interaction)
```

There are `r length(keep_genes)` significant genes and `r length(keep_metabo)` significant metabolites.

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("RNASeq", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M"]}))

rna_seq_metabo_m_gfcv_data <- multi_exp[c(keep_genes, keep_metabo), 
                                intersect_samples, 
                                c("RNASeq", "Metabolomics")]

all(colnames(assay(experiments(rna_seq_metabo_m_gfcv_data)$RNASeq)) == 
colnames(assay(experiments(rna_seq_metabo_m_gfcv_data)$Metabolomics)))
```

#### Correlations between RNASeq-Metabolites

```{r}
cor_res <- cor(t(assay(experiments(rna_seq_metabo_m_gfcv_data)$RNASeq)), 
               t(assay(experiments(rna_seq_metabo_m_gfcv_data)$Metabolomics)))
cor_gene_met <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)
cor_gene_met <- data.frame("gene" = rownames(cor_res)[cor_gene_met[,1]], "feature" = colnames(cor_res)[cor_gene_met[,2]])
```

##### Results Annotation

```{r}
mapped_genes <- mapIds(x = org.Mm.eg.db, keys = cor_gene_met$gene, column = c("GENENAME"), keytype = "ENSEMBL")
# dmr_information <- lapply(cor_gene_dmr$DMR, function(dmr_region){
#     dna_epi_data[rownames(dna_epi_data) == dmr_region,]
# })
# dmr_information <- do.call(rbind, dmr_information)
putative_metabolites <- sapply(cor_gene_met$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})

res_annotated_gene_met <- cbind(cor_gene_met, data.frame("Gene_Name" = mapped_genes, 
                                            "Putative_Metabolites" = putative_metabolites))
```

#### Correlations between metabolites and microorganisms

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("x16S_Metagenomics", "Metabolomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M" & exp_data$gfcv == "C"]}))

metabo_meta_m_gfcv_data <- multi_exp[c(keep_metabo, rownames(experiments(multi_exp)$x16S_Metagenomics)), 
                                intersect_samples, 
                                c("x16S_Metagenomics", "Metabolomics")]

all(colnames(assay(experiments(metabo_meta_m_gfcv_data)$x16S_Metagenomics)) == 
colnames(assay(experiments(metabo_meta_m_gfcv_data)$Metabolomics)))

cor_res <- cor(t(assay(experiments(metabo_meta_m_gfcv_data)$Metabolomics)),
               t(as.matrix(assay(experiments(metabo_meta_m_gfcv_data)$x16S_Metagenomics))))

metabo_org_cor <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)

cor_metag_metab <- data.frame("feature" = rownames(cor_res)[metabo_org_cor[,1]], 
                              "microorganism" = colnames(cor_res)[metabo_org_cor[,2]])
```

##### Results Annotation

```{r}

putative_metabolites <- sapply(cor_metag_metab$feature, function(feature_name){
    metabo_results_merged$Interaction$PutativeMetabolites[rownames(metabo_results_merged$Interaction) == feature_name]
})
res_annotated_met_metag <- cbind(cor_metag_metab, putative_metabolites)
```

#### Merged results data

```{r}
rna_met_meta <- left_join(x = res_annotated_gene_met, y = res_annotated_met_metag, "feature", multiple = "all")

rna_met_meta_complete_all_vary <- rna_met_meta[apply(rna_met_meta, 1, function(x){!any(is.na(x))}),]

full_correlation_data <- list("Independent genes" = rna_met_meta_complete_all_genes, 
                             "Female only genes" = rna_met_meta_complete_female_genes, 
                             "Male only genes" = rna_met_meta_complete_male_genes,
                             "Female/Male different change" = rna_met_meta_complete_all_vary)

contrs_result_sig_list <- full_correlation_data
contrs_names <- names(contrs_result_sig_list)
```

#### Correlation Results {.tabset}

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = c(
    "##### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`

### Enrichment Results {.tabset}

```{r, cache=TRUE}
dt_to_pathway <- lapply(contrs_result_sig_list, function(ctrs){
    unique(ctrs$gene)
})

enrich_results <- multi_enrich(dt_to_pathway = dt_to_pathway, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05)
enrich_plot_data <- enrich_plots(enrich_results)
```

```{r, results='asis', fig.width=8, fig.height=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### Metagenomics plots {.tabset}

```{r}
pieMetagenomicsPlot <- function(hierarchy_data, hierarchy_name, palette_colors){
    if(nrow(hierarchy_data) == 0){return(NA)}
    pie_plot <- ggplot(hierarchy_data) + 
        geom_bar(aes(x = "", y = Freq, fill = Var1), stat = "identity", color = "black") +
        coord_polar("y") +  theme_void() + xlab(NULL) + ggtitle(hierarchy_name) + 
        scale_fill_manual(values = palette_colors) + 
        theme(legend.title = element_blank(), 
            plot.title = element_text(hjust = 0.5), legend.text = element_text(size = 8))
    pie_plot
}

metag_data_list <- lapply(contrs_result_sig_list, function(ctrs){
    cecum_cor <- ctrs[grepl(pattern = "cecum", ctrs$microorganism),]
    colon_cor <- ctrs[grepl(pattern = "colon", ctrs$microorganism),]
    metag_data_list <- list("Cecum" = cecum_cor, "Colon" = colon_cor)
    metag_data_list <- lapply(metag_data_list, function(org_data){
        
        distinct_data <- distinct(org_data[c("gene", "microorganism")])
        hierarchy_ids <- c("D_0", "D_1", "D_2", "D_3", "D_4", "D_5", "D_6")

        hierarchy_data <- strsplit(org_data$microorganism, ";")
        hierarchy_data <- sapply(hierarchy_data, function(hierarchy_names){
            found_d <- setdiff(grep("D_", hierarchy_names), 
                           grep("uncultured", hierarchy_names, ignore.case = TRUE))
            hierarchy_names[max(found_d)]
        })
        hierarchy_data_list <- lapply(hierarchy_ids, function(id){
            hierarchy <- as.data.frame(table(hierarchy_data[grepl(id,
                                                                hierarchy_data)]))
        })
        names(hierarchy_data_list) <-  c("Kingdom", "Phylum", "Class", 
                                        "Order", "Family", "Genus", "Specie")
        hierarchy_data_list
    })
    metag_data_list
})

palette_fun <- colorRampPalette(brewer.pal(12, "Set3"))
palette_colors_list <- lapply(seq_along(metag_data_list[[1]][[1]]), function(hier_id){
    ctrs_species <- lapply(metag_data_list, function(org_data){
        rbind(org_data[[1]][[hier_id]], org_data[[2]][[hier_id]])
    })
    ctrs_species <- do.call(rbind, ctrs_species)
    unique_species <- unique(ctrs_species$Var1)
    
    ctrs_species_colors <- palette_fun(length(unique_species))
    names(ctrs_species_colors) <- unique_species
    ctrs_species_colors
})

pie_results <- lapply(metag_data_list, function(ctrs){
    lapply(ctrs, function(organ){
        hierarchy_plots <- lapply(seq_along(organ), function(hierarchy_id){
            pieMetagenomicsPlot(organ[[hierarchy_id]], 
                                palette_colors = palette_colors_list[[hierarchy_id]], 
                                hierarchy_name = names(organ)[hierarchy_id]) 
        })
        names(hierarchy_plots) <- names(organ)
        hierarchy_plots
    })
})
```

```{r, results='asis'}
for(i in seq_along(pie_results)){
    cat("#### ", names(pie_results)[i], "{.tabset} \n");
    for(z in seq_along(pie_results[[i]])){
        cat("##### ", names(pie_results[[i]])[z], "{.tabset} \n")
        for(k in seq_along(pie_results[[i]][[z]])){
            cat("###### ", names(pie_results[[i]][[z]])[k], " \n")
            if(!any(is.na(pie_results[[i]][[z]][[k]]))){
                knit_print(pie_results[[i]][[z]][[k]])
                cat("\n")
            } else{
                cat("No correlated species \n")
            }
            cat("\n")
        }
        cat("\n")
    }
}
```


### Metallomics correlations {.tabset}

```{r}
intersect_samples <- Reduce(intersect, 
                            lapply(experiments(multi_exp[,,c("x16S_Metagenomics", "Metallomics")]),
                                   function(exp_data){exp_data$samplename[exp_data$sex == "M" & exp_data$gfcv == "C"]}))

metallo_meta_m_gfcv_data <- multi_exp[, intersect_samples,
                                    c("x16S_Metagenomics", "Metallomics")]

all(colnames(assay(experiments(metallo_meta_m_gfcv_data)$x16S_Metagenomics)) == 
colnames(assay(experiments(metallo_meta_m_gfcv_data)$Metallomics)))

cor_res <- cor(t(assay(experiments(metallo_meta_m_gfcv_data)$Metallomics)),
               t(as.matrix(assay(experiments(metallo_meta_m_gfcv_data)$x16S_Metagenomics))))

metallo_org_cor <- which(abs(cor_res) > cor_threshold, arr.ind = TRUE, useNames = FALSE)

metal_metag_cor <- data.frame("metal" = rownames(cor_res)[metallo_org_cor[,1]], 
                              "microorganism" = colnames(cor_res)[metallo_org_cor[,2]])
```

```{r}
contrs_result_sig_list <- list("Metal_Microorganism_Correlations" = metal_metag_cor)
contrs_names <- names(contrs_result_sig_list)
```

#### Correlation Results {.tabset}

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = c(
    "##### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`

#### Merged Correlations {.tabset}

```{r}
contrs_result_sig_list <- lapply(full_correlation_data, function(ctrs){
    ctrs_merged <- left_join(ctrs, metal_metag_cor, by = "microorganism")
    ctrs_merged[apply(ctrs_merged, 1, function(x){!any(is.na(x))}),]
})
contrs_names <- names(contrs_result_sig_list)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = c(
    "##### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`