---
title: "RNA seq Analysis"
output: html_document
date: "2023-02-01"
params:
    input_dir: "../../inputs/RNA_Sequencing/"
    output_dir: "../../outputs/RNA_Sequencing/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

```{r}
library(AnnotationDbi)
library(org.Mm.eg.db)
library(stringi)
library(DESeq2)
library(dplyr)
library(mixOmics)
library(sva)
library(edgeR)
library(IHW)
library(biomaRt)
library(plotly)
library(ggplot2)

source("../Utility_Functions/utilities.R")
```

## Loading the data

```{r,include=TRUE, results='hide'}
myFactors <- read.csv(paste0(params$input_dir, "MERGED_gfcv&LIMA_metadata.csv"),
                         stringsAsFactors = T,
                         colClasses = "factor" ,
                         row.names = 1)

myCounts <- read.csv(paste0(params$input_dir, "MERGED_gfcv&LIMA_raw_counts.csv"),
                         stringsAsFactors = F,
                         row.names = 1)
```

### Model data frames and metadata
Create 2 different data frames for each model, with counts and metadata
```{r}
#counts GFCV
myGFCVcounts <- dplyr::select(myCounts, -contains("X"))

myGFCVcounts <- as.data.frame(t(myGFCVcounts))

#metadata GFCV
myGFCVfactors <- dplyr::slice(myFactors,37:74)

#write.csv("gfcv_metadata.csv")
```

### Length function
Create a function to calculate trasncript length from GTF file (Sonia Tarazona)
```{r}
gtf2length <-  function (mydir, myfile) {
  
  mygtf = read.delim(file.path(mydir, myfile), header = FALSE, as.is = TRUE)
  
  colnames(mygtf) = c("entity", "start", "end", "gene", "transcript")
  
  mygtf = mygtf[which(mygtf[,"entity"] == "exon"),-1]
  
  mylength = abs(mygtf[,"end"]-mygtf[,"start"]) + 1
  
  trans.length = aggregate(mylength, 
                           by = list("gene" = mygtf[,"gene"],
                                     "transcript" = mygtf[,"transcript"]), sum)
  
  gene.length = aggregate(trans.length[,"x"],
                          by = list("gene" = trans.length[,"gene"]), median)
  
  colnames(gene.length)[2] = "length"

  gene.length
  
}
```

### Transcript length calculation
Calculate transcript length from GTF file
```{r}
mydir <- params$input_dir
myfile <- "mmuGTF.txt"

lengthGTF <- data.frame(gtf2length(mydir,myfile))

lengthGTF <- data.frame(ensembl_gene_id=lengthGTF$gene,
                        lengthGTF=lengthGTF$length)

lengthGTF$ensembl_gene_id <- stri_sub(lengthGTF$ensembl_gene_id, 1,18)
```


### TPM calculation
Calculate TPM
```{r}
tpm <- function(counts, lengths) {
  rate <- counts / lengths
  rate / sum(rate) * 1e6
}

myGFCVcounts_t <- as.data.frame(t(myGFCVcounts))

myGFCVcounts_t$ensembl_gene_id <- rownames(myGFCVcounts_t)

lengthGTF <- dplyr::inner_join(lengthGTF,
                               myGFCVcounts_t,
                               by="ensembl_gene_id")

lengthGTF <- dplyr::select(lengthGTF,
                           ensembl_gene_id,
                           lengthGTF)

myGFCVcounts_t <- dplyr::select(myGFCVcounts_t,
                                -ensembl_gene_id)

rownames(lengthGTF) <- lengthGTF$ensembl_gene_id

lengthGTF <- dplyr::select(lengthGTF, lengthGTF)

dim(lengthGTF)
dim(myGFCVcounts_t)

gfcvTPM <- data.frame(ensembl_gene_id=rownames(myGFCVcounts_t),NA)

for (i in 1:ncol(myGFCVcounts_t)){ 
  sample <- myGFCVcounts_t[,i,drop=T]
  TPM <- tpm(myGFCVcounts_t[,i],lengthGTF)
  gfcvTPM[,i] <- cbind(data.frame(TPM))
  colnames(gfcvTPM) <- colnames(myGFCVcounts_t[,i])
}

colnames(gfcvTPM) <- colnames(myGFCVcounts_t)

rownames(gfcvTPM) <- rownames(myGFCVcounts_t)
```


### Filtering 80% and TPM > 0.1

Filtering out genes with TPM > 0.1, output in counts 

```{r}
# filtering 80% from TPM
dim(gfcvTPM)

gfcvTPMall_t <- as.data.frame(t(gfcvTPM))

dim(gfcvTPMall_t)

myGFCVfactors$group <- factor(myGFCVfactors$groups_m, 
                                 levels = c("CVM", "CVF", 
                                            "GFM", "GFF"))
myGFCVfactors$sex <- factor(myGFCVfactors$sex_m, 
                                 levels = c("M", "F"))
myGFCVfactors$gfcv <- factor(myGFCVfactors$gfcv, 
                                 levels = c("C", "G"))
    
groups <- levels(myGFCVfactors$group)
cutoff <- 0.2
thresholds <- round(table(myGFCVfactors$group) * cutoff)

selected_cols <- lapply(groups, function(i){
  gr <- rownames(myGFCVfactors[myGFCVfactors$group == i,])
  ceros <- gfcvTPMall_t[gr,] == 0
  selc_gr <- which(colSums(ceros) < thresholds[i])
  return(selc_gr)
})

selected_index <- unique(unlist(selected_cols))
myGFCVcountsTPM_80 <- gfcvTPMall_t[,selected_index]

dim(myGFCVcountsTPM_80)
myGFCVcountsTPM_80 <- as.data.frame(t(myGFCVcountsTPM_80))
dim(myGFCVcountsTPM_80)

# Filtering TPM>0.1 after the 80% filter

dim(myGFCVcountsTPM_80)
myGFCVcountsTPM_80 <- myGFCVcountsTPM_80 %>% dplyr::mutate(aver = rowSums(.)/38)

myGFCVcountsTPM_80ord <- dplyr::arrange(myGFCVcountsTPM_80, aver)

dim(myGFCVcountsTPM_80ord)
  
myGFCVcountsTPM_80ord_f <- dplyr::filter(myGFCVcountsTPM_80ord, aver > 1)

dim(myGFCVcountsTPM_80ord_f) # same results, the order does not matter

myGFCVcountsTPM_80ord_f <- dplyr::select(myGFCVcountsTPM_80ord_f, -aver)

dim(myGFCVcountsTPM_80ord_f)
```

### 80% and 0.1 filtered, raw counts

Only keep genes that appear in 80% of the samples in at least one group. 

```{r}
myGFCVcountsTPM_80_f_id <- data.frame(ensembl_gene_id=rownames(myGFCVcountsTPM_80ord_f))


dim(myGFCVcounts_t)
myGFCVcounts_t$ensembl_gene_id <- rownames(myGFCVcounts_t)

# keep raw counts for genes with 80% filter
myGFCVcounts80 <- dplyr::inner_join(myGFCVcountsTPM_80_f_id,
                                    myGFCVcounts_t,
                                    by="ensembl_gene_id")

dim(myGFCVcounts80)

rownames(myGFCVcounts80) <- myGFCVcounts80$ensembl_gene_id

myGFCVcounts80 <- dplyr::select(myGFCVcounts80, -ensembl_gene_id)

#Eliminate genes with variance 0
# dim(myGFCVcounts80)
# gene_var <- data.frame(gene_var=apply(myGFCVcounts80, 1, var))
# dim(gene_var)
# gene_var_id <- dplyr::filter(gene_var,gene_var>0) # Michael Love against that
# dim(gene_var_id)



##################################################
##################################################
##### Matrix to use for DESeq2 and downstream# ###
                dim(myGFCVcounts80)############# #
##################################################
##################################################

#write_xlsx(myGFCVcounts80, "myGFCVcounts80.xlsx")
# write.csv(myGFCVcounts80, file = "GFCV_transcriptomics_filtered.csv", 
#           row.names = TRUE)

dim(myGFCVcounts80)

myGFCVcounts80_t <- as.data.frame(t(myGFCVcounts80))

dim(myGFCVcounts80_t)


# p_rna_raw <-mixOmics::pca(myGFCVcounts80_t, ncomp = 6, center = T, scale = T,
#          max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
#          V = NULL, multilevel = NULL)
# 
# 
# pdf("GFCV_liv_PCAmixomics_raw80_jun2022.pdf")
# 
# plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$gfcv, legend = T, title= "rna tpm liv by gfcv")
# 
# plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$group, legend = T, title= "rna tpm liv by group",ellipse = T, ellipse.level = 0.99)
# 
# plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$sex, legend = T, title= "rna tpm liv by sex",ellipse = T, ellipse.level = 0.99)
# 
# plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$batch,comp = c(1,2), legend = T, title= "rna tpm liv by batch",ellipse = T, ellipse.level = 0.99)
# plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$batch,comp = c(1,3), legend = T, title= "rna tpm liv by batch",ellipse = T, ellipse.level = 0.99)
# plotIndiv(p_rna_raw, ind.names = T, group = myGFCVfactors$batch,comp = c(2,3), legend = T, title= "rna tpm liv by batch",ellipse = T, ellipse.level = 0.99)
# 
# dev.off()
# 
# metapei_id <- data.frame(ensembl_gene_id=rownames(myGFCVcounts80))
# 
# 
# save(metapei_id, file="GFCV_metaepi1_20k_ensembl_gene_id_80filter.Rdata")

#write_xlsx(metapei_id, "GFCV_metaepi1_20k_ensembl_gene_id_80filter.xlsx")

```

### ComBat batch correction

There is a mild batch effect that we already saw in other omics and we opted to correct this batch effect with the same algorithm (ComBat)

```{r}
# Combat on 80% and 0.1 filtered data
dim(myGFCVcounts80)
class(myGFCVcounts80)

myGFCVcounts80_m <- as.matrix(myGFCVcounts80)

myGFCVcounts80_CB <-  ComBat(myGFCVcounts80_m, myGFCVfactors$batch)

sum(table(myGFCVcounts80_CB[myGFCVcounts80_CB<0]))


myGFCVcounts80_CB[myGFCVcounts80_CB<0] <- 0


sum(table(myGFCVcounts80_CB[myGFCVcounts80_CB<0]))


myGFCVcounts80_CB <- as.data.frame(round(myGFCVcounts80_CB, digits = 0))
myGFCVcounts80_CB <- as.data.frame(t(myGFCVcounts80_CB))


dim(myGFCVcounts80_CB)
dim(myGFCVcounts80_t)

```

```{r}
### Raw
raw_lima_DESeq <- DESeqDataSetFromMatrix(countData = t(myGFCVcounts),
                                      colData = myGFCVfactors,
                                      design= ~ 0 + group)

vst_dt <- t(assay(vst(raw_lima_DESeq, blind = FALSE)))
vst_pca <- prcomp(vst_dt, scale. = FALSE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
        group_var = myGFCVfactors$group, 
        ellipse_var = myGFCVfactors$group, allElipse = TRUE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
         group_var = myGFCVfactors$batch, 
         ellipse_var = myGFCVfactors$batch, allElipse = TRUE)

pcvar <- round(vst_pca$sdev^2/sum(vst_pca$sdev^2)*100, 2)

d3plot <- plot_ly(x = vst_pca$x[,1], y = vst_pca$x[,2], z = vst_pca$x[,3], color = myGFCVfactors$group)
d3plot <- d3plot %>% add_markers()
d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")),
                     yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")),
                     zaxis = list(title = paste0('PC3', "(", pcvar[3], "%)"))))
d3plot

##################
## Filtered
f80_lima_DESeq <- DESeqDataSetFromMatrix(countData = myGFCVcounts80,
                                      colData = myGFCVfactors,
                                      design= ~ 0 + group)
vst_dt <- t(assay(vst(f80_lima_DESeq, blind = FALSE)))
vst_pca <- prcomp(vst_dt, scale. = FALSE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
         group_var = myGFCVfactors$group, 
         ellipse_var = myGFCVfactors$group, allElipse = TRUE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
         group_var = myGFCVfactors$batch, 
         ellipse_var = myGFCVfactors$batch, allElipse = TRUE)

pcvar <- round(vst_pca$sdev^2/sum(vst_pca$sdev^2)*100, 2)

d3plot <- plot_ly(x = vst_pca$x[,1], y = vst_pca$x[,2], z = vst_pca$x[,3],
                    color = myGFCVfactors$group)
d3plot <- d3plot %>% add_markers()
d3plot <- d3plot %>% layout(scene = list(
    xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")),
    yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")),
    zaxis = list(title = paste0('PC3', "(", pcvar[3], "%)"))))
d3plot

## Batch corrected
f80_lima_DESeq <- DESeqDataSetFromMatrix(countData = t(myGFCVcounts80_CB),
                                      colData = myGFCVfactors,
                                      design= ~ 0 + group)
vst_dt <- t(assay(vst(f80_lima_DESeq, blind = FALSE)))
vst_pca <- prcomp(vst_dt, scale. = FALSE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
         group_var = myGFCVfactors$group, 
         ellipse_var = myGFCVfactors$group, allElipse = TRUE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
         group_var = myGFCVfactors$batch, 
         ellipse_var = myGFCVfactors$batch, allElipse = TRUE)

pcvar <- round(vst_pca$sdev^2/sum(vst_pca$sdev^2)*100, 2)

d3plot <- plot_ly(x = vst_pca$x[,1], y = vst_pca$x[,2], z = vst_pca$x[,3],
                    color = myGFCVfactors$group)
d3plot <- d3plot %>% add_markers()
d3plot <- d3plot %>% layout(scene = list(
    xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")),
    yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")),
    zaxis = list(title = paste0('PC3', "(", pcvar[3], "%)"))))
d3plot
```


## GFCV DESeq2

### DESeq

<!-- #### Simplified model parametrization -->

<!-- ```{r} -->
<!-- gfcv_80_dt <-  -->


<!-- gfcv_DESeq <- DESeqDataSetFromMatrix(countData = gfcv_80_dt, -->
<!--                                      colData = myGFCVfactors, -->
<!--                                      design = ~ batch + gfcv + sex + gfcv:sex) -->
<!-- gfcv_DESeq$gfcv <- relevel(gfcv_DESeq$gfcv, ref = "C") -->
<!-- gfcv_DESeq$sex <- relevel(gfcv_DESeq$sex, ref = "M") -->

<!-- gfcv_DESeq <- DESeq(gfcv_DESeq, test = "Wald") -->

<!-- ctrs_vec <- resultsNames(gfcv_DESeq) -->
<!-- ctrs_vec -->
<!-- gfcv_DESeq_results <- lapply(ctrs_vec[5], function(ctrs){results(gfcv_DESeq, name = ctrs, filterFun = ihw)}) -->
<!-- sapply(gfcv_DESeq_results, summary) -->
<!-- names(gfcv_DESeq_results) <- ctrs_vec[5] -->

<!-- saveRDS(gfcv_DESeq_results, "RNASeq_gfcv_siplified_model_results_list.rds") -->
<!-- ``` -->

#### Complete model parametrization

```{r}
gfcv_80_exp <- SummarizedExperiment(assays = myGFCVcounts80, 
                                    colData = myGFCVfactors)


gfcv_DESeq <- DESeqDataSetFromMatrix(countData = assay(gfcv_80_exp),
                                      colData = colData(gfcv_80_exp),
                                      design= ~ 0 + group + batch)

gfcv_DESeq <- DESeq(gfcv_DESeq, test = "Wald")

ctrs_vec <- resultsNames(gfcv_DESeq)
ctrs_vec_list <- list(c("group", "GFM", "CVM"),c("group", "GFF", "CVF"), 
                      c("group", "CVF", "CVM"), c("group", "GFF", "GFM"),
                      c("group", "GFM", "CVF"), c("group", "GFF", "CVM"))

gfcv_DESeq_results_complete <- lapply(ctrs_vec_list, function(ctrs){
    results(gfcv_DESeq, contrast = ctrs)
})
lapply(gfcv_DESeq_results_complete, summary)

names(gfcv_DESeq_results_complete) <- c("GFM_CVM", "GFF_CVF", "CVF_CVM", 
                                        "GFF_GFM", "GFM_CVF", "GFF_CVM")

# saveRDS(gfcv_DESeq_results_complete, "RNASeq_gfcv_complete_model_ihw_results_list.rds")
```

#### Simplified model parametrization without interecept

```{r}
gfcv_DESeq <- DESeqDataSetFromMatrix(countData = assay(gfcv_80_exp),
                                      colData = colData(gfcv_80_exp),
                                      design= ~ batch + gfcv + sex)
gfcv_DESeq$gfcv <- relevel(gfcv_DESeq$gfcv, ref = "C")
gfcv_DESeq$sex <- relevel(gfcv_DESeq$sex, ref = "M")

gfcv_DESeq <- DESeq(gfcv_DESeq, test = "Wald")

ctrs_vec <- resultsNames(gfcv_DESeq)
ctrs_vec

gfcv_DESeq_results <- lapply(ctrs_vec[3], function(ctrs){results(gfcv_DESeq, name = ctrs)})
sapply(gfcv_DESeq_results, summary)
names(gfcv_DESeq_results) <- "GF_CV"
```

```{r}
gfcv_DESeq_results <- lapply(gfcv_DESeq_results, function(ctrs_res){
    colnames(ctrs_res)[c(2,5:6)] <- c("log2FC","p.value", "adj.p.value")
    ctrs_res
})

gfcv_DESeq_results_complete <- lapply(gfcv_DESeq_results_complete, function(ctrs_res){
    colnames(ctrs_res)[c(2,5:6)] <- c("log2FC","p.value", "adj.p.value")
    ctrs_res
})

rowData(gfcv_80_exp) <- cbind(do.call(cbind, gfcv_DESeq_results), 
                              do.call(cbind, gfcv_DESeq_results_complete))

rowData(gfcv_80_exp)$gene_name <- mapIds(org.Mm.eg.db, keys = rownames(gfcv_80_exp), 
                                    column = c("GENENAME"), 
                         keytype = "ENSEMBL", multiVals = "first")
rowData(gfcv_80_exp)$entrezID <- mapIds(org.Mm.eg.db, keys = rownames(gfcv_80_exp), 
                                    column = c("ENTREZID"), 
                         keytype = "ENSEMBL", multiVals = "first")
rowData(gfcv_80_exp)$symbol <- mapIds(org.Mm.eg.db, keys = rownames(gfcv_80_exp), 
                                    column = c("SYMBOL"), 
                         keytype = "ENSEMBL", multiVals = "first")
rowData(gfcv_80_exp)$gene_lab <- ifelse(!is.na(rowData(gfcv_80_exp)$gene_name),
                                   rowData(gfcv_80_exp)$gene_name, 
                                   rownames(gfcv_80_exp))

saveRDS(gfcv_80_exp, paste0(params$output_dir, "RNA_Sequencing_exp_gfcv.rds"))
```

