---
title: "RNA Seq LIMA Enrichments"
output: html_document
date: "2023-04-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(knitr.duplicate.label = "allow")
```

```{r}
library(AnnotationDbi)
library(org.Mm.eg.db)
library(DESeq2)
library(sva)
library(mixOmics)
library(knitr)
library(DT)
library(clusterProfiler)
library(dplyr)
library(msigdbr)
library(MultiAssayExperiment)
source("../../Utility Functions/utilities.R")
```

## Loading the RNASeq results

```{r}
p_val_thr <- 0.05
rna_seq_results <- readRDS("../Multi_Omics_Analysis/Multivariate_Merged_Multi_Omics.rds")
rna_seq_results <- experiments(rna_seq_results)[["RNASeq.liver"]]

contrs_names <- colnames(rowData(rna_seq_results))
contrs_names <- gsub(".adj.p.value", "", 
                    contrs_names[grep("adj.p.value", contrs_names)])

sig_tables <- lapply(contrs_names, function(ctrs){
    ctrs_col <- paste0(ctrs, ".adj.p.value")
    ctrs_exprs <- parse(text = paste0("!is.na(`", ctrs_col, "`) & `", 
                                       ctrs_col, "` <= ", p_val_thr))
    row_dt <- rowData(rna_seq_results)
    
    subset(row_dt, eval(ctrs_exprs, envir = row_dt))
})
names(sig_tables) <- contrs_names

sig_genes <- lapply(sig_tables, function(ctrs){
    rownames(ctrs)
})

contrs_result_sig_list <- lapply(sig_tables, annotate_data, annotationPackage = org.Mm.eg.db)


genes_gsea <- lapply(contrs_names, function(ctrs_name){
    ctrs <- rowData(rna_seq_results)
    fc_list <- ctrs[[paste0(ctrs_name, ".log2FC")]]
    names(fc_list) <- rownames(ctrs)
    fc_list
})
names(genes_gsea) <- contrs_names
```

# Enrichment analysis {.tabset}

## Enrichment results full lists {.tabset}

### Table results {.tabset}

```{r}
contrs_result_sig_list <- lapply(names(contrs_result_sig_list), function(ctrs_name){
    ctrs <- contrs_result_sig_list[[ctrs_name]]
    as.data.frame(ctrs[order(ctrs[[grep(pattern = paste0(ctrs_name, ".adj.p.value"), 
                                    colnames(ctrs))]]),])
})
names(contrs_result_sig_list) <- contrs_names

tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE)) %>% formatRound(columns = colnames(contrs_result_sig_list[[{{i}}]])[sapply(contrs_result_sig_list[[{{i}}]], is.numeric)], digits = 3)",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`

### ORA {.tabset}

```{r, cache=TRUE}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify_res = TRUE)

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_gsea)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### GSEA {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = genes_gsea, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify_res = TRUE, enrichFun = "gsea")

enrich_results <- lapply(enrich_results, function(ctrs){
    lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
})

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_gsea)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```


### KEGG {.tabset}

```{r}
genes_entrez_ids <- lapply(genes_gsea, function(ctrs){
    names(ctrs) <- mapIds(x = org.Mm.eg.db, names(ctrs), column = "ENTREZID", keytype = "ENSEMBL")
    ctrs[names(ctrs)[!is.na(names(ctrs))]]
})

set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               keggOrg = "mmu", keggKeyType = "kegg",
                               pvalueCutoff = 0.05, simplify_res = TRUE, enrichFun = "kegg")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```


### WP {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr",  wpOrg = "Mus musculus",
                               pvalueCutoff = 0.05, simplify_res = TRUE, enrichFun = "wp")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### MSIGDBR ORA {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr",  
                               msigdbCategory = "H", msigdbSpc = "Mus musculus",
                               pvalueCutoff = 0.05, simplify_res = TRUE, 
                               enrichFun = "msigdbrORA")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### MSIGDBR GSEA {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = genes_gsea, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               msigdbCategory = "H", msigdbSpc = "Mus musculus",
                               pvalueCutoff = 0.05, simplify_res = TRUE, 
                               enrichFun = "msigdbrGSEA")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

## Enrichment results subsets {.tabset}

```{r}
expression_list <- list(
    "HFD Altered" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05)
    ),
        "INT Restored Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `INT-CNT.adj.p.value` > 0.05 & `HFD-CNT.log2FC` > 0
    ), 
    "INT Restored Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `INT-CNT.adj.p.value` > 0.05 & `HFD-CNT.log2FC` < 0
    ),
    "INT Restored Over Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` > 0 &
        `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` > 0
    ),
    "INT Restored Over Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` > 0 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0 &
        `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` < 0
    ),
    "INT Resistant Over Up" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` > 0 & 
        `INT-CNT.adj.p.value` <= 0.05 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` > 0)
    ),
    "INT Resistant Over Down" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
        `INT-CNT.adj.p.value` <= 0.05 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0)
    ),
    "INT Resistant Up" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` > 0 & 
        `INT-CNT.adj.p.value` <= 0.05 & 
        `INT-HFD.adj.p.value` > 0.05) 
    ),
    "INT Resistant Down" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
        `INT-CNT.adj.p.value` <= 0.05 &
        `INT-HFD.adj.p.value` > 0.05)
    ),
    "INT Resistant Half Up" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` > 0 & 
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0 & 
        `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` > 0)
    ),
    "INT Resistant Half Down" = quote(
        (`HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` > 0 &
        `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` < 0)
    ),
    "HFE Unnaffected" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` > 0.05 & `HFE-HFD.adj.p.value` <= 0.05 
    ),
    "HFE Affected" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` <= 0.05 & `HFE-HFD.adj.p.value` > 0.05
    ),
    "HFE Other" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & ((`HFE-CNT.adj.p.value` <= 0.05 & `HFE-HFD.adj.p.value` <= 0.05) |
            (`HFE-CNT.adj.p.value` > 0.05 & `HFE-HFD.adj.p.value` > 0.05))
    )
)
```

```{r}
expression_list_sub <- list("Restored Genes" = grep("Restored", names(expression_list), value = TRUE), 
                            "Resistant Genes" = grep("Resistant", names(expression_list), value = TRUE))
# hfd_affected_cols <- grep("Affected", names(expression_list))


condition_OR <- function(eval_exprs){
    Reduce(f = function(x, y){expr(!!x | !!y)},
           x = eval_exprs)    
}

expression_list_sub <- lapply(expression_list_sub, function(eval_exprs){
    eval_exprs <- expression_list[eval_exprs]
    condition_OR(eval_exprs)
})

expression_list_sub$`HFE Unnaffected Genes` <- expression_list[[grep("HFE Unnaffected", names(expression_list), value = TRUE)]]
expression_list_sub$`HFE Affected Genes` <- expression_list[[grep("HFE Affected", names(expression_list), value = TRUE)]]


sig_tables <- lapply(expression_list_sub, function(ctrs){
    row_dt <- rowData(rna_seq_results)
    subset(row_dt, eval(ctrs, envir = row_dt))
})

sig_genes <- lapply(sig_tables, function(ctrs){
    rownames(ctrs)
})

contrs_result_sig_list <- lapply(sig_tables, annotate_data, annotationPackage = org.Mm.eg.db)
```


### Table results {.tabset}

```{r}
contrs_result_sig_list <- lapply(contrs_result_sig_list, 
                                function(x){as.data.frame(x)})

tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names[c(1,1,5,3)], template = c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE)) %>% formatRound(columns = colnames(contrs_result_sig_list[[{{i}}]])[sapply(contrs_result_sig_list[[{{i}}]], is.numeric)], digits = 3)",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`

### ORA {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify_res = TRUE)
enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_gsea[contrs_names[c(1,1,5,3)]])
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### KEGG {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               keggOrg = "mmu", keggKeyType = "kegg",
                               pvalueCutoff = 0.05, simplify_res = TRUE, enrichFun = "kegg")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids[contrs_names[c(1,1,5,3)]])
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### WP {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr",  wpOrg = "Mus musculus",
                               pvalueCutoff = 0.05, simplify_res = TRUE, enrichFun = "wp")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids[contrs_names[c(1,1,5,3)]])
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### MSIGDBR ORA {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr",  
                               msigdbCategory = "H", msigdbSpc = "Mus musculus",
                               pvalueCutoff = 0.05, simplify_res = TRUE, 
                               enrichFun = "msigdbrORA")

# enrich_results <- lapply(enrich_results, function(ctrs){
#     lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
# })

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_entrez_ids)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

## Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```