---
title: "LIMA model analysis"
output: html_document
date: "2023-04-12"
params:
    input_dir: "../../inputs/RNA_Sequencing/"
    output_dir: "../../outputs/RNA_Sequencing/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

```{r}
library(AnnotationDbi)
library(org.Mm.eg.db)
library(stringi)
library(DESeq2)
library(dplyr)
library(mixOmics)
library(sva)
library(edgeR)
library(IHW)
library(biomaRt)
library(plotly)
library(ggplot2)
source("../Utility_Functions/utilities.R")
```

## Loading the data

```{r,include=TRUE, results='hide'}
myFactors <- read.csv(paste0(params$input_dir, "MERGED_gfcv&LIMA_metadata.csv"),
                         stringsAsFactors = T,
                         colClasses = "factor" ,
                         row.names = 1)

myCounts <- read.csv(paste0(params$input_dir, "MERGED_gfcv&LIMA_raw_counts.csv"),
                         stringsAsFactors = F,
                         colClasses = c("character",rep("numeric",74)),
                         row.names = 1)
```

### Model data frames and metadata
Create 2 different data frames for each model, with counts and metadata
```{r}
#counts GFCV
myLIMAcounts <- dplyr::select(myCounts, contains("X"))

myLIMAcounts <- as.data.frame(t(myLIMAcounts))

#metadata GFCV
myLIMAfactors <- subset(myFactors, gfcv == "")

all(gsub("X", "", rownames(myLIMAcounts)) == myLIMAfactors$samplename)
myLIMAfactors$samplename_old <- myLIMAfactors$samplename
myLIMAfactors$samplename <- make.names(myLIMAfactors$samplename)

rownames(myLIMAfactors) <- myLIMAfactors$samplename

# write.csv(myLIMAfactors, file = "RNASeq_lima_metadata.csv")
```

### Length function

Create a function to calculate transcript length from GTF file (Sonia Tarazona)

```{r}
gtf2length <-  function (mydir, myfile) {
  
  mygtf = read.delim(file.path(mydir, myfile), header = FALSE, as.is = TRUE)
  
  colnames(mygtf) = c("entity", "start", "end", "gene", "transcript")
  
  mygtf = mygtf[which(mygtf[,"entity"] == "exon"),-1]
  
  mylength = abs(mygtf[,"end"]-mygtf[,"start"]) + 1
  
  trans.length = aggregate(mylength, 
                           by = list("gene" = mygtf[,"gene"],
                                     "transcript" = mygtf[,"transcript"]), sum)
  
  gene.length = aggregate(trans.length[,"x"],
                          by = list("gene" = trans.length[,"gene"]), median)
  
  colnames(gene.length)[2] = "length"

  gene.length
  
}
```

### Transcript length calculation
Calculate transcript length from GTF file
```{r}
mydir <- "./"
myfile <- paste0(params$input_dir, "mmuGTF.txt")

lengthGTF <- data.frame(gtf2length(mydir, myfile))

lengthGTF <- data.frame(ensembl_gene_id=lengthGTF$gene,
                        lengthGTF=lengthGTF$length)

lengthGTF$ensembl_gene_id <- stri_sub(lengthGTF$ensembl_gene_id, 1,18)
```


### TPM calculation

Calculate TPM

```{r}
tpm <- function(counts, lengths) {
  rate <- counts / lengths
  rate / sum(rate) * 1e6
}

myLIMAcounts_t <- as.data.frame(t(myLIMAcounts))

myLIMAcounts_t$ensembl_gene_id <- rownames(myLIMAcounts_t)

lengthGTF <- dplyr::inner_join(lengthGTF,
                               myLIMAcounts_t,
                               by="ensembl_gene_id")

lengthGTF <- dplyr::select(lengthGTF,
                           ensembl_gene_id,
                           lengthGTF)

myLIMAcounts_t <- dplyr::select(myLIMAcounts_t,
                                -ensembl_gene_id)

rownames(lengthGTF) <- lengthGTF$ensembl_gene_id

lengthGTF <- dplyr::select(lengthGTF, lengthGTF)

dim(lengthGTF)
dim(myLIMAcounts_t)

limaTPM <- data.frame(ensembl_gene_id=rownames(myLIMAcounts_t),NA)

for (i in seq_len(ncol(myLIMAcounts_t))){ 
  TPM <- tpm(myLIMAcounts_t[,i],lengthGTF)
  limaTPM[,i] <- cbind(data.frame(TPM))
  colnames(limaTPM) <- colnames(myLIMAcounts_t[,i])
}

colnames(limaTPM) <- colnames(myLIMAcounts_t)

rownames(limaTPM) <- rownames(myLIMAcounts_t)
```


### Filtering 80% and TPM > 0.1

Filtering out genes with TPM > 0.1, output in counts 

```{r}
# filtering 80% from TPM
dim(limaTPM)

limaTPMall_t <- as.data.frame(t(limaTPM))

dim(limaTPMall_t)

myLIMAfactors$groups_m <- factor(myLIMAfactors$groups_m, 
                                 levels = c("CNT","HFD", 
                                            "INT", "HFE"))
myLIMAfactors$group <- factor(myLIMAfactors$groups_m, 
                              levels = c("CNT", "HFD", "INT", "HFE"))

groups <- levels(myLIMAfactors$groups_m)
cutoff <- 0.2
thresholds <- round(table(myLIMAfactors$groups_m) * cutoff)
selected_cols <- lapply(groups, function(i){
  gr <- rownames(myLIMAfactors[myLIMAfactors$groups_m == i,])
  ceros <- limaTPMall_t[gr,] == 0
  selc_gr <- which(colSums(ceros) < thresholds[i])
  return(selc_gr)
})
selected_index <- unique(unlist(selected_cols))
myLIMAcountsTPM_80 <- limaTPMall_t[,selected_index]

dim(myLIMAcountsTPM_80)
myLIMAcountsTPM_80 <- as.data.frame(t(myLIMAcountsTPM_80))
dim(myLIMAcountsTPM_80)

# Filtering TPM>0.1 after the 80% filter

dim(myLIMAcountsTPM_80)
myLIMAcountsTPM_80 <- myLIMAcountsTPM_80 %>% dplyr::mutate(aver = rowSums(.)/ncol(myLIMAcountsTPM_80))

myLIMAcountsTPM_80ord <- dplyr::arrange(myLIMAcountsTPM_80, aver)

dim(myLIMAcountsTPM_80ord)
  
myLIMAcountsTPM_80ord_f <- dplyr::filter(myLIMAcountsTPM_80ord, aver > 0.1)

dim(myLIMAcountsTPM_80ord_f) # same results, the order does not matter

myLIMAcountsTPM_80ord_f <- dplyr::select(myLIMAcountsTPM_80ord_f, -aver)

dim(myLIMAcountsTPM_80ord_f)
```

```{r}
myLIMAcountsTPM_80_f_id <- data.frame(ensembl_gene_id=rownames(myLIMAcountsTPM_80ord_f))


dim(myLIMAcounts_t)
myLIMAcounts_t$ensembl_gene_id <- rownames(myLIMAcounts_t)



# keep raw counts for genes with 80% filter
myLIMAcounts80 <- dplyr::inner_join(myLIMAcountsTPM_80_f_id,
                                    myLIMAcounts_t,
                                    by="ensembl_gene_id")

dim(myLIMAcounts80)

rownames(myLIMAcounts80) <- myLIMAcounts80$ensembl_gene_id

myLIMAcounts80 <- dplyr::select(myLIMAcounts80, -ensembl_gene_id)


#Eliminate genes with variance 0
# dim(myLIMAcounts80)
# gene_var <- data.frame(gene_var=apply(myLIMAcounts80, 1, var))
# dim(gene_var)
# gene_var_id <- dplyr::filter(gene_var,gene_var>0) # Michael Love against that
# dim(gene_var_id)

##################################################
##################################################
##### Matrix to use for DESeq2 and downstream ####
                dim(myLIMAcounts80)###############
##################################################
##################################################

#write_xlsx(myLIMAcounts80, "myLIMAcounts80.xlsx")

rna_seq_exp <- SummarizedExperiment(assays = myLIMAcounts80, 
                                    colData = myLIMAfactors)

# saveRDS(rna_seq_exp, paste0(params$output_dir, "rna_seq_lima_proc.rds"))
# dim(myLIMAcounts80)
# 
# myLIMAcounts80_t <- as.data.frame(t(myLIMAcounts80))
# 
# dim(myLIMAcounts80_t)
# 
# 
# p_rna_raw <-mixOmics::pca(myLIMAcounts80_t, ncomp = 6, center = T, scale = T,
#          max.iter = 500, tol = 1e-09, logratio = 'none',   ilr.offset = 0.001,
#          V = NULL, multilevel = NULL)
# 
# metapei_id <- data.frame(ensembl_gene_id=rownames(myLIMAcounts80))
# 
# 
# save(metapei_id, file="GFCV_metaepi1_20k_ensembl_gene_id_80filter.Rdata")

#write_xlsx(metapei_id, "GFCV_metaepi1_20k_ensembl_gene_id_80filter.xlsx")

```

```{r}
raw_lima_DESeq <- DESeqDataSetFromMatrix(countData = t(myLIMAcounts),
                                      colData = myLIMAfactors,
                                      design= ~ 0 + groups_m)

vst_dt <- t(assay(vst(raw_lima_DESeq, blind = FALSE)))
vst_pca <- prcomp(vst_dt, scale. = FALSE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
        group_var = myLIMAfactors$groups_m, 
        ellipse_var = myLIMAfactors$groups_m, allElipse = TRUE)

pcvar <- round(vst_pca$sdev^2/sum(vst_pca$sdev^2)*100, 2)

d3plot <- plot_ly(x = vst_pca$x[,1], y = vst_pca$x[,2], z = vst_pca$x[,3], color = myLIMAfactors$groups_m)
d3plot <- d3plot %>% add_markers()
d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")),
                     yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")),
                     zaxis = list(title = paste0('PC3', "(", pcvar[3], "%)"))))
d3plot

##################
f80_lima_DESeq <- DESeqDataSetFromMatrix(countData = assay(rna_seq_exp),
                                      colData = colData(rna_seq_exp),
                                      design= ~ 0 + groups_m)
vst_dt <- t(assay(vst(f80_lima_DESeq, blind = FALSE)))
vst_pca <- prcomp(vst_dt, scale. = FALSE)

pca_plot(pca_data = vst_pca, pc_x = 1, pc_y = 2, 
         group_var = rna_seq_exp$groups_m, 
         ellipse_var = rna_seq_exp$groups_m, allElipse = TRUE)

pcvar <- round(vst_pca$sdev^2/sum(vst_pca$sdev^2)*100, 2)

d3plot <- plot_ly(x = vst_pca$x[,1], y = vst_pca$x[,2], z = vst_pca$x[,3],
                    color = myLIMAfactors$groups_m)
d3plot <- d3plot %>% add_markers()
d3plot <- d3plot %>% layout(scene = list(
    xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")),
    yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")),
    zaxis = list(title = paste0('PC3', "(", pcvar[3], "%)"))))
d3plot
```

## Statistical Analysis

```{r}
lima_DESeq <- DESeqDataSetFromMatrix(countData = assay(rna_seq_exp),
                                      colData = colData(rna_seq_exp),
                                      design= ~ 0 + group)

lima_DESeq <- DESeq(lima_DESeq, test = "Wald")

ctrs_vec <- t(combn(levels(myLIMAfactors$group), m = 2))
ctrs_vec <- ctrs_vec[,2:1]

ctrs_vec_nm <- apply(ctrs_vec, 1, paste, collapse = "-")

ctrs_vec <- cbind("group", ctrs_vec)

ctrs_vec <- lapply(seq_len(nrow(ctrs_vec)), function(ctrs){
    ctrs_vec[ctrs,]
})

lima_DESeq_results <- lapply(ctrs_vec, function(ctrs){
    results(lima_DESeq, contrast = ctrs, pAdjustMethod = "fdr")
})

sapply(lima_DESeq_results, summary)
names(lima_DESeq_results) <- ctrs_vec_nm

lima_DESeq_results <- lapply(lima_DESeq_results, function(ctrs){
    ctrs$padj[is.na(ctrs$padj)] <- 1
    ctrs
})

####

lima_DESeq_results <- lapply(lima_DESeq_results, function(ctrs_res){
    colnames(ctrs_res)[c(2,5:6)] <- c("log2FC","p.value", "adj.p.value")
    ctrs_res
})

lima_DESeq_results <- do.call(cbind, lima_DESeq_results)

lima_DESeq_results$gene_name <- mapIds(org.Mm.eg.db, keys = rownames(lima_DESeq_results), 
                                    column = c("GENENAME"), 
                         keytype = "ENSEMBL", multiVals = "first")
lima_DESeq_results$entrezID <- mapIds(org.Mm.eg.db, keys = rownames(lima_DESeq_results), 
                                    column = c("ENTREZID"), 
                         keytype = "ENSEMBL", multiVals = "first")
lima_DESeq_results$symbol <- mapIds(org.Mm.eg.db, keys = rownames(lima_DESeq_results), 
                                    column = c("SYMBOL"), 
                         keytype = "ENSEMBL", multiVals = "first")
lima_DESeq_results$gene_lab <- ifelse(!is.na(lima_DESeq_results$gene_name),
                                   lima_DESeq_results$gene_name, 
                                   rownames(lima_DESeq_results))

rowData(rna_seq_exp) <- as.data.frame(lima_DESeq_results)

saveRDS(rna_seq_exp, paste0(params$output_dir, "RNA_Sequencing_exp_lima.rds"))
```

