---
title: "Enrichment Analysis of genes"
output: html_document
date: "2023-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(knitr.duplicate.label = "allow")
```

```{r}
library(AnnotationDbi)
library(org.Mm.eg.db)
library(DESeq2)
library(sva)
library(mixOmics)
library(knitr)
library(DT)
library(clusterProfiler)
source("../../Utility Functions/utilities.R")
```


## Loading the RNASeq results

```{r}
rna_seq_results <- readRDS("../RNA_seq/RNASeq_gfcv_siplified_model_results_list.rds")
rna_seq_full_results <- readRDS("../RNA_seq/RNASeq_gfcv_complete_model_ihw_results_list.rds")
rna_seq_full_results_no_int <- readRDS("../RNA_seq/RNASeq_gfcv_no_interaction_model_results_list.rds")

rna_seq_full_results <- c(rna_seq_results, rna_seq_full_results, rna_seq_full_results_no_int)

sig_tables <- lapply(rna_seq_full_results, function(ctrs){
    subset(ctrs ,!is.na(padj) & padj <= 0.1)
})

sig_genes <- lapply(sig_tables, function(ctrs){
    rownames(ctrs)
})

contrs_result_sig_list <- lapply(sig_tables, annotate_data, annotationPackage = org.Mm.eg.db)
```
# Enrichment analysis {.tabset}

## Enrichment results full lists {.tabset}

### Table results {.tabset}

```{r}
contrs_result_sig_list <- lapply(contrs_result_sig_list, function(ctrs){
    ctrs <- as.data.frame(ctrs)
    ctrs[order(ctrs$padj),]
})
contrs_names <- names(contrs_result_sig_list)
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE)) %>% formatRound(columns = colnames(contrs_result_sig_list[[{{i}}]])[-c(8,9,10)], digits = 3)",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`

### ORA {.tabset}

```{r}
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify = TRUE)
enrich_plot_data <- enrich_plots(enrich_results)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### GSEA {.tabset}

```{r}
genes_gsea <- lapply(rna_seq_full_results, function(ctrs){
    fc_list <- ctrs$log2FoldChange
    names(fc_list) <- rownames(ctrs)
    fc_list
})
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = genes_gsea, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify = TRUE, enrichFun = "gsea")

enrich_results <- lapply(enrich_results, function(ctrs){
    lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
})

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_gsea)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

## Enrichment results filtered lists {.tabset}

```{r}

independent_genes <- subset(contrs_result_sig_list$gfcv_G_vs_C, 
                            rownames(contrs_result_sig_list$gfcv_G_vs_C) %in% setdiff(sig_genes$gfcv_G_vs_C,
                                                                                      sig_genes$gfcvG.sex_mF))
male_only_genes <- subset(contrs_result_sig_list$GFM_CVM, 
                          rownames(contrs_result_sig_list$GFM_CVM) %in% setdiff(intersect(sig_genes$GFM_CVM,
                                                                                          sig_genes$gfcvG.sex_mF),
                                                                                sig_genes$GFF_CVF))
female_only_genes <- subset(contrs_result_sig_list$GFF_CVF, 
                            rownames(contrs_result_sig_list$GFF_CVF) %in% setdiff(intersect(sig_genes$GFF_CVF,
                                                                                            sig_genes$gfcvG.sex_mF),
                                                                                  sig_genes$GFM_CVM))
female_male_genes <- subset(contrs_result_sig_list$gfcvG.sex_mF, 
                            rownames(contrs_result_sig_list$gfcvG.sex_mF) %in% intersect(intersect(sig_genes$gfcvG.sex_mF, 
                                                                                                   sig_genes$GFM_CVM),
                                                                                         sig_genes$GFF_CVF))
feminized_masculinized_genes <- subset(contrs_result_sig_list$gfcvG.sex_mF, 
                                    rownames(contrs_result_sig_list$gfcvG.sex_mF) %in% setdiff(setdiff(sig_genes$gfcvG.sex_mF, 
                                                                                                   sig_genes$GFM_CVF),
                                                                                         sig_genes$GFF_CVM))

feminized_genes <- subset(contrs_result_sig_list$gfcvG.sex_mF, 
                                    rownames(contrs_result_sig_list$gfcvG.sex_mF) %in% intersect(setdiff(sig_genes$gfcvG.sex_mF, 
                                                                                                   sig_genes$GFM_CVF),
                                                                                         sig_genes$GFF_CVM))

masculinized_genes <- subset(contrs_result_sig_list$gfcvG.sex_mF, 
                                    rownames(contrs_result_sig_list$gfcvG.sex_mF) %in% setdiff(intersect(sig_genes$gfcvG.sex_mF, 
                                                                                                   sig_genes$GFM_CVF),
                                                                                         sig_genes$GFF_CVM))


dt_to_tables <- list("Sex independent DE genes" = independent_genes, 
                      "Male only DE genes" = male_only_genes, 
                      "Female only DE genes" = female_only_genes, 
                      "Female/Male DE genes" = female_male_genes,
                     "Feminized only genes" = feminized_genes,
                     "Masculinized only genes" = masculinized_genes,
                     "Feminized/Masculinized DE genes" = feminized_masculinized_genes)
```

### Table results {.tabset}

```{r}
contrs_result_sig_list <- dt_to_tables
contrs_result_sig_list <- lapply(contrs_result_sig_list, function(ctrs){
    ctrs <- as.data.frame(ctrs)
    ctrs[order(ctrs$padj),]
})
contrs_names <- names(contrs_result_sig_list)

tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE)) %>% formatRound(columns = colnames(contrs_result_sig_list[[{{i}}]])[-c(8,9,10)], digits = 3)",
    "```\n",
    "\n"
))
```

`r knit(text=unlist(tables_chunk))`

### ORA {.tabset}

```{r}
dt_to_pathway <- lapply(dt_to_tables, function(ctrs){rownames(ctrs)})
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = dt_to_pathway, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify = TRUE)
enrich_plot_data <- enrich_plots(enrich_results)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```

### GSEA {.tabset}

```{r}
sig_genes_gsea <- lapply(contrs_result_sig_list, function(ctrs){
    fc_list <- ctrs$log2FoldChange
    names(fc_list) <- rownames(ctrs)
    fc_list
})
set.seed(123)
enrich_results <- multi_enrich(dt_to_pathway = sig_genes_gsea, universe = NULL, 
                               OrgDb = org.Mm.eg.db, keyType = "ENSEMBL", 
                               qvalueCutoff = 0.05, pAdjustMthd = "fdr", 
                               pvalueCutoff = 0.05, simplify = TRUE, enrichFun = "gsea")

enrich_results <- lapply(enrich_results, function(ctrs){
    lapply(ctrs, setReadable, OrgDb = "org.Mm.eg.db", keyType = "ENSEMBL")
})

enrich_plot_data <- enrich_plots(enrich_results, fold_change_list = genes_gsea)
```

```{r, results='asis', fig.height=8, fig.width=8}
enrich_chunks(enrichData_all = enrich_results, enrich_plot_data = enrich_plot_data)
```