---
title: "Multi_Omics_Analysis"
output: html_document
date: "2023-05-13"
params:
    input_dir: "../../inputs/"
    output_dir: "../../outputs/Multi_Omics_Analysis/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.duplicate.label = "allow")
```

## Required Packages

```{r}
library(MOFA2)
library(dplyr)
library(knitr)
library(ggplot2)
library(MultiAssayExperiment)
library(grDevices)
library(rGREAT)
library(msigdbr)
library(DT)
library(TxDb.Mmusculus.UCSC.mm39.refGene)
source("../Utility_Functions/utilities.R")
library(TreeSummarizedExperiment)
library(GGally)
```

## Loaded matrices

```{r}
multi_omics_raw <- readRDS(paste0(params$output_dir, "Multi_Omics_Analysis_exp_lima.rds"))
multi_omics <- readRDS(paste0(params$output_dir, "Multi_Omics_Analysis_MA_exp_lima.rds"))
names(experiments(multi_omics))

filter_list <- c("16S", "Shotgun.cecum", "Shotgun.colon")
```

## Preparing the omics matrices 

```{r}
multi_omics_names <- names(experiments(multi_omics))

multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|unsupervised", multi_omics_names)]

# multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|unsupervised|Targeted|Metallomics|Histone", multi_omics_names)]

# multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)])
multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("Species|metaerg_kegg|cogs", multi_omics_names)])
# |EC|metaer_kegg|cogs
multi_omics_selected
```




```{r}
# CV filter criteria
#
# filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
#     multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]
# 
#     if(grepl("RNA_Seq|DNA|Untargeted", multi_omics_names)){
#         if(grepl("RNA_Seq", multi_omics_names)){
#             cd_thr <- 0.0
#         } else if(grepl("unsupervised", multi_omics_names)){
#             cd_thr <- 0.99
#         } else if(grepl("DNA", multi_omics_names)){
#             cd_thr <- 0
#         } else{
#             cd_thr <- 0
#         }
#         var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){
#             sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE)
#         })
#         multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, cd_thr)],]
#     }
#     multi_omics_exp
# })
# names(filtered_exps) <- names(multi_omics)
# experiments(multi_omics) <- ExperimentList(filtered_exps)
```

```{r}
filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
    multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]
    p_val_dt <- rowData(multi_omics_exp)[,grepl("adj.p.value", colnames(rowData(multi_omics_exp)))]

    multi_omics_exp <- multi_omics_exp[apply(as.matrix(p_val_dt) <= 0.05, 1, any),]
    multi_omics_exp
})
names(filtered_exps) <- names(multi_omics)
experiments(multi_omics) <- ExperimentList(filtered_exps)
```

## Run MOFA2

```{r}
num_factors <- 3

# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,,multi_omics_selected],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "specie_metaerg_pathway_mofa",
#     color_vec = c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd",
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"),
#                                 params = param_list
#                             )
#                     )

# & !multi_omics$samplename %in% c("X643C", "X644C", "X658C")

param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
    output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
    multi_omics_exp = multi_omics[,colData(multi_omics)$group != "HFE", multi_omics_selected],
    group_var = "group",
    num_factors = num_factors,
    results_name = "any_sig_005_mofa_int",
    color_vec = c("CNT"="black", "HFD"="red", "INT"="blue")
)
# c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
xfun::Rscript_call(rmarkdown::render,
                    args = list(input = "./MOFA_Models.Rmd", 
                                output_file = paste0(param_list$output_dir,
                                                     param_list$results_name, ".html"), 
                                params = param_list
                            )
                    )

# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "INT" & !multi_omics$samplename %in% c("X643C", "X644C", "X658C"),multi_omics_selected],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "specie_metaerg_pathway_any_sig_mofa_hfe",
#     color_vec = c("CNT"="black", "HFD"="red", "HFE"="green")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd", 
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"), 
#                                 params = param_list
#                             )
#                     )
```




```{r}
# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "HFE"],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "mofa_int",
#     color_vec = c("CNT"="black", "HFD"="red", "INT"="blue")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd", 
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"), 
#                                 params = param_list
#                             )
#                     )

param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
    output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
    multi_omics_exp = multi_omics[,colData(multi_omics)$group != "INT"],
    group_var = "group",
    num_factors = num_factors,
    results_name = "mofa_hfe",
    color_vec = c("CNT"="black", "HFD"="red", "HFE"="green")
)
# c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
xfun::Rscript_call(rmarkdown::render,
                    args = list(input = "./MOFA_Models.Rmd", 
                                output_file = paste0(param_list$output_dir,
                                                     param_list$results_name, ".html"), 
                                params = param_list
                            )
                    )
```




```{r}
mice_dt <- openxlsx::read.xlsx(paste0(params$input_dir, "AnimalData_OscarYanes.xlsx"))

# mice_num <- mice_dt[,c("BW(g)", "WATep(g)", "Liver(g)")]
mice_num <- mice_dt
mice_s <- as.data.frame(mice_num)

dict_vect <- c("Ctrl" = "CNT", "Ctrl ext" = "CNT", "HFD" = "HFD", "INT2" = "INT", "HFD excl." = "HFE")
mice_s$group <- factor(dict_vect[mice_dt$Group], levels = c("CNT", "HFD", "INT", "HFE"))

dict_vect_g <- c("CNT" = "C", "HFD" = "H", "INT" = "I", "HFE" = "E")

mice_s$samplename <- paste0(make.names(mice_dt$ID), dict_vect_g[mice_s$group])


mice_s$samplename[!mice_s$samplename %in% multi_omics$samplename]
multi_omics$samplename[!multi_omics$samplename %in% mice_s$samplename]

mice_s$samplename[mice_s$samplename == "X686I"] <- "X687I"

mice_s <- dplyr::left_join(as.data.frame(colData(multi_omics)), mice_s, by = "samplename")
mice_s
```


```{r}
d3plot <- plot_ly(x = mice_s$`BW(g)`, y = mice_s$`WATep(g)`, z = mice_s$`Liver(g)`,
                    color = as.factor(mice_s$Sacrifice))
d3plot <- d3plot %>% add_markers()
d3plot <- d3plot %>% layout(scene = list(
    xaxis = list(title = "BW (g)"),
    yaxis = list(title = "WATep (g)"),
    zaxis = list(title = "Liver (g)")))
d3plot
```

```{r}
factor_dt <- readRDS(paste0(params$output_dir, "/MOFA_Results/specie_metaerg_pathway_any_sig_mofa_int/factor_data.rds"))

mice_sub <- mice_s[mice_s$group.x != "HFE",]

all(names(factor_dt$group1[,3]) == mice_sub$samplename)

cor(mice_sub$`BW(g)`, factor_dt$group1[,3])
cor(mice_sub$`WATep(g)`, factor_dt$group1[,3])
cor(mice_sub$`Liver(g)`, factor_dt$group1[,3])


ctr_sub <- mice_sub[mice_sub$group.x == "CNT",]
ctr_fact <- factor_dt$group1[mice_sub$group.x == "CNT",]

cor(ctr_sub$`BW(g)`, ctr_fact[,3])
cor(ctr_sub$`WATep(g)`, ctr_fact[,3])
cor(ctr_sub$`Liver(g)`, ctr_fact[,3])

plot(ctr_sub$`BW(g)`, ctr_fact[,3])
plot(ctr_sub$`WATep(g)`, ctr_fact[,3])
plot(ctr_sub$`Liver(g)`, ctr_fact[,3])


```


### ORA analysis {.tabset}

```{r}
output_dir <- paste0(params$output_dir, "MOFA_Results/any_sig_005_mofa_int/")
omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))
rna_seq_rd <- rowData(experiments(omic_dt)$RNA_Sequencing)
    
rna_seq_list <- lapply(seq_len(num_factors), function(factor_id){
    list("All_Weights" = rna_seq_rd,
        "Positive_Weights" = rna_seq_rd[rna_seq_rd[[paste0("Factor_", factor_id, "_weights")]] >= 0,],
        "Negative_Weights" = rna_seq_rd[rna_seq_rd[[paste0("Factor_", factor_id, "_weights")]] < 0,])
})
names(rna_seq_list) <- paste0("Factor_", seq_len(num_factors))

for(rna_seq_name in names(rna_seq_list)){
    
    rna_seq_exp <- rna_seq_list[[rna_seq_name]]
    
    gsea_cols_list <- lapply(names(rna_seq_exp), function(x){paste0(rna_seq_name, "_weights")})
    names(gsea_cols_list) <- names(rna_seq_exp)
    
    param_list <- list(
        output_dir = output_dir,
        exp_list = rna_seq_exp,
        gsea_cols_list = gsea_cols_list,
        gsea = TRUE,
        ora_filter = parse(text = paste0(rna_seq_name, "_weights_sig")),
        order_column = paste0(rna_seq_name, "_weights"),
        results_name = rna_seq_name,
        legend_name = paste0(rna_seq_name, "_weights")
    )
    xfun::Rscript_call(rmarkdown::render,
                        args = list(input = "./RNA_Sequencing_Enrichments.Rmd", 
                                    output_file = paste0(param_list$output_dir, 
                                                         "RNA_Sequencing_Enrichment_",
                                                         rna_seq_name, ".html"), 
                                    params = param_list
                                )
                        )
}
```

### rGREAT Enrichment & Region Analysis {.tabset}

```{r}
dna_met_list <- lapply(seq_len(num_factors), function(factor_id){
    dna_met_dt <- experiments(omic_dt)$WGB_DNA_Methylation
    list("All_Weights" = dna_met_dt,
        "Positive_Weights" = dna_met_dt[rowData(dna_met_dt)[[paste0("Factor_", factor_id, "_weights")]] >= 0,],
        "Negative_Weights" = dna_met_dt[rowData(dna_met_dt)[[paste0("Factor_", factor_id, "_weights")]] < 0,])
})

names(dna_met_list) <- paste0("Factor_", seq_len(num_factors))


for(dna_seq_name in names(dna_met_list)){
    
    dna_seq_exp <- dna_met_list[[dna_seq_name]]
    
    gsea_cols_list <- lapply(names(dna_seq_exp), function(x){paste0(dna_seq_name, "_weights")})
    names(gsea_cols_list) <- names(dna_seq_exp)
    
    param_list <- list(
        dmr_exp_list = dna_seq_exp,
        txdb_id = "TxDb.Mmusculus.UCSC.mm39.refGene",
        msigdbr_specie = "Mus musculus",
        ora_filter = parse(text = paste0(dna_seq_name, "_weights_sig")),
        random_seed = 1234,
        output_dir = output_dir,
        results_name = dna_seq_name
    )
    xfun::Rscript_call(rmarkdown::render,
                        args = list(input = "../WGB_DNA_Methylation/DMR_enrichment_analysis.Rmd", 
                                    output_file = paste0(param_list$output_dir, 
                                                         "WGB_Sequencing_Enrichment_",
                                                         dna_seq_name, ".html"), 
                                    params = param_list
                                )
                        )
}
```


```{r}
library(msigdbr)
txdb <- "TxDb.Mmusculus.UCSC.mm39.refGene"
h_gene_sets = msigdbr(species = "Mus musculus", category = "H")

h_gene_sets = split(h_gene_sets$entrez_gene, h_gene_sets$gs_name)
h_gene_sets = lapply(h_gene_sets, as.character)  # just to make sure gene IDs are all in character.
```

```{r}
sig_tables <- lapply(factor_ids, function(id){
    multi_omics[keep_var_list[[id]],,"DNA_Methylation.liver"][["DNA_Methylation.liver"]]
})
names(sig_tables) <- paste("Factor", seq_along(sig_tables))
    
set.seed(1234)
great_results <- lapply(sig_tables, function(ctrs_res){
    if(nrow(ctrs_res) == 0){return(NA)}
    sub_granges <- rowRanges(ctrs_res)
    mcols(sub_granges) <- mcols(sub_granges)[!colnames(mcols(sub_granges)) %in% 
                                               c("seqnames", "ranges", "strand", 
                                                "seqlevels", "seqlengths",
                                                "isCircular", "start", "end", 
                                                "width", "element")]  
    return(great(gr = sub_granges, gene_sets = h_gene_sets, 
                tss_source = txdb))
})

contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    return(getEnrichmentTable(gr_res))
})
contrs_names <- names(contrs_result_sig_list)
template <- c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

### Region Annotation

```{r}
omic_dt_list <- c("All Results" = multi_omics[["DNA_Methylation.liver"]], sig_tables[1:2])

dmr_annotation_plot(omic_dt = omic_dt_list)
# test_dt <- factor_tables[1:2,]
# fisher.test(factor_tables[1,])
```

### Chromosome Plots

```{r, fig.height=6, fig.width=8}
possible_chromosomes  <- unique(seqnames(rowRanges(multi_omics[["DNA_Methylation.liver"]])))
dmr_annotation_plot(omic_dt_list, annotation_labels = possible_chromosomes, 
                    annotation_metadata_var = "seqnames", x_lab = "Chromosome Proportions")
```


```{r}
# possible_annotations <- c("Intron", "Promoter", "Distal Intergenic", "3' UTR", "5' UTR", "Exon", "Downstream")
# 
# factor_tables <- lapply(c("All Results" = multi_omics[["DNA_Methylation.liver"]], sig_tables[1:3]), function(exp_dt){
#     exp_annotation <- sapply(possible_annotations, function(x){
#         found_anno <- grepl(pattern = x, rowData(exp_dt)$annotation)
#         sum(found_anno)
#     })
#     exp_annotation
#     # table(unlist(exp_annotation))
# })
# factor_tables <- do.call(rbind, factor_tables)
# factor_annotation_dt <-factor_tables/rowSums(factor_tables)
# 
# plot_dt <- reshape2::melt(factor_annotation_dt)
# plot_dt$Var1 <- factor(plot_dt$Var1, levels = rev(levels(plot_dt$Var1)))
# ggplot(plot_dt, aes(x = value, y = Var1, fill = Var2)) + geom_col() + scale_fill_brewer(palette = "Set2") + theme_minimal() + xlab("Region Proportions") + ylab(NULL) + theme(legend.title = element_blank())
```

<!-- ## LIMA full unsupervised -->

<!-- ```{r} -->
<!-- multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|DNA_Methylation.liver", multi_omics_names)] -->
<!-- # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)]) -->
<!-- multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)]) -->

<!-- multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected]) -->
<!-- plot_data_overview(multi_omics_mofa) -->
<!-- ``` -->

<!-- ## Parameters -->

<!-- ```{r} -->
<!-- data_opts <- get_default_data_options(multi_omics_mofa) -->
<!-- head(data_opts) -->

<!-- model_opts <- get_default_model_options(multi_omics_mofa) -->
<!-- head(model_opts) -->
<!-- model_opts$num_factors <- 8 -->

<!-- train_opts <- get_default_training_options(multi_omics_mofa) -->
<!-- head(train_opts) -->
<!-- ``` -->

<!-- ## Model and training -->

<!-- ```{r} -->
<!-- multi_omics_mofa <- prepare_mofa( -->
<!--   object = multi_omics_mofa, -->
<!--   data_options = data_opts, -->
<!--   model_options = model_opts, -->
<!--   training_options = train_opts -->
<!-- ) -->

<!-- outfile <- file.path(getwd(),"model_us.hdf5") -->
<!-- multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE) -->

<!-- ``` -->

<!-- ## Downstream Analysis -->

<!-- ```{r} -->
<!-- # file_path <- "./model.hdf5" -->
<!-- # model <- load_model(file_path) -->
<!-- colData(multi_omics)$sample <- colData(multi_omics)$samplename -->

<!-- samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample") -->

<!-- plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90)) -->

<!-- p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5,  -->
<!--                  color_by = "group.y",  -->
<!--                  add_violin = TRUE) -->

<!-- p <- p +  -->
<!--   scale_color_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) + -->
<!--   scale_fill_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) -->
<!-- print(p) -->
<!-- ``` -->

<!-- ```{r, fig.height=12, fig.width=12} -->
<!-- plot_factors(multi_omics_mofa,  -->
<!--   factors = 1:6, -->
<!--   color_by = "group.y" -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pca_dt <- MOFA2::get_factors(object = multi_omics_mofa, factors = 1:3) -->

<!-- MOFA2::get_variance_explained(object = multi_omics_mofa, factors = 1:3) -->

<!-- library(plotly) -->

<!-- d3plot <- plot_ly(x = pca_dt$group1[,1], y = pca_dt$group1[,2], z = pca_dt$group1[,3], color = multi_omics_mofa@samples_metadata$group.y) -->
<!-- d3plot <- d3plot %>% add_markers() -->
<!-- # d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")), -->
<!-- #                      yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")), -->
<!-- #                      zaxis = list(title = paste0('PC4', "(", pcvar[4], "%)")))) -->
<!-- d3plot -->
<!-- ``` -->

<!-- ```{r} -->
<!-- plot_weights(multi_omics_mofa,  -->
<!--   view = 1, -->
<!--   factor = 2, -->
<!--   nfeatures = 10,     # Number of features to highlight -->
<!--   scale = T,          # Scale weights from -1 to 1 -->
<!--   abs = F            # Take the absolute value? -->
<!-- ) -->

<!-- weight_dt <- get_weights(object = multi_omics_mofa, factors = 2, as.data.frame = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(view_id in seq_len(length(multi_omics_selected))){ -->
<!--     plot_data_heatmap(multi_omics_mofa, -->
<!--       view = view_id,         # view of interest -->
<!--       factor = 2,             # factor of interest -->
<!--       features = 40,          # number of features to plot (they are selected by weight) -->

<!--       # extra arguments that are passed to the `pheatmap` function -->
<!--       cluster_rows = TRUE, cluster_cols = TRUE, -->
<!--       show_rownames = TRUE, show_colnames = FALSE, annotation_samples = "group.y", -->
<!--     ) -->
<!-- } -->
<!-- ``` -->


## GFCV

## Loaded matrices

```{r}
library(MOFA2)
multi_omics_raw <- readRDS("./Merged_Multi_Omics_gfcv.rds")
multi_omics <- readRDS("./Multivariate_Merged_Multi_Omics_gfcv.rds")
names(experiments(multi_omics))

filter_list <- c("16S")

selected_omics <- lapply(filter_list, function(dt_name){
    omic_dt_names <- grep(dt_name, names(experiments(multi_omics)))
    omic_exp_list <- experiments(multi_omics[,,omic_dt_names])
    
    sig_vars_num <- sapply(omic_exp_list, function(omic_exp){
        var_dt <- rowData(omic_exp)
        p_val_dt <- var_dt[,grep("adj.p.value", colnames(var_dt))]
        sig_vars <- apply(p_val_dt, 1, function(p_vals){
            any(p_vals <= 0.05, na.rm = TRUE)
        })
        sum(unlist(sig_vars))
    })
    sig_vars_num
})
names(selected_omics) <- filter_list

a <- do.call(cbind, selected_omics)

selected_omics <- lapply(filter_list, function(dt_name){
    omic_dt_names <- grep(dt_name, names(experiments(multi_omics)))
    omic_exp_list <- experiments(multi_omics[,,omic_dt_names])
    
    sig_vars_num <- sapply(omic_exp_list, function(omic_exp){
        nrow(omic_exp)
    })
    sig_vars_num
})
names(selected_omics) <- filter_list

b <- do.call(cbind, selected_omics)
```

## Preparing the omics matrices 

```{r}
multi_omics_names <- names(experiments(multi_omics))

multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S", multi_omics_names)]
# multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)])
multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)])
multi_omics_selected
```

```{r}
filtered_exps <- lapply(multi_omics_selected, function(multi_omics_names){
    multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]

    if(grepl("RNASeq", multi_omics_names)){
        var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){
            sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE)
        })
        multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, 0)],]
    }
    multi_omics_exp
})
names(filtered_exps) <- multi_omics_selected
experiments(multi_omics) <- ExperimentList(filtered_exps)
```
```{r}
multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics)
plot_data_overview(multi_omics_mofa)
```


```{r}
multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected])
plot_data_overview(multi_omics_mofa)
```

## Parameters

```{r}
data_opts <- get_default_data_options(multi_omics_mofa)
head(data_opts)

model_opts <- get_default_model_options(multi_omics_mofa)
head(model_opts)
model_opts$num_factors <- 8

train_opts <- get_default_training_options(multi_omics_mofa)
head(train_opts)
```

## Model and training

```{r}
multi_omics_mofa <- prepare_mofa(
  object = multi_omics_mofa,
  data_options = data_opts,
  model_options = model_opts,
  training_options = train_opts
)

outfile <- file.path(getwd(),"model.hdf5")
multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE)

```

## Downstream Analysis

```{r}
# file_path <- "./model.hdf5"
# model <- load_model(file_path)
colData(multi_omics)$sample <- colData(multi_omics)$samplename

samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample")

plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90))

p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5, 
                 color_by = "group.y", 
                 add_violin = TRUE)

p <- p + 
  scale_color_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green")) +
  scale_fill_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green"))
print(p)
```

## Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```