---
title: "Multi_Omics_Analysis"
output: html_document
date: "2023-05-13"
params:
    input_dir: "../../inputs/"
    output_dir: "../../outputs/Multi_Omics_Analysis/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.duplicate.label = "allow")
```

## Required Packages

```{r}
library(MOFA2)
library(dplyr)
library(knitr)
library(ggplot2)
library(MultiAssayExperiment)
library(grDevices)
library(rGREAT)
library(msigdbr)
library(DT)
library(TxDb.Mmusculus.UCSC.mm39.refGene)
source("../Utility_Functions/utilities.R")
library(TreeSummarizedExperiment)
library(GGally)
library(igraph)
```



## Loaded matrices

```{r}
multi_omics_raw <- readRDS(paste0(params$output_dir, "Multi_Omics_Analysis_exp_lima.rds"))
multi_omics <- readRDS(paste0(params$output_dir, "Multi_Omics_Analysis_MA_exp_lima.rds"))
names(experiments(multi_omics))

filter_list <- c("16S", "Shotgun.cecum", "Shotgun.colon")
```

## Preparing the omics matrices 

```{r}
multi_omics_names <- names(experiments(multi_omics))

multi_omics_selected <- multi_omics_names[!grepl("16S|Shotgun|unsupervised", multi_omics_names)]

# multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|unsupervised|Targeted|Metallomics|Histone", multi_omics_names)]

# multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)])
multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("Species|metaerg_kegg|cogs", multi_omics_names)])
# |EC|metaer_kegg|cogs
multi_omics_selected
```




```{r}
# CV filter criteria
#
# filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
#     multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]
# 
#     if(grepl("RNA_Seq|DNA|Untargeted", multi_omics_names)){
#         if(grepl("RNA_Seq", multi_omics_names)){
#             cd_thr <- 0.0
#         } else if(grepl("unsupervised", multi_omics_names)){
#             cd_thr <- 0.99
#         } else if(grepl("DNA", multi_omics_names)){
#             cd_thr <- 0
#         } else{
#             cd_thr <- 0
#         }
#         var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){
#             sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE)
#         })
#         multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, cd_thr)],]
#     }
#     multi_omics_exp
# })
# names(filtered_exps) <- names(multi_omics)
# experiments(multi_omics) <- ExperimentList(filtered_exps)
```

```{r}
filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
    multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]
    p_val_dt <- rowData(multi_omics_exp)[,grepl("adj.p.value", colnames(rowData(multi_omics_exp)))]
    multi_omics_exp <- multi_omics_exp[apply(as.matrix(p_val_dt) <= 0.05, 1, any),]
    multi_omics_exp
})
names(filtered_exps) <- names(multi_omics)
experiments(multi_omics) <- ExperimentList(filtered_exps)
```

## Run MOFA2

We have a multi-omic dataset with over 35 datasets, including the different levels of hierarchy for the 16S and Shotgun Metagenomics and the different annotations for the functional metagenomics results.

We visualized the principal component analysis for each individual omic to have a vision of the data and select which datasets from the hierarchical levels and the functional metagenomics. With the results we choose to keep the Species hierarchy level and for the functional metagenomics the COGS and metaerg-kegg annotations.

We also applied a p-value threshold of 0.05 in any of the contrast to filter variables and match the results from the whole genome bisulphite that, because we used the DSS method that finds DMRs from clusters of statistically significant CpGs were already biased.

Thus, the MOFA model will help us find correlated feature patterns in the statistically significant data.

```{r}
num_factors <- 3

# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,,multi_omics_selected],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "specie_metaerg_pathway_mofa",
#     color_vec = c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd",
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"),
#                                 params = param_list
#                             )
#                     )

# & !multi_omics$samplename %in% c("X643C", "X644C", "X658C")

param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
    output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
    multi_omics_exp = multi_omics[,colData(multi_omics)$group != "HFE", multi_omics_selected],
    group_var = "group",
    num_factors = num_factors,
    results_name = "any_sig_005_mofa_int",
    color_vec = c("CNT"="black", "HFD"="red", "INT"="blue")
)
# c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
xfun::Rscript_call(rmarkdown::render,
                    args = list(input = "./MOFA_Models.Rmd", 
                                output_file = paste0(param_list$output_dir,
                                                     param_list$results_name, ".html"), 
                                params = param_list
                            )
                    )



# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "INT" & !multi_omics$samplename %in% c("X643C", "X644C", "X658C"),multi_omics_selected],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "specie_metaerg_pathway_any_sig_mofa_hfe",
#     color_vec = c("CNT"="black", "HFD"="red", "HFE"="green")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd", 
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"), 
#                                 params = param_list
#                             )
#                     )
```

We can look at the resulting html file and we can the that all most of the variance for all the omics data is captured by the two first factors and the third one is biased towards the RNA-Seq. This makes sense because we have big differences in variable size between all the omics, this phenomena is difficult to avoid because we have many omics with very low values, even if we didn't apply a filter, thus, as the first two factors, surprisingly don't seem to be affected by this effect, and we have all the omics represented at least with a 20% of explained variance in these, we decided to proceed with it.

The factor boxplots show quite clearly two different patterns in the data, the first factor, represents features which change by the high fat diet (HFD) and are restored to similar levels by the intervention (INT); the second one shows features that are affected by HFD but the intervention fails to restore to control levels, keeping their levels changed.

Correlating these factors with the explained variance of each omic, we can see many omics biased towards one of both factors thus, we can propose that the liver histone hodifications, metallomics, targeted metabolomics and methylomics (WGB); serum targeted metabolomics and the cecum and colon functional metagenomics KEGG have a higher variance explained in the first factor thus, are the omics with the most significant features that are restored by the intervention and the cecum metallomics and targeted metabolomics are biased toward the second factor, thus they are changed by the HFD and kept mostly unchanged by the intervention.

These correlations with the first factor correlation reveal that epigenetic signatures, when affected by the high fat diet, are mostly restored to control levels after a life-style intervention, and the histone modifications and targeted metabolomics, which focused on epigenetically relevant metabolites.

All the remaining omics: the liver untargeted metabolomics, rna-sequencing and the colon and cecum metagenomics and functional clusters of orthologous groups of proteins, have even weights indicating that they contain features that can explain the variance in both factors.

Interestingly most of the metagenomics data, except the kegg functions show this pattern with the transcriptomics and untargeted metabolomics.

Looking at the weights of each feature in these factors we can see which of them are most important in explaining these patterns of the data. We looked at the absolute value distribution of weights for each omic and used as threshold for variable selection a percentage of features proportional to the variance explained of each omic in each factor.

We applied an overrepresentation analysis (ORA) to the significant genes and differentially methylated regions from each weight, dividing them into three subset: all the weights and dividing it into positive and negative weights.

### ORA analysis {.tabset}

```{r}
output_dir <- paste0(params$output_dir, "MOFA_Results/any_sig_005_mofa_int/")
omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))
rna_seq_rd <- rowData(experiments(omic_dt)$RNA_Sequencing)
    
rna_seq_list <- lapply(seq_len(num_factors), function(factor_id){
    list("All_Weights" = rna_seq_rd,
        "Positive_Weights" = rna_seq_rd[rna_seq_rd[[paste0("Factor_", factor_id, "_weights")]] >= 0,],
        "Negative_Weights" = rna_seq_rd[rna_seq_rd[[paste0("Factor_", factor_id, "_weights")]] < 0,])
})
names(rna_seq_list) <- paste0("Factor_", seq_len(num_factors))

enrich_results_dir <- paste0(output_dir, "enrich_results/")

if(!dir.exists(enrich_results_dir)){dir.create(enrich_results_dir)}

for(rna_seq_name in names(rna_seq_list)){
    
    rna_seq_exp <- rna_seq_list[[rna_seq_name]]
    
    gsea_cols_list <- lapply(names(rna_seq_exp), function(x){paste0(rna_seq_name, "_weights")})
    names(gsea_cols_list) <- names(rna_seq_exp)
    
    param_list <- list(
        output_dir = enrich_results_dir,
        exp_list = rna_seq_exp,
        gsea_cols_list = gsea_cols_list,
        gsea = TRUE,
        ora_filter = parse(text = paste0(rna_seq_name, "_weights_sig")),
        order_column = paste0(rna_seq_name, "_weights"),
        results_name = rna_seq_name,
        legend_name = paste0(rna_seq_name, "_weights")
    )
    xfun::Rscript_call(rmarkdown::render,
                        args = list(input = "./RNA_Sequencing_Enrichments.Rmd", 
                                    output_file = paste0(param_list$output_dir, 
                                                         "RNA_Sequencing_Enrichment_",
                                                         rna_seq_name, ".html"), 
                                    params = param_list
                                )
                        )
}
```

The positive weights from the factor 1, genes that increase with the high fat diet and get reduced by the intervention, are enriched with GO terms related to the immune system such as response to beta interferon, chemotaxis, phagocytosis... These can be very well related with the known inflammatory problems that arise from the high fat diet, the subsequent infiltration of of immune cells into the liver and its reversibility before reaching the liver cyrrosis and loss of function. The negative weights, which includes genes that are reduced by the high fat diet and increased with the intervention, indicate epigenetic changes showing GO terms such as demethylation, that inversely correlates with the WGB results where there was an increased methylation state in the high fat diet (could check if all the weights are positive in the WGB), meaning that the demethylation was reduced (Could plot all the genes related to demethylation and correlate them to the DMRs) and chromatin remodelling; and metabolic changes such as catabolic processes which makes sense to be reduced when there is an excess of nutrients.

The positive weights from factor 2, genes that are increased by the high fat diet and they keep their expression levels, show changes in the metabolism, focused on lipids and small molecules      


### Metalomics correlation with the diet

```{r}
omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))

diet_metals <- openxlsx::read.xlsx("../../inputs/Metallomics/metallomics_raw.xlsx", 
                                    sheet = "diets", rowNames = TRUE)
rownames(diet_metals) <- gsub("CTR", "CNT", rownames(diet_metals))

dt_metals <- experiments(omic_dt[,,c("Metallomics.cecum", "Metallomics.liver")])

cor_mat <- lapply(dt_metals, function(exp_dt){
    diet_metals_long <- diet_metals[as.character(exp_dt$group), rownames(exp_dt)]
    # plotly::plot_ly(z = cor(t(assay(exp_dt)), diet_metals_long))
    
    return(diag(cor(t(assay(exp_dt)), diet_metals_long)))
})
cor_mat

```


### rGREAT Enrichment & Region Analysis {.tabset}

```{r}
dna_met_list <- lapply(seq_len(num_factors), function(factor_id){
    dna_met_dt <- experiments(omic_dt)$WGB_DNA_Methylation
    list("All_Weights" = dna_met_dt,
        "Positive_Weights" = dna_met_dt[rowData(dna_met_dt)[[paste0("Factor_", factor_id, "_weights")]] >= 0,],
        "Negative_Weights" = dna_met_dt[rowData(dna_met_dt)[[paste0("Factor_", factor_id, "_weights")]] < 0,])
})

names(dna_met_list) <- paste0("Factor_", seq_len(num_factors))

rgreat_output_dir <- paste0(output_dir, "rGreat_results/")

if(!dir.exists(rgreat_output_dir)){dir.create(rgreat_output_dir)}

for(dna_seq_name in names(dna_met_list)){
    
    dna_seq_exp <- dna_met_list[[dna_seq_name]]
    
    gsea_cols_list <- lapply(names(dna_seq_exp), function(x){paste0(dna_seq_name, "_weights")})
    names(gsea_cols_list) <- names(dna_seq_exp)
    
    param_list <- list(
        dmr_exp_list = dna_seq_exp,
        txdb_id = "TxDb.Mmusculus.UCSC.mm39.refGene",
        msigdbr_specie = "Mus musculus",
        ora_filter = parse(text = paste0(dna_seq_name, "_weights_sig")),
        random_seed = 1234,
        output_dir = output_dir,
        results_name = dna_seq_name
    )
    xfun::Rscript_call(rmarkdown::render,
                        args = list(input = "../WGB_DNA_Methylation/DMR_enrichment_analysis.Rmd", 
                                    output_file = paste0(rgreat_output_dir, 
                                                         "WGB_Sequencing_Enrichment_",
                                                         dna_seq_name, ".html"), 
                                    params = param_list
                                )
                        )
}
```

### MetaboAnalyst Mummichog analysis

```{r}
library(MetaboAnalystR)
library(SummarizedExperiment)

unt_met_list <- lapply(seq_len(num_factors), function(factor_id){
    unt_met_dt <- experiments(omic_dt)$Untargeted_Metabolomics
    
    list("All_Weights" = unt_met_dt,
        "Positive_Weights" = unt_met_dt[rowData(unt_met_dt)[[paste0("Factor_", factor_id, "_weights")]] >= 0,],
        "Negative_Weights" = unt_met_dt[rowData(unt_met_dt)[[paste0("Factor_", factor_id, "_weights")]] < 0,])
})

names(unt_met_list) <- paste0("Factor_", seq_len(num_factors))

unt_met_dt <- lapply(names(unt_met_list), function(factor_id){
    factor_dt <- unt_met_list[[factor_id]]
    lapply(factor_dt, function(weight_dt){
        feature_dt <- as.data.frame(rowData(weight_dt))[,c("mzmed", "rtmed",
                                                        paste0(factor_id, "_weights"))]
        feature_dt$p.value <- ifelse(rowData(weight_dt)[,paste0(factor_id, 
                                                                "_weights_sig")], 0.05, 1)
        feature_dt$mode <- tolower(gsub("^.*_", "", gsub("\\..*$", "", 
                                                        rownames(feature_dt))))
        feature_dt <- feature_dt[c("mzmed", "rtmed", paste0(factor_id, "_weights"), 
                                    "p.value", "mode")]
        colnames(feature_dt) <- c("mz", "rt", "t.score", "p.value", "mode")
        feature_dt$t.score <- abs(feature_dt$t.score)
        feature_dt
    })
})
names(unt_met_dt) <- names(unt_met_list)

old_dir <- getwd()
mummichog_dir <- paste0(output_dir, "mummichog_metaboanalyst/")

if(!dir.exists(mummichog_dir)){dir.create(mummichog_dir)}

for(factor_id in names(unt_met_dt)){
    for(weight_id in names(unt_met_dt[[factor_id]])){
        results_output <- paste0(mummichog_dir, factor_id, "_", weight_id, "/")    
        if(!dir.exists(results_output)){dir.create(results_output)}
        setwd(results_output)
        write.table(unt_met_dt[[factor_id]][[weight_id]], row.names = FALSE, 
                    file = "./mummichog_dt.txt", sep = ",")
        
        mSet <- InitDataObjects("mass_all", "mummichog", FALSE)
        
        mSet <- UpdateInstrumentParameters(mSet, 10.0, "mixed", "yes", 0.05)
        
        mSet <- Read.PeakListData(mSet, "mummichog_dt.txt")
        
        mSet <- SetRTincluded(mSet, "seconds")
        
        mSet <- SanityCheckMummichogData(mSet)
        
        mSet <- SetPeakEnrichMethod(mSet, "mum", "v2")
        
        mSet <- SetMummichogPval(mSet, cutoff = 0.05)
        
        mSet <- PerformPSEA(mSet, "mmu_kegg", "current", 3 , 100)
        
        mSet<-PlotPeaks2Paths(mSet, "peaks_to_paths_0_", "png", 72, width=NA)
        saveRDS(mSet, file = "./metaboanalyst_obj.rds")
        rm(mSet)
        setwd(old_dir)
    }
}

# setwd(old_dir)
```




### Comparing DMR and Transcriptomics enrichment results

```{r}
library(GOSemSim)
library(org.Mm.eg.db)
library(AnnotationDbi)
library(DOSE)


overlap_list <- list()
for(factor_id in paste0("Factor_", seq_len(num_factors))){
    rna_seq_enrich <- readRDS(paste0(enrich_results_dir, factor_id, "_rna_seq_enrich_results.rds"))
    rna_seq_enrich <- rna_seq_enrich$ORA
    dna_seq_enrich <- readRDS(paste0(rgreat_output_dir, factor_id, "_dmr_enrich_results.rds"))
    dna_seq_enrich <- dna_seq_enrich$ORA
    
    overlap_list[[factor_id]] <- list()
    
    for(type_enrich in names(rna_seq_enrich)){
        rna_enrich_list <- rna_seq_enrich[[type_enrich]]
        dna_enrich_list <- dna_seq_enrich[[type_enrich]]
        
        overlap_list[[factor_id]][[type_enrich]] <- list()
        
        for(enrich_id in names(rna_enrich_list)){
            rna_obj <- rna_enrich_list[[enrich_id]]
            rna_dt <- as.data.frame(rna_obj)
            
            dna_obj <- dna_enrich_list[[enrich_id]]
            dna_dt <- dna_obj@table
            
            # if(enrich_id == "kegg"){
            #     dna_dt$description <- gsub("^ ", "", sapply(strsplit(dna_dt$id, ":"), function(x){x[2]}))
            #     dna_dt$id <- sapply(strsplit(dna_dt$id, ":"), function(x){x[1]})
            # }
            # if(enrich_id == "wp"){
            #     dna_dt$description <- dna_dt$id
            #     dna_dt$id <- sapply(strsplit(dna_dt$id, "%"), 
            #                         grep, value = TRUE, pattern = "WP[[:digit:]]+")
            #     
            # }
            if(enrich_id %in% c("BP", "CC", "MF")){
                go_sem <- godata(OrgDb = org.Mm.eg.db, ont = tolower(enrich_id))
                
                go_sim <- go_similarity_matrix(rna_dt$ID, dna_dt$id, 
                                                semData = go_sem, measure = "Wang")
                rna_sim <- rownames(go_sim)[apply(go_sim >= 0.7, 1, any)]
                dna_sim <- colnames(go_sim)[apply(go_sim >= 0.7, 2, any)]
                
                rna_filt <- rna_dt[rna_dt$ID %in% rna_sim,]
                dna_filt <- dna_dt[dna_dt$id %in% dna_sim,]
                
                rna_gene_list <- strsplit(rna_filt$geneID, "/")
                
                overlap_list[[factor_id]][[type_enrich]][[enrich_id]] <- list(
                    "go_rna_dt" = rna_filt, 
                    "go_dna_dt" = dna_filt)
                
            } else{
                rna_filt <- rna_dt[rna_dt$ID %in% intersect(rna_dt$ID, dna_dt$id),]
                dna_filt <- dna_dt[dna_dt$id %in% intersect(rna_dt$ID, dna_dt$id),]
                
                rna_gene_list <- strsplit(rna_filt$geneID, "/")
                rna_gene_list <- lapply(rna_gene_list, mapIds, x = org.Mm.eg.db, keytype = "ENTREZID",
                                        column = "SYMBOL", multiVals = "first")
                    
                overlap_list[[factor_id]][[type_enrich]][[enrich_id]] <- list(
                    "go_rna_dt" = rna_filt, 
                    "go_dna_dt" = dna_filt)   
            }
            
            names(rna_gene_list) <- rna_filt$ID
            
            dna_annot <- getRegionGeneAssociations(dna_obj)
            dmr_gene_list <- as.list(mcols(dna_annot)$annotated_genes)
            names(dmr_gene_list) <- rownames(mcols(dna_annot))
            
            found_gsets <- dna_obj@gene_sets[dna_filt$id]
            
            dna_gene_list <- lapply(found_gsets, function(x){
                found_g <- lapply(dmr_gene_list, function(dmr_ids){
                    dmr_ids[x]
                })
                found_g <- unlist(found_g)
                unique(found_g[!is.na(found_g)])
            })
            gene_overlap <- intersect(unique(unlist(dna_gene_list)),
                                      unique(unlist(rna_gene_list)))
            found_pairs <- data.frame("transcript" = gene_overlap, 
                       "dmr" = sapply(gene_overlap, function(x){
                           found_dmr <- sapply(dmr_gene_list, function(dmr_info){
                               any(dmr_info %in% x)
                           })
                           paste(names(found_dmr)[found_dmr], collapse = ";")
                       }), 
                       "dmr_GO" = sapply(gene_overlap, function(x){
                           found_dmr <- sapply(dna_gene_list, function(dmr_info){
                               any(dmr_info %in% x)
                           })
                           paste(names(found_dmr)[found_dmr], collapse = ";")
                       }), 
                       "gene_GO" = sapply(gene_overlap, function(x){
                           found_dmr <- sapply(rna_gene_list, function(dmr_info){
                               any(dmr_info %in% x)
                           })
                           paste(names(found_dmr)[found_dmr], collapse = ";")
                       }))
            overlap_list[[factor_id]][[type_enrich]][[enrich_id]]$gene_overlap <- found_pairs
            
            
            
            # overlap_list[[factor_id]][[type_enrich]][[enrich_id]] <- length(intersect(rna_dt, dna_dt))/length(union(rna_dt, dna_dt))
            
            
        }
        
    }
}

saveRDS(overlap_list, file = paste0(output_dir, "go_overlap_results.rds"))

overlap_list <- readRDS(paste0(output_dir, "go_overlap_results.rds"))
```

```{r}
lapply(names(overlap_list), function(factor_id){
    factor_dt <- overlap_list[[factor_id]]
    
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        
        lapply(names(weight_dt), function(enrich_id){
            enrich_dt <- weight_dt[[enrich_id]]
            
            if(nrow(enrich_dt$gene_overlap) <= 2){return(NULL)}
            rna_genes <- enrich_dt$gene_overlap$transcript
            rna_seq_dt <- experiments(omic_dt)$RNA_Sequencing
            rna_seq_dt <- rna_seq_dt[,rna_seq_dt$group != "HFE"]
    
            p1 <- heatMapFun(dt_list = as.matrix(assay(rna_seq_dt[rowData(rna_seq_dt)$symbol %in% rna_genes,])),
                       metadata = colData(rna_seq_dt), stratificationVar = "group", 
                       clustering_distance_rows = "euclidean", show_colnames = T, 
                       plot_name = paste(factor_id, weight_id, enrich_id))
            
            dmr_genes <- enrich_dt$gene_overlap$transcript
            dmr_seq_dt <- experiments(omic_dt)$RNA_Sequencing
            dmr_seq_dt <- dmr_seq_dt[,dmr_seq_dt$group != "HFE"]
    
            p2 <-heatMapFun(dt_list = as.matrix(assay(dmr_seq_dt[rowData(dmr_seq_dt)$symbol %in% rna_genes,])),
                       metadata = colData(dmr_seq_dt), stratificationVar = "group", 
                       clustering_distance_rows = "euclidean", show_colnames = T, 
                       plot_name = paste(factor_id, weight_id, enrich_id))
            
            list(p1, p2)
        })
    })
})
```

### KEGG joint pathway analysis

```{r}
output_dir <- paste0(params$output_dir, "MOFA_Results/any_sig_005_mofa_int/")
omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))

experiments(omic_dt) <- ExperimentList(lapply(experiments(omic_dt), function(omic_exp){
    
    omic_rd <- rowData(omic_exp)[,grepl("Factor_[1,2,3]_weights$", colnames(rowData(omic_exp)))]
    
    rowData(omic_exp)$max_weight_id <- apply(abs(as.matrix(omic_rd)), 1, which.max)
    omic_exp
}))


rna_seq_rd <- lapply(experiments(omic_dt)[c("RNA_Sequencing", "WGB_DNA_Methylation", "Untargeted_Metabolomics")], rowData)
num_factors <- 2




library(stringr)
kegg_db_path <- paste0(params$output_dir, "mmu_kegg_pathway_db.rds")
test <- rGREAT::getKEGGPathways("mmu")

pathway_ids <- sapply(strsplit(names(test), ":"), function(x){x[[1]]})
names(test) <- pathway_ids

if(!file.exists(kegg_db_path)){
    library(KEGGREST)
    
    kegg_db_info <- lapply(pathway_ids, function(x){keggGet(x)[[1]]})
    names(kegg_db_info) <- pathway_ids
    saveRDS(kegg_db_info, kegg_db_path)    
} else{
    kegg_db_info <- readRDS(kegg_db_path)    
}

gene_enrich_db <- lapply(names(kegg_db_info), function(kegg_db_id){
    gene_info <- as.numeric(kegg_db_info[[kegg_db_id]]$GENE)
    
    gene_info <- gene_info[!is.na(gene_info)]    
    as.character(gene_info)
})
names(gene_enrich_db) <- names(kegg_db_info)

integ_enrich_db <- lapply(names(kegg_db_info), function(kegg_db_id){
    gene_info <- as.numeric(kegg_db_info[[kegg_db_id]]$GENE)
    
    gene_info <- gene_info[!is.na(gene_info)]    
    compound_info <- names(kegg_db_info[[kegg_db_id]]$COMPOUND)
    db_info <- c(gene_info, compound_info)
    if(length(db_info) == 0){return(NA)}    
    data.frame("term" = kegg_db_id, "gene" = db_info)        
})

names(integ_enrich_db) <- names(kegg_db_info)


for(pathway_id in names(integ_enrich_db[is.na(integ_enrich_db)])){
    integ_enrich_db[[pathway_id]] <- data.frame("term" = pathway_id, "gene" = test[[pathway_id]])
}
integ_enrich_db <- do.call(rbind, integ_enrich_db)

term2name_dt <- lapply(names(kegg_db_info), function(kegg_id){
    data.frame("term" = kegg_id, "name" = kegg_db_info[[kegg_id]]$NAME)
})
term2name_dt <- do.call(rbind, term2name_dt)

omic_enrich <- c("Untargeted_Metabolomics", "RNA_Sequencing", "WGB_DNA_Methylation")
ids_enrich_names <- c("metabo_analyst_cp", "entrezID", "geneId")

library(SummarizedExperiment)

omic_dict_list <- mapply(function(omic_id, enrich_var){
    
    omic_data <- omic_dt[[omic_id]]
    omic_dict <- rowData(omic_data)[[enrich_var]]
    
    names(omic_dict) <- paste0(omic_id, ".", rownames(omic_data))
    
    omic_dict
    
}, omic_id = omic_enrich, enrich_var = ids_enrich_names)

```


```{r}
feature_seq_list <- lapply(seq_len(num_factors), function(factor_id){
    
    # factor_col_sig <- paste0("Factor_", factor_id, "_weights_sig")
    factor_col_sig <- paste0("max_weight_id == ", factor_id)
    
    factor_col <- paste0("Factor_", factor_id, "_weights")
    
    weight_expr_list <- list("All_Weights" = parse(text = factor_col_sig),
                            "Positive_Weights" = parse(text = paste0(factor_col_sig, " & ", factor_col, " > 0")),
                            "Negative_Weights" = parse(text = paste0(factor_col_sig, " & ", factor_col, " < 0")))
    
    lapply(weight_expr_list, function(weight_expr){
        filtered_sig_features <- lapply(rna_seq_rd, function(x){
            x <- as.data.frame(x)
            rownames(x)[eval(expr = weight_expr, envir = x)]
        })
        # filtered_sig_features <- Reduce(c, filtered_sig_features)
        filtered_sig_features
    })
})
names(feature_seq_list) <- paste0("Factor_", seq_len(num_factors))
```

```{r}
factor_ids_enrich <- lapply(feature_seq_list, function(factor_dt){
    
    lapply(factor_dt, function(weight_type_dt){
        
        metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][paste0("Untargeted_Metabolomics", 
                                                                        ".", weight_type_dt[["Untargeted_Metabolomics"]])]
        
        metabo_dt <- metabo_dt[!is.null(metabo_dt)]
        metabo_dt <- strsplit(metabo_dt, ";")
        metabo_dt <- metabo_dt[!is.na(metabo_dt)]
        
        wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][paste0("WGB_DNA_Methylation", 
                                                                ".", weight_type_dt[[ "WGB_DNA_Methylation"]])]
        wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
        
        rna_dt <- omic_dict_list[["RNA_Sequencing"]][paste0("RNA_Sequencing",
                                                            ".", weight_type_dt[["RNA_Sequencing"]])]
        rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
        
        unlist(c(metabo_dt, rna_dt, wgb_dt))
    })
})

universe_list_ids <- lapply(names(rna_seq_rd), function(omic_name){
    omic_ids <- omic_dict_list[[omic_name]][paste0(omic_name, ".", rownames(rna_seq_rd[[omic_name]]))]
    omic_ids <- omic_ids[!is.null(omic_ids)]
    omic_ids <- unlist(strsplit(omic_ids, ";"))
    omic_ids[!is.na(omic_ids)]

})
names(universe_list_ids) <- names(rna_seq_rd)
universe_list_ids <- unlist(Reduce(c, universe_list_ids))
```

```{r}
library(clusterProfiler)
enrich_results_list_full <- lapply(factor_ids_enrich, function(factor_dt){
    lapply(factor_dt, function(weight_dt){
        enricher(gene = weight_dt, universe = universe_list_ids, 
                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)
    })
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(output_dir, "Kegg_Joint_Pathway_Enrichment.pdf"))
lapply(names(enrich_results_list_full), function(factor_id){
    factor_dt <- enrich_results_list_full[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        
        plot_name <- gsub("_", " ", paste(factor_id, weight_id))
        
        dotplot(weight_dt) + ggtitle(paste("Kegg joint pathway enrichment" , plot_name))
    })
})
dev.off()
```

## MOFA variables correlations (any significant)

```{r}
output_dir <- paste0(params$output_dir, "MOFA_Results/any_sig_005_mofa_int/")
corr_output_dir <- paste0(output_dir, "correlation_results_max_weight/")

omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))

sapply(names(experiments(omic_dt)), function(x){
    x_dt <- experiments(omic_dt)[[x]]
    hist(rowData(x_dt)$Factor_1_weights, main = x)
})

experiments(omic_dt) <- ExperimentList(lapply(experiments(omic_dt), function(omic_exp){
    
    omic_rd <- rowData(omic_exp)[,grepl("Factor_[1,2,3]_weights$", colnames(rowData(omic_exp)))]
    
    rowData(omic_exp)$max_weight_id <- apply(abs(as.matrix(omic_rd)), 1, which.max)
    omic_exp
}))

network_omics <- list("All Omics" = names(assays(omic_dt)))

# network_omics$`All Omics` <- grep("Untargeted|Targeted|RNASeq|species|metaerg_kegg|Metallomics|PTM|DNA_Methylation.liver",
#                                   network_omics$`All Omics`, value = TRUE)

# lapply(experiments(multi_omic_dt), function(x){colnames(rowData(x))})

variables_dicts <- lapply(network_omics[[1]], function(omic_name){
    if(omic_name == "RNA_Sequencing"){
        return("gene_lab")
    } else if(omic_name == "Untargeted_Metabolomics"){
        return("metaboanalyst_compound_names")
        #return("HFD-CNT.formula")
    } else if(grepl("DNA_Methylation", omic_name)){
        return("geneName")
    } else if(grepl("Shotgun|Metallomics|16S|Targeted|Histone", omic_name)){
        return("readable_name")  
    } else{
        return(NA)
    }
})
names(variables_dicts) <- network_omics[[1]]
variables_dicts[is.na(variables_dicts)] <- NULL

add_var_names <- grep("weights|FC", colnames(rowData(omic_dt@ExperimentList$RNA_Sequencing)), value = T)

# corr_method_vect <- c("kendall", "spearman", "pearson")
corr_method_vect <- c("spearman")

for(corr_method in corr_method_vect){
    xfun::Rscript_call(rmarkdown::render,
                            args = list(input = "../Multi_Omics_Correlations/Correlations_Script.Rmd",
                              output_file = output_dir,
                              params = list(
                                corr_threshold = 0,
                                corr_method = corr_method,
                                multi_omic_path = omic_dt,
                                group_metadata_var = "group",
                                sample_levels_2_correlate = c("CNT", "HFD", "INT"),
                                filter_variable_2_correlate = list(
                                    "Factor_1" = quote(max_weight_id == 1),
                                    "Factor_2" = quote(max_weight_id == 2)
                                ),
                                p_adj_thr = 1,
                                omic_list_2_correlate = network_omics,
                                rowData_variable_name_list = variables_dicts,
                                draw_networks = TRUE,
                                draw_cytoscape = FALSE,
                                add_autocorrelations = FALSE,
                                keep_full_correlations = TRUE,
                                output_dir = corr_output_dir,
                                output_suffix = paste0("0_005_chi_f1f2_", corr_method),
                                add_variables = TRUE,
                                variables_names = add_var_names,
                                draw_cluster_scatter_plots = FALSE))
                      )
    # xfun::Rscript_call(rmarkdown::render,
    #                         args = list(input = "../Multi_Omics_Correlations/Correlations_Script.Rmd",
    #                           output_file = output_dir,
    #                           params = list(
    #                             corr_threshold = 0,
    #                             corr_method = corr_method,
    #                             multi_omic_path = omic_dt,
    #                             group_metadata_var = "group",
    #                             sample_levels_2_correlate = c("CNT", "HFD", "INT"),
    #                             filter_variable_2_correlate = list(
    #                                 "Factor_1" = quote(Factor_1_weights_sig),
    #                                 "Factor_2" = quote(Factor_2_weights_sig)
    #                             ),
    #                             p_adj_thr = 1,
    #                             omic_list_2_correlate = network_omics,
    #                             rowData_variable_name_list = variables_dicts,
    #                             draw_networks = TRUE,
    #                             draw_cytoscape = FALSE,
    #                             add_autocorrelations = FALSE,
    #                             keep_full_correlations = TRUE,
    #                             output_dir = corr_output_dir,
    #                             output_suffix = paste0("0_005_chi_f1f2_", corr_method),
    #                             add_variables = TRUE,
    #                             variables_names = add_var_names,
    #                             draw_cluster_scatter_plots = FALSE))
    #                   )


    # xfun::Rscript_call(rmarkdown::render,
    #                             args = list(input = "../Multi_Omics_Correlations/Correlations_Script.Rmd",
    #                               output_file = output_dir,
    #                               params = list(
    #                                 corr_threshold = 0,
    #                                 corr_method = corr_method,
    #                                 multi_omic_path = omic_dt,
    #                                 group_metadata_var = "group",
    #                                 sample_levels_2_correlate = c("CNT", "HFD", "INT"),
    #                                 filter_variable_2_correlate = list(
    #                                     "All" = NA
    #                                 ),
    #                                 omic_list_2_correlate = network_omics,
    #                                 rowData_variable_name_list = variables_dicts,
    #                                 draw_networks = FALSE,
    #                                 draw_cytoscape = FALSE,
    #                                 add_autocorrelations = FALSE,
    #                                 keep_full_correlations = TRUE,
    #                                 output_dir = corr_output_dir,
    #                                 output_suffix = paste0("0_005_chi", corr_method),
    #                                 add_variables = TRUE,
    #                                 variables_names = add_var_names,
    #                                 draw_cluster_scatter_plots = FALSE))
    #                       )
}
# rmarkdown::render(input = "../Multi_Omics_Correlations/Correlations_Script.Rmd",
#                           output_file = corr_output_dir,
#                           params = list(
#                             corr_threshold = 0.9,
#                             multi_omic_path = omic_dt,
#                             group_metadata_var = "group",
#                             sample_levels_2_correlate = c("CNT", "HFD", "INT"),
#                             filter_variable_2_correlate = list(
#                                 "All" = NA
#                             ),
#                             omic_list_2_correlate = network_omics,
#                             rowData_variable_name_list = variables_dicts,
#                             draw_networks = TRUE,
#                             add_autocorrelations = FALSE,
#                             keep_full_correlations = TRUE,
#                             output_dir = corr_output_dir,
#                             output_suffix = "09_chi"))
```

### Network downstream analysis

```{r}
library(ProteoNet)
library(igraph)
library(dplyr)
library(org.Mm.eg.db)
output_dir <- paste0(params$output_dir, "MOFA_Results/any_sig_005_mofa_int/")
corr_output_dir <- paste0(output_dir, "correlation_results_max_weight/")

omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))

nw_paths <- list.files(corr_output_dir, pattern = "\\.rds$", full.names = T)
nw_graphs <- lapply(nw_paths, readRDS)


variables_dicts <- lapply(names(assays(omic_dt)), function(omic_name){
    if(omic_name == "RNA_Sequencing"){
        return("gene_lab")
    } else if(omic_name == "Untargeted_Metabolomics"){
        return("metaboanalyst_compound_names")
        #return("HFD-CNT.formula")
    } else if(grepl("DNA_Methylation", omic_name)){
        return("geneName")
    } else if(grepl("Shotgun|Metallomics|16S|Targeted|Histone", omic_name)){
        return("readable_name")  
    } else{
        return(NA)
    }
})
names(variables_dicts) <- names(assays(omic_dt))
variables_dicts[is.na(variables_dicts)] <- NULL

score_names <- gsub("\\.rds", "", sapply(strsplit(basename(nw_paths), "_"),
                                            function(x){x[length(x)]}))

# score_filters <- c(0.6, 0.8, 0.8)
# names(score_filters) <- score_names
# names(nw_graphs) <- score_names
# 
# 
# score_filters$spearman$Factor_2
score_filters <- c(0.8)
names(score_filters) <- score_names
names(nw_graphs) <- score_names

# score_filters <- list("spearman" = list("Factor_1" = 0.85, "Factor_2" = 0.8))
# names(nw_graphs) <- score_names

# rowData(omic_dt[["Shotgun_Metagenomics.cecum.Species"]])$readable_name <- paste("Shotgun", rowData(omic_dt[["Shotgun_Metagenomics.cecum.Species"]])$readable_name)
# 
# rowData(omic_dt[["Shotgun_Metagenomics.colon.Species"]])$readable_name <- paste("Shotgun", rowData(omic_dt[["Shotgun_Metagenomics.colon.Species"]])$readable_name)
# 
# 
# rowData(omic_dt[["16S_Metagenomics.cecum.Species"]])$readable_name <- paste("16S", 
#                                                                             rowData(omic_dt[["16S_Metagenomics.cecum.Species"]])$readable_name)
# 
# rowData(omic_dt[["16S_Metagenomics.c.Species"]])$readable_name <- paste("16S", rowData(omic_dt[["16S_Metagenomics.cecum.Species"]])$readable_name)
```

```{r}
variable_dictionary <- lapply(names(variables_dicts), function(dataset_name){
        
    dataset_var <- rowData(omic_dt[[dataset_name]])[variables_dicts[[dataset_name]]]
    dataset_vector <- unlist(dataset_var[[1]])
    names(dataset_vector) <- paste0(dataset_name, ".", rownames(dataset_var))
    
    dataset_vector
})
names(variable_dictionary) <- names(variables_dicts)

nw_graphs <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(factor_dt){
        vertex_attr(factor_dt[["All Omics"]], "Readable_Name") <- sapply(names(V(factor_dt[["All Omics"]])), 
                                                                    function(v_name){
                omic_id <- vertex_attr(factor_dt[["All Omics"]], name = "Node_Omic", index = v_name)
                if(omic_id %in% names(variable_dictionary)){
                    variable_dictionary[[omic_id]][v_name]
                } else{
                    return(v_name)
                }
            })
        factor_dt
    })
})

```

```{r}
# filter_scores <- seq(0.2, 0.9, by = 0.05)
# 
# nw_filter_score <- lapply(names(nw_graphs), function(score_id){
#     score_dt <- nw_graphs[[score_id]]
# 
#     score_res <- lapply(names(score_dt), function(weight_id){
#         weight_dt <- score_dt[[weight_id]][["All Omics"]]
# 
#         pdf(paste0(corr_output_dir, "degree_distribution_", score_id, "_", weight_id, ".pdf"))
# 
#         power_fit_pval <- sapply(filter_scores, function(x){
#             if(x == 1 || x == 0){return(NA)}
#             graph_weights <- abs(edge_attr(weight_dt,
#                                     name = "weight"))
#             sub_g <- subgraph.edges(weight_dt,
#                         E(weight_dt)[graph_weights >= x],
#                         delete.vertices = TRUE)
#             if(length(sub_g) == 0){return(NA)}
#             hist(degree(sub_g), breaks = 100, main = x)
#             return(power.law.fit(degree(sub_g))$KS.stat)
#             degree(sub_g)
#         })
# 
#         dev.off()
# 
#         names(power_fit_pval) <- filter_scores
#         power_fit_pval
#         # return(names(power_fit_pval)[which.min(power_fit_pval)])
#     })
#     names(score_res) <- names(score_dt)
#     score_res
# })
# names(nw_filter_score) <- names(nw_graphs)
# score_filters <- lapply(nw_filter_score, function(score_dt){
#     lapply(score_dt, function(weight_dt){
#         # lapply(weight_dt, function(filter_scores){
#         #     if(all(is.na(filter_scores))){return(NA)}
#         #     return(power.law.fit(filter_scores)$KS.stat)
#         # })
#         # plot(x = names(weight_dt), y = weight_dt)
#         return(names(weight_dt)[which.min(weight_dt)])
#     })
# })
```

```{r}
rand_seed <- 1234
universe_list <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(weight_dt){
        lapply(weight_dt, function(graph_dt){
            names(V(graph_dt))
        })
    })
})
# 
# 
# score_filters[[score_id]]]

nw_graphs <- lapply(names(nw_graphs), function(score_id){
    score_dt <- nw_graphs[[score_id]]
    score_res <- lapply(names(score_dt), function(weight_id){
        weight_dt <- score_dt[[weight_id]]

        lapply(weight_dt, function(graph_dt){
            graph_weights <- abs(edge_attr(graph_dt,
                            name = "weight"))
            graph_adj.p.val <- edge_attr(graph_dt, name = "adj.p.value")
            subgraph.edges(graph_dt,
                        E(graph_dt)[graph_weights >= score_filters[score_id] & E(graph_dt)[graph_adj.p.val <= 0.05]],
                        delete.vertices = TRUE)
            # subgraph.edges(graph_dt,
                        # E(graph_dt)[graph_weights >= as.numeric(score_filters[[score_id]][[weight_id]] & E(graph_dt)[graph_adj.p.val <= 0.05])],
                        # delete.vertices = TRUE)

        })
    })
    names(score_res) <- names(score_dt)
    score_res
})
names(nw_graphs) <- score_names
# graph_weights <- abs(edge_attr(nw_dt, name = "weight"))
#
# hist(graph_weights)
# hist(degree.distribution(nw_dt), breaks = 100)

# set.seed(rand_seed)
nw_graphs <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(weight_dt){
        lapply(weight_dt, function(graph_dt){
            membership_dt <- membership(infomaps_website_clustering(
                                        graph = graph_dt, weights = abs(edge_attr(graph_dt, name = "weight")),
                                        rand.seed = rand_seed, num_trials = 10))
            vertex_attr(graph_dt, name = "Infomaps.web") <- paste0("CL_", membership_dt)
            return(graph_dt)
        })
    })
})


lapply(names(nw_graphs), function(score_id){
    score_dt <- nw_graphs[[score_id]]
    lapply(names(score_dt), function(weight_id){

        weight_dt <- score_dt[[weight_id]]
        lapply(weight_dt, function(graph_dt){
            hist(table(vertex_attr(graph_dt, name = "Infomaps.web")), breaks = 100)
        })
    })
})

```

```{r}
# table_list <- lapply(seq_along(clu_vertex), function(id){
#     clu_ids <- clu_vertex[[id]]
#     tx <- sapply(strsplit(clu_ids, "\\."), function(x){x[[1]]})
#     return(data.frame("Omic_Name" = tx, "clu_id" = id))
# })
# table_list <- do.call(rbind, table_list)
# table(table_list)
```

```{r}
## DMR & RNA_Seq enrich

# library(org.Mm.eg.db)
# library(GOSemSim)
# library(rGREAT) 
# enrich_multi_omic_clusters <- function(graph, membership_attribute, universe, min_n = 25){
#     
#     membership_vector <- vertex_attr(graph, membership_attribute)
#     clu_vertex <- lapply(unique(membership_vector), function(cl_id){
#         names(V(graph))[membership_vector == cl_id]
#     })
#     clu_vertex <- clu_vertex[lengths(clu_vertex) >= min_n]
#     rna_seq_tag <- "RNA_Sequencing\\."
#     dna_seq_tag <- "WGB_DNA_Methylation\\."
#     
#     
#     
#     general_tag <- paste0(rna_seq_tag, "|", dna_seq_tag)
#     
#     enrich_res <- lapply(clu_vertex, function(clu_ids){
#         rna_seq_ids <- grep(rna_seq_tag, clu_ids, value = TRUE)
#         rna_seq_ids <- gsub(rna_seq_tag, "", rna_seq_ids)
#         
#         clusterProfiler::enrichGO(
#             gene = rna_seq_ids, OrgDb = org.Mm.eg.db, 
#             keyType = "ENSEMBL", ont = "BP", 
#             pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
#             universe = universe_rna_seq)
#     }) 
#         
#     # universe_rna_seq <- grep(rna_seq_tag, universe, value = TRUE)
#     # universe_rna_seq <- gsub(rna_seq_tag, "", universe_rna_seq)
#         
#     # universe_dna_seq <- grep(dna_seq_tag, universe, value = TRUE)
#     # universe_dna_seq <- gsub(dna_seq_tag, "", universe_dna_seq)
#     # universe_dna_seq <- rowRanges(omic_dt[["WGB_DNA_Methylation"]][universe_dna_seq,])
#     # bp_data <- godata(OrgDb = org.Mm.eg.db, ont = "BP")
#     
#     # enrich_res <- lapply(clu_vertex, function(clu_ids){
#     #     rna_seq_ids <- grep(rna_seq_tag, clu_ids, value = TRUE)
#     #     rna_seq_ids <- gsub(rna_seq_tag, "", rna_seq_ids)
#     #     rna_seq_enrich <- clusterProfiler::enrichGO(
#     #         gene = rna_seq_ids, OrgDb = org.Mm.eg.db, 
#     #         keyType = "ENSEMBL", ont = "BP", 
#     #         pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
#     #         universe = universe_rna_seq)
#     #     rna_seq_enrich <- as.data.frame(rna_seq_enrich)
#         
#         # dna_seq_ids <- grep(dna_seq_tag, clu_ids, value = TRUE)
#         # dna_seq_ids <- gsub(dna_seq_tag, "", dna_seq_ids)
#         # dna_seq_ids <- rowRanges(omic_dt[["WGB_DNA_Methylation"]][dna_seq_ids,])
#         # 
#         # dna_seq_enrich <- great(gr = dna_seq_ids, 
#         #                         gene_sets = "GO:BP",  
#         #                         tss_source = "TxDb.Mmusculus.UCSC.mm39.refGene", 
#         #                         background = universe_dna_seq)
#         # 
#         # dna_seq_enrich <- getEnrichmentTable(dna_seq_enrich)
#         # dna_seq_enrich <- subset(dna_seq_enrich, p_adjust <= 0.05)
#         
#         # return(list("RNA_Seq_Enrich" = rna_seq_enrich, 
#         #             "DMR_Enrich" = dna_seq_enrich))
#     #     rna_seq_enrich
#     # }) 
#     enrich_res
# }
membership_attribute <- "Infomaps.web"
min_n = 10;

rna_seq_tag <- "RNA_Sequencing\\."
dna_seq_tag <- "WGB_DNA_Methylation\\."
general_tag <- paste0(rna_seq_tag, "|", dna_seq_tag)


dna_seq_dict <- rowData(omic_dt[["WGB_DNA_Methylation"]])$geneId
names(dna_seq_dict) <- rownames(omic_dt[["WGB_DNA_Methylation"]])
rna_seq_dict <- rowData(omic_dt[["RNA_Sequencing"]])$entrezID
names(rna_seq_dict) <- rownames(omic_dt[["RNA_Sequencing"]])

universe_dict <- c(dna_seq_dict, rna_seq_dict)
universe_dict


# enrich_results <- lapply(names(nw_graphs), function(score_id){
#     score_dt <- nw_graphs[[score_id]]
#     score_results <- lapply(names(score_dt), function(weight_name){
#         weight_dt <- score_dt[[weight_name]][["All Omics"]]
#         universe_dt <- universe_list[[score_id]][[weight_name]][["All Omics"]]
#         
#         universe <- grep(general_tag, universe_dt, value = TRUE)
#         universe <- gsub(general_tag, "", universe)
#         
#         universe <- universe_dict[universe]
#         
#         membership_vector <- vertex_attr(weight_dt, membership_attribute)
#         clu_vertex <- lapply(unique(membership_vector), function(cl_id){
#             names(V(weight_dt))[membership_vector == cl_id]
#         })
#         names(clu_vertex) <- unique(membership_vector)
#         clu_vertex <- clu_vertex[lengths(clu_vertex) >= min_n]
# 
#         enrich_res <- lapply(clu_vertex, function(clu_ids){
#             clu_ids_t <- grep(general_tag, clu_ids, value = TRUE)
#             clu_ids_t <- gsub(general_tag, "", clu_ids_t)
#             clu_ids_t <- universe_dict[clu_ids_t]
#         
#             enrich_clu <- clusterProfiler::enrichGO(
#                 gene = clu_ids_t, OrgDb = org.Mm.eg.db, 
#                 keyType = "ENTREZID", ont = "BP", 
#                 pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
#                 universe = universe)
#             if(nrow(as.data.frame(enrich_clu)) == 0){return(NA)}
#             clusterProfiler::simplify(enrich_clu)
#         }) 
#         enrich_res[!is.na(enrich_res)]
#     })
#     names(score_results) <- names(score_dt)
#     score_results
# })
# names(enrich_results) <- names(nw_graphs)
```

```{r}
library(clusterProfiler)
# 
# 
# 
# graph <- nw_graphs$pearson$Factor_1$`All Omics`
# 
# integ_res_dir <- paste0(corr_output_dir, "/", score_id, "/")
# if(!dir.exists(integ_res_dir)){dir.create(integ_res_dir)}
# 
# output_path <- paste0(integ_res_dir, "/Factor_1/")
# results_name <- "Factor_1_All"


clu_ids_enrich <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(factor_dt){
        graph <- factor_dt[["All Omics"]]
        membership_vector <- vertex_attr(graph, membership_attribute)
        clu_vertex <- lapply(unique(membership_vector), function(cl_id){
            names(V(graph))[membership_vector == cl_id]
        })
        names(clu_vertex) <- unique(membership_vector)
        cluster_id_list <-  lapply(clu_vertex, function(clu_ids){
            if(length(clu_ids) <= min_n){return(NA)}
            metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
                                                                            clu_ids, value = T)]
            metabo_dt <- metabo_dt[!is.null(metabo_dt)]
            metabo_dt <- strsplit(metabo_dt, ";")
            metabo_dt <- metabo_dt[!is.na(metabo_dt)]
            
            wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
                                                                            clu_ids, value = T)]
            wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
            
            rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
                                                                            clu_ids, value = T)]
            rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
            
            return(c(metabo_dt, rna_dt, wgb_dt))
            # sapply(clu, function(clu_ids){
            #     browser()
            #     
            #     
            #     
            #     omic_dt <- strsplit(clu_ids, "\\.")
            # })
        })
        cluster_id_list[!is.na(cluster_id_list)]
    })
})

universe_list_ids <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(factor_dt){
        graph <- factor_dt[["All Omics"]]
        
        universe_ids <- names(V(graph))
    
        metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
                                                                            universe_ids, value = T)]
        metabo_dt <- metabo_dt[!is.null(metabo_dt)]
        metabo_dt <- unlist(strsplit(metabo_dt, ";"))
        metabo_dt <- metabo_dt[!is.na(metabo_dt)]
        
        wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
                                                                        universe_ids, value = T)]
        wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
        
        rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
                                                                        universe_ids, value = T)]
        rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
            
        universe_ids_f <- c(metabo_dt, wgb_dt, rna_dt)
        universe_ids_f
    })
})

enrich_results_list <- lapply(names(clu_ids_enrich), function(score_id){
    factor_results <- lapply(names(clu_ids_enrich[[score_id]]), function(factor_id){
        enrich_results <- lapply(clu_ids_enrich[[score_id]][[factor_id]], function(clu_ids){
            
            clu_ids <- unlist(clu_ids)
            
            enrich_res <- enricher(gene = clu_ids, universe = universe_list_ids[[score_id]][[factor_id]], 
                                    pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                                    TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)
            if(nrow(as.data.frame(enrich_res)) == 0){return(NA)}
            return(enrich_res)
        })
        return(enrich_results[!is.na(enrich_results)])
    })
    names(factor_results) <- names(clu_ids_enrich[[score_id]])
    factor_results
})
names(enrich_results_list) <- names(clu_ids_enrich)
```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        
        weight_dt <- factor_dt[[weight_id]]
        nw_output_dir <- paste0(corr_output_dir, factor_id, "/")
        
        if(!dir.exists(nw_output_dir)){dir.create(nw_output_dir)}
        pdf(paste0(nw_output_dir, "Clu_Kegg_Joint_Pathway_Enrichment_", 
                    factor_id, "_", weight_id, ".pdf"))
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            plot_name <- gsub("_", " ", clu_id)
            
            print(dotplot(clu_dt) + ggtitle(paste("Kegg joint pathway enrichment" , plot_name)))
        })
        dev.off()
    })
})
```


```{r}
## All MOFA2 background
# full_mofa_features <- unlist(rownames(omic_dt))
# full_mofa_features <- paste0(names(full_mofa_features), ".", full_mofa_features)
# 
# metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
#                                                                     full_mofa_features, value = T)]
# metabo_dt <- metabo_dt[!is.null(metabo_dt)]
# metabo_dt <- unlist(strsplit(metabo_dt, ";"))
# metabo_dt <- metabo_dt[!is.na(metabo_dt)]
# 
# wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
#                                                                 full_mofa_features, value = T)]
# wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
# 
# rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
#                                                                 full_mofa_features, value = T)]
# rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
# 
# full_universe_ids <- c(metabo_dt, wgb_dt, rna_dt)
# 
# enrich_results_list_full <- lapply(names(clu_ids_enrich), function(score_id){
#     factor_results <- lapply(names(clu_ids_enrich[[score_id]]), function(factor_id){
#         enrich_results <- lapply(clu_ids_enrich[[score_id]][[factor_id]], function(clu_ids){
#             enrich_res <- enricher(gene = clu_ids, universe = full_universe_ids, 
#                                     pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
#                                     TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)  
#             if(nrow(as.data.frame(enrich_res)) == 0){return(NA)}
#             return(enrich_res)
#         })
#         return(enrich_results[!is.na(enrich_results)])
#     })
#     names(factor_results) <- names(clu_ids_enrich[[score_id]])
#     factor_results
# })
# names(enrich_results_list_full) <- names(clu_ids_enrich)

# clu_dict_list <- lapply(clu_ids_enrich, function(score_dt){
#     lapply(score_dt, function(weight_dt){
#         lapply(weight_dt, function(clu_dt){
#             dict_res <- names(clu_dt)
#             names(dict_res) <- clu_dt
#             dict_res    
#         })
#     })
# })

```

```{r}
enrich_results_selected <- enrich_results_list

for(score_id in names(nw_graphs)){
    for(weight_id in names(nw_graphs[[score_id]])){
        dict_dt <- clu_ids_enrich[[score_id]][[weight_id]]
        
        term_list <- lapply(names(enrich_results_selected[[score_id]][[weight_id]]), 
                            function(clu_id){
                                res_enrich <- as.data.frame(enrich_results_selected[[score_id]][[weight_id]][[clu_id]])
                                if(nrow(res_enrich) == 0){return(NA)}
                                res_enrich$ID <- paste0("kegg_pathway.", clu_id, ".", res_enrich$ID)
                                res_enrich$Infomaps.web <- clu_id
                                res_enrich
                            })
        names(term_list) <- names(enrich_results_selected[[score_id]][[weight_id]])

        term_list <- term_list[!is.na(term_list)]
        
        
        
        ## Define nodes n1
        term_dt <- lapply(names(term_list), function(clu_id){
            geneID_list <- strsplit(term_list[[clu_id]]$geneID, "/")
            names(geneID_list) <- term_list[[clu_id]]$ID
            
            found_ids <- lapply(names(geneID_list), function(pathway_ids){
                ids <- geneID_list[[pathway_ids]]
                
                found <- names(dict_dt[[clu_id]])[sapply(dict_dt[[clu_id]], function(ids_dict){any(ids_dict %in% ids)})]
                
                # found <- neighborhood(graph = nw_graphs[[score_id]][[weight_id]][["All Omics"]],
                #                     order = 1, nodes = found)
                # browser()
                # unique(unlist(sapply(found, names)))
                unique(unlist(found))
                # unlist(strsplit(paste(pathway_ids,  unique(found), sep = "\t"), "\t"))
            })
            names(found_ids) <- names(geneID_list)
            found_ids
            # Reduce(c, found_ids)
        })
        names(term_dt) <- names(term_list) 
        
        ## Define edges n1
        
        term_dt_n1 <- lapply(term_dt, function(clu_dt){
            
            found_ids <- lapply(names(clu_dt), function(pathway_ids){
                ids <- clu_dt[[pathway_ids]]
                ids <- unique(unlist(ids))
                unlist(strsplit(paste(pathway_ids,  unique(ids), sep = "\t"), "\t"))
            })
            names(found_ids) <- names(clu_dt)
            # found_ids
            Reduce(c, found_ids)
        })
        names(term_dt_n1) <- names(term_list) 
        
        term_dt_n1 <- Reduce(c, term_dt_n1)
        
        
        
        ## Define edges n2
        term_dt_n2 <- lapply(term_dt, function(clu_dt){
            
            found_ids <- lapply(names(clu_dt), function(pathway_ids){
                ids <- clu_dt[[pathway_ids]]
                
                found <- neighborhood(graph = nw_graphs[[score_id]][[weight_id]][["All Omics"]],
                                    order = 1, nodes = ids)
                found <- unique(unlist(sapply(found, names)))
                unlist(strsplit(paste(pathway_ids,  unique(found), sep = "\t"), "\t"))
            })
            names(found_ids) <- names(clu_dt)
            # found_ids
            Reduce(c, found_ids)
        })
        names(term_dt_n2) <- names(term_list) 
        
        term_dt_n2 <- Reduce(c, term_dt_n2)
        
        # term_description <- term_list$Description
        # names(term_description) <- term_list$ID
        
        ## Define Nodes
        term_list <- do.call(rbind, term_list)
        
        nw_graphs[[score_id]][[weight_id]][["All Omics"]] <- add.vertices(
            graph = nw_graphs[[score_id]][[weight_id]][["All Omics"]], 
            nv = nrow(term_list), 
            attr = list("name" = term_list$ID, 
                        "Readable_Name" = term_list$Description, 
                        "Node_Omic" = "kegg pathway", 
                        "Infomaps.web" = term_list$Infomaps.web))
        
        nw_graphs[[score_id]][[weight_id]][["All Omics"]] <- add.edges(
            graph = nw_graphs[[score_id]][[weight_id]][["All Omics"]], 
            edges = term_dt_n1, attr = list("weight" = 0, "Type" = "kegg_n1"))
        
        nw_graphs[[score_id]][[weight_id]][["All Omics"]] <- add.edges(
            graph = nw_graphs[[score_id]][[weight_id]][["All Omics"]], 
            edges = term_dt_n2, attr = list("weight" = 0, "Type" = "kegg_n2"))
    }
}
```


```{r}
library(MultiAssayExperiment)
library(RCy3)
omic_names_dup <- sapply(strsplit(names(experiments(omic_dt)), "\\."), function(x){x[[1]]})

unique_omics <- unique(omic_names_dup)

omic_colors <- RColorBrewer::brewer.pal(n = length(unique_omics), name = "Set3")

col_list <- lapply(seq_along(unique_omics), function(omic_id){
    omic_name <- unique_omics[omic_id]
    omic_num <- length(omic_names_dup[omic_names_dup == omic_name])
    omic_col <- omic_colors[omic_id]
    
    col_ramp <- colorRampPalette(colors = c(omic_col, "white"))(omic_num + 8)[seq_len(omic_num)]
    return(col_ramp)
})
color_vect <- unlist(col_list)
names(color_vect) <- names(experiments(omic_dt))

cytoscapePing()

shape_vect <- RCy3::getNodeShapes()[seq_along(unique_omics)]
shape_vect[is.na(shape_vect)] <- "ELLIPSE"

shape_vect_ext <- unlist(lapply(seq_along(shape_vect), function(omic_id){
    omic_name <- unique_omics[omic_id]
    shape_omic <- shape_vect[omic_id]
    omic_num <- length(omic_names_dup[omic_names_dup == omic_name])
    
    return(rep(shape_omic, omic_num))
}))
names(shape_vect_ext) <- names(experiments(omic_dt))

term_network_list <- list()

for(score_id in names(nw_graphs)){
    score_dt <- nw_graphs[[score_id]]
    
    term_network_list[[score_id]] <- list()
    
    for(weight_id in names(score_dt)){
    
        term_network_list[[score_id]][[weight_id]] <- list()

        weight_dt <- score_dt[[weight_id]][["All Omics"]]
        
        clu_dt <- vertex_attr(weight_dt, name = "Infomaps.web")
        # vertex.color = vertex_attr(test_sub, name = "Node_Omic")
        # plot.igraph(omic_network, vertex.size = 0.5, edge.size = 0.1, 
        #             vertex.label.cex = 0.2, edge.arrow.size = 0.1, 
        #             layout = layout_as_tree(omic_network))
        sufix_id <- paste0(score_id, ".", weight_id)
        
        RCy3::createNetworkFromIgraph(weight_dt)
        RCy3::renameNetwork(title = sufix_id)
        
        RCy3::copyVisualStyle(from.style = "default", 
                              to.style = sufix_id)
        RCy3::setVisualStyle(sufix_id)
        
        set.seed(1234)
        clu_colors <- RCy3::paletteColorRandom(value.count = length(unique(clu_dt)))
        
        
        # RCy3::setNodeColorMapping(table.column = "Infomaps.web",
        #                           table.column.values = unique(clu_dt),
        #                           colors = clu_colors,
        #                           mapping.type = "discrete", style.name = sufix_id)
        RCy3::setNodeColorMapping(table.column = "Node_Omic",
                                  table.column.values = names(color_vect),
                                  colors = color_vect,
                                  mapping.type = "discrete", style.name = sufix_id)
        
        RCy3::setNodeShapeMapping(table.column = "Node_Omic", 
                                  table.column.values = names(shape_vect_ext), 
                                  shapes = shape_vect_ext, style.name = sufix_id)
        
        RCy3::setNodeLabelMapping(table.column = "Readable_Name", style.name = sufix_id)
        
        RCy3::setNodeFontSizeDefault(new.size = 6, style.name = sufix_id)
        
        small_node <- names(V(weight_dt))[grepl("Shotgun_Metagenomics_Functional", 
                                                vertex_attr(weight_dt, name = "Node_Omic"))]
        
        RCy3::setNodePropertyBypass(node.names = small_node, new.values = 60, 
                                    visual.property = "NODE_LABEL_WIDTH")
        RCy3::setNodePropertyBypass(node.names = small_node, new.values = 4, 
                                    visual.property = "NODE_LABEL_FONT_SIZE")
        
        # RCy3::setNodeColorMapping(table.column = make.names(params$fold_change_variable), 
        #                         colors = colorRampPalette(colors = c("blue", 
        #                                                             "white", 
        #                                                             "red")),
        #                         style.name = params$output_suffix)
        RCy3::setEdgeColorMapping(table.column = "weight", 
                                colors = colorRampPalette(colors = c("#B3CDE3", 
                                                                    "#FBB4AE")), 
                                style.name = sufix_id)
        
        
        for(clu_id in names(enrich_results_selected[[score_id]][[weight_id]])){
            
            term_network_list[[score_id]][[weight_id]][[clu_id]] <- list()
            
            clu_name <- paste0(sufix_id, ".", clu_id)
            
            
            RCy3::createSubnetwork(nodes = clu_id, nodes.by.col = "Infomaps.web", 
                                   subnetwork.name = clu_name, network = sufix_id)
            
            clu_nw <- subgraph(graph = weight_dt, 
                                vids = V(weight_dt)[vertex_attr(graph = weight_dt, 
                                                                name = "Infomaps.web") == clu_id])
            
            clu_nw_n2 <- subgraph.edges(clu_nw, 
                                        E(clu_nw)[grepl("kegg_n2", edge_attr(clu_nw, name = "Type"))])
            
            clu_vertex <- V(clu_nw_n2)[vertex_attr(clu_nw_n2, name = "Node_Omic") == "kegg pathway"]
            
            term_nw_list <- lapply(names(clu_vertex), function(term_name){
                neighbors_ids <- names(neighbors(clu_nw_n2, term_name))
                neighbors_ids <- neighbors_ids[!grepl("RNA_Sequencing|Untargeted_Metabolomics|WGB_DNA|kegg_pathway",
                                                    neighbors_ids)]
                c(neighbors_ids, term_name)
            })
            names(term_nw_list) <- names(clu_vertex)

            
            # term_neighbors <- lapply(seq_along(clu_vertex), function(x){
            #     neighborhood_2 <- names(neighborhood(weight_dt, order = 2, 
            #                                         nodes = clu_vertex[x])[[1]])
            #     neighborhood_1 <- names(neighborhood(weight_dt, order = 1, 
            #                                         nodes = clu_vertex[x])[[1]])
            #     return(neighborhood_2[!neighborhood_2 %in% neighborhood_1])
            # })
            # names(term_neighbors) <- names(clu_vertex)
            # 
            # term_neighbors_filtered <- lapply(names(term_neighbors), function(term_names){
            #     term_ids <- term_neighbors[[term_names]]
            #     filtered_ids <- term_ids[!grepl("RNA_Sequencing|Untargeted_Metabolomics|WGB_DNA|kegg_pathway",
            #                                     term_ids)]
            #     c(filtered_ids, term_names)
            # })
            
            # names(term_neighbors_filtered) <- names(clu_vertex)
            # term_neighbors_n2 <- lapply(seq_along(clu_vertex), function(x){
            #     igraph::V
            # })
            
            
            term_description <- vertex_attr(weight_dt, name = "Readable_Name", names(clu_vertex))
            
            for(term_id in seq_along(term_nw_list)){
                
                term_network_list[[score_id]][[weight_id]][[clu_id]][[term_description[term_id]]] <- subgraph(clu_nw_n2, 
                                                                              vids = term_nw_list[[term_id]])
                
                term_name <- paste0(clu_name, ".", term_description[term_id])
                
                RCy3::createSubnetwork(nodes = term_nw_list[[term_id]], nodes.by.col = "name", 
                                   subnetwork.name = term_name, network = clu_name)
                RCy3::layoutNetwork(layout.name = "hierarchical", network = term_name)
                
                network_output_dir <- paste0(corr_output_dir, score_id, "/")
                
                if(!dir.exists(network_output_dir)){
                    dir.create(network_output_dir)
                }
                if(length(term_nw_list[[term_id]]) <= 1){next}
                Sys.sleep(time = 2.5)
                
                RCy3::exportSVG(filename = paste0(network_output_dir, 
                                                make.names(term_name), ".svg"),
                                network = term_name)
            }
         }
        
    }
}
saveSession(paste0(corr_output_dir, "correlation_results_f1f2_network_full"))
```

```{r}
omic_names_dict <- c("Histone Modifications", "Liver Metallomics", 
                            "Cecum Metallomics", "RNA Sequencing", 
                            "Cecum Targeted Metabolomics", "Liver Targeted Metabolomics", 
                            "Serum Targeted Metabolomics", "Untargeted Metabolomics", 
                            "Liver DNA Methylation", 
                            "Cecum 16S Metagenomics", "Colon 16S Metagenomics", 
                            "Cecum Shotgun Metagenomics", "Colon Shotgun Metagenomics", 
                            "Cecum Shotgun Functional Metaerg KEGG", "Shotgun Functional Colon Metaerg KEGG", 
                            "Cecum Shotgun Functional COGS", "Shotgun Functional Colon Metaerg COGS")
names(omic_names_dict) <- names(experiments(omic_dt))

feature_ids_dict <- lapply(names(variables_dicts), function(dataset_name){
        
    dataset_var <- rowData(omic_dt[[dataset_name]])[variables_dicts[[dataset_name]]]
    dataset_vector <- unlist(dataset_var[[1]])
    names(dataset_vector) <- paste0(dataset_name, ".", rownames(dataset_var))
    
    dataset_vector
})
names(feature_ids_dict) <- names(variables_dicts)

# feature_ids <- term_nw_list[[term_id]]

# plot_title <- term_name

if(!is.null(omic_names_dict)){
    names(color_vect) <- omic_names_dict[as.character(names(color_vect))]
}

group_color_dict <- RColorBrewer::brewer.pal(n = length(levels(omic_dt$group)), 
                                             name = "Set1")
names(group_color_dict) <- levels(omic_dt$group)

plot_nw_term <- function(omic_dt, feature_ids, group_var = "group", 
                        sample_var = "samplename", feature_ids_dict = NULL, 
                        omic_names_dict = NULL, group_color_dict, plot_title){
    plot_dt <- lapply(names(experiments(omic_dt)), function(omic_id){
        omic_filt <- omic_dt[[omic_id]]
        found <- paste0(omic_id, ".", rownames(omic_filt)) %in% feature_ids
        if(any(found)){
            return(omic_filt[found,])
        } else{
            return(NA)
        }
    })
    names(plot_dt) <- names(experiments(omic_dt))
    plot_dt <- plot_dt[!is.na(plot_dt)]
    
    library(reshape2)
    
    group_var_dict <- colData(omic_dt)[[group_var]]
    names(group_var_dict) <- colData(omic_dt)[[sample_var]]
    
    # Melt data to plot and add omic name and group variable
    plot_dt <- lapply(names(plot_dt), function(omic_exp_name){
        omic_exp <- plot_dt[[omic_exp_name]]
        melted_dt <- melt(as.matrix(assay(omic_exp)))
        
        colnames(melted_dt)[1:2] <- c("var_id", "sample_name")
        # melted_dt$var_id <- factor(gsub(paste0(omic_exp_name, "."), "", melted_dt$var_id))
         melted_dt$var_id <- factor(paste(omic_exp_name, 
                                        melted_dt$var_id, sep = "."))
        melted_dt$omic_name <- factor(omic_exp_name)
        
        melted_dt$group <- factor(group_var_dict[as.character(melted_dt$sample_name)])
        
        if(!is.null(omic_names_dict)){
            melted_dt$omic_name <- factor(omic_names_dict[as.character(melted_dt$omic_name)])
        }
        if(!is.null(feature_ids_dict)){
            melted_dt$var_id <- factor(feature_ids_dict[[omic_exp_name]][as.character(melted_dt$var_id)])
        }
        
        melted_dt
    })
    
    plot_dt <- do.call(rbind, plot_dt)
    
    plot_dt$var_id <- factor(str_wrap(plot_dt$var_id, width = 40))
    
    sample_levels <- levels(plot_dt$sample_name)
    reordered_sample_levels <- sample_levels[order(group_var_dict[as.character(sample_levels)])]
    
    plot_dt$sample_name <- reorder(plot_dt$sample_name, 
                                    new.order = reordered_sample_levels)
    
    x_colors <- group_color_dict[as.character(group_var_dict[levels(plot_dt$sample_name)])]
    
    if(all(is.null(plot_dt))){browser()}
    labels_dt <- distinct(plot_dt[,c("var_id", "omic_name")])
    # labels_dt <- labels_dt[order(labels_dt$omic_name),]
    
    var_omic_dict <- labels_dt$omic_name
    names(var_omic_dict) <- labels_dt$var_id
    
    var_levels <- levels(plot_dt$var_id)
    reordered_var_levels <- var_levels[order(var_omic_dict[as.character(var_levels)])]
    
    plot_dt$var_id <- reorder(plot_dt$var_id, 
                            new.order = reordered_var_levels)
    
    legend_colors <- color_vect[as.character(sapply(levels(plot_dt$var_id), function(x){
        labels_dt$omic_name[as.character(labels_dt$var_id) == x]
    }))]

    names(legend_colors) <- levels(plot_dt$var_id)
    legend_colors
    
    p <- ggplot(plot_dt, aes(x = sample_name, y = value, color = omic_name, 
                             group = var_id, shape = var_id)) + 
        geom_point() + geom_smooth(aes(color = omic_name, fill = omic_name),
                                   se = FALSE, span = 0.5, alpha = 0.1) + theme_minimal() + 
        theme(axis.text.x = element_text(angle = 90, colour = x_colors), 
                legend.position = "right", 
                legend.justification = "top", 
                legend.title = element_blank(), 
                plot.title = element_text(hjust = 0.5),
                legend.spacing.y = unit(0.1, "cm")) + 
        ggtitle(plot_title) + 
        xlab(NULL) + ylab("Normalized Values") + 
        scale_color_manual(values = color_vect)  + 
        guides(shape = guide_legend(label.theme = element_text(size = 7.5), 
                                    order = 2,  byrow = TRUE,
                                    override.aes = list(
                                        color = legend_colors
                                    ),
                                    keyheight = 0.05), 
               col = guide_legend(order = 1), fill = "none") + 
        scale_shape_manual(values = rep(16, length(unique(plot_dt$var_id)))) 
    return(p)
}
```


```{r}
lapply(names(term_network_list), function(score_id){
    score_nw <- term_network_list[[score_id]]
    lapply(names(score_nw), function(factor_id){
        factor_nw <- score_nw[[factor_id]]
        
        pdf(paste0(corr_output_dir, "Correlation_Plots_", 
                        score_id, "_", factor_id, ".pdf"), width = 12, height = 8)
        lapply(names(factor_nw), function(clu_id){
            
            clu_nw <- factor_nw[[clu_id]]
            lapply(names(clu_nw), function(term_id){
                term_nw <- clu_nw[[term_id]]
                feature_ids <- names(V(term_nw))
                
                if(length(feature_ids) == 1){return(NULL)}
                
                print(plot_nw_term(omic_dt = omic_dt, feature_ids = feature_ids,
                            feature_ids_dict = feature_ids_dict, 
                            omic_names_dict = omic_names_dict,
                            group_color_dict = group_color_dict,
                            plot_title = paste(clu_id, term_id)))
            })
            
        })
        dev.off()
    })
})
```



```{r}
# annotate(x = 20, y = 10, geom = "text", label = labels_dt$var_id)
# print(p)
# 
# labels_dt <- distinct(plot_dt[,c("omic_var_id_readable", "omic_name_readable", "omic_var_id")])
# labels_dt <- labels_dt[order(labels_dt$omic_name_readable),]
# 
# ###
# 
# legend_colors <- color_vect[as.character(sapply(levels(plot_dt$omic_var_id_readable), function(x){
#     labels_dt$omic_var_id_readable[as.character(labels_dt$omic_var_id_readable) == x]
# }))]
# 
# 
# p <- ggplot(plot_dt, aes(x = sample_name, y = value, color = omic_name_readable, 
#                          group = omic_var_id_readable, 
#                          shape = omic_var_id_readable)) + 
#     geom_point() + geom_smooth(aes(color = omic_name_readable, fill = omic_name_readable),
#                                se = TRUE, span = 0.5, alpha = 0.1) + theme_minimal() + 
#     theme(axis.text.x = element_text(angle = 90), 
#             legend.position = "right", 
#             legend.justification = "top", 
#             legend.title = element_blank(), 
#             plot.title = element_text(hjust = 0.5),
#             legend.spacing.y = unit(0.1, "cm")) + 
#     ggtitle(term_name) + 
#     xlab(NULL) + ylab("Normalized Values") + 
#     scale_color_manual(values = color_vect) +
#     scale_fill_manual(values = color_vect) + 
#     guides(shape = guide_legend(label.theme = element_text(size = 7.5), 
#                                 order = 2,  byrow = TRUE,
#                                 override.aes = list(
#                                     color = legend_colors
#                                 ),
#                                 keyheight = 0.05), 
#            col = guide_legend(order = 1), fill = "none") + 
#     scale_shape_manual(values = rep(16, length(unique(plot_dt$omic_var_id))))
#     # annotate(x = 20, y = 10, geom = "text", label = labels_dt$var_id)
# print(p)
# 
# #######
# # test <- reshape2::dcast(plot_dt, formula = 1 ~ omic_var_id, fun.aggregate = mean)
# library(dplyr)
# av_dt <- plot_dt %>% group_by(group, omic_var_id) %>% summarise(avg = mean(value), dt_sd = sd(value), omic_name = unique(omic_name))
# 
# p2 <- ggplot(av_dt, aes(x = group, y = avg, color = omic_name, 
#                          group = omic_var_id, shape = omic_var_id)) + 
#     geom_point() + geom_line() + geom_ribbon(aes(ymax = avg + dt_sd, 
#                        ymin = avg - dt_sd, fill = omic_name),
#                        alpha=0.2) + theme_minimal() + 
#     theme(axis.text.x = element_text(angle = 90), 
#             legend.position = "right", 
#             legend.justification = "top", 
#             legend.title = element_blank(), 
#             plot.title = element_text(hjust = 0.5),
#             legend.spacing.y = unit(0.1, "cm")) + 
#     ggtitle(term_name) + 
#     xlab(NULL) + ylab("Normalized Values") + 
#     scale_color_manual(values = color_vect) + 
#     scale_fill_manual(values = color_vect) +
#     guides(shape = guide_legend(label.theme = element_text(size = 7.5), 
#                                 order = 2,  byrow = TRUE,
#                                 override.aes = list(
#                                     color = legend_colors
#                                 ),
#                                 keyheight = 0.05), 
#            col = guide_legend(order = 1), fill = "none") + 
#     scale_shape_manual(values = rep(16, length(unique(plot_dt$omic_var_id))))
#     # annotate(x = 20, y = 10, geom = "text", label = labels_dt$var_id)
# print(p2)

```



```{r}
# enrich_graph_clusters <- function(cluster_id_list, membership_attribute, 
#                                   universe, min_n = 25){
#     membership_vector <- vertex_attr(graph, membership_attribute)
#     clu_vertex <- lapply(unique(membership_vector), function(cl_id){
#         names(V(graph))[membership_vector == cl_id]
#     })
#     clu_vertex <- clu_vertex[lengths(clu_vertex) >= min_n]
#     rna_seq_tag <- "RNA_Sequencing\\."
#     dna_seq_tag <- "WGB_DNA_Methylation\\."
#     
#     universe_rna_seq <- grep(rna_seq_tag, universe, value = TRUE)
#     universe_rna_seq <- gsub(rna_seq_tag, "", universe_rna_seq)
#     
#     universe_dna_seq <- grep(dna_seq_tag, universe, value = TRUE)
#     universe_dna_seq <- gsub(dna_seq_tag, "", universe_dna_seq)
#     universe_dna_seq <- rowRanges(omic_dt[["WGB_DNA_Methylation"]][universe_dna_seq,])
# }


 # KEGGREST::keggFind(database = "compound", query = universe_features_kegg[1])[[1]]$PATHWAY

# mSet<-InitDataObjects("conc", "pathinteg", FALSE)
# mSet<-SetOrganism(mSet, "mmu")
# 
# geneListFile<-"replace_with_your_file_name"
# geneList<-readChar(geneListFile, file.info(geneListFile)$size)
# 
# mSet<-PerformGeneMapping(mSet, geneList, "hsa", "symbol");
# 
# cmpdListFile<-"replace_with_your_file_name"
# cmpdList<-readChar(cmpdListFile, file.info(cmpdListFile)$size)
# mSet<-PerformCmpdMapping(mSet, cmpdList, "hsa", "name");
# mSet<-CreateMappingResultTable(mSet)
# 
# mSet<-PrepareIntegData(mSet);
# mSet<-PerformIntegPathwayAnalysis(mSet, "dc", "hyper", "all", "query");
# 
# mSet<-PlotPathSummary(mSet, F, "path_view_0_", "png", 72, width=NA, NA, NA )
# mSet<-CreateIntegMatchingTable(mSet);
# 
# saveRDS(mSet, paste0(output_path, results_name))
```


<!-- ## LIMA full unsupervised -->

<!-- ```{r} -->
<!-- multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|DNA_Methylation.liver", multi_omics_names)] -->
<!-- # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)]) -->
<!-- multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)]) -->

<!-- multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected]) -->
<!-- plot_data_overview(multi_omics_mofa) -->
<!-- ``` -->

<!-- ## Parameters -->

<!-- ```{r} -->
<!-- data_opts <- get_default_data_options(multi_omics_mofa) -->
<!-- head(data_opts) -->

<!-- model_opts <- get_default_model_options(multi_omics_mofa) -->
<!-- head(model_opts) -->
<!-- model_opts$num_factors <- 8 -->

<!-- train_opts <- get_default_training_options(multi_omics_mofa) -->
<!-- head(train_opts) -->
<!-- ``` -->

<!-- ## Model and training -->

<!-- ```{r} -->
<!-- multi_omics_mofa <- prepare_mofa( -->
<!--   object = multi_omics_mofa, -->
<!--   data_options = data_opts, -->
<!--   model_options = model_opts, -->
<!--   training_options = train_opts -->
<!-- ) -->

<!-- outfile <- file.path(getwd(),"model_us.hdf5") -->
<!-- multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE) -->

<!-- ``` -->

<!-- ## Downstream Analysis -->

<!-- ```{r} -->
<!-- # file_path <- "./model.hdf5" -->
<!-- # model <- load_model(file_path) -->
<!-- colData(multi_omics)$sample <- colData(multi_omics)$samplename -->

<!-- samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample") -->

<!-- plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90)) -->

<!-- p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5,  -->
<!--                  color_by = "group.y",  -->
<!--                  add_violin = TRUE) -->

<!-- p <- p +  -->
<!--   scale_color_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) + -->
<!--   scale_fill_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) -->
<!-- print(p) -->
<!-- ``` -->

<!-- ```{r, fig.height=12, fig.width=12} -->
<!-- plot_factors(multi_omics_mofa,  -->
<!--   factors = 1:6, -->
<!--   color_by = "group.y" -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pca_dt <- MOFA2::get_factors(object = multi_omics_mofa, factors = 1:3) -->

<!-- MOFA2::get_variance_explained(object = multi_omics_mofa, factors = 1:3) -->

<!-- library(plotly) -->

<!-- d3plot <- plot_ly(x = pca_dt$group1[,1], y = pca_dt$group1[,2], z = pca_dt$group1[,3], color = multi_omics_mofa@samples_metadata$group.y) -->
<!-- d3plot <- d3plot %>% add_markers() -->
<!-- # d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")), -->
<!-- #                      yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")), -->
<!-- #                      zaxis = list(title = paste0('PC4', "(", pcvar[4], "%)")))) -->
<!-- d3plot -->
<!-- ``` -->

<!-- ```{r} -->
<!-- plot_weights(multi_omics_mofa,  -->
<!--   view = 1, -->
<!--   factor = 2, -->
<!--   nfeatures = 10,     # Number of features to highlight -->
<!--   scale = T,          # Scale weights from -1 to 1 -->
<!--   abs = F            # Take the absolute value? -->
<!-- ) -->

<!-- weight_dt <- get_weights(object = multi_omics_mofa, factors = 2, as.data.frame = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(view_id in seq_len(length(multi_omics_selected))){ -->
<!--     plot_data_heatmap(multi_omics_mofa, -->
<!--       view = view_id,         # view of interest -->
<!--       factor = 2,             # factor of interest -->
<!--       features = 40,          # number of features to plot (they are selected by weight) -->

<!--       # extra arguments that are passed to the `pheatmap` function -->
<!--       cluster_rows = TRUE, cluster_cols = TRUE, -->
<!--       show_rownames = TRUE, show_colnames = FALSE, annotation_samples = "group.y", -->
<!--     ) -->
<!-- } -->
<!-- ``` -->


<!-- # ```{r} -->
<!-- # # param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/", -->
<!-- # #     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/", -->
<!-- # #     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "HFE"], -->
<!-- # #     group_var = "group", -->
<!-- # #     num_factors = num_factors, -->
<!-- # #     results_name = "mofa_int", -->
<!-- # #     color_vec = c("CNT"="black", "HFD"="red", "INT"="blue") -->
<!-- # # ) -->
<!-- # # # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green") -->
<!-- # # xfun::Rscript_call(rmarkdown::render, -->
<!-- # #                     args = list(input = "./MOFA_Models.Rmd",  -->
<!-- # #                                 output_file = paste0(param_list$output_dir, -->
<!-- # #                                                      param_list$results_name, ".html"),  -->
<!-- # #                                 params = param_list -->
<!-- # #                             ) -->
<!-- # #                     ) -->
<!-- #  -->
<!-- # param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/", -->
<!-- #     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/", -->
<!-- #     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "INT"], -->
<!-- #     group_var = "group", -->
<!-- #     num_factors = num_factors, -->
<!-- #     results_name = "mofa_hfe", -->
<!-- #     color_vec = c("CNT"="black", "HFD"="red", "HFE"="green") -->
<!-- # ) -->
<!-- # # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green") -->
<!-- # xfun::Rscript_call(rmarkdown::render, -->
<!-- #                     args = list(input = "./MOFA_Models.Rmd",  -->
<!-- #                                 output_file = paste0(param_list$output_dir, -->
<!-- #                                                      param_list$results_name, ".html"),  -->
<!-- #                                 params = param_list -->
<!-- #                             ) -->
<!-- #                     ) -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ## Pre-Analysis -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # mice_dt <- openxlsx::read.xlsx(paste0(params$input_dir, "AnimalData_OscarYanes.xlsx")) -->
<!-- #  -->
<!-- # # mice_num <- mice_dt[,c("BW(g)", "WATep(g)", "Liver(g)")] -->
<!-- # mice_num <- mice_dt -->
<!-- # mice_s <- as.data.frame(mice_num) -->
<!-- #  -->
<!-- # dict_vect <- c("Ctrl" = "CNT", "Ctrl ext" = "CNT", "HFD" = "HFD", "INT2" = "INT", "HFD excl." = "HFE") -->
<!-- # mice_s$group <- factor(dict_vect[mice_dt$Group], levels = c("CNT", "HFD", "INT", "HFE")) -->
<!-- #  -->
<!-- # dict_vect_g <- c("CNT" = "C", "HFD" = "H", "INT" = "I", "HFE" = "E") -->
<!-- #  -->
<!-- # mice_s$samplename <- paste0(make.names(mice_dt$ID), dict_vect_g[mice_s$group]) -->
<!-- #  -->
<!-- #  -->
<!-- # mice_s$samplename[!mice_s$samplename %in% multi_omics$samplename] -->
<!-- # multi_omics$samplename[!multi_omics$samplename %in% mice_s$samplename] -->
<!-- #  -->
<!-- # mice_s$samplename[mice_s$samplename == "X686I"] <- "X687I" -->
<!-- #  -->
<!-- # mice_s <- dplyr::left_join(as.data.frame(colData(multi_omics)), mice_s, by = "samplename") -->
<!-- # mice_s -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # d3plot <- plot_ly(x = mice_s$`BW(g)`, y = mice_s$`WATep(g)`, z = mice_s$`Liver(g)`, -->
<!-- #                     color = as.factor(mice_s$Sacrifice)) -->
<!-- # d3plot <- d3plot %>% add_markers() -->
<!-- # d3plot <- d3plot %>% layout(scene = list( -->
<!-- #     xaxis = list(title = "BW (g)"), -->
<!-- #     yaxis = list(title = "WATep (g)"), -->
<!-- #     zaxis = list(title = "Liver (g)"))) -->
<!-- # d3plot -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # factor_dt <- readRDS(paste0(params$output_dir, "/MOFA_Results/any_sig_005_mofa_int_all_meta/factor_data.rds")) -->
<!-- #  -->
<!-- # mice_sub <- mice_s[mice_s$group.x != "HFE",] -->
<!-- #  -->
<!-- # all(names(factor_dt$group1[,3]) == mice_sub$samplename) -->
<!-- #  -->
<!-- # cor(mice_sub$`BW(g)`, factor_dt$group1[,3]) -->
<!-- # cor(mice_sub$`WATep(g)`, factor_dt$group1[,3]) -->
<!-- # cor(mice_sub$`Liver(g)`, factor_dt$group1[,3]) -->
<!-- #  -->
<!-- #  -->
<!-- # ctr_sub <- mice_sub[mice_sub$group.x == "CNT",] -->
<!-- # ctr_fact <- factor_dt$group1[mice_sub$group.x == "CNT",] -->
<!-- #  -->
<!-- # cor(ctr_sub$`BW(g)`, ctr_fact[,3]) -->
<!-- # cor(ctr_sub$`WATep(g)`, ctr_fact[,3]) -->
<!-- # cor(ctr_sub$`Liver(g)`, ctr_fact[,3]) -->
<!-- #  -->
<!-- # plot(ctr_sub$`BW(g)`, ctr_fact[,3]) -->
<!-- # plot(ctr_sub$`WATep(g)`, ctr_fact[,3]) -->
<!-- # plot(ctr_sub$`Liver(g)`, ctr_fact[,3]) -->
<!-- #  -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ## GFCV -->
<!-- #  -->
<!-- # ## Loaded matrices -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # library(MOFA2) -->
<!-- # multi_omics_raw <- readRDS("./Merged_Multi_Omics_gfcv.rds") -->
<!-- # multi_omics <- readRDS("./Multivariate_Merged_Multi_Omics_gfcv.rds") -->
<!-- # names(experiments(multi_omics)) -->
<!-- #  -->
<!-- # filter_list <- c("16S") -->
<!-- #  -->
<!-- # selected_omics <- lapply(filter_list, function(dt_name){ -->
<!-- #     omic_dt_names <- grep(dt_name, names(experiments(multi_omics))) -->
<!-- #     omic_exp_list <- experiments(multi_omics[,,omic_dt_names]) -->
<!-- #      -->
<!-- #     sig_vars_num <- sapply(omic_exp_list, function(omic_exp){ -->
<!-- #         var_dt <- rowData(omic_exp) -->
<!-- #         p_val_dt <- var_dt[,grep("adj.p.value", colnames(var_dt))] -->
<!-- #         sig_vars <- apply(p_val_dt, 1, function(p_vals){ -->
<!-- #             any(p_vals <= 0.05, na.rm = TRUE) -->
<!-- #         }) -->
<!-- #         sum(unlist(sig_vars)) -->
<!-- #     }) -->
<!-- #     sig_vars_num -->
<!-- # }) -->
<!-- # names(selected_omics) <- filter_list -->
<!-- #  -->
<!-- # a <- do.call(cbind, selected_omics) -->
<!-- #  -->
<!-- # selected_omics <- lapply(filter_list, function(dt_name){ -->
<!-- #     omic_dt_names <- grep(dt_name, names(experiments(multi_omics))) -->
<!-- #     omic_exp_list <- experiments(multi_omics[,,omic_dt_names]) -->
<!-- #      -->
<!-- #     sig_vars_num <- sapply(omic_exp_list, function(omic_exp){ -->
<!-- #         nrow(omic_exp) -->
<!-- #     }) -->
<!-- #     sig_vars_num -->
<!-- # }) -->
<!-- # names(selected_omics) <- filter_list -->
<!-- #  -->
<!-- # b <- do.call(cbind, selected_omics) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Preparing the omics matrices  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # multi_omics_names <- names(experiments(multi_omics)) -->
<!-- #  -->
<!-- # multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S", multi_omics_names)] -->
<!-- # # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)]) -->
<!-- # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)]) -->
<!-- # multi_omics_selected -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # filtered_exps <- lapply(multi_omics_selected, function(multi_omics_names){ -->
<!-- #     multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]] -->
<!-- #  -->
<!-- #     if(grepl("RNASeq", multi_omics_names)){ -->
<!-- #         var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){ -->
<!-- #             sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE) -->
<!-- #         }) -->
<!-- #         multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, 0)],] -->
<!-- #     } -->
<!-- #     multi_omics_exp -->
<!-- # }) -->
<!-- # names(filtered_exps) <- multi_omics_selected -->
<!-- # experiments(multi_omics) <- ExperimentList(filtered_exps) -->
<!-- # ``` -->
<!-- # ```{r} -->
<!-- # multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics) -->
<!-- # plot_data_overview(multi_omics_mofa) -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected]) -->
<!-- # plot_data_overview(multi_omics_mofa) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Parameters -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # data_opts <- get_default_data_options(multi_omics_mofa) -->
<!-- # head(data_opts) -->
<!-- #  -->
<!-- # model_opts <- get_default_model_options(multi_omics_mofa) -->
<!-- # head(model_opts) -->
<!-- # model_opts$num_factors <- 8 -->
<!-- #  -->
<!-- # train_opts <- get_default_training_options(multi_omics_mofa) -->
<!-- # head(train_opts) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Model and training -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # multi_omics_mofa <- prepare_mofa( -->
<!-- #   object = multi_omics_mofa, -->
<!-- #   data_options = data_opts, -->
<!-- #   model_options = model_opts, -->
<!-- #   training_options = train_opts -->
<!-- # ) -->
<!-- #  -->
<!-- # outfile <- file.path(getwd(),"model.hdf5") -->
<!-- # multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE) -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Downstream Analysis -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # # file_path <- "./model.hdf5" -->
<!-- # # model <- load_model(file_path) -->
<!-- # colData(multi_omics)$sample <- colData(multi_omics)$samplename -->
<!-- #  -->
<!-- # samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample") -->
<!-- #  -->
<!-- # plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90)) -->
<!-- #  -->
<!-- # p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5,  -->
<!-- #                  color_by = "group.y",  -->
<!-- #                  add_violin = TRUE) -->
<!-- #  -->
<!-- # p <- p +  -->
<!-- #   scale_color_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green")) + -->
<!-- #   scale_fill_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green")) -->
<!-- # print(p) -->
<!-- # ``` -->

## Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```