---
title: "Multi_Omics_Analysis"
output: html_document
date: "2023-05-13"
params:
    input_dir: "../../inputs/"
    output_dir: "../../outputs/Multi_Omics_Analysis/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.duplicate.label = "allow")
```

## Required Packages

```{r, }
library(MOFA2)
library(dplyr)
library(knitr)
library(ggplot2)
library(MultiAssayExperiment)
library(grDevices)
library(rGREAT)
library(msigdbr)
library(DT)
library(TxDb.Mmusculus.UCSC.mm39.refGene)
source("../Utility_Functions/utilities.R")
library(TreeSummarizedExperiment)
library(GGally)
library(igraph)
library(survcomp)
```



## Loaded matrices

```{r}
multi_omics_raw <- readRDS(paste0(params$output_dir, "Multi_Omics_Analysis_exp_lima.rds"))
multi_omics <- readRDS(paste0(params$output_dir, "Multi_Omics_Analysis_MA_exp_lima.rds"))
names(experiments(multi_omics))

filter_list <- c("16S", "Shotgun.cecum", "Shotgun.colon")
```

## Preparing the omics matrices 

```{r}
multi_omics_names <- names(experiments(multi_omics))

multi_omics_selected <- multi_omics_names[!grepl("16S|Shotgun|unsupervised", multi_omics_names)]

# multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|unsupervised|Targeted|Metallomics|Histone", multi_omics_names)]

# multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)])
multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("Species|metaerg_kegg|cogs", multi_omics_names)])
# |EC|metaer_kegg|cogs
multi_omics_selected
```


```{r}
omic_names_dup <- sapply(strsplit(multi_omics_selected, "\\."), 
                        function(x){x[[1]]})

unique_omics <- unique(omic_names_dup)

omic_colors <- RColorBrewer::brewer.pal(n = length(unique_omics), name = "Set3")

col_list <- lapply(seq_along(unique_omics), function(omic_id){
    omic_name <- unique_omics[omic_id]
    omic_num <- length(omic_names_dup[omic_names_dup == omic_name])
    omic_col <- omic_colors[omic_id]
    
    col_ramp <- colorRampPalette(colors = c(omic_col, "white"))(omic_num + 8)[seq_len(omic_num)]
    return(col_ramp)
})
color_vect <- unlist(col_list)
names(color_vect) <- multi_omics_selected

omic_color_dict <- color_vect

shape_vect <- RCy3::getNodeShapes()[seq_along(unique_omics)]
shape_vect[is.na(shape_vect)] <- "ELLIPSE"

shape_vect_ext <- unlist(lapply(seq_along(shape_vect), function(omic_id){
    omic_name <- unique_omics[omic_id]
    shape_omic <- shape_vect[omic_id]
    omic_num <- length(omic_names_dup[omic_names_dup == omic_name])
    
    return(rep(shape_omic, omic_num))
}))
names(shape_vect_ext) <- multi_omics_selected

```

```{r}
# CV filter criteria
#
# filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
#     multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]
# 
#     if(grepl("RNA_Seq|DNA|Untargeted", multi_omics_names)){
#         if(grepl("RNA_Seq", multi_omics_names)){
#             cd_thr <- 0.0
#         } else if(grepl("unsupervised", multi_omics_names)){
#             cd_thr <- 0.99
#         } else if(grepl("DNA", multi_omics_names)){
#             cd_thr <- 0
#         } else{
#             cd_thr <- 0
#         }
#         var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){
#             sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE)
#         })
#         multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, cd_thr)],]
#     }
#     multi_omics_exp
# })
# names(filtered_exps) <- names(multi_omics)
# experiments(multi_omics) <- ExperimentList(filtered_exps)
```

```{r}
filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
    multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]
    p_val_dt <- rowData(multi_omics_exp)[,grepl("adj.p.value", colnames(rowData(multi_omics_exp)))]
    p_val_dt <- p_val_dt[,grepl("HFD-CNT|INT-HFD|INT-CNT", colnames(p_val_dt))]
    
    multi_omics_exp <- multi_omics_exp[apply(as.matrix(p_val_dt) <= 0.05, 1, any),]
    multi_omics_exp
})
names(filtered_exps) <- names(multi_omics)
experiments(multi_omics) <- ExperimentList(filtered_exps)
```

```{r}
omic_names_dict <- c("Histone Modifications", "Liver Metallomics", 
                            "Cecum Metallomics", "RNA Sequencing", 
                            "Cecum Targeted Metabolomics", "Liver Targeted Metabolomics", 
                            "Serum Targeted Metabolomics", "Untargeted Metabolomics", 
                            "Liver DNA Methylation", 
                            "Cecum 16S Metagenomics", "Colon 16S Metagenomics", 
                            "Cecum Shotgun Metagenomics", "Colon Shotgun Metagenomics", 
                            "Cecum Shotgun Functional Metaerg KEGG", "Shotgun Functional Colon Metaerg KEGG", 
                            "Cecum Shotgun Functional COGS", "Shotgun Functional Colon Metaerg COGS")
names(omic_names_dict) <- names(experiments(multi_omics[,,multi_omics_selected]))

group_color_dict <- RColorBrewer::brewer.pal(n = length(levels(multi_omics$group)), 
                                                 name = "Set1")
names(group_color_dict) <- levels(droplevels(multi_omics$group))

# group_color_dict <- c("CNT"="black", "HFD"="red", "INT"="blue")
```

```{r}
group_var <- "group"
sample_var <- "samplename"

selected_exps <- multi_omics[,colData(multi_omics)$group != "HFE", multi_omics_selected]

plot_dt <- lapply(names(experiments(selected_exps)), function(omic_name){
    omic_cd <- data.frame("sample_name" = colnames(experiments(selected_exps)[[omic_name]]))
    omic_cd$var_id <- omic_name
    omic_cd$value <- 1
    omic_cd
})
plot_dt <- do.call(rbind, plot_dt)
plot_dt$sample_name <- factor(plot_dt$sample_name)

group_var_dict <- colData(multi_omics)[[group_var]]
names(group_var_dict) <- colData(multi_omics)[[sample_var]]

sample_levels <- levels(plot_dt$sample_name)
reordered_sample_levels <- sample_levels[order(group_var_dict[as.character(sample_levels)])]

plot_dt$sample_name <- reorder(plot_dt$sample_name, 
                                new.order = reordered_sample_levels)

names(omic_color_dict) <- omic_names_dict[as.character(names(omic_color_dict))]

plot_dt$var_id <- factor(omic_names_dict[as.character(plot_dt$var_id)])
x_colors <- group_color_dict[as.character(group_var_dict[levels(plot_dt$sample_name)])]

# plot_nw_term(omic_dt = plot_dt)
ggplot(plot_dt, aes(x = sample_name, fill = var_id, y = var_id)) + 
            geom_tile() + 
            theme_minimal() + 
            theme(axis.text.x = element_text(angle = 90, 
                                             colour = x_colors),
                  axis.text.y = element_text(size = 5),
                    legend.position = "none", 
                    legend.justification = "top", 
                    legend.title = element_blank(), 
                    plot.title = element_text(hjust = 0.5),
                    legend.spacing.y = unit(0.1, "cm"), 
                  legend.text = element_text(size = 6), legend.key.size = unit(1, "line")) + 
            ggtitle("Omic Overview") + 
            xlab(NULL) + ylab(NULL) +
            scale_fill_manual(values = omic_color_dict)
```


## Metalomics correlation with the diet

```{r}
# omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))
# 
# diet_metals <- openxlsx::read.xlsx("../../inputs/Metallomics/metallomics_raw.xlsx",
#                                     sheet = "diets", rowNames = TRUE)
# rownames(diet_metals) <- gsub("CTR", "CNT", rownames(diet_metals))
# 
# dt_metals <- experiments(omic_dt[,,c("Metallomics.cecum", "Metallomics.liver")])
# 
# cor_mat <- lapply(dt_metals, function(exp_dt){
#     diet_metals_long <- diet_metals[as.character(exp_dt$group), rownames(exp_dt)]
#     # plotly::plot_ly(z = cor(t(assay(exp_dt)), diet_metals_long))
# 
#     return(diag(cor(t(assay(exp_dt)), diet_metals_long)))
# })
# cor_mat

```


## Run MOFA2

We have a multi-omic dataset with over 35 datasets, including the different levels of hierarchy for the 16S and Shotgun Metagenomics and the different annotations for the functional metagenomics results.

We visualized the principal component analysis for each individual omic to have a vision of the data and select which datasets from the hierarchical levels and the functional metagenomics. With the results we choose to keep the Species hierarchy level and for the functional metagenomics the COGS and metaerg-kegg annotations.

We also applied a p-value threshold of 0.05 in any of the contrast to filter variables and match the results from the whole genome bisulphite that, because we used the DSS method that finds DMRs from clusters of statistically significant CpGs were already biased.

Thus, the MOFA model will help us find correlated feature patterns in the statistically significant data.

```{r}
num_factors <- 3

# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,,multi_omics_selected],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "specie_metaerg_pathway_mofa",
#     color_vec = c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd",
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"),
#                                 params = param_list
#                             )
#                     )

# & !multi_omics$samplename %in% c("X643C", "X644C", "X658C")

mofa_dir_name <- "sig2_int_thr07"

param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
    output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
    multi_omics_exp = multi_omics[,colData(multi_omics)$group != "HFE", multi_omics_selected],
    group_var = "group",
    num_factors = num_factors,
    results_name = mofa_dir_name,
    color_vec = group_color_dict)
# c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
xfun::Rscript_call(rmarkdown::render,
                    args = list(input = "./MOFA_Models.Rmd", 
                                output_file = paste0(param_list$output_dir,
                                                     param_list$results_name, ".html"), 
                                params = param_list
                            )
                    )



# param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/",
#     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/",
#     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "INT" & !multi_omics$samplename %in% c("X643C", "X644C", "X658C"),multi_omics_selected],
#     group_var = "group",
#     num_factors = num_factors,
#     results_name = "specie_metaerg_pathway_any_sig_mofa_hfe",
#     color_vec = c("CNT"="black", "HFD"="red", "HFE"="green")
# )
# # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")
# xfun::Rscript_call(rmarkdown::render,
#                     args = list(input = "./MOFA_Models.Rmd", 
#                                 output_file = paste0(param_list$output_dir,
#                                                      param_list$results_name, ".html"), 
#                                 params = param_list
#                             )
#                     )
```

We can look at the resulting html file and we can the that all most of the variance for all the omics data is captured by the two first factors and the third one is biased towards the RNA-Seq. This makes sense because we have big differences in variable size between all the omics, this phenomena is difficult to avoid because we have many omics with very low values, even if we didn't apply a filter, thus, as the first two factors, surprisingly don't seem to be affected by this effect, and we have all the omics represented at least with a 20% of explained variance in these, we decided to proceed with it.

The factor boxplots show quite clearly two different patterns in the data, the first factor, represents features which change by the high fat diet (HFD) and are restored to similar levels by the intervention (INT); the second one shows features that are affected by HFD but the intervention fails to restore to control levels, keeping their levels changed.

Correlating these factors with the explained variance of each omic, we can see many omics biased towards one of both factors thus, we can propose that the liver histone hodifications, metallomics, targeted metabolomics and methylomics (WGB); serum targeted metabolomics and the cecum and colon functional metagenomics KEGG have a higher variance explained in the first factor thus, are the omics with the most significant features that are restored by the intervention and the cecum metallomics and targeted metabolomics are biased toward the second factor, thus they are changed by the HFD and kept mostly unchanged by the intervention.

These correlations with the first factor correlation reveal that epigenetic signatures, when affected by the high fat diet, are mostly restored to control levels after a life-style intervention, and the histone modifications and targeted metabolomics, which focused on epigenetically relevant metabolites.

All the remaining omics: the liver untargeted metabolomics, rna-sequencing and the colon and cecum metagenomics and functional clusters of orthologous groups of proteins, have even weights indicating that they contain features that can explain the variance in both factors.

Interestingly most of the metagenomics data, except the kegg functions show this pattern with the transcriptomics and untargeted metabolomics.

Looking at the weights of each feature in these factors we can see which of them are most important in explaining these patterns of the data. We looked at the absolute value distribution of weights for each omic and used as threshold for variable selection a percentage of features proportional to the variance explained of each omic in each factor.

We applied an overrepresentation analysis (ORA) to the significant genes and differentially methylated regions from each weight, dividing them into three subset: all the weights and dividing it into positive and negative weights.

## Variable correlations

```{r}
output_dir <- paste0(params$output_dir, "MOFA_Results/", mofa_dir_name, "/")
corr_output_dir <- paste0(output_dir, "correlation_results/")

omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))

sapply(names(experiments(omic_dt)), function(x){
    x_dt <- experiments(omic_dt)[[x]]
    hist(rowData(x_dt)$Factor_1_weights, main = x)
})

experiments(omic_dt) <- ExperimentList(lapply(experiments(omic_dt), function(omic_exp){
    
    omic_rd <- rowData(omic_exp)[,grepl("Factor_[1,2,3]_weights$", colnames(rowData(omic_exp)))]
    
    rowData(omic_exp)$max_weight_id <- apply(abs(as.matrix(omic_rd)), 1, which.max)
    omic_exp
}))

network_omics <- list("All Omics" = names(assays(omic_dt)))

# network_omics$`All Omics` <- grep("Untargeted|Targeted|RNASeq|species|metaerg_kegg|Metallomics|PTM|DNA_Methylation.liver",
#                                   network_omics$`All Omics`, value = TRUE)

# lapply(experiments(multi_omic_dt), function(x){colnames(rowData(x))})

variables_dicts <- lapply(network_omics[[1]], function(omic_name){
    if(omic_name == "RNA_Sequencing"){
        return("gene_lab")
    } else if(omic_name == "Untargeted_Metabolomics"){
        return("metaboanalyst_compound_names")
        #return("HFD-CNT.formula")
    } else if(grepl("DNA_Methylation", omic_name)){
        return("geneName")
    } else if(grepl("Shotgun|Metallomics|16S|Targeted|Histone", omic_name)){
        return("readable_name")  
    } else{
        return(NA)
    }
})
names(variables_dicts) <- network_omics[[1]]
variables_dicts[is.na(variables_dicts)] <- NULL

add_var_names <- grep("weights|FC", colnames(rowData(omic_dt@ExperimentList$RNA_Sequencing)), value = T)

# corr_method_vect <- c("kendall", "spearman", "pearson")
corr_method_vect <- c("spearman")

for(corr_method in corr_method_vect){
    xfun::Rscript_call(rmarkdown::render,
                            args = list(input = "../Multi_Omics_Correlations/Correlations_Script.Rmd",
                              output_file = output_dir,
                              params = list(
                                corr_threshold = 0,
                                corr_method = corr_method,
                                multi_omic_path = omic_dt,
                                group_metadata_var = "group",
                                sample_levels_2_correlate = c("CNT", "HFD", "INT"),
                                filter_variable_2_correlate = list(
                                    "Factor_1" = quote(Factor_1_weights_sig),
                                    "Factor_2" = quote(Factor_2_weights_sig)
                                ),
                                p_adj_thr = 1,
                                omic_list_2_correlate = network_omics,
                                rowData_variable_name_list = variables_dicts,
                                draw_networks = TRUE,
                                draw_cytoscape = FALSE,
                                add_autocorrelations = FALSE,
                                keep_full_correlations = TRUE,
                                output_dir = corr_output_dir,
                                output_suffix = paste0("0_005_chi_f1f2_", corr_method),
                                add_variables = TRUE,
                                variables_names = add_var_names,
                                draw_cluster_scatter_plots = FALSE))
                      )
    # xfun::Rscript_call(rmarkdown::render,
    #                             args = list(input = "../Multi_Omics_Correlations/Correlations_Script.Rmd",
    #                               output_file = output_dir,
    #                               params = list(
    #                                 corr_threshold = 0,
    #                                 corr_method = corr_method,
    #                                 multi_omic_path = omic_dt,
    #                                 group_metadata_var = "group",
    #                                 sample_levels_2_correlate = c("CNT", "HFD", "INT"),
    #                                 filter_variable_2_correlate = list(
    #                                     "All" = NA
    #                                 ),
    #                                 omic_list_2_correlate = network_omics,
    #                                 rowData_variable_name_list = variables_dicts,
    #                                 draw_networks = FALSE,
    #                                 draw_cytoscape = FALSE,
    #                                 add_autocorrelations = FALSE,
    #                                 keep_full_correlations = TRUE,
    #                                 output_dir = corr_output_dir,
    #                                 output_suffix = paste0("0_005_chi", corr_method),
    #                                 add_variables = TRUE,
    #                                 variables_names = add_var_names,
    #                                 draw_cluster_scatter_plots = FALSE))
    #                       )
}
```

## Correlation Analysis

```{r}
library(ProteoNet)
library(igraph)
library(dplyr)
library(org.Mm.eg.db)
output_dir <- paste0(params$output_dir, "MOFA_Results/", mofa_dir_name, "/")
corr_output_dir <- paste0(output_dir, "correlation_results/")

omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds"))

nw_paths <- list.files(corr_output_dir, pattern = "network_results.*\\.rds$", full.names = T)
nw_graphs <- lapply(nw_paths, readRDS)

nw_graphs <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(weight_dt){
        weight_dt[["All Omics"]]
    })
})
# table(vertex_attr(nw_graphs$spearman$Factor_1, "Node_Omic"))
# table(vertex_attr(nw_graphs$spearman$Factor_2, "Node_Omic"))

score_names <- gsub("\\.rds", "", sapply(strsplit(grep("network_results", 
                                                        basename(nw_paths), 
                                                        value = TRUE), "_"),
                                            function(x){x[length(x)]}))
score_filters <- c(0.8)
names(score_filters) <- score_names
names(nw_graphs) <- score_names
```

### Enrich information

```{r}
omic_enrich <- c("Untargeted_Metabolomics", "RNA_Sequencing", "WGB_DNA_Methylation")
ids_enrich_names <- c("metabo_analyst_cp", "entrezID", "geneId")

library(SummarizedExperiment)

## Enrich IDs dictionary
omic_dict_list <- mapply(function(omic_id, enrich_var){
    
    omic_data <- omic_dt[[omic_id]]
    omic_dict <- rowData(omic_data)[[enrich_var]]
    
    names(omic_dict) <- paste0(omic_id, ".", rownames(omic_data))
    
    omic_dict
    
}, omic_id = omic_enrich, enrich_var = ids_enrich_names)

rna_seq_rd <- lapply(experiments(omic_dt)[c("RNA_Sequencing", "WGB_DNA_Methylation", "Untargeted_Metabolomics")], rowData)
num_factors <- 2

## Universe IDs mapped
universe_list_ids <- lapply(names(rna_seq_rd), function(omic_name){
    omic_ids <- omic_dict_list[[omic_name]][paste0(omic_name, ".", rownames(rna_seq_rd[[omic_name]]))]
    omic_ids <- omic_ids[!is.null(omic_ids)]
    omic_ids <- unlist(strsplit(omic_ids, ";"))
    omic_ids[!is.na(omic_ids)]

})
names(universe_list_ids) <- names(rna_seq_rd)
```

### Correlation networks input

```{r}
feature_ids_dict <- lapply(names(variables_dicts), function(dataset_name){
    
    dataset_var <- rowData(omic_dt[[dataset_name]])[variables_dicts[[dataset_name]]]
    dataset_vector <- unlist(dataset_var[[1]])
    dataset_vector[is.na(dataset_vector)] <- paste(dataset_name, "No annotation")
    dataset_vector <- make.unique(dataset_vector, sep = ".")
    
    if(dataset_name == "WGB_DNA_Methylation"){
        dataset_vector <- paste("DMR", dataset_vector)
    }
    names(dataset_vector) <- paste0(dataset_name, ".", rownames(dataset_var))
    dataset_vector
})
names(feature_ids_dict) <- names(variables_dicts)

var_order_col_list <- lapply(nw_graphs, function(score_nw){
    factor_col <- lapply(names(score_nw), function(factor_id){
        paste0(factor_id, "_weights")    
    })
    names(factor_col) <- names(score_nw)
    factor_col
})
```

```{r}
# variable_dictionary <- lapply(names(variables_dicts), function(dataset_name){
#         

#     dataset_var <- rowData(omic_dt[[dataset_name]])[variables_dicts[[dataset_name]]]
#     dataset_vector <- unlist(dataset_var[[1]])
#     names(dataset_vector) <- paste0(dataset_name, ".", rownames(dataset_var))
#     
#     dataset_vector
# })
# names(variable_dictionary) <- names(variables_dicts)
# 
# nw_graphs <- lapply(nw_graphs, function(score_dt){
#     lapply(score_dt, function(factor_dt){
#         vertex_attr(factor_dt[["All Omics"]], "Readable_Name") <- sapply(names(V(factor_dt[["All Omics"]])), 
#                                                                     function(v_name){
#                 omic_id <- vertex_attr(factor_dt[["All Omics"]], name = "Node_Omic", index = v_name)
#                 if(omic_id %in% names(variable_dictionary)){
#                     variable_dictionary[[omic_id]][v_name]
#                 } else{
#                     return(v_name)
#                 }
#             })
#         factor_dt
#     })
# })

```

```{r}
# filter_scores <- seq(0.2, 0.9, by = 0.05)
# 
# nw_filter_score <- lapply(names(nw_graphs), function(score_id){
#     score_dt <- nw_graphs[[score_id]]
# 
#     score_res <- lapply(names(score_dt), function(weight_id){
#         weight_dt <- score_dt[[weight_id]][["All Omics"]]
# 
#         pdf(paste0(corr_output_dir, "degree_distribution_", score_id, "_", weight_id, ".pdf"))
# 
#         power_fit_pval <- sapply(filter_scores, function(x){
#             if(x == 1 || x == 0){return(NA)}
#             graph_weights <- abs(edge_attr(weight_dt,
#                                     name = "weight"))
#             sub_g <- subgraph.edges(weight_dt,
#                         E(weight_dt)[graph_weights >= x],
#                         delete.vertices = TRUE)
#             if(length(sub_g) == 0){return(NA)}
#             hist(degree(sub_g), breaks = 100, main = x)
#             return(power.law.fit(degree(sub_g))$KS.stat)
#             degree(sub_g)
#         })
# 
#         dev.off()
# 
#         names(power_fit_pval) <- filter_scores
#         power_fit_pval
#         # return(names(power_fit_pval)[which.min(power_fit_pval)])
#     })
#     names(score_res) <- names(score_dt)
#     score_res
# })
# names(nw_filter_score) <- names(nw_graphs)
# score_filters <- lapply(nw_filter_score, function(score_dt){
#     lapply(score_dt, function(weight_dt){
#         # lapply(weight_dt, function(filter_scores){
#         #     if(all(is.na(filter_scores))){return(NA)}
#         #     return(power.law.fit(filter_scores)$KS.stat)
#         # })
#         # plot(x = names(weight_dt), y = weight_dt)
#         return(names(weight_dt)[which.min(weight_dt)])
#     })
# })
```

### Correlation filter

```{r}
rand_seed <- 1234
universe_list <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(graph_dt){
        names(V(graph_dt))
    })
})

nw_graphs <- lapply(names(nw_graphs), function(score_id){
    score_dt <- nw_graphs[[score_id]]
    score_res <- lapply(names(score_dt), function(graph_id){
        graph_dt <- score_dt[[graph_id]]

        graph_weights <- abs(edge_attr(graph_dt,
                        name = "weight"))
        graph_adj.p.val <- edge_attr(graph_dt, name = "adj.p.value")
        subgraph.edges(graph_dt,
                    E(graph_dt)[graph_weights >= score_filters[score_id] & graph_adj.p.val <= 0.05],
                    delete.vertices = TRUE)
    })
    names(score_res) <- names(score_dt)
    score_res
})
names(nw_graphs) <- score_names
```

#### Full network

```{r}
membership_attribute <- "cluster_ids"
min_n <- 10
nw_graphs <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(graph_dt){
        # graph_dt <- subgraph(graph_dt, clusters(graph_dt))
        # graph_dt <- decompose(graph_dt, max.comps = 1)[[1]]
        vertex_attr(graph_dt, name = membership_attribute) <- "CL_1"
        return(graph_dt)
    })
})

```

#### Full enrichment

```{r}
clu_ids_enrich <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(graph_dt){
        membership_vector <- vertex_attr(graph_dt, membership_attribute)
        clu_vertex <- lapply(unique(membership_vector), function(cl_id){
            names(V(graph_dt))[membership_vector == cl_id]
        })
        names(clu_vertex) <- unique(membership_vector)
        cluster_id_list <-  lapply(clu_vertex, function(clu_ids){
            if(length(clu_ids) <= min_n){return(NA)}
            metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
                                                                            clu_ids, value = T)]
            metabo_dt <- metabo_dt[!is.null(metabo_dt)]
            metabo_dt <- strsplit(metabo_dt, ";")
            metabo_dt <- metabo_dt[!is.na(metabo_dt)]
            
            wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
                                                                            clu_ids, value = T)]
            wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
            
            rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
                                                                            clu_ids, value = T)]
            rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
            
            return(list("Untargeted_Metabolomics" = metabo_dt, 
                        "RNA_Sequencing" = rna_dt, 
                        "WGB_DNA_Methylation" = wgb_dt))
        })
        cluster_id_list[!is.na(cluster_id_list)]
    })
})
```

##### Partial kegg joint pathway

```{r}
library(clusterProfiler)
kegg_db_enrich_path <- paste0(params$output_dir, "kegg_enrich_db.rds")
kegg_db_list <- readRDS(kegg_db_enrich_path)
term2name_dt <- kegg_db_list$term2name
integ_enrich_db <- kegg_db_list$term2gene
membership_attribute <- "cluster_ids"
min_n <- 10
enrich_name <- "factor_part_kegg"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}
```


```{r}
weight_table <- prop.table(table(integ_enrich_db$term, integ_enrich_db$type), margin = 1)

universe_m <- unlist(Reduce(c, universe_list_ids["Untargeted_Metabolomics"]))
universe_g <- unlist(Reduce(c, universe_list_ids[c("RNA_Sequencing", "WGB_DNA_Methylation")]))

enrich_results_partial <- lapply(names(clu_ids_enrich), function(factor_id){
    factor_dt <- clu_ids_enrich[[factor_id]]
    factor_res <- lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(weight_dt, function(clu_dt){
            ids_m <- Reduce(c, clu_dt["Untargeted_Metabolomics"])
            ids_g <- Reduce(c, clu_dt[c("RNA_Sequencing", 
                                        "WGB_DNA_Methylation")])
            
            enrich_m <- enricher(gene = unlist(ids_m), universe = NULL, 
                                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                                TERM2GENE = integ_enrich_db[integ_enrich_db$type == "metabolite_id",], 
                                TERM2NAME = term2name_dt)
            enrich_g <- enricher(gene = unlist(ids_g), universe = NULL, 
                                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                                TERM2GENE = integ_enrich_db[integ_enrich_db$type == "gene_id",], 
                                TERM2NAME = term2name_dt)
            
            if(is.null(enrich_m) & is.null(enrich_g)){
                return(NULL)
            } else if(!is.null(enrich_m) & is.null(enrich_g)){
                enrich_m <- setReadable_custom(enrich_m, featureDict = ids_m)
                return(enrich_m)
            } else if(!is.null(enrich_g) & is.null(enrich_m)){
                enrich_g <- setReadable_custom(enrich_g, featureDict = ids_g)
                return(enrich_g)
            }
            enrich_m <- setReadable_custom(enrich_m, featureDict = ids_m)
            enrich_g <- setReadable_custom(enrich_g, featureDict = ids_g)
    
            merged_enrich <- dplyr::full_join(x = enrich_g@result, y = enrich_m@result, by = "ID")
             
            merged_enrich$geneID <- paste(merged_enrich$geneID.x, 
                                          merged_enrich$geneID.y, 
                                        sep = "/")
            merged_enrich$pvalue <- sapply(merged_enrich$ID, function(id){
                p_vals <- merged_enrich[merged_enrich$ID == id, 
                                        c("pvalue.x", "pvalue.y")]
                survcomp::combine.test(p = p_vals, 
                                       weight = weight_table[id,], 
                                       method = "logit", 
                                       na.rm = T)    
            })
            merged_enrich$p.adjust <- p.adjust(merged_enrich$pvalue, 
                                                        method = "fdr")
            merged_enrich$qvalue <- enrich_g@qvalueCutoff
            merged_enrich$Count <- rowSums(cbind(merged_enrich$Count.x,
                                                merged_enrich$Count.y), 
                                            na.rm = TRUE)
            
            merged_enrich$Description <- merged_enrich$Description.x
            merged_enrich$Description[is.na(merged_enrich$Description)] <- merged_enrich$Description.y[is.na(merged_enrich$Description)]
            
            merged_enrich$GeneRatio <- mapply(FUN = function(x, y){
                if(any(is.na(x))){return(paste(y, collapse = "/"))}
                if(any(is.na(y))){return(paste(x, collapse = "/"))}
                paste(sum(as.numeric(x[1]), as.numeric(y[1]), na.rm = TRUE),
                    sum(as.numeric(x[2]), as.numeric(y[2]), na.rm = TRUE), sep = "/")
            }, x = strsplit(merged_enrich$GeneRatio.x, "/"), 
                y = strsplit(merged_enrich$GeneRatio.y, "/"))
            
            enrich_g@result <- merged_enrich
            enrich_g
            # merged_enrich[merged_enrich$p.adjust <= 0.05,]
        })
    })
    names(factor_res) <- names(factor_dt)
    factor_res
})
names(enrich_results_partial) <- names(clu_ids_enrich)

results_wb <- openxlsx::createWorkbook()
lapply(names(enrich_results_partial), function(factor_id){
    factor_dt <- enrich_results_partial[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            sheet_name <- paste(factor_id, weight_id, clu_id, sep = ".")
            openxlsx::addWorksheet(wb = results_wb,
                                    sheetName = sheet_name)
            openxlsx::writeData(wb = results_wb, sheet = sheet_name,
                                x = as.data.frame(clu_dt))
        })
    })
    openxlsx::saveWorkbook(results_wb, 
                            file = paste0(enrich_nw_dir, "enrich_results.xlsx"), 
                            overwrite = TRUE)
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(enrich_nw_dir, enrich_name, "_Pathway_Enrich.pdf"))
lapply(names(enrich_results_partial), function(factor_id){
    factor_dt <- enrich_results_partial[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            if(nrow(as.data.frame(clu_dt)) == 0){return(NULL)}
            plot_name <- gsub("_", " ", paste(factor_id, weight_id, clu_id))
    
            dotplot(clu_dt) + ggtitle(paste("Kegg joint pathway enrichment" , plot_name))
        })
    })
})
dev.off()
```

###### Network creation

```{r}
xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_partial,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_color_dict = omic_color_dict,
            omic_shape_dict = shape_vect_ext,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```


##### DMR & RNASeq GO term enrichment

```{r}
library(clusterProfiler)
kegg_db_enrich_path <- paste0(params$output_dir, "kegg_enrich_db.rds")
kegg_db_list <- readRDS(kegg_db_enrich_path)
term2name_dt <- kegg_db_list$term2name
integ_enrich_db <- kegg_db_list$term2gene
min_n <- 10
enrich_name <- "factor_full_go"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}
```


```{r}
## DMR & RNA_Seq enrich
library(org.Mm.eg.db)
library(GOSemSim)
library(rGREAT)

enrich_results_list <- lapply(nw_graphs, function(factor_dt){
    lapply(factor_dt, function(weight_dt){
        enrich_multi_omic_clusters(graph = weight_dt, 
                                feature_dict = Reduce(c, omic_dict_list[c("RNA_Sequencing",
                                                                        "WGB_DNA_Methylation")]),
                                membership_attribute = membership_attribute, 
                                universe = NULL, min_n = 10)    
    })
})
    
results_wb <- openxlsx::createWorkbook()
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            sheet_name <- paste(factor_id, weight_id, clu_id, sep = ".")
            openxlsx::addWorksheet(wb = results_wb,
                                    sheetName = sheet_name)
            openxlsx::writeData(wb = results_wb, sheet = sheet_name,
                                x = as.data.frame(clu_dt))
        })
    })
    openxlsx::saveWorkbook(results_wb, 
                            file = paste0(enrich_nw_dir, "enrich_results.xlsx"), 
                            overwrite = TRUE)
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(enrich_nw_dir, enrich_name," _Pathway_Enrich.pdf"))
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            if(nrow(as.data.frame(clu_dt)) == 0){return(NULL)}
            plot_name <- gsub("_", " ", paste(factor_id, weight_id))
            
            dotplot(clu_dt) + ggtitle(paste("Enriched Kegg Pathways" , plot_name, clu_id))
        })
        
    })
})
dev.off()
```

###### Network creation

```{r}
xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_list,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_color_dict = omic_color_dict,
            omic_shape_dict = shape_vect_ext,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```



##### Full kegg joint pathway analysis

```{r}
library(clusterProfiler)
kegg_db_enrich_path <- paste0(params$output_dir, "kegg_enrich_db.rds")
kegg_db_list <- readRDS(kegg_db_enrich_path)
term2name_dt <- kegg_db_list$term2name
integ_enrich_db <- kegg_db_list$term2gene
min_n <- 10
enrich_name <- "factor_full_kegg"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}
```

```{r}
enrich_results_list <- lapply(names(clu_ids_enrich), function(score_id){
    factor_results <- lapply(names(clu_ids_enrich[[score_id]]), function(factor_id){
        enrich_results <- lapply(names(clu_ids_enrich[[score_id]][[factor_id]]), function(clu_id){
            clu_ids_list <- clu_ids_enrich[[score_id]][[factor_id]][[clu_id]]
            enrich_res <- enricher(
                gene = unlist(clu_ids_list), 
                universe = NULL, 
                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)
            # if(nrow(as.data.frame(enrich_res)) == 0){return(NA)}
            # unlist(universe_list_ids)
            enrich_res <- setReadable_custom(enrich_res, 
                                            featureDict = Reduce(c, clu_ids_list))
            return(enrich_res)
        })
        names(enrich_results) <- names(clu_ids_enrich[[score_id]][[factor_id]])
        return(enrich_results[!is.na(enrich_results)])
    })
    names(factor_results) <- names(clu_ids_enrich[[score_id]])
    factor_results
})
names(enrich_results_list) <- names(clu_ids_enrich)

results_wb <- openxlsx::createWorkbook()
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            sheet_name <- paste(factor_id, weight_id, clu_id, sep = ".")
            openxlsx::addWorksheet(wb = results_wb,
                                    sheetName = sheet_name)
            openxlsx::writeData(wb = results_wb, sheet = sheet_name,
                                x = as.data.frame(clu_dt))
        })
    })
    openxlsx::saveWorkbook(results_wb, 
                            file = paste0(enrich_nw_dir, "enrich_results.xlsx"), 
                            overwrite = TRUE)
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(enrich_nw_dir, enrich_name," _Pathway_Enrich.pdf"))
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            if(nrow(as.data.frame(clu_dt)) == 0){return(NULL)}
            plot_name <- gsub("_", " ", paste(factor_id, weight_id))
            
            dotplot(clu_dt) + ggtitle(paste("Enriched Kegg Pathways" , plot_name, clu_id))
        })
        
    })
})
dev.off()
```

###### Network creation

```{r}
xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_list,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_color_dict = omic_color_dict,
            omic_shape_dict = shape_vect_ext,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```

###### Network creation test

```{r}
enrich_name <- "factor_full_kegg_2"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}

omic_dt_mod <- omic_dt
mofa_model <- load_model(paste0(output_dir, "model.hdf5"))
exp_pred <- predict(mofa_model)

new_exps <- lapply(names(exp_pred), function(omic_name){
    exp_dt <- exp_pred[[omic_name]][[1]]
    rownames(exp_dt) <- gsub(paste0("_", omic_name, "$"), "", rownames(exp_dt))
    
    omic_exp <- experiments(omic_dt)[[omic_name]]
    SummarizedExperiment(assays = exp_dt, rowData = rowData(omic_exp), 
                         colData = colData(experiments(omic_dt)[["RNA_Sequencing"]])[colnames(exp_dt),])
    
})
names(new_exps) <- names(exp_pred)

omic_dt_mod <- MultiAssayExperiment(experiments = ExperimentList(new_exps), colData = colData(new_exps$RNA_Sequencing))

xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_list,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt_mod,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_shape_dict = shape_vect_ext,
            omic_color_dict = omic_color_dict,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```



#### Cluster network

```{r}
nw_graphs <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(graph_dt){
        # graph_dt <- subgraph(graph_dt, clusters(graph_dt))
        graph_dt <- decompose(graph_dt, max.comps = 1)[[1]]
        membership_dt <- membership(infomaps_website_clustering(
                                    graph = graph_dt, weights = abs(edge_attr(graph_dt, name = "weight")),
                                    rand.seed = rand_seed, num_trials = 100))
        vertex_attr(graph_dt, name = "cluster_ids") <- paste0("CL_", membership_dt)
        return(graph_dt)
    })
})

table(vertex_attr(nw_graphs$spearman$Factor_1, "cluster_ids"))
table(vertex_attr(nw_graphs$spearman$Factor_2, "cluster_ids"))

lapply(names(nw_graphs), function(score_id){
    score_dt <- nw_graphs[[score_id]]
    lapply(names(score_dt), function(weight_id){
        graph_dt <- score_dt[[weight_id]]
        hist(table(vertex_attr(graph_dt, name = "cluster_ids")), breaks = 100)
    })
})
```

#### Cluster enrichment

```{r}
membership_attribute <- "cluster_ids"
min_n = 10;

library(clusterProfiler)

clu_ids_enrich <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(graph){
        membership_vector <- vertex_attr(graph, membership_attribute)
        clu_vertex <- lapply(unique(membership_vector), function(cl_id){
            names(V(graph))[membership_vector == cl_id]
        })
        names(clu_vertex) <- unique(membership_vector)
        cluster_id_list <-  lapply(clu_vertex, function(clu_ids){
            if(length(clu_ids) <= min_n){return(NA)}
            metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
                                                                            clu_ids, value = T)]
            metabo_dt <- metabo_dt[!is.null(metabo_dt)]
            metabo_dt <- strsplit(metabo_dt, ";")
            metabo_dt <- metabo_dt[!is.na(metabo_dt)]
            
            wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
                                                                            clu_ids, value = T)]
            wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
            
            rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
                                                                            clu_ids, value = T)]
            rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
            
            return(list("Untargeted_Metabolomics" = metabo_dt, 
                        "RNA_Sequencing" = rna_dt,
                        "WGB_DNA_Methylation" = wgb_dt))
        })
        cluster_id_list[!is.na(cluster_id_list)]
    })
})

universe_list_ids <- lapply(nw_graphs, function(score_dt){
    lapply(score_dt, function(graph){
        
        universe_ids <- names(V(graph))
    
        metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
                                                                            universe_ids, value = T)]
        metabo_dt <- metabo_dt[!is.null(metabo_dt)]
        metabo_dt <- unlist(strsplit(metabo_dt, ";"))
        metabo_dt <- metabo_dt[!is.na(metabo_dt)]
        
        wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
                                                                        universe_ids, value = T)]
        wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
        
        rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
                                                                        universe_ids, value = T)]
        rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
            
        universe_ids_f <- list("Untargeted_Metabolomics" = metabo_dt, 
                               "WGB_DNA_Methylation" = wgb_dt, 
                               "RNA_Sequencing" = rna_dt)
        universe_ids_f
    })
})
```

##### Partial pathway enrichment

```{r}
library(clusterProfiler)
kegg_db_enrich_path <- paste0(params$output_dir, "kegg_enrich_db.rds")
kegg_db_list <- readRDS(kegg_db_enrich_path)
term2name_dt <- kegg_db_list$term2name
integ_enrich_db <- kegg_db_list$term2gene
membership_attribute <- "cluster_ids"
min_n <- 10
enrich_name <- "clu_part_kegg"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}
```


```{r}
weight_table <- prop.table(table(integ_enrich_db$term, integ_enrich_db$type), margin = 1)

enrich_results_partial <- lapply(names(clu_ids_enrich), function(factor_id){
    factor_dt <- clu_ids_enrich[[factor_id]]
    factor_res <- lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        weight_res <- lapply(weight_dt, function(clu_dt){
            universe_dt <- universe_list_ids[[factor_id]][[weight_id]]
            
            ids_m <- Reduce(c, clu_dt["Untargeted_Metabolomics"])
            ids_g <- Reduce(c, clu_dt[c("RNA_Sequencing", 
                                        "WGB_DNA_Methylation")])
            
            universe_m <- unlist(Reduce(c, universe_dt["Untargeted_Metabolomics"]))
            universe_g <- unlist(Reduce(c, universe_dt[c("RNA_Sequencing",
                                                        "WGB_DNA_Methylation")]))
                    
            enrich_m <- enricher(gene = unlist(ids_m), universe = universe_m, 
                                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                                TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)
            
            enrich_g <- enricher(gene = unlist(ids_g), universe = universe_g, 
                                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                                TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)
            
            if(is.null(enrich_m) & is.null(enrich_g)){
                return(NULL)
            } else if(!is.null(enrich_m) & is.null(enrich_g)){
                enrich_m <- setReadable_custom(enrich_m, featureDict = ids_m)
                return(enrich_m)
            } else if(!is.null(enrich_g) & is.null(enrich_m)){
                enrich_g <- setReadable_custom(enrich_g, featureDict = ids_g)
                return(enrich_g)
            }
            enrich_m <- setReadable_custom(enrich_m, featureDict = ids_m)
            enrich_g <- setReadable_custom(enrich_g, featureDict = ids_g)

            merged_enrich <- dplyr::full_join(x = enrich_g@result, 
                                            y = enrich_m@result, 
                                            by = "ID")
            merged_enrich$geneID <- paste(merged_enrich$geneID.x, 
                                          merged_enrich$geneID.y, sep = "/")
            
            merged_enrich$pvalue <- sapply(merged_enrich$ID, function(id){
                p_vals <- merged_enrich[merged_enrich$ID == id, 
                                        c("pvalue.x", "pvalue.y")]
                if(any(is.na(p_vals))){
                    return(p_vals[!is.na(p_vals)])
                }
                survcomp::combine.test(p = p_vals, 
                                       weight = weight_table[id,], method = "logit")    
            })
            merged_enrich$p.adjust <- p.adjust(merged_enrich$pvalue, 
                                                        method = "fdr")
            merged_enrich$qvalue <- enrich_g@qvalueCutoff
            
            merged_enrich$Count <- rowSums(cbind(merged_enrich$Count.x,
                                                merged_enrich$Count.y), 
                                            na.rm = TRUE)
            
            merged_enrich$Description <- merged_enrich$Description.x
            merged_enrich$Description[is.na(merged_enrich$Description)] <- merged_enrich$Description.y[is.na(merged_enrich$Description)]
            
            merged_enrich$GeneRatio <- mapply(FUN = function(x, y){
                if(any(is.na(x))){return(paste(y, collapse = "/"))}
                if(any(is.na(y))){return(paste(x, collapse = "/"))}
                paste(sum(as.numeric(x[1]), as.numeric(y[1]), na.rm = TRUE),
                    sum(as.numeric(x[2]), as.numeric(y[2]), na.rm = TRUE), sep = "/")
            }, x = strsplit(merged_enrich$GeneRatio.x, "/"), 
                y = strsplit(merged_enrich$GeneRatio.y, "/"))
            
            
            merged_enrich$BgRatio <- mapply(FUN = function(x, y){
                if(any(is.na(x))){return(paste(y, collapse = "/"))}
                if(any(is.na(y))){return(paste(x, collapse = "/"))}
                paste(sum(as.numeric(x[1]), as.numeric(y[1]), na.rm = TRUE),
                    sum(as.numeric(x[2]), as.numeric(y[2]), na.rm = TRUE), sep = "/")
            }, x = strsplit(merged_enrich$BgRatio.x, "/"), 
                y = strsplit(merged_enrich$BgRatio.y, "/"))
            enrich_g@result <- merged_enrich[,c("ID", "Description", 
                                                "GeneRatio", "BgRatio", 
                                                "pvalue", "p.adjust", "qvalue", 
                                                "geneID", "Count")]
            enrich_g
        })
        weight_res[!sapply(weight_res, is.null)]
    })
    names(factor_res) <- names(factor_dt)
    factor_res
})

names(enrich_results_partial) <- names(clu_ids_enrich)

results_wb <- openxlsx::createWorkbook()
lapply(names(enrich_results_partial), function(factor_id){
    factor_dt <- enrich_results_partial[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            sheet_name <- paste(factor_id, weight_id, clu_id, sep = ".")
            openxlsx::addWorksheet(wb = results_wb,
                                    sheetName = sheet_name)
            openxlsx::writeData(wb = results_wb, sheet = sheet_name,
                                x = as.data.frame(clu_dt))
        })
    })
    openxlsx::saveWorkbook(results_wb, 
                            file = paste0(enrich_nw_dir, "enrich_results.xlsx"), 
                            overwrite = TRUE)
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(enrich_nw_dir, enrich_name, "_Pathway_Enrich.pdf"))
lapply(names(enrich_results_partial), function(factor_id){
    factor_dt <- enrich_results_partial[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]

            if(nrow(as.data.frame(clu_dt)) == 0){return(NULL)}
            plot_name <- gsub("_", " ", paste(factor_id, weight_id, clu_id))
            dotplot(clu_dt) + ggtitle(paste("Kegg joint pathway enrichment" , plot_name))
        })
    })
})
dev.off()
```


###### Network creation

```{r}
xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_partial,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_shape_dict = shape_vect_ext,
            omic_color_dict = omic_color_dict,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```



##### DMR & RNASeq GO term enrichment

```{r}
library(clusterProfiler)
min_n <- 10
enrich_name <- "clu_full_go"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}
```


```{r}
## DMR & RNA_Seq enrich
library(org.Mm.eg.db)
library(GOSemSim)
library(rGREAT)

enrich_results_list <- lapply(names(nw_graphs), function(factor_id){
    factor_dt <- nw_graphs[[factor_id]]
    factor_res <- lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        
        weight_universe <- universe_list_ids[[factor_id]][[weight_id]]
        
        enrich_multi_omic_clusters(graph = weight_dt, 
                                feature_dict = Reduce(c, omic_dict_list[c("RNA_Sequencing",
                                                                    "WGB_DNA_Methylation")]),
                                membership_attribute = membership_attribute, 
                                universe = unlist(weight_universe[c("RNA_Sequencing",
                                                                    "WGB_DNA_Methylation")]), 
                                min_n = 10)    
    })
    names(factor_res) <- names(factor_dt)
    factor_res
})
names(enrich_results_list) <- names(nw_graphs)
results_wb <- openxlsx::createWorkbook()
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            sheet_name <- paste(factor_id, weight_id, clu_id, sep = ".")
            openxlsx::addWorksheet(wb = results_wb,
                                    sheetName = sheet_name)
            openxlsx::writeData(wb = results_wb, sheet = sheet_name,
                                x = as.data.frame(clu_dt))
        })
    })
    openxlsx::saveWorkbook(results_wb, 
                            file = paste0(enrich_nw_dir, "enrich_results.xlsx"), 
                            overwrite = TRUE)
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(enrich_nw_dir, enrich_name," _Pathway_Enrich.pdf"))
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            if(nrow(as.data.frame(clu_dt)) == 0){return(NULL)}
            plot_name <- gsub("_", " ", paste(factor_id, weight_id))
            
            dotplot(clu_dt) + ggtitle(paste("Enriched Kegg Pathways" , plot_name, clu_id))
        })
        
    })
})
dev.off()
```

###### Network creation

```{r}
xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_list,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_color_dict = omic_color_dict,
            omic_shape_dict = shape_vect_ext,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```




##### Full Joint Pathway Enrichment

```{r}
library(clusterProfiler)
kegg_db_enrich_path <- paste0(params$output_dir, "kegg_enrich_db.rds")
kegg_db_list <- readRDS(kegg_db_enrich_path)
term2name_dt <- kegg_db_list$term2name
integ_enrich_db <- kegg_db_list$term2gene
membership_attribute <- "cluster_ids"
min_n <- 10
enrich_name <- "clu_full_kegg"
enrich_nw_dir <- paste0(corr_output_dir, enrich_name, "/")

if(!dir.exists(enrich_nw_dir)){dir.create(enrich_nw_dir)}
```


```{r}
enrich_results_list <- lapply(names(clu_ids_enrich), function(score_id){
    factor_results <- lapply(names(clu_ids_enrich[[score_id]]), function(factor_id){
        enrich_results <- lapply(names(clu_ids_enrich[[score_id]][[factor_id]]), function(clu_id){
            clu_ids_list <- clu_ids_enrich[[score_id]][[factor_id]][[clu_id]]
            enrich_res <- enricher(
                gene = unlist(clu_ids_list), 
                universe = unlist(universe_list_ids), 
                pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
                TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)
            # if(nrow(as.data.frame(enrich_res)) == 0){return(NA)}
            
            enrich_res <- setReadable_custom(enrich_res, 
                                            featureDict = Reduce(c, clu_ids_list))
            return(enrich_res)
        })
        names(enrich_results) <- names(clu_ids_enrich[[score_id]][[factor_id]])
        return(enrich_results[!is.na(enrich_results)])
    })
    names(factor_results) <- names(clu_ids_enrich[[score_id]])
    factor_results
})
names(enrich_results_list) <- names(clu_ids_enrich)

results_wb <- openxlsx::createWorkbook()
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            sheet_name <- paste(factor_id, weight_id, clu_id, sep = ".")
            openxlsx::addWorksheet(wb = results_wb,
                                    sheetName = sheet_name)
            openxlsx::writeData(wb = results_wb, sheet = sheet_name,
                                x = as.data.frame(clu_dt))
        })
    })
    openxlsx::saveWorkbook(results_wb, 
                            file = paste0(enrich_nw_dir, "enrich_results.xlsx"), 
                            overwrite = TRUE)
})

```

```{r, fig.height=6, fig.width=8}
library(ggplot2)
pdf(paste0(enrich_nw_dir, enrich_name, "_Pathway_Enrich.pdf"))
lapply(names(enrich_results_list), function(factor_id){
    factor_dt <- enrich_results_list[[factor_id]]
    lapply(names(factor_dt), function(weight_id){
        weight_dt <- factor_dt[[weight_id]]
        lapply(names(weight_dt), function(clu_id){
            clu_dt <- weight_dt[[clu_id]]
            if(nrow(as.data.frame(clu_dt)) == 0){return(NULL)}
            plot_name <- gsub("_", " ", paste(factor_id, weight_id))
            
            dotplot(clu_dt) + ggtitle(paste("Enriched Kegg Pathways" , plot_name, clu_id))
        })
        
    })
})
dev.off()
```

###### Network creation

```{r}
xfun::Rscript_call(rmarkdown::render,
    args = list(
        input = "../Multi_Omics_Correlations/Correlations_Network_Enrichment_Plots.Rmd",
        output_file = output_dir,
        params = list(
            nw_graphs = nw_graphs,
            enrich_results_list = enrich_results_list,
            output_dir = enrich_nw_dir,
            omic_dt = omic_dt,
            group_var = "group",
            universe_ids_list = universe_list_ids,
            var_order_col_list = var_order_col_list,
            omic_names_dict = omic_names_dict,
            omic_color_dict = omic_color_dict,
            omic_shape_dict = shape_vect_ext,
            group_color_dict = group_color_dict,
            feature_ids_dict = feature_ids_dict)
    )
)
```


```{r}
## All MOFA2 background
# full_mofa_features <- unlist(rownames(omic_dt))
# full_mofa_features <- paste0(names(full_mofa_features), ".", full_mofa_features)
# 
# metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][grep("Untargeted_Metabolomics", 
#                                                                     full_mofa_features, value = T)]
# metabo_dt <- metabo_dt[!is.null(metabo_dt)]
# metabo_dt <- unlist(strsplit(metabo_dt, ";"))
# metabo_dt <- metabo_dt[!is.na(metabo_dt)]
# 
# wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][grep("WGB_DNA_Methylation", 
#                                                                 full_mofa_features, value = T)]
# wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)]
# 
# rna_dt <- omic_dict_list[["RNA_Sequencing"]][grep("RNA_Sequencing", 
#                                                                 full_mofa_features, value = T)]
# rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)]
# 
# full_universe_ids <- c(metabo_dt, wgb_dt, rna_dt)
# 
# enrich_results_list_full <- lapply(names(clu_ids_enrich), function(score_id){
#     factor_results <- lapply(names(clu_ids_enrich[[score_id]]), function(factor_id){
#         enrich_results <- lapply(clu_ids_enrich[[score_id]][[factor_id]], function(clu_ids){
#             enrich_res <- enricher(gene = clu_ids, universe = full_universe_ids, 
#                                     pvalueCutoff = 0.05, pAdjustMethod = "fdr", 
#                                     TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt)  
#             if(nrow(as.data.frame(enrich_res)) == 0){return(NA)}
#             return(enrich_res)
#         })
#         return(enrich_results[!is.na(enrich_results)])
#     })
#     names(factor_results) <- names(clu_ids_enrich[[score_id]])
#     factor_results
# })
# names(enrich_results_list_full) <- names(clu_ids_enrich)

# clu_dict_list <- lapply(clu_ids_enrich, function(score_dt){
#     lapply(score_dt, function(weight_dt){
#         lapply(weight_dt, function(clu_dt){
#             dict_res <- names(clu_dt)
#             names(dict_res) <- clu_dt
#             dict_res    
#         })
#     })
# })

```

```{r}
# annotate(x = 20, y = 10, geom = "text", label = labels_dt$var_id)
# print(p)
# 
# labels_dt <- distinct(plot_dt[,c("omic_var_id_readable", "omic_name_readable", "omic_var_id")])
# labels_dt <- labels_dt[order(labels_dt$omic_name_readable),]
# 
# ###
# 
# legend_colors <- color_vect[as.character(sapply(levels(plot_dt$omic_var_id_readable), function(x){
#     labels_dt$omic_var_id_readable[as.character(labels_dt$omic_var_id_readable) == x]
# }))]
# 
# 
# p <- ggplot(plot_dt, aes(x = sample_name, y = value, color = omic_name_readable, 
#                          group = omic_var_id_readable, 
#                          shape = omic_var_id_readable)) + 
#     geom_point() + geom_smooth(aes(color = omic_name_readable, fill = omic_name_readable),
#                                se = TRUE, span = 0.5, alpha = 0.1) + theme_minimal() + 
#     theme(axis.text.x = element_text(angle = 90), 
#             legend.position = "right", 
#             legend.justification = "top", 
#             legend.title = element_blank(), 
#             plot.title = element_text(hjust = 0.5),
#             legend.spacing.y = unit(0.1, "cm")) + 
#     ggtitle(term_name) + 
#     xlab(NULL) + ylab("Normalized Values") + 
#     scale_color_manual(values = color_vect) +
#     scale_fill_manual(values = color_vect) + 
#     guides(shape = guide_legend(label.theme = element_text(size = 7.5), 
#                                 order = 2,  byrow = TRUE,
#                                 override.aes = list(
#                                     color = legend_colors
#                                 ),
#                                 keyheight = 0.05), 
#            col = guide_legend(order = 1), fill = "none") + 
#     scale_shape_manual(values = rep(16, length(unique(plot_dt$omic_var_id))))
#     # annotate(x = 20, y = 10, geom = "text", label = labels_dt$var_id)
# print(p)
# 
# #######
# # test <- reshape2::dcast(plot_dt, formula = 1 ~ omic_var_id, fun.aggregate = mean)
# library(dplyr)
# av_dt <- plot_dt %>% group_by(group, omic_var_id) %>% summarise(avg = mean(value), dt_sd = sd(value), omic_name = unique(omic_name))
# 
# p2 <- ggplot(av_dt, aes(x = group, y = avg, color = omic_name, 
#                          group = omic_var_id, shape = omic_var_id)) + 
#     geom_point() + geom_line() + geom_ribbon(aes(ymax = avg + dt_sd, 
#                        ymin = avg - dt_sd, fill = omic_name),
#                        alpha=0.2) + theme_minimal() + 
#     theme(axis.text.x = element_text(angle = 90), 
#             legend.position = "right", 
#             legend.justification = "top", 
#             legend.title = element_blank(), 
#             plot.title = element_text(hjust = 0.5),
#             legend.spacing.y = unit(0.1, "cm")) + 
#     ggtitle(term_name) + 
#     xlab(NULL) + ylab("Normalized Values") + 
#     scale_color_manual(values = color_vect) + 
#     scale_fill_manual(values = color_vect) +
#     guides(shape = guide_legend(label.theme = element_text(size = 7.5), 
#                                 order = 2,  byrow = TRUE,
#                                 override.aes = list(
#                                     color = legend_colors
#                                 ),
#                                 keyheight = 0.05), 
#            col = guide_legend(order = 1), fill = "none") + 
#     scale_shape_manual(values = rep(16, length(unique(plot_dt$omic_var_id))))
#     # annotate(x = 20, y = 10, geom = "text", label = labels_dt$var_id)
# print(p2)

```



```{r}
# enrich_graph_clusters <- function(cluster_id_list, membership_attribute, 
#                                   universe, min_n = 25){
#     membership_vector <- vertex_attr(graph, membership_attribute)
#     clu_vertex <- lapply(unique(membership_vector), function(cl_id){
#         names(V(graph))[membership_vector == cl_id]
#     })
#     clu_vertex <- clu_vertex[lengths(clu_vertex) >= min_n]
#     rna_seq_tag <- "RNA_Sequencing\\."
#     dna_seq_tag <- "WGB_DNA_Methylation\\."
#     
#     universe_rna_seq <- grep(rna_seq_tag, universe, value = TRUE)
#     universe_rna_seq <- gsub(rna_seq_tag, "", universe_rna_seq)
#     
#     universe_dna_seq <- grep(dna_seq_tag, universe, value = TRUE)
#     universe_dna_seq <- gsub(dna_seq_tag, "", universe_dna_seq)
#     universe_dna_seq <- rowRanges(omic_dt[["WGB_DNA_Methylation"]][universe_dna_seq,])
# }


 # KEGGREST::keggFind(database = "compound", query = universe_features_kegg[1])[[1]]$PATHWAY

# mSet<-InitDataObjects("conc", "pathinteg", FALSE)
# mSet<-SetOrganism(mSet, "mmu")
# 
# geneListFile<-"replace_with_your_file_name"
# geneList<-readChar(geneListFile, file.info(geneListFile)$size)
# 
# mSet<-PerformGeneMapping(mSet, geneList, "hsa", "symbol");
# 
# cmpdListFile<-"replace_with_your_file_name"
# cmpdList<-readChar(cmpdListFile, file.info(cmpdListFile)$size)
# mSet<-PerformCmpdMapping(mSet, cmpdList, "hsa", "name");
# mSet<-CreateMappingResultTable(mSet)
# 
# mSet<-PrepareIntegData(mSet);
# mSet<-PerformIntegPathwayAnalysis(mSet, "dc", "hyper", "all", "query");
# 
# mSet<-PlotPathSummary(mSet, F, "path_view_0_", "png", 72, width=NA, NA, NA )
# mSet<-CreateIntegMatchingTable(mSet);
# 
# saveRDS(mSet, paste0(output_path, results_name))
```


<!-- ## LIMA full unsupervised -->

<!-- ```{r} -->
<!-- multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|DNA_Methylation.liver", multi_omics_names)] -->
<!-- # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)]) -->
<!-- multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)]) -->

<!-- multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected]) -->
<!-- plot_data_overview(multi_omics_mofa) -->
<!-- ``` -->

<!-- ## Parameters -->

<!-- ```{r} -->
<!-- data_opts <- get_default_data_options(multi_omics_mofa) -->
<!-- head(data_opts) -->

<!-- model_opts <- get_default_model_options(multi_omics_mofa) -->
<!-- head(model_opts) -->
<!-- model_opts$num_factors <- 8 -->

<!-- train_opts <- get_default_training_options(multi_omics_mofa) -->
<!-- head(train_opts) -->
<!-- ``` -->

<!-- ## Model and training -->

<!-- ```{r} -->
<!-- multi_omics_mofa <- prepare_mofa( -->
<!--   object = multi_omics_mofa, -->
<!--   data_options = data_opts, -->
<!--   model_options = model_opts, -->
<!--   training_options = train_opts -->
<!-- ) -->

<!-- outfile <- file.path(getwd(),"model_us.hdf5") -->
<!-- multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE) -->

<!-- ``` -->

<!-- ## Downstream Analysis -->

<!-- ```{r} -->
<!-- # file_path <- "./model.hdf5" -->
<!-- # model <- load_model(file_path) -->
<!-- colData(multi_omics)$sample <- colData(multi_omics)$samplename -->

<!-- samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample") -->

<!-- plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90)) -->

<!-- p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5,  -->
<!--                  color_by = "group.y",  -->
<!--                  add_violin = TRUE) -->

<!-- p <- p +  -->
<!--   scale_color_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) + -->
<!--   scale_fill_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) -->
<!-- print(p) -->
<!-- ``` -->

<!-- ```{r, fig.height=12, fig.width=12} -->
<!-- plot_factors(multi_omics_mofa,  -->
<!--   factors = 1:6, -->
<!--   color_by = "group.y" -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pca_dt <- MOFA2::get_factors(object = multi_omics_mofa, factors = 1:3) -->

<!-- MOFA2::get_variance_explained(object = multi_omics_mofa, factors = 1:3) -->

<!-- library(plotly) -->

<!-- d3plot <- plot_ly(x = pca_dt$group1[,1], y = pca_dt$group1[,2], z = pca_dt$group1[,3], color = multi_omics_mofa@samples_metadata$group.y) -->
<!-- d3plot <- d3plot %>% add_markers() -->
<!-- # d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")), -->
<!-- #                      yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")), -->
<!-- #                      zaxis = list(title = paste0('PC4', "(", pcvar[4], "%)")))) -->
<!-- d3plot -->
<!-- ``` -->

<!-- ```{r} -->
<!-- plot_weights(multi_omics_mofa,  -->
<!--   view = 1, -->
<!--   factor = 2, -->
<!--   nfeatures = 10,     # Number of features to highlight -->
<!--   scale = T,          # Scale weights from -1 to 1 -->
<!--   abs = F            # Take the absolute value? -->
<!-- ) -->

<!-- weight_dt <- get_weights(object = multi_omics_mofa, factors = 2, as.data.frame = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(view_id in seq_len(length(multi_omics_selected))){ -->
<!--     plot_data_heatmap(multi_omics_mofa, -->
<!--       view = view_id,         # view of interest -->
<!--       factor = 2,             # factor of interest -->
<!--       features = 40,          # number of features to plot (they are selected by weight) -->

<!--       # extra arguments that are passed to the `pheatmap` function -->
<!--       cluster_rows = TRUE, cluster_cols = TRUE, -->
<!--       show_rownames = TRUE, show_colnames = FALSE, annotation_samples = "group.y", -->
<!--     ) -->
<!-- } -->
<!-- ``` -->


<!-- # ```{r} -->
<!-- # # param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/", -->
<!-- # #     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/", -->
<!-- # #     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "HFE"], -->
<!-- # #     group_var = "group", -->
<!-- # #     num_factors = num_factors, -->
<!-- # #     results_name = "mofa_int", -->
<!-- # #     color_vec = c("CNT"="black", "HFD"="red", "INT"="blue") -->
<!-- # # ) -->
<!-- # # # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green") -->
<!-- # # xfun::Rscript_call(rmarkdown::render, -->
<!-- # #                     args = list(input = "./MOFA_Models.Rmd",  -->
<!-- # #                                 output_file = paste0(param_list$output_dir, -->
<!-- # #                                                      param_list$results_name, ".html"),  -->
<!-- # #                                 params = param_list -->
<!-- # #                             ) -->
<!-- # #                     ) -->
<!-- #  -->
<!-- # param_list <- list(input_dir = "../../outputs/Multi_Omics_Analysis/", -->
<!-- #     output_dir = "../../outputs/Multi_Omics_Analysis/MOFA_Results/", -->
<!-- #     multi_omics_exp = multi_omics[,colData(multi_omics)$group != "INT"], -->
<!-- #     group_var = "group", -->
<!-- #     num_factors = num_factors, -->
<!-- #     results_name = "mofa_hfe", -->
<!-- #     color_vec = c("CNT"="black", "HFD"="red", "HFE"="green") -->
<!-- # ) -->
<!-- # # c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green") -->
<!-- # xfun::Rscript_call(rmarkdown::render, -->
<!-- #                     args = list(input = "./MOFA_Models.Rmd",  -->
<!-- #                                 output_file = paste0(param_list$output_dir, -->
<!-- #                                                      param_list$results_name, ".html"),  -->
<!-- #                                 params = param_list -->
<!-- #                             ) -->
<!-- #                     ) -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ## Pre-Analysis -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # mice_dt <- openxlsx::read.xlsx(paste0(params$input_dir, "AnimalData_OscarYanes.xlsx")) -->
<!-- #  -->
<!-- # # mice_num <- mice_dt[,c("BW(g)", "WATep(g)", "Liver(g)")] -->
<!-- # mice_num <- mice_dt -->
<!-- # mice_s <- as.data.frame(mice_num) -->
<!-- #  -->
<!-- # dict_vect <- c("Ctrl" = "CNT", "Ctrl ext" = "CNT", "HFD" = "HFD", "INT2" = "INT", "HFD excl." = "HFE") -->
<!-- # mice_s$group <- factor(dict_vect[mice_dt$Group], levels = c("CNT", "HFD", "INT", "HFE")) -->
<!-- #  -->
<!-- # dict_vect_g <- c("CNT" = "C", "HFD" = "H", "INT" = "I", "HFE" = "E") -->
<!-- #  -->
<!-- # mice_s$samplename <- paste0(make.names(mice_dt$ID), dict_vect_g[mice_s$group]) -->
<!-- #  -->
<!-- #  -->
<!-- # mice_s$samplename[!mice_s$samplename %in% multi_omics$samplename] -->
<!-- # multi_omics$samplename[!multi_omics$samplename %in% mice_s$samplename] -->
<!-- #  -->
<!-- # mice_s$samplename[mice_s$samplename == "X686I"] <- "X687I" -->
<!-- #  -->
<!-- # mice_s <- dplyr::left_join(as.data.frame(colData(multi_omics)), mice_s, by = "samplename") -->
<!-- # mice_s -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # d3plot <- plot_ly(x = mice_s$`BW(g)`, y = mice_s$`WATep(g)`, z = mice_s$`Liver(g)`, -->
<!-- #                     color = as.factor(mice_s$Sacrifice)) -->
<!-- # d3plot <- d3plot %>% add_markers() -->
<!-- # d3plot <- d3plot %>% layout(scene = list( -->
<!-- #     xaxis = list(title = "BW (g)"), -->
<!-- #     yaxis = list(title = "WATep (g)"), -->
<!-- #     zaxis = list(title = "Liver (g)"))) -->
<!-- # d3plot -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # factor_dt <- readRDS(paste0(params$output_dir, "/MOFA_Results/any_sig_005_mofa_int_all_meta/factor_data.rds")) -->
<!-- #  -->
<!-- # mice_sub <- mice_s[mice_s$group.x != "HFE",] -->
<!-- #  -->
<!-- # all(names(factor_dt$group1[,3]) == mice_sub$samplename) -->
<!-- #  -->
<!-- # cor(mice_sub$`BW(g)`, factor_dt$group1[,3]) -->
<!-- # cor(mice_sub$`WATep(g)`, factor_dt$group1[,3]) -->
<!-- # cor(mice_sub$`Liver(g)`, factor_dt$group1[,3]) -->
<!-- #  -->
<!-- #  -->
<!-- # ctr_sub <- mice_sub[mice_sub$group.x == "CNT",] -->
<!-- # ctr_fact <- factor_dt$group1[mice_sub$group.x == "CNT",] -->
<!-- #  -->
<!-- # cor(ctr_sub$`BW(g)`, ctr_fact[,3]) -->
<!-- # cor(ctr_sub$`WATep(g)`, ctr_fact[,3]) -->
<!-- # cor(ctr_sub$`Liver(g)`, ctr_fact[,3]) -->
<!-- #  -->
<!-- # plot(ctr_sub$`BW(g)`, ctr_fact[,3]) -->
<!-- # plot(ctr_sub$`WATep(g)`, ctr_fact[,3]) -->
<!-- # plot(ctr_sub$`Liver(g)`, ctr_fact[,3]) -->
<!-- #  -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ## GFCV -->
<!-- #  -->
<!-- # ## Loaded matrices -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # library(MOFA2) -->
<!-- # multi_omics_raw <- readRDS("./Merged_Multi_Omics_gfcv.rds") -->
<!-- # multi_omics <- readRDS("./Multivariate_Merged_Multi_Omics_gfcv.rds") -->
<!-- # names(experiments(multi_omics)) -->
<!-- #  -->
<!-- # filter_list <- c("16S") -->
<!-- #  -->
<!-- # selected_omics <- lapply(filter_list, function(dt_name){ -->
<!-- #     omic_dt_names <- grep(dt_name, names(experiments(multi_omics))) -->
<!-- #     omic_exp_list <- experiments(multi_omics[,,omic_dt_names]) -->
<!-- #      -->
<!-- #     sig_vars_num <- sapply(omic_exp_list, function(omic_exp){ -->
<!-- #         var_dt <- rowData(omic_exp) -->
<!-- #         p_val_dt <- var_dt[,grep("adj.p.value", colnames(var_dt))] -->
<!-- #         sig_vars <- apply(p_val_dt, 1, function(p_vals){ -->
<!-- #             any(p_vals <= 0.05, na.rm = TRUE) -->
<!-- #         }) -->
<!-- #         sum(unlist(sig_vars)) -->
<!-- #     }) -->
<!-- #     sig_vars_num -->
<!-- # }) -->
<!-- # names(selected_omics) <- filter_list -->
<!-- #  -->
<!-- # a <- do.call(cbind, selected_omics) -->
<!-- #  -->
<!-- # selected_omics <- lapply(filter_list, function(dt_name){ -->
<!-- #     omic_dt_names <- grep(dt_name, names(experiments(multi_omics))) -->
<!-- #     omic_exp_list <- experiments(multi_omics[,,omic_dt_names]) -->
<!-- #      -->
<!-- #     sig_vars_num <- sapply(omic_exp_list, function(omic_exp){ -->
<!-- #         nrow(omic_exp) -->
<!-- #     }) -->
<!-- #     sig_vars_num -->
<!-- # }) -->
<!-- # names(selected_omics) <- filter_list -->
<!-- #  -->
<!-- # b <- do.call(cbind, selected_omics) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Preparing the omics matrices  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # multi_omics_names <- names(experiments(multi_omics)) -->
<!-- #  -->
<!-- # multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S", multi_omics_names)] -->
<!-- # # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)]) -->
<!-- # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)]) -->
<!-- # multi_omics_selected -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # filtered_exps <- lapply(multi_omics_selected, function(multi_omics_names){ -->
<!-- #     multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]] -->
<!-- #  -->
<!-- #     if(grepl("RNASeq", multi_omics_names)){ -->
<!-- #         var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){ -->
<!-- #             sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE) -->
<!-- #         }) -->
<!-- #         multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, 0)],] -->
<!-- #     } -->
<!-- #     multi_omics_exp -->
<!-- # }) -->
<!-- # names(filtered_exps) <- multi_omics_selected -->
<!-- # experiments(multi_omics) <- ExperimentList(filtered_exps) -->
<!-- # ``` -->
<!-- # ```{r} -->
<!-- # multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics) -->
<!-- # plot_data_overview(multi_omics_mofa) -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected]) -->
<!-- # plot_data_overview(multi_omics_mofa) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Parameters -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # data_opts <- get_default_data_options(multi_omics_mofa) -->
<!-- # head(data_opts) -->
<!-- #  -->
<!-- # model_opts <- get_default_model_options(multi_omics_mofa) -->
<!-- # head(model_opts) -->
<!-- # model_opts$num_factors <- 8 -->
<!-- #  -->
<!-- # train_opts <- get_default_training_options(multi_omics_mofa) -->
<!-- # head(train_opts) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Model and training -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # multi_omics_mofa <- prepare_mofa( -->
<!-- #   object = multi_omics_mofa, -->
<!-- #   data_options = data_opts, -->
<!-- #   model_options = model_opts, -->
<!-- #   training_options = train_opts -->
<!-- # ) -->
<!-- #  -->
<!-- # outfile <- file.path(getwd(),"model.hdf5") -->
<!-- # multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE) -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- # ## Downstream Analysis -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # # file_path <- "./model.hdf5" -->
<!-- # # model <- load_model(file_path) -->
<!-- # colData(multi_omics)$sample <- colData(multi_omics)$samplename -->
<!-- #  -->
<!-- # samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample") -->
<!-- #  -->
<!-- # plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90)) -->
<!-- #  -->
<!-- # p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5,  -->
<!-- #                  color_by = "group.y",  -->
<!-- #                  add_violin = TRUE) -->
<!-- #  -->
<!-- # p <- p +  -->
<!-- #   scale_color_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green")) + -->
<!-- #   scale_fill_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green")) -->
<!-- # print(p) -->
<!-- # ``` -->

## Old Code

<!-- ### ORA analysis {.tabset} -->

<!-- ```{r} -->
<!-- output_dir <- paste0(params$output_dir, "MOFA_Results/any_sig_005_mofa_int/") -->
<!-- omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds")) -->
<!-- rna_seq_rd <- rowData(experiments(omic_dt)$RNA_Sequencing) -->

<!-- rna_seq_list <- lapply(seq_len(num_factors), function(factor_id){ -->
<!--     list("All_Weights" = rna_seq_rd, -->
<!--         "Positive_Weights" = rna_seq_rd[rna_seq_rd[[paste0("Factor_", factor_id, "_weights")]] >= 0,], -->
<!--         "Negative_Weights" = rna_seq_rd[rna_seq_rd[[paste0("Factor_", factor_id, "_weights")]] < 0,]) -->
<!-- }) -->
<!-- names(rna_seq_list) <- paste0("Factor_", seq_len(num_factors)) -->

<!-- enrich_results_dir <- paste0(output_dir, "enrich_results/") -->

<!-- if(!dir.exists(enrich_results_dir)){dir.create(enrich_results_dir)} -->

<!-- for(rna_seq_name in names(rna_seq_list)){ -->

<!--     rna_seq_exp <- rna_seq_list[[rna_seq_name]] -->

<!--     gsea_cols_list <- lapply(names(rna_seq_exp), function(x){paste0(rna_seq_name, "_weights")}) -->
<!--     names(gsea_cols_list) <- names(rna_seq_exp) -->

<!--     param_list <- list( -->
<!--         output_dir = enrich_results_dir, -->
<!--         exp_list = rna_seq_exp, -->
<!--         gsea_cols_list = gsea_cols_list, -->
<!--         gsea = TRUE, -->
<!--         ora_filter = parse(text = paste0(rna_seq_name, "_weights_sig")), -->
<!--         order_column = paste0(rna_seq_name, "_weights"), -->
<!--         results_name = rna_seq_name, -->
<!--         legend_name = paste0(rna_seq_name, "_weights") -->
<!--     ) -->
<!--     xfun::Rscript_call(rmarkdown::render, -->
<!--                         args = list(input = "./RNA_Sequencing_Enrichments.Rmd",  -->
<!--                                     output_file = paste0(param_list$output_dir,  -->
<!--                                                          "RNA_Sequencing_Enrichment_", -->
<!--                                                          rna_seq_name, ".html"),  -->
<!--                                     params = param_list -->
<!--                                 ) -->
<!--                         ) -->
<!-- } -->
<!-- ``` -->

<!-- The positive weights from the factor 1, genes that increase with the high fat diet and get reduced by the intervention, are enriched with GO terms related to the immune system such as response to beta interferon, chemotaxis, phagocytosis... These can be very well related with the known inflammatory problems that arise from the high fat diet, the subsequent infiltration of of immune cells into the liver and its reversibility before reaching the liver cyrrosis and loss of function. The negative weights, which includes genes that are reduced by the high fat diet and increased with the intervention, indicate epigenetic changes showing GO terms such as demethylation, that inversely correlates with the WGB results where there was an increased methylation state in the high fat diet (could check if all the weights are positive in the WGB), meaning that the demethylation was reduced (Could plot all the genes related to demethylation and correlate them to the DMRs) and chromatin remodelling; and metabolic changes such as catabolic processes which makes sense to be reduced when there is an excess of nutrients. -->

<!-- The positive weights from factor 2, genes that are increased by the high fat diet and they keep their expression levels, show changes in the metabolism, focused on lipids and small molecules       -->



<!-- ### rGREAT Enrichment & Region Analysis {.tabset} -->

<!-- ```{r} -->
<!-- dna_met_list <- lapply(seq_len(num_factors), function(factor_id){ -->
<!--     dna_met_dt <- experiments(omic_dt)$WGB_DNA_Methylation -->
<!--     list("All_Weights" = dna_met_dt, -->
<!--         "Positive_Weights" = dna_met_dt[rowData(dna_met_dt)[[paste0("Factor_", factor_id, "_weights")]] >= 0,], -->
<!--         "Negative_Weights" = dna_met_dt[rowData(dna_met_dt)[[paste0("Factor_", factor_id, "_weights")]] < 0,]) -->
<!-- }) -->

<!-- names(dna_met_list) <- paste0("Factor_", seq_len(num_factors)) -->

<!-- rgreat_output_dir <- paste0(output_dir, "rGreat_results/") -->

<!-- if(!dir.exists(rgreat_output_dir)){dir.create(rgreat_output_dir)} -->

<!-- for(dna_seq_name in names(dna_met_list)){ -->

<!--     dna_seq_exp <- dna_met_list[[dna_seq_name]] -->

<!--     gsea_cols_list <- lapply(names(dna_seq_exp), function(x){paste0(dna_seq_name, "_weights")}) -->
<!--     names(gsea_cols_list) <- names(dna_seq_exp) -->

<!--     param_list <- list( -->
<!--         dmr_exp_list = dna_seq_exp, -->
<!--         txdb_id = "TxDb.Mmusculus.UCSC.mm39.refGene", -->
<!--         msigdbr_specie = "Mus musculus", -->
<!--         ora_filter = parse(text = paste0(dna_seq_name, "_weights_sig")), -->
<!--         random_seed = 1234, -->
<!--         output_dir = output_dir, -->
<!--         results_name = dna_seq_name -->
<!--     ) -->
<!--     xfun::Rscript_call(rmarkdown::render, -->
<!--                         args = list(input = "../WGB_DNA_Methylation/DMR_enrichment_analysis.Rmd",  -->
<!--                                     output_file = paste0(rgreat_output_dir,  -->
<!--                                                          "WGB_Sequencing_Enrichment_", -->
<!--                                                          dna_seq_name, ".html"),  -->
<!--                                     params = param_list -->
<!--                                 ) -->
<!--                         ) -->
<!-- } -->
<!-- ``` -->

<!-- ### MetaboAnalyst Mummichog analysis -->

<!-- ```{r} -->
<!-- library(MetaboAnalystR) -->
<!-- library(SummarizedExperiment) -->

<!-- unt_met_list <- lapply(seq_len(num_factors), function(factor_id){ -->
<!--     unt_met_dt <- experiments(omic_dt)$Untargeted_Metabolomics -->

<!--     list("All_Weights" = unt_met_dt, -->
<!--         "Positive_Weights" = unt_met_dt[rowData(unt_met_dt)[[paste0("Factor_", factor_id, "_weights")]] >= 0,], -->
<!--         "Negative_Weights" = unt_met_dt[rowData(unt_met_dt)[[paste0("Factor_", factor_id, "_weights")]] < 0,]) -->
<!-- }) -->

<!-- names(unt_met_list) <- paste0("Factor_", seq_len(num_factors)) -->

<!-- unt_met_dt <- lapply(names(unt_met_list), function(factor_id){ -->
<!--     factor_dt <- unt_met_list[[factor_id]] -->
<!--     lapply(factor_dt, function(weight_dt){ -->
<!--         feature_dt <- as.data.frame(rowData(weight_dt))[,c("mzmed", "rtmed", -->
<!--                                                         paste0(factor_id, "_weights"))] -->
<!--         feature_dt$p.value <- ifelse(rowData(weight_dt)[,paste0(factor_id,  -->
<!--                                                                 "_weights_sig")], 0.05, 1) -->
<!--         feature_dt$mode <- tolower(gsub("^.*_", "", gsub("\\..*$", "",  -->
<!--                                                         rownames(feature_dt)))) -->
<!--         feature_dt <- feature_dt[c("mzmed", "rtmed", paste0(factor_id, "_weights"),  -->
<!--                                     "p.value", "mode")] -->
<!--         colnames(feature_dt) <- c("mz", "rt", "t.score", "p.value", "mode") -->
<!--         feature_dt$t.score <- abs(feature_dt$t.score) -->
<!--         feature_dt -->
<!--     }) -->
<!-- }) -->
<!-- names(unt_met_dt) <- names(unt_met_list) -->

<!-- old_dir <- getwd() -->
<!-- mummichog_dir <- paste0(output_dir, "mummichog_metaboanalyst/") -->

<!-- if(!dir.exists(mummichog_dir)){dir.create(mummichog_dir)} -->

<!-- for(factor_id in names(unt_met_dt)){ -->
<!--     for(weight_id in names(unt_met_dt[[factor_id]])){ -->
<!--         results_output <- paste0(mummichog_dir, factor_id, "_", weight_id, "/")     -->
<!--         if(!dir.exists(results_output)){dir.create(results_output)} -->
<!--         setwd(results_output) -->
<!--         write.table(unt_met_dt[[factor_id]][[weight_id]], row.names = FALSE,  -->
<!--                     file = "./mummichog_dt.txt", sep = ",") -->

<!--         mSet <- InitDataObjects("mass_all", "mummichog", FALSE) -->

<!--         mSet <- UpdateInstrumentParameters(mSet, 10.0, "mixed", "yes", 0.05) -->

<!--         mSet <- Read.PeakListData(mSet, "mummichog_dt.txt") -->

<!--         mSet <- SetRTincluded(mSet, "seconds") -->

<!--         mSet <- SanityCheckMummichogData(mSet) -->

<!--         mSet <- SetPeakEnrichMethod(mSet, "mum", "v2") -->

<!--         mSet <- SetMummichogPval(mSet, cutoff = 0.05) -->

<!--         mSet <- PerformPSEA(mSet, "mmu_kegg", "current", 3 , 100) -->

<!--         mSet<-PlotPeaks2Paths(mSet, "peaks_to_paths_0_", "png", 72, width=NA) -->
<!--         saveRDS(mSet, file = "./metaboanalyst_obj.rds") -->
<!--         rm(mSet) -->
<!--         setwd(old_dir) -->
<!--     } -->
<!-- } -->

<!-- # setwd(old_dir) -->
<!-- ``` -->

<!-- ### Comparing DMR and Transcriptomics enrichment results -->

<!-- ```{r} -->
<!-- library(GOSemSim) -->
<!-- library(org.Mm.eg.db) -->
<!-- library(AnnotationDbi) -->
<!-- library(DOSE) -->


<!-- overlap_list <- list() -->
<!-- for(factor_id in paste0("Factor_", seq_len(num_factors))){ -->
<!--     rna_seq_enrich <- readRDS(paste0(enrich_results_dir, factor_id, "_rna_seq_enrich_results.rds")) -->
<!--     rna_seq_enrich <- rna_seq_enrich$ORA -->
<!--     dna_seq_enrich <- readRDS(paste0(rgreat_output_dir, factor_id, "_dmr_enrich_results.rds")) -->
<!--     dna_seq_enrich <- dna_seq_enrich$ORA -->

<!--     overlap_list[[factor_id]] <- list() -->

<!--     for(type_enrich in names(rna_seq_enrich)){ -->
<!--         rna_enrich_list <- rna_seq_enrich[[type_enrich]] -->
<!--         dna_enrich_list <- dna_seq_enrich[[type_enrich]] -->

<!--         overlap_list[[factor_id]][[type_enrich]] <- list() -->

<!--         for(enrich_id in names(rna_enrich_list)){ -->
<!--             rna_obj <- rna_enrich_list[[enrich_id]] -->
<!--             rna_dt <- as.data.frame(rna_obj) -->

<!--             dna_obj <- dna_enrich_list[[enrich_id]] -->
<!--             dna_dt <- dna_obj@table -->

<!--             # if(enrich_id == "kegg"){ -->
<!--             #     dna_dt$description <- gsub("^ ", "", sapply(strsplit(dna_dt$id, ":"), function(x){x[2]})) -->
<!--             #     dna_dt$id <- sapply(strsplit(dna_dt$id, ":"), function(x){x[1]}) -->
<!--             # } -->
<!--             # if(enrich_id == "wp"){ -->
<!--             #     dna_dt$description <- dna_dt$id -->
<!--             #     dna_dt$id <- sapply(strsplit(dna_dt$id, "%"),  -->
<!--             #                         grep, value = TRUE, pattern = "WP[[:digit:]]+") -->
<!--             #      -->
<!--             # } -->
<!--             if(enrich_id %in% c("BP", "CC", "MF")){ -->
<!--                 go_sem <- godata(OrgDb = org.Mm.eg.db, ont = tolower(enrich_id)) -->

<!--                 go_sim <- go_similarity_matrix(rna_dt$ID, dna_dt$id,  -->
<!--                                                 semData = go_sem, measure = "Wang") -->
<!--                 rna_sim <- rownames(go_sim)[apply(go_sim >= 0.7, 1, any)] -->
<!--                 dna_sim <- colnames(go_sim)[apply(go_sim >= 0.7, 2, any)] -->

<!--                 rna_filt <- rna_dt[rna_dt$ID %in% rna_sim,] -->
<!--                 dna_filt <- dna_dt[dna_dt$id %in% dna_sim,] -->

<!--                 rna_gene_list <- strsplit(rna_filt$geneID, "/") -->

<!--                 overlap_list[[factor_id]][[type_enrich]][[enrich_id]] <- list( -->
<!--                     "go_rna_dt" = rna_filt,  -->
<!--                     "go_dna_dt" = dna_filt) -->

<!--             } else{ -->
<!--                 rna_filt <- rna_dt[rna_dt$ID %in% intersect(rna_dt$ID, dna_dt$id),] -->
<!--                 dna_filt <- dna_dt[dna_dt$id %in% intersect(rna_dt$ID, dna_dt$id),] -->

<!--                 rna_gene_list <- strsplit(rna_filt$geneID, "/") -->
<!--                 rna_gene_list <- lapply(rna_gene_list, mapIds, x = org.Mm.eg.db, keytype = "ENTREZID", -->
<!--                                         column = "SYMBOL", multiVals = "first") -->

<!--                 overlap_list[[factor_id]][[type_enrich]][[enrich_id]] <- list( -->
<!--                     "go_rna_dt" = rna_filt,  -->
<!--                     "go_dna_dt" = dna_filt)    -->
<!--             } -->

<!--             names(rna_gene_list) <- rna_filt$ID -->

<!--             dna_annot <- getRegionGeneAssociations(dna_obj) -->
<!--             dmr_gene_list <- as.list(mcols(dna_annot)$annotated_genes) -->
<!--             names(dmr_gene_list) <- rownames(mcols(dna_annot)) -->

<!--             found_gsets <- dna_obj@gene_sets[dna_filt$id] -->

<!--             dna_gene_list <- lapply(found_gsets, function(x){ -->
<!--                 found_g <- lapply(dmr_gene_list, function(dmr_ids){ -->
<!--                     dmr_ids[x] -->
<!--                 }) -->
<!--                 found_g <- unlist(found_g) -->
<!--                 unique(found_g[!is.na(found_g)]) -->
<!--             }) -->
<!--             gene_overlap <- intersect(unique(unlist(dna_gene_list)), -->
<!--                                       unique(unlist(rna_gene_list))) -->
<!--             found_pairs <- data.frame("transcript" = gene_overlap,  -->
<!--                        "dmr" = sapply(gene_overlap, function(x){ -->
<!--                            found_dmr <- sapply(dmr_gene_list, function(dmr_info){ -->
<!--                                any(dmr_info %in% x) -->
<!--                            }) -->
<!--                            paste(names(found_dmr)[found_dmr], collapse = ";") -->
<!--                        }),  -->
<!--                        "dmr_GO" = sapply(gene_overlap, function(x){ -->
<!--                            found_dmr <- sapply(dna_gene_list, function(dmr_info){ -->
<!--                                any(dmr_info %in% x) -->
<!--                            }) -->
<!--                            paste(names(found_dmr)[found_dmr], collapse = ";") -->
<!--                        }),  -->
<!--                        "gene_GO" = sapply(gene_overlap, function(x){ -->
<!--                            found_dmr <- sapply(rna_gene_list, function(dmr_info){ -->
<!--                                any(dmr_info %in% x) -->
<!--                            }) -->
<!--                            paste(names(found_dmr)[found_dmr], collapse = ";") -->
<!--                        })) -->
<!--             overlap_list[[factor_id]][[type_enrich]][[enrich_id]]$gene_overlap <- found_pairs -->



<!--             # overlap_list[[factor_id]][[type_enrich]][[enrich_id]] <- length(intersect(rna_dt, dna_dt))/length(union(rna_dt, dna_dt)) -->


<!--         } -->

<!--     } -->
<!-- } -->

<!-- saveRDS(overlap_list, file = paste0(output_dir, "go_overlap_results.rds")) -->

<!-- overlap_list <- readRDS(paste0(output_dir, "go_overlap_results.rds")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- lapply(names(overlap_list), function(factor_id){ -->
<!--     factor_dt <- overlap_list[[factor_id]] -->

<!--     lapply(names(factor_dt), function(weight_id){ -->
<!--         weight_dt <- factor_dt[[weight_id]] -->

<!--         lapply(names(weight_dt), function(enrich_id){ -->
<!--             enrich_dt <- weight_dt[[enrich_id]] -->

<!--             if(nrow(enrich_dt$gene_overlap) <= 2){return(NULL)} -->
<!--             rna_genes <- enrich_dt$gene_overlap$transcript -->
<!--             rna_seq_dt <- experiments(omic_dt)$RNA_Sequencing -->
<!--             rna_seq_dt <- rna_seq_dt[,rna_seq_dt$group != "HFE"] -->

<!--             p1 <- heatMapFun(dt_list = as.matrix(assay(rna_seq_dt[rowData(rna_seq_dt)$symbol %in% rna_genes,])), -->
<!--                        metadata = colData(rna_seq_dt), stratificationVar = "group",  -->
<!--                        clustering_distance_rows = "euclidean", show_colnames = T,  -->
<!--                        plot_name = paste(factor_id, weight_id, enrich_id)) -->

<!--             dmr_genes <- enrich_dt$gene_overlap$transcript -->
<!--             dmr_seq_dt <- experiments(omic_dt)$RNA_Sequencing -->
<!--             dmr_seq_dt <- dmr_seq_dt[,dmr_seq_dt$group != "HFE"] -->

<!--             p2 <-heatMapFun(dt_list = as.matrix(assay(dmr_seq_dt[rowData(dmr_seq_dt)$symbol %in% rna_genes,])), -->
<!--                        metadata = colData(dmr_seq_dt), stratificationVar = "group",  -->
<!--                        clustering_distance_rows = "euclidean", show_colnames = T,  -->
<!--                        plot_name = paste(factor_id, weight_id, enrich_id)) -->

<!--             list(p1, p2) -->
<!--         }) -->
<!--     }) -->
<!-- }) -->
<!-- ``` -->

<!-- ### KEGG joint pathway analysis -->

<!-- ```{r} -->
<!-- output_dir <- paste0(params$output_dir, "MOFA_Results/sig_int_thr08/") -->
<!-- omic_dt <- readRDS(paste0(output_dir, "multi_omics_filtered.rds")) -->

<!-- # experiments(omic_dt) <- ExperimentList(lapply(experiments(omic_dt), function(omic_exp){ -->
<!-- #      -->
<!-- #     omic_rd <- rowData(omic_exp)[,grepl("Factor_[1,2,3]_weights$", colnames(rowData(omic_exp)))] -->
<!-- #      -->
<!-- #     rowData(omic_exp)$max_weight_id <- apply(abs(as.matrix(omic_rd)), 1, which.max) -->
<!-- #     omic_exp -->
<!-- # })) -->

<!-- kegg_db_enrich_path <- paste0(params$output_dir, "kegg_enrich_db.rds") -->

<!-- if(!file.exists(kegg_db_enrich_path)){ -->

<!--     library(stringr) -->
<!--     kegg_db_path <- paste0(params$output_dir, "mmu_kegg_pathway_db.rds") -->
<!--     test <- rGREAT::getKEGGPathways("mmu") -->

<!--     pathway_ids <- sapply(strsplit(names(test), ":"), function(x){x[[1]]}) -->
<!--     names(test) <- pathway_ids -->

<!--     if(!file.exists(kegg_db_path)){ -->
<!--         library(KEGGREST) -->

<!--         kegg_db_info <- lapply(pathway_ids, function(x){keggGet(x)[[1]]}) -->
<!--         names(kegg_db_info) <- pathway_ids -->
<!--         saveRDS(kegg_db_info, kegg_db_path)     -->
<!--     } else{ -->
<!--         kegg_db_info <- readRDS(kegg_db_path)     -->
<!--     } -->

<!--     gene_enrich_db <- lapply(names(kegg_db_info), function(kegg_db_id){ -->
<!--         gene_info <- as.numeric(kegg_db_info[[kegg_db_id]]$GENE) -->

<!--         gene_info <- gene_info[!is.na(gene_info)]     -->
<!--         as.character(gene_info) -->
<!--     }) -->
<!--     names(gene_enrich_db) <- names(kegg_db_info) -->

<!--     integ_enrich_db <- lapply(names(kegg_db_info), function(kegg_db_id){ -->
<!--         gene_info <- as.numeric(kegg_db_info[[kegg_db_id]]$GENE) -->

<!--         gene_info <- gene_info[!is.na(gene_info)]     -->
<!--         compound_info <- names(kegg_db_info[[kegg_db_id]]$COMPOUND) -->
<!--         db_info <- c(gene_info, compound_info) -->


<!--         if(length(db_info) == 0){return(NA)}     -->
<!--         data.frame("term" = kegg_db_id, "gene" = db_info,  -->
<!--                     "type" = c(rep("gene_id", length(gene_info)), -->
<!--                             rep("metabolite_id", length(compound_info))))         -->
<!--     }) -->

<!--     names(integ_enrich_db) <- names(kegg_db_info) -->


<!--     for(pathway_id in names(integ_enrich_db[is.na(integ_enrich_db)])){ -->
<!--         integ_enrich_db[[pathway_id]] <- data.frame("term" = pathway_id,  -->
<!--                                                     "gene" = test[[pathway_id]],  -->
<!--                                                     "type" = "gene_id") -->
<!--     } -->
<!--     integ_enrich_db <- do.call(rbind, integ_enrich_db) -->

<!--     term2name_dt <- lapply(names(kegg_db_info), function(kegg_id){ -->
<!--         data.frame("term" = kegg_id, "name" = kegg_db_info[[kegg_id]]$NAME) -->
<!--     }) -->
<!--     term2name_dt <- do.call(rbind, term2name_dt) -->

<!--     saveRDS(list("term2name" = term2name_dt, "term2gene" = integ_enrich_db),  -->
<!--             file = kegg_db_enrich_path) -->
<!-- } else{ -->
<!--     kegg_db_list <- readRDS(kegg_db_enrich_path) -->
<!--     term2name_dt <- kegg_db_list$term2name -->
<!--     integ_enrich_db <- kegg_db_list$term2gene -->
<!-- } -->

<!-- omic_enrich <- c("Untargeted_Metabolomics", "RNA_Sequencing", "WGB_DNA_Methylation") -->
<!-- ids_enrich_names <- c("metabo_analyst_cp", "entrezID", "geneId") -->

<!-- library(SummarizedExperiment) -->

<!-- ## Enrich IDs dictionary -->
<!-- omic_dict_list <- mapply(function(omic_id, enrich_var){ -->

<!--     omic_data <- omic_dt[[omic_id]] -->
<!--     omic_dict <- rowData(omic_data)[[enrich_var]] -->

<!--     names(omic_dict) <- paste0(omic_id, ".", rownames(omic_data)) -->

<!--     omic_dict -->

<!-- }, omic_id = omic_enrich, enrich_var = ids_enrich_names) -->

<!-- rna_seq_rd <- lapply(experiments(omic_dt)[c("RNA_Sequencing", "WGB_DNA_Methylation", "Untargeted_Metabolomics")], rowData) -->
<!-- num_factors <- 2 -->

<!-- ## Universe IDs mapped -->
<!-- universe_list_ids <- lapply(names(rna_seq_rd), function(omic_name){ -->
<!--     omic_ids <- omic_dict_list[[omic_name]][paste0(omic_name, ".", rownames(rna_seq_rd[[omic_name]]))] -->
<!--     omic_ids <- omic_ids[!is.null(omic_ids)] -->
<!--     omic_ids <- unlist(strsplit(omic_ids, ";")) -->
<!--     omic_ids[!is.na(omic_ids)] -->

<!-- }) -->
<!-- names(universe_list_ids) <- names(rna_seq_rd) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- feature_seq_list <- lapply(seq_len(num_factors), function(factor_id){ -->

<!--     factor_col_sig <- paste0("Factor_", factor_id, "_weights_sig") -->
<!--     # factor_col_sig <- paste0("max_weight_id == ", factor_id) -->

<!--     factor_col <- paste0("Factor_", factor_id, "_weights") -->

<!--     weight_expr_list <- list("All_Weights" = parse(text = factor_col_sig), -->
<!--                             "Positive_Weights" = parse(text = paste0(factor_col_sig, " & ", factor_col, " > 0")), -->
<!--                             "Negative_Weights" = parse(text = paste0(factor_col_sig, " & ", factor_col, " < 0"))) -->

<!--     lapply(weight_expr_list, function(weight_expr){ -->
<!--         filtered_sig_features <- lapply(rna_seq_rd, function(x){ -->
<!--             x <- as.data.frame(x) -->
<!--             rownames(x)[eval(expr = weight_expr, envir = x)] -->
<!--         }) -->
<!--         # filtered_sig_features <- Reduce(c, filtered_sig_features) -->
<!--         filtered_sig_features -->
<!--     }) -->
<!-- }) -->
<!-- names(feature_seq_list) <- paste0("Factor_", seq_len(num_factors)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- factor_ids_enrich <- lapply(feature_seq_list, function(factor_dt){ -->

<!--     lapply(factor_dt, function(weight_type_dt){ -->

<!--         metabo_dt <- omic_dict_list[["Untargeted_Metabolomics"]][paste0("Untargeted_Metabolomics",  -->
<!--                                                                         ".", weight_type_dt[["Untargeted_Metabolomics"]])] -->

<!--         metabo_dt <- metabo_dt[!is.null(metabo_dt)] -->
<!--         metabo_dt <- strsplit(metabo_dt, ";") -->
<!--         metabo_dt <- metabo_dt[!is.na(metabo_dt)] -->

<!--         wgb_dt <- omic_dict_list[["WGB_DNA_Methylation"]][paste0("WGB_DNA_Methylation",  -->
<!--                                                                 ".", weight_type_dt[[ "WGB_DNA_Methylation"]])] -->
<!--         wgb_dt <- wgb_dt[!is.null(wgb_dt) & !is.na(wgb_dt)] -->

<!--         rna_dt <- omic_dict_list[["RNA_Sequencing"]][paste0("RNA_Sequencing", -->
<!--                                                             ".", weight_type_dt[["RNA_Sequencing"]])] -->
<!--         rna_dt <- rna_dt[!is.null(rna_dt) & !is.na(rna_dt)] -->

<!--         list("Untargeted_Metabolomics" = metabo_dt,  -->
<!--             "RNA_Sequencing" = rna_dt,  -->
<!--             "WGB_DNA_Methylation" = wgb_dt) -->
<!--     }) -->
<!-- }) -->


<!-- # universe_list_ids <- unlist(Reduce(c, universe_list_ids)) -->
<!-- ``` -->

<!-- #### Full Joint -->

<!-- ```{r} -->
<!-- universe <- unlist(Reduce(c, universe_list_ids)) -->
<!-- factor_ids <- lapply(factor_ids_enrich, function(factor_dt){ -->
<!--     lapply(factor_dt, function(weight_type_dt){ -->
<!--         unlist(Reduce(c, weight_type_dt)) -->
<!--     }) -->
<!-- }) -->

<!-- library(clusterProfiler) -->
<!-- enrich_results_list_full <- lapply(factor_ids, function(factor_dt){ -->
<!--     lapply(factor_dt, function(weight_dt){ -->
<!--         enricher(gene = weight_dt, universe = universe,  -->
<!--                 pvalueCutoff = 0.05, pAdjustMethod = "fdr",  -->
<!--                 TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt) -->
<!--     }) -->
<!-- }) -->

<!-- results_wb <- openxlsx::createWorkbook() -->
<!-- lapply(names(factor_ids), function(factor_id){ -->
<!--     factor_dt <- enrich_results_list_full[[factor_id]] -->

<!--     lapply(names(factor_dt), function(weight_id){ -->

<!--         sheet_name <- paste(factor_id, ".", weight_id) -->
<!--         openxlsx::addWorksheet(wb = results_wb,  -->
<!--                                 sheetName = sheet_name) -->
<!--         openxlsx::writeData(wb = results_wb, sheet = sheet_name,  -->
<!--                             x = as.data.frame(factor_dt[[weight_id]])) -->

<!--     }) -->
<!--     openxlsx::saveWorkbook(results_wb,  -->
<!--                             file = paste0(output_dir, "enrich_results.xlsx"),  -->
<!--                             overwrite = TRUE) -->
<!-- }) -->

<!-- ``` -->

<!-- ```{r, fig.height=6, fig.width=8} -->
<!-- library(ggplot2) -->
<!-- pdf(paste0(output_dir, "Kegg_Joint_Pathway_Enrichment.pdf")) -->
<!-- lapply(names(enrich_results_list_full), function(factor_id){ -->
<!--     factor_dt <- enrich_results_list_full[[factor_id]] -->
<!--     lapply(names(factor_dt), function(weight_id){ -->
<!--         weight_dt <- factor_dt[[weight_id]] -->

<!--         if(nrow(as.data.frame(weight_dt)) == 0){return(NULL)} -->
<!--         plot_name <- gsub("_", " ", paste(factor_id, weight_id)) -->

<!--         dotplot(weight_dt) + ggtitle(paste("Kegg joint pathway enrichment" , plot_name)) -->
<!--     }) -->
<!-- }) -->
<!-- dev.off() -->
<!-- ``` -->

<!-- #### Partial Joint -->

<!-- ```{r} -->
<!-- weight_table <- prop.table(table(integ_enrich_db$term, integ_enrich_db$type), margin = 1) -->

<!-- universe_m <- unlist(Reduce(c, universe_list_ids["Untargeted_Metabolomics"])) -->
<!-- universe_g <- unlist(Reduce(c, universe_list_ids[c("RNA_Sequencing", "WGB_DNA_Methylation")])) -->

<!-- enrich_results_partial <- lapply(names(factor_ids_enrich), function(factor_id){ -->
<!--     factor_dt <- factor_ids_enrich[[factor_id]] -->
<!--     lapply(names(factor_dt), function(weight_id){ -->
<!--         weight_dt <- factor_dt[[weight_id]] -->

<!--         ids_m <- unlist(Reduce(c, weight_dt["Untargeted_Metabolomics"])) -->
<!--         ids_g <- unlist(Reduce(c, weight_dt[c("RNA_Sequencing",  -->
<!--                                             "WGB_DNA_Methylation")])) -->

<!--         enrich_m <- enricher(gene = ids_m, universe = universe_m,  -->
<!--                             pvalueCutoff = 0.05, pAdjustMethod = "fdr",  -->
<!--                             TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt) -->
<!--         enrich_g <- enricher(gene = ids_g, universe = universe_g,  -->
<!--                             pvalueCutoff = 0.05, pAdjustMethod = "fdr",  -->
<!--                             TERM2GENE = integ_enrich_db, TERM2NAME = term2name_dt) -->



<!--         merged_enrich <- dplyr::full_join(x = enrich_g@result, y = enrich_m@result, by = "ID") -->

<!--         combined_pvals <- sapply(merged_enrich$ID, function(id){ -->
<!--             p_vals <- merged_enrich[merged_enrich$ID == id,  -->
<!--                                     c("pvalue.x", "pvalue.y")] -->
<!--             if(any(is.na(p_vals))){ -->
<!--                 return(p_vals[!is.na(p_vals)]) -->
<!--             } -->
<!--             survcomp::combine.test(p = p_vals,  -->
<!--                                    weight = weight_table[id,], method = "logit")     -->
<!--         }) -->
<!--         merged_enrich$combined_adj.p.value <- p.adjust(combined_pvals,  -->
<!--                                                     method = "fdr") -->
<!--         merged_enrich[merged_enrich$combined_adj.p.value <= 0.05,] -->
<!--     }) -->
<!-- }) -->
<!-- names(enrich_results_partial) <- names(factor_ids_enrich) -->

<!-- results_wb <- openxlsx::createWorkbook() -->
<!-- lapply(names(factor_ids), function(factor_id){ -->
<!--     factor_dt <- enrich_results_partial[[factor_id]] -->

<!--     lapply(names(factor_dt), function(weight_id){ -->

<!--         sheet_name <- paste(factor_id, ".", weight_id) -->
<!--         openxlsx::addWorksheet(wb = results_wb,  -->
<!--                                 sheetName = sheet_name) -->
<!--         openxlsx::writeData(wb = results_wb, sheet = sheet_name,  -->
<!--                             x = as.data.frame(factor_dt[[weight_id]])) -->

<!--     }) -->
<!--     openxlsx::saveWorkbook(results_wb,  -->
<!--                             file = paste0(output_dir, "enrich_results.xlsx"),  -->
<!--                             overwrite = TRUE) -->
<!-- }) -->

<!-- ``` -->

<!-- ```{r, fig.height=6, fig.width=8} -->
<!-- library(ggplot2) -->
<!-- pdf(paste0(output_dir, "Kegg_Joint_Pathway_Enrichment.pdf")) -->
<!-- lapply(names(enrich_results_partial), function(factor_id){ -->
<!--     factor_dt <- enrich_results_partial[[factor_id]] -->
<!--     lapply(names(factor_dt), function(weight_id){ -->
<!--         weight_dt <- factor_dt[[weight_id]] -->

<!--         if(nrow(as.data.frame(weight_dt)) == 0){return(NULL)} -->
<!--         plot_name <- gsub("_", " ", paste(factor_id, weight_id)) -->

<!--         dotplot(weight_dt) + ggtitle(paste("Kegg joint pathway enrichment" , plot_name)) -->
<!--     }) -->
<!-- }) -->
<!-- dev.off() -->
<!-- ``` -->



## Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```