---
title: "Individual Omics Analysis"
output: html_document
date: "2023-05-12"
params:
    p.value: 0.05
    input_dir: "../../outputs/Multi_Omics_Analysis/"
    output_dir: "../../outputs/Multi_Omics_Analysis/"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required Packages

```{r}
library(MultiAssayExperiment)
library(xcms)
library(TreeSummarizedExperiment)
library(DESeq2)
library(SummarizedExperiment)
library(dplyr)
library(vsn)
library(compositions)
library(ggplot2)
library(plotly)
library(impute)
library(ggpubr)
library(RColorBrewer)
library(reshape2)
source("../Utility_Functions/utilities.R")
```

## Analysis functions

```{r}
data_boxplots <- function(omic_dt, omic_name, output_dir, 
                        ylab = "Intensity", xlab_var = "group", 
                        main_title = "rownames", ...){
    var_dt <- assay(omic_dt)
    if (main_title == "rownames" || !(main_title %in% colnames(rowData(omic_dt)))){
        main_title <- rownames(var_dt)
    } else {
        main_title <- make.names(rowData(omic_dt)[[main_title]])
    }
    pdf(paste0(output_dir, omic_name, "_boxplots.pdf"))
    for(dt_var_id in seq_len(nrow(var_dt))){
        
        plot_dt <- cbind(colData(omic_dt)[xlab_var], t(var_dt[dt_var_id,]))
            
        print(ggboxplot(data = as.data.frame(plot_dt), x = xlab_var, 
                    y = make.names(rownames(var_dt)[dt_var_id]), ylab = ylab,
                    title = main_title[dt_var_id], 
                    color = xlab_var, add = "jitter", palette = brewer.pal(n = length(unique(plot_dt[[xlab_var]])), name = "Dark2")) + theme(legend.position = "none", plot.title = element_text(hjust = 0.5)))
    }
    dev.off()
}

data_pca <- function(dt_list, metadata, stratificationVar, pc_x = 1, pc_y = 2, 
                     scale. = FALSE, center = FALSE, plot_name, color_vect, ...){
    group_var <- metadata[[stratificationVar]]
    if(any(is.na(dt_list))){
        dt_list <- t(impute::impute.knn(as.matrix(dt_list))$data)    
    } else{
        dt_list <- t(dt_list)
    }
    
    pca_dt <- prcomp(dt_list, center = center, scale. = scale.)
    
    pca_plot(pca_dt, pc_x = pc_x, pc_y = pc_y, group_var = group_var, 
            ellipse_var = group_var, interactive = FALSE, color_vect = color_vect) + 
        ggtitle(plot_name)
}
```


## Introduction
    
## Objectives

- Study the life-style intervention model (lima) to find:
    - What are the effects of the high fat diet (HFD) and the life-style intervention
        - Which components/organs are most affected by the high fat diet (HFD)
        - Which components/organs are restored by the posterior life-style intervention (INT)
        - Which components/organs are resistant to the life-style intervention (INT)
        - Is the microbiota diversity reduced with the high fat diet (HFD)?  Is it restored with the life-style intervention (INT)?
    - Why is the high fat diet excluded (HFE) group resistant to the high fat diet (HFD)?
        - Which components/organs are affected differently in the high fat diet excluded (HFE) group and the high fat diet (HFD) group?
        - Is it more similar to the control group (CNT) than the life-style intervention group (INT)?
        - Do they have a specific composition of the microbiota?
- Integrate the multi omics data from the lima model to find how the microbiota metabolites affects mice through epigenetic modifications.
    - Which epigenetically relevant metabolites correlate/are knownly produced/induced by the microbiota
    - Which epigenetic modifications related to the above mentioned metabolites are found on the data
    - Which genes related to these epigenetic modications are actually affected

Necessary contrasts:
    - Sig. HFD-CTR (1)
    - Sig. HFD-CTR / No Sig. INT-CTR (2)
    - (1) - (2)
    
    
## Loading the data

```{r}
output_dir <- paste0(params$output_dir, "Results/")
if(!dir.exists(output_dir)){dir.create(output_dir)}

multi_omics_raw <- readRDS(paste0(params$input_dir, "Multi_Omics_Analysis_exp_lima.rds"))
multi_omics_ma <- readRDS(paste0(params$input_dir, "Multi_Omics_Analysis_MA_exp_lima.rds"))
data_list <- list("Raw_Results" = multi_omics_raw[,colData(multi_omics_raw)$samplename[colData(multi_omics_raw)$group != "HFE"]], 
                  "Multivariate_Normalized" = multi_omics_ma[,])

multi_omic_any_sig <- lapply(data_list, function(multi_omics_dt){
    keep_rows_list <- lapply(experiments(multi_omics_dt), function(omic_dt){
        omic_rowdata <- rowData(omic_dt)
        p_val_dt <- as.matrix(omic_rowdata[,grep("adj.p.value", colnames(omic_rowdata))])
        p_val_dt <- p_val_dt[,grepl("HFD-CNT|INT-HFD|INT-CNT", colnames(p_val_dt))]
        keep_vars <- rowSums(p_val_dt <= 0.05, na.rm = TRUE) != 0
        
        omic_mat <- as.matrix(assay(omic_dt)) 
        omic_mat <- (omic_mat != 0 & !is.na(omic_mat))

        rownames(omic_rowdata)[keep_vars]
    })
    multi_omics_dt[keep_rows_list,,]
})

data_list <- multi_omic_any_sig
names(data_list) <- paste0(names(data_list), "Any Sig")
```

```{r}
group_color_dict <- RColorBrewer::brewer.pal(n = length(levels(droplevels(multi_omic_any_sig$Multivariate_Normalized$group))), 
                                                 name = "Set2")
names(group_color_dict) <- levels(droplevels(multi_omic_any_sig$Multivariate_Normalized$group))
group_color_dict
```


## PCA and Heatmap Outputs

```{r}
all_omics_functions <- list("PCA_Results" = data_pca, 
                            "Heatmap_Results" = heatMapFun)

for(data_mae_name in names(data_list)){
    experiment_list <- experiments(data_list[[data_mae_name]])
    
    plot_output_dir <- paste0(output_dir, data_mae_name, "/")
    if(!dir.exists(plot_output_dir)){dir.create(plot_output_dir)}
    
    for(fun_name in names(all_omics_functions)){
        param_list <- list()
        
        pdf(paste0(plot_output_dir, fun_name, ".pdf"))
        for(experiment_name in names(experiment_list)){
            param_list$plot_name <- experiment_name
            param_list$dt_list <- assay(experiment_list[[experiment_name]])
            
            if(nrow(param_list$dt_list) < 2){next}
            
            param_list$metadata <- colData(experiment_list[[experiment_name]])
            param_list$stratificationVar <- "group"
            param_list$colo
            
            if(fun_name == "Heatmap_Results"){
                param_list$cdQuantile <- ifelse(nrow(param_list$dt_list) >= 4000, 0, 0)        
                param_list$show_colnames <- ifelse(nrow(param_list$dt_list) > 100, FALSE, TRUE)
                
                param_list$color_list <- list("group" = group_color_dict)
                
            }
            if(fun_name == "PCA_Results"){
                param_list$color_vect <- group_color_dict
            }
            if(experiment_name %in% c("DNA_Methylation_unsupervised.liver", "DNA_Methylation.liver")){
                param_list$cdQuantile <- 0
                set.seed(123)
                param_list$dt_list <- param_list$dt_list[sample(x = seq_len(nrow(param_list$dt_list)), size = 1000),]
            }
            if(data_mae_name == "Raw_Results"){
                param_list$clustering_distance_rows <- dist(scale(t(param_list$dt_list)))
            } else{
                param_list$clustering_distance_rows <-  dist(t(param_list$dt_list))
            }
            
            
            print(do.call(all_omics_functions[[fun_name]], param_list))
        }
        dev.off()
    }    
}

rm(param_list)
```


```{r}
rna_seq_test <- multi_omic_any_sig$Raw_Results@ExperimentList$RNA_Sequencing

plot_dt <- scale(t(as.matrix(assay(rna_seq_test))))

rna_seq_test_n <- scale_assay(count_vst_deseq2(rna_seq_test))

pheatmap::pheatmap(t(as.matrix(assay(rna_seq_test_n))), show_colnames = FALSE, 
                   clustering_method = "ward.D2")

pheatmap::pheatmap(t(as.matrix(assay(rna_seq_test_n))), show_colnames = FALSE, 
                   clustering_method = "median")

pheatmap::pheatmap(t(as.matrix(assay(rna_seq_test_n))), show_colnames = FALSE, 
                   clustering_method = "average")

pheatmap::pheatmap(t(as.matrix(assay(rna_seq_test_n))), show_colnames = FALSE, 
                   clustering_method = "centroid")

```


## Other plots

### DNA Methylation density

```{r}
wgb_dt <- multi_omic_any_sig$Multivariate_Normalized[["WGB_DNA_Methylation"]]

plot_dt <- assay(wgb_dt)

plot_dt <- t(apply(plot_dt, 1, function(x){tapply(x, INDEX = droplevels(wgb_dt$group), FUN = median)}))
plot_dt <- reshape2::melt(plot_dt)
# plot_dt$group <- colData(wgb_dt)[plot_dt$variable, "group"]

library(ggplot2)
```

```{r, fig.height=6, fig.width=4}
pdf(paste0(output_dir, "median_zscore_dmr_group_methylation.pdf"))
ggplot(plot_dt) + geom_histogram(aes(x = value, fill = Var2)) + 
    theme_minimal() + xlab("Median DMR group methylation (Z score)") + ylab(NULL) + 
    theme(legend.position = "none") + facet_grid(vars(Var2)) + 
    scale_fill_manual(values = group_color_dict)
dev.off()
```

```{r}
wgb_dt <- multi_omic_any_sig$Multivariate_Normalized[["WGB_DNA_Methylation"]]

plot_dt <- assay(wgb_dt)

plot_dt <- t(apply(plot_dt, 1, function(x){tapply(x, INDEX = droplevels(wgb_dt$group), FUN = mean)}))
plot_dt <- reshape2::melt(plot_dt)
# plot_dt$group <- colData(wgb_dt)[plot_dt$variable, "group"]
```

```{r, fig.height=6, fig.width=4}
pdf(paste0(output_dir, "mean_zscore_dmr_group_methylation.pdf"))
ggplot(plot_dt) + geom_histogram(aes(x = value, fill = Var2)) + 
    theme_minimal() + xlab("Mean DMR group methylation (Z score)") + ylab(NULL) + 
    theme(legend.position = "none") + facet_grid(vars(Var2)) + 
    scale_fill_manual(values = group_color_dict)
dev.off()
```

### RNA Sequencing density

```{r}
rna_dt <- multi_omic_any_sig$Multivariate_Normalized[["RNA_Sequencing"]]

plot_dt <- assay(rna_dt)

plot_dt <- t(apply(plot_dt, 1, function(x){tapply(x, INDEX = droplevels(rna_dt$group), FUN = median)}))
plot_dt <- reshape2::melt(plot_dt)
# plot_dt$group <- colData(wgb_dt)[plot_dt$variable, "group"]

library(ggplot2)
```

```{r, fig.height=6, fig.width=4}
pdf(paste0(output_dir, "median_zscore_group_rna_seq.pdf"))
ggplot(plot_dt) + geom_histogram(aes(x = value, fill = Var2)) + 
    theme_minimal() + xlab("Median RNA seq group expression (Z score)") + ylab(NULL) + 
    theme(legend.position = "none") + facet_grid(vars(Var2)) + 
    scale_fill_manual(values = group_color_dict)
dev.off()
```

```{r}
rna_dt <- multi_omic_any_sig$Multivariate_Normalized[["RNA_Sequencing"]]

plot_dt <- assay(rna_dt)

plot_dt <- t(apply(plot_dt, 1, function(x){tapply(x, INDEX = droplevels(rna_dt$group), FUN = mean)}))
plot_dt <- reshape2::melt(plot_dt)
# plot_dt$group <- colData(wgb_dt)[plot_dt$variable, "group"]
```

```{r, fig.height=6, fig.width=4}
pdf(paste0(output_dir, "mean_zscore_group_rna_seq.pdf"))
ggplot(plot_dt) + geom_histogram(aes(x = value, fill = Var2)) + 
    theme_minimal() + xlab("Mean RNA seq group expression (Z score)") + ylab(NULL) + 
    theme(legend.position = "none") + facet_grid(vars(Var2)) + 
    scale_fill_manual(values = group_color_dict)
dev.off()
```

### Untargeted Metabolomics density

```{r}
met_dt <- multi_omic_any_sig$Multivariate_Normalized[["Untargeted_Metabolomics"]]

plot_dt <- assay(met_dt)

plot_dt <- t(apply(plot_dt, 1, function(x){tapply(x, INDEX = droplevels(met_dt$group), FUN = median)}))
plot_dt <- reshape2::melt(plot_dt)
# plot_dt$group <- colData(wgb_dt)[plot_dt$variable, "group"]

library(ggplot2)
```

```{r, fig.height=6, fig.width=4}
pdf(paste0(output_dir, "median_zscore_dmr_group_metabolites.pdf"))
ggplot(plot_dt) + geom_histogram(aes(x = value, fill = Var2)) + 
    theme_minimal() + xlab("Median DMR group metabolites (Z score)") + ylab(NULL) + 
    theme(legend.position = "none") + facet_grid(vars(Var2)) + 
    scale_fill_manual(values = group_color_dict)
dev.off()
```

```{r}
met_dt <- multi_omic_any_sig$Multivariate_Normalized[["Untargeted_Metabolomics"]]

plot_dt <- assay(met_dt)

plot_dt <- t(apply(plot_dt, 1, function(x){tapply(x, INDEX = droplevels(met_dt$group), FUN = mean)}))
plot_dt <- reshape2::melt(plot_dt)
# plot_dt$group <- colData(wgb_dt)[plot_dt$variable, "group"]
```

```{r, fig.height=6, fig.width=4}
pdf(paste0(output_dir, "mean_zscore_dmr_group_metabolites.pdf"))
ggplot(plot_dt) + geom_histogram(aes(x = value, fill = Var2)) + 
    theme_minimal() + xlab("Mean DMR group metabolites (Z score)") + ylab(NULL) + 
    theme(legend.position = "none") + facet_grid(vars(Var2)) + 
    scale_fill_manual(values = group_color_dict)
dev.off()
```



## Pre-Analysis

```{r}
mice_dt <- openxlsx::read.xlsx("../../inputs/AnimalData_OscarYanes.xlsx")
# mice_num <- mice_dt[,c("BW(g)", "WATep(g)", "Liver(g)")]
mice_num <- mice_dt
mice_s <- as.data.frame(mice_num)
dict_vect <- c("Ctrl" = "CNT", "Ctrl ext" = "CNT", "HFD" = "HFD", 
                "INT2" = "INT", "HFD excl." = "HFE")
mice_s$group <- factor(dict_vect[mice_dt$Group], levels = c("CNT", "HFD", "INT", "HFE"))
dict_vect_g <- c("CNT" = "C", "HFD" = "H", "INT" = "I", "HFE" = "E")
mice_s$samplename <- paste0(make.names(mice_dt$ID), dict_vect_g[mice_s$group])

mice_s$samplename[!mice_s$samplename %in% multi_omic_any_sig$Multivariate_Normalized$samplename]

multi_omic_any_sig$samplename[!multi_omic_any_sig$Multivariate_Normalized$samplename %in% mice_s$samplename]

mice_s$samplename[mice_s$samplename == "X686I"] <- "X687I"

mice_s <- dplyr::left_join(as.data.frame(colData(multi_omic_any_sig$Multivariate_Normalized)), mice_s, by = "samplename")
mice_s

# write.table(mice_s, "../../outputs/Multi_Omics_Analysis/mouse_info.csv")

rna_seq_dt <- assay(multi_omic_any_sig$Multivariate_Normalized@ExperimentList$RNA_Sequencing)
test_cor <- cor(t(rna_seq_dt), mice_s[,c("WATep(g)", "Liver(g)", "BW(g)")])
```

## Component List

```{r}
# expression_list <- list(
#     "HFD Altered" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05
#     ),
#     "INT Restored Up" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` > 0 & 
#         `INT-HFD.adj.p.value` <= 0.05 & 
#         `INT-CNT.adj.p.value` > 0.05
#     ), 
#     "INT Restored Down" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
#         `INT-HFD.adj.p.value` <= 0.05 & 
#         `INT-CNT.adj.p.value` > 0.05
#     ),
#     "INT Restored Over Down" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 &
#         `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` > 0 &
#         `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` > 0
#     ),
#     "INT Restored Over Up" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 &
#         `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0 &
#         `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` < 0
#     ),
#     "INT Resistant Over Up" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 & 
#         `INT-CNT.adj.p.value` <= 0.05 &
#         `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` >= 0
#     ),
#     "INT Resistant Over Down" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
#         `INT-CNT.adj.p.value` <= 0.05 &
#         `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0
#     ),
#     "INT Resistant Up" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 & 
#         `INT-CNT.adj.p.value` <= 0.05 & 
#         `INT-HFD.adj.p.value` > 0.05 
#     ),
#     "INT Resistant Down" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
#         `INT-CNT.adj.p.value` <= 0.05 &
#         `INT-HFD.adj.p.value` > 0.05
#     ),
#     "INT Resistant Half Up" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 & 
#         ((`INT-HFD.adj.p.value` <= 0.05 & `INT-CNT.adj.p.value` <= 0.05) |
#         (`INT-HFD.adj.p.value` > 0.05 & `INT-CNT.adj.p.value` > 0.05)) &
#         `INT-HFD.log2FC` < 0 & 
#         `INT-CNT.log2FC` >= 0
#     ),
#     "INT Resistant Half Down" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 &
#         ((`INT-HFD.adj.p.value` <= 0.05 & `INT-CNT.adj.p.value` <= 0.05) |
#         (`INT-HFD.adj.p.value` > 0.05 & `INT-CNT.adj.p.value` > 0.05)) & 
#         `INT-HFD.log2FC` >= 0 &
#         `INT-CNT.log2FC` < 0
#     ), "INT Resistant Weird" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 &
#         `INT-HFD.adj.p.value` > 0.05 & 
#         `INT-CNT.adj.p.value` > 0.05 & 
#         ((`INT-HFD.log2FC` >= 0 & `INT-CNT.log2FC` >= 0) | 
#             (`INT-HFD.log2FC` < 0 & `INT-CNT.log2FC` < 0))
#     ),
#     "HFE Unaffected" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` > 0.05 & `HFE-HFD.adj.p.value` <= 0.05
#     ),
#     "HFE Affected equal HFD" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` <= 0.05 & `HFE-HFD.adj.p.value` > 0.05
#     ),
#     "HFE Affected diff HFD" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` <= 0.05 & `HFE-HFD.adj.p.value` <= 0.05
#     ),
#     "HFE Other" = quote(
#         `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` > 0.05 & `HFE-HFD.adj.p.value` > 0.05
#     )
# )

expression_list <- list(
    "HFD Altered" = quote(
        `HFD-CNT.adj.p.value` <= 0.05
    ),
    "INT Restored Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` > 0 & 
        ((`INT-HFD.adj.p.value` <= 0.05 & 
        `INT-CNT.adj.p.value` > 0.05) | 
            ((`INT-HFD.adj.p.value` > 0.05 & 
            `INT-CNT.adj.p.value` > 0.05) & 
                ((`INT-HFD.adj.p.value` < `INT-CNT.adj.p.value`) | 
                     ((`INT-HFD.adj.p.value` == `INT-CNT.adj.p.value`) & (abs(`INT-HFD.log2FC`) - abs(`INT-CNT.log2FC`)) > 0))))
    ), 
    "INT Restored Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
        ((`INT-HFD.adj.p.value` <= 0.05 & 
        `INT-CNT.adj.p.value` > 0.05) | 
            ((`INT-HFD.adj.p.value` > 0.05 & 
            `INT-CNT.adj.p.value` > 0.05) & 
                ((`INT-HFD.adj.p.value` < `INT-CNT.adj.p.value`) | 
                     ((`INT-HFD.adj.p.value` == `INT-CNT.adj.p.value`) & (abs(`INT-HFD.log2FC`) - abs(`INT-CNT.log2FC`)) > 0))))
    ),
    "INT Restored Over Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` > 0 &
        `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` > 0
    ),
    "INT Restored Over Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0 &
        `INT-CNT.adj.p.value` <= 0.05 & `INT-CNT.log2FC` < 0
    ),
    "INT Resistant Over Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 & 
        `INT-CNT.adj.p.value` <= 0.05 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` >= 0
    ),
    "INT Resistant Over Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
        `INT-CNT.adj.p.value` <= 0.05 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-HFD.log2FC` < 0
    ),
    "INT Resistant Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 & 
        ((`INT-CNT.adj.p.value` <= 0.05 & 
        `INT-HFD.adj.p.value` > 0.05) | 
            ((`INT-HFD.adj.p.value` > 0.05 & 
            `INT-CNT.adj.p.value` > 0.05) & 
                ((`INT-HFD.adj.p.value` > `INT-CNT.adj.p.value`) | 
                     ((`INT-HFD.adj.p.value` == `INT-CNT.adj.p.value`) & (abs(`INT-HFD.log2FC`) - abs(`INT-CNT.log2FC`)) < 0))))
    ),
    "INT Resistant Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 & 
        ((`INT-CNT.adj.p.value` <= 0.05 & 
        `INT-HFD.adj.p.value` > 0.05) | 
            ((`INT-HFD.adj.p.value` > 0.05 & 
            `INT-CNT.adj.p.value` > 0.05) & 
                ((`INT-HFD.adj.p.value` > `INT-CNT.adj.p.value`) | 
                     ((`INT-HFD.adj.p.value` == `INT-CNT.adj.p.value`) & (abs(`INT-HFD.log2FC`) - abs(`INT-CNT.log2FC`)) < 0))))
    ),
    "INT Resistant Half Up" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` >= 0 & 
        `INT-HFD.adj.p.value` <= 0.05 & `INT-CNT.adj.p.value` <= 0.05 &
        `INT-HFD.log2FC` < 0 & 
        `INT-CNT.log2FC` >= 0
    ),
    "INT Resistant Half Down" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFD-CNT.log2FC` < 0 &
        `INT-HFD.adj.p.value` <= 0.05 & `INT-CNT.adj.p.value` <= 0.05 & 
        `INT-HFD.log2FC` >= 0 &
        `INT-CNT.log2FC` < 0
    ),
    "HFE Unaffected" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` > 0.05 & `HFE-HFD.adj.p.value` <= 0.05
    ),
    "HFE Affected equal HFD" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` <= 0.05 & `HFE-HFD.adj.p.value` > 0.05
    ),
    "HFE Affected diff HFD" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` <= 0.05 & `HFE-HFD.adj.p.value` <= 0.05
    ),
    "HFE Other" = quote(
        `HFD-CNT.adj.p.value` <= 0.05 & `HFE-CNT.adj.p.value` > 0.05 & `HFE-HFD.adj.p.value` > 0.05
    )
)

multi_omic_filtered_list <- lapply(data_list, function(multi_omics_dt){
    lapply(expression_list, function(ctrs_exprs){
        
        keep_rows_list <- lapply(experiments(multi_omics_dt), function(omic_dt){
            omic_rowdata <- rowData(omic_dt)
            rownames(omic_rowdata)[which(eval(ctrs_exprs, envir = omic_rowdata))]
        })
        
        multi_omics_dt[keep_rows_list,,]
    })
})
```

## Restored/Resistant Table

```{r}
ctrs_sig <- lapply(multi_omic_filtered_list[[2]], function(omic_ctrs){
    lengths(rownames(omic_ctrs))
})

ctrs_sig <- do.call(cbind, ctrs_sig)

restored_cols <- grep("Restored", colnames(ctrs_sig)) 
hfd_affected_cols <- grep("Altered", colnames(ctrs_sig))
resistant_cols <- grep("Resistant", colnames(ctrs_sig))

affected_col <- grep("Affected", colnames(ctrs_sig))
unaffected_col <- grep("Unaffected", colnames(ctrs_sig))
other_col <- grep("Other", colnames(ctrs_sig))


dt_table <- t(apply(ctrs_sig, 1, function(omic_var){
   c("% Restored" = sum(omic_var[restored_cols])/omic_var[hfd_affected_cols], 
     "% Resistant" = sum(omic_var[resistant_cols])/omic_var[hfd_affected_cols], 
     "Total Altered" = omic_var["HFD Altered"])
}))


hfe_table <- t(apply(ctrs_sig, 1, function(omic_var){
   c("% HFE Affected" = sum(omic_var[affected_col])/omic_var[hfd_affected_cols], 
     "% HFE Unaffected" = sum(omic_var[unaffected_col])/omic_var[hfd_affected_cols],
     "% HFE Other" = sum(omic_var[other_col])/omic_var[hfd_affected_cols],
     "Total Altered" = omic_var["HFD Altered"])
}))

```

## Boxplots

```{r}
lapply(names(multi_omic_filtered_list), function(multi_omic_filtered_name){
    multi_omic_filtered <- multi_omic_filtered_list[[multi_omic_filtered_name]]

    lapply(names(multi_omic_filtered)[2:15], function(sig_ctrs_name){
        sig_ctrs <- multi_omic_filtered[[sig_ctrs_name]]
        
        ctrs_dir <- paste0("./Results/", multi_omic_filtered_name, "/",
                        make.names(sig_ctrs_name), "/")
        
        if(!dir.exists(ctrs_dir)){dir.create(ctrs_dir)}
        
        for(omic_name in names(sig_ctrs)[1]){
            if(grepl("RNASeq", omic_name)){
                main_title <- "gene_lab" 
                ylab <- ifelse(multi_omic_filtered_name == "Raw_Results", 
                               "Counts", "VST Scaled Counts")
            } else if(grepl("16S", omic_name)){
                main_title <- "specie_name_short"
                ylab <- ifelse(multi_omic_filtered_name == "Raw_Results", 
                               "% of abundance", "CLR Transformed abundance")
                
            } else if(grepl("Shotgun", omic_name)){
                main_title <- "specie_name_short"
                ylab <- ifelse(multi_omic_filtered_name == "Raw_Results", 
                               "Counts", "CLR Transformed counts")
                
            } else if(grepl("PTM|DNA", omic_name)){
                ylab <- ifelse(multi_omic_filtered_name == "Raw_Results", 
                               "B Value", "M Value")
                main_title <- "rownames"
            } else if(grepl("Untargeted", omic_name)){
                ylab <- ifelse(multi_omic_filtered_name == "Raw_Results", 
                               "Intensity", "VST Transformed intensity")
                main_title <- "rownames"
            } else{
                main_title <- "rownames"
                ylab <- ifelse(multi_omic_filtered_name == "Raw_Results", 
                               "Intensity", "Scaled Intensity")
            }
            if(nrow(experiments(sig_ctrs)[[omic_name]]) == 0){next}
            data_boxplots(omic_dt = experiments(sig_ctrs)[[omic_name]], 
                          omic_name = omic_name, 
                          output_dir = ctrs_dir, ylab = ylab, 
                          main_title = main_title)   
        }
    })
})
```

## Data analysis

```{r, results='asis'}
for(data_list_names in names(complete_data_list)){
    cat("### ", data_list_names, "{.tabset} \n")
    for(fun_name in names(function_list)){
        cat("#### ", fun_name, "{.tabset} \n")
        fun_param <- list(dt_list = complete_data_list[[data_list_names]],
                          metadata = complete_metadata_list[[data_list_names]],
                          stratificationVar = "Recurrencia")
        if(fun_name == "PCA Analysis"){
            fun_param$stratificationVar <- c("Recurrencia",
                                             "Grado Histológico tumoral (Low o High Grade)")
        }
        if(fun_name == "Image Complications"){
            fun_param$metadata <- complete_hist_metadata_list[[data_list_names]]
        }
        if(fun_name == "Univariate Results" | fun_name == "pls-DA Caret"){
            fun_param$file_name <- data_list_names
        }
        cat("\n")
        do.call(function_list[[fun_name]], fun_param)
        cat("\n")
    }
    cat("\n")
}
```

