---
title: "Multi_Omics_Analysis"
output: html_document
date: "2023-05-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.duplicate.label = "allow")
```

## Required Packages

```{r}
library(MOFA2)
library(dplyr)
library(knitr)
library(ggplot2)
library(MultiAssayExperiment)
library(grDevices)
library(rGREAT)
library(msigdbr)
library(DT)
library(TxDb.Mmusculus.UCSC.mm39.refGene)
source("../../Utility Functions/utilities.R")
library(TreeSummarizedExperiment)
```

## Loaded matrices

```{r}
multi_omics_raw <- readRDS("./Merged_Multi_Omics.rds")
multi_omics <- readRDS("./Multivariate_Merged_Multi_Omics.rds")
names(experiments(multi_omics))

filter_list <- c("16S", "Shotgun.cecum", "Shotgun.colon")

# selected_omics <- lapply(filter_list, function(dt_name){
#     omic_dt_names <- grep(dt_name, names(experiments(multi_omics)))
#     omic_exp_list <- experiments(multi_omics[,,omic_dt_names])
#     
#     sig_vars_num <- sapply(omic_exp_list, function(omic_exp){
#         var_dt <- rowData(omic_exp)
#         p_val_dt <- var_dt[,grep("adj.p.value", colnames(var_dt))]
#         sig_vars <- apply(p_val_dt, 1, function(p_vals){
#             any(p_vals <= 0.05, na.rm = TRUE)
#         })
#         sum(unlist(sig_vars))
#     })
#     sig_vars_num
# })
# names(selected_omics) <- filter_list
# 
# a <- do.call(cbind, selected_omics)

# selected_omics <- lapply(filter_list, function(dt_name){
#     omic_dt_names <- grep(dt_name, names(experiments(multi_omics)))
#     omic_exp_list <- experiments(multi_omics[,,omic_dt_names])
#     
#     sig_vars_num <- sapply(omic_exp_list, function(omic_exp){
#         nrow(omic_exp)
#     })
#     sig_vars_num
# })
# names(selected_omics) <- filter_list
# 
# b <- do.call(cbind, selected_omics)
```

## Preparing the omics matrices 

```{r}
multi_omics_names <- names(experiments(multi_omics))

multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|unsupervised", multi_omics_names)]
# multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)])
multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)])
```

```{r}
filtered_exps <- lapply(names(multi_omics), function(multi_omics_names){
    multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]

    if(grepl("RNASeq|DNA|Untargeted", multi_omics_names)){
        if(grepl("RNASeq", multi_omics_names)){
            cd_thr <- 0.85
        } else if(grepl("unsupervised", multi_omics_names)){
            cd_thr <- 0.99
        } else if(grepl("DNA", multi_omics_names)){
            cd_thr <- 0.5
        } else{
            cd_thr <- 0
        }
        var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){
            sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE)
        })
        multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, cd_thr)],]
    }
    multi_omics_exp
})
names(filtered_exps) <- names(multi_omics)
experiments(multi_omics) <- ExperimentList(filtered_exps)
```

```{r}
multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics)
plot_data_overview(multi_omics_mofa)
```

X642C / X652C

```{r}
multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected])
plot_data_overview(multi_omics_mofa)
```

## Parameters

```{r}
data_opts <- get_default_data_options(multi_omics_mofa)
head(data_opts)

model_opts <- get_default_model_options(multi_omics_mofa)
head(model_opts)
model_opts$num_factors <- 8

train_opts <- get_default_training_options(multi_omics_mofa)
head(train_opts)
```

## LIMA model {.tabset}

```{r}
multi_omics_mofa <- prepare_mofa(
  object = multi_omics_mofa,
  data_options = data_opts,
  model_options = model_opts,
  training_options = train_opts
)

outfile <- file.path(getwd(),"model.hdf5")
multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE)

```

### Downstream Analysis

```{r}
# file_path <- "./model.hdf5"
# model <- load_model(file_path)
colData(multi_omics)$sample <- colData(multi_omics)$samplename

samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample")

plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90))

p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5, 
                 color_by = "group.y", 
                 add_violin = TRUE)

p <- p + 
  scale_color_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) +
  scale_fill_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green"))
print(p)
```

```{r, fig.height=12, fig.width=12}
plot_factors(multi_omics_mofa, 
  factors = 1:6,
  color_by = "group.y"
)
```

```{r}
pca_dt <- MOFA2::get_factors(object = multi_omics_mofa, factors = 1:3)

MOFA2::get_variance_explained(object = multi_omics_mofa, factors = 1:3)

library(plotly)

d3plot <- plot_ly(x = pca_dt$group1[,1], y = pca_dt$group1[,2], z = pca_dt$group1[,3], color = multi_omics_mofa@samples_metadata$group.y)
d3plot <- d3plot %>% add_markers()
# d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")),
#                      yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")),
#                      zaxis = list(title = paste0('PC4', "(", pcvar[4], "%)"))))
d3plot
```



```{r}
weight_dt <- get_weights(object = multi_omics_mofa, factors = 2, as.data.frame = TRUE)
```

```{r}
factor_ids <- c(1,2,3,4,5)
output_dir <- "D:/Jordi/Doctorat/GNOTOBIOMICS/Data_Analysis/Multi_Omics_Analysis/Results/MOFA2_Results/"

var_exp_dt <- MOFA2::get_variance_explained(multi_omics_mofa)[[2]][[1]]/100

keep_var_list <- list()

for(id in factor_ids){
    
    keep_var_list[id] <- list(c())
    
    weight_dt <- get_weights(multi_omics_mofa, factors = id)

    pdf(paste0(output_dir, "factor_", id, "_top_variable_heatmaps.pdf"))
    for(view_id in MOFA2::views_names(multi_omics_mofa)){
        exp_dt <- MOFA2::get_data(multi_omics_mofa, views = view_id)[[1]][[1]]
        exp_metadata <- MOFA2::samples_metadata(multi_omics_mofa)
        
        feature_n_percent <- var_exp_dt[id,view_id]
        feature_n <- ceiling(max(min(nrow(exp_dt)*feature_n_percent, 2000), 2))
        factor_weight_dt <- weight_dt[[view_id]]
        
        # hist(abs(factor_weight_dt[,1]))
        # sum(abs(factor_weight_dt[,1]) > 1)
        # 
        samples_found <- !matrixStats::colAlls(x = exp_dt, value = NA)
        rownames(exp_metadata) <- exp_metadata$samplename
        
        keep_vars <- rownames(factor_weight_dt)[order(abs(factor_weight_dt[,1]), decreasing = TRUE)][seq_len(feature_n)]
        
        keep_var_list[[id]] <- c(keep_var_list[[id]], gsub(paste0("_", view_id), "", keep_vars)) 
                                 
        heatMapFun(exp_dt[keep_vars,samples_found], metadata = exp_metadata[samples_found,], stratificationVar = "group.y", 
                cdQuantile = 0, 
                plot_name = paste(view_id, "Top", feature_n, "Variables in factor", id, "(By absolute weight)"), 
                show_colnames = ifelse(length(keep_vars) >= 100, FALSE, TRUE), show_rownames = TRUE,
                clustering_distance_rows = "euclidean")
        
    }
    dev.off()
}

```



### GREAT Enrichment & Region Analysis {.tabset}

```{r}
library(msigdbr)
txdb <- "TxDb.Mmusculus.UCSC.mm39.refGene"
h_gene_sets = msigdbr(species = "Mus musculus", category = "H")

h_gene_sets = split(h_gene_sets$entrez_gene, h_gene_sets$gs_name)
h_gene_sets = lapply(h_gene_sets, as.character)  # just to make sure gene IDs are all in character.
```

```{r}
sig_tables <- lapply(factor_ids, function(id){
    multi_omics[keep_var_list[[id]],,"DNA_Methylation.liver"][["DNA_Methylation.liver"]]
})
names(sig_tables) <- paste("Factor", seq_along(sig_tables))
    
set.seed(1234)
great_results <- lapply(sig_tables, function(ctrs_res){
    if(nrow(ctrs_res) == 0){return(NA)}
    sub_granges <- rowRanges(ctrs_res)
    mcols(sub_granges) <- mcols(sub_granges)[!colnames(mcols(sub_granges)) %in% 
                                               c("seqnames", "ranges", "strand", 
                                                "seqlevels", "seqlengths",
                                                "isCircular", "start", "end", 
                                                "width", "element")]  
    return(great(gr = sub_granges, gene_sets = h_gene_sets, 
                tss_source = txdb))
})

contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    return(getEnrichmentTable(gr_res))
})
contrs_names <- names(contrs_result_sig_list)
template <- c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

### Region Annotation

```{r}
omic_dt_list <- c("All Results" = multi_omics[["DNA_Methylation.liver"]], sig_tables[1:2])

dmr_annotation_plot(omic_dt = omic_dt_list)
# test_dt <- factor_tables[1:2,]
# fisher.test(factor_tables[1,])
```

### Chromosome Plots

```{r, fig.height=6, fig.width=8}
possible_chromosomes  <- unique(seqnames(rowRanges(multi_omics[["DNA_Methylation.liver"]])))
dmr_annotation_plot(omic_dt_list, annotation_labels = possible_chromosomes, 
                    annotation_metadata_var = "seqnames", x_lab = "Chromosome Proportions")
```


```{r}
# possible_annotations <- c("Intron", "Promoter", "Distal Intergenic", "3' UTR", "5' UTR", "Exon", "Downstream")
# 
# factor_tables <- lapply(c("All Results" = multi_omics[["DNA_Methylation.liver"]], sig_tables[1:3]), function(exp_dt){
#     exp_annotation <- sapply(possible_annotations, function(x){
#         found_anno <- grepl(pattern = x, rowData(exp_dt)$annotation)
#         sum(found_anno)
#     })
#     exp_annotation
#     # table(unlist(exp_annotation))
# })
# factor_tables <- do.call(rbind, factor_tables)
# factor_annotation_dt <-factor_tables/rowSums(factor_tables)
# 
# plot_dt <- reshape2::melt(factor_annotation_dt)
# plot_dt$Var1 <- factor(plot_dt$Var1, levels = rev(levels(plot_dt$Var1)))
# ggplot(plot_dt, aes(x = value, y = Var1, fill = Var2)) + geom_col() + scale_fill_brewer(palette = "Set2") + theme_minimal() + xlab("Region Proportions") + ylab(NULL) + theme(legend.title = element_blank())
```

<!-- ## LIMA full unsupervised -->

<!-- ```{r} -->
<!-- multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S|DNA_Methylation.liver", multi_omics_names)] -->
<!-- # multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)]) -->
<!-- multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)]) -->

<!-- multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected]) -->
<!-- plot_data_overview(multi_omics_mofa) -->
<!-- ``` -->

<!-- ## Parameters -->

<!-- ```{r} -->
<!-- data_opts <- get_default_data_options(multi_omics_mofa) -->
<!-- head(data_opts) -->

<!-- model_opts <- get_default_model_options(multi_omics_mofa) -->
<!-- head(model_opts) -->
<!-- model_opts$num_factors <- 8 -->

<!-- train_opts <- get_default_training_options(multi_omics_mofa) -->
<!-- head(train_opts) -->
<!-- ``` -->

<!-- ## Model and training -->

<!-- ```{r} -->
<!-- multi_omics_mofa <- prepare_mofa( -->
<!--   object = multi_omics_mofa, -->
<!--   data_options = data_opts, -->
<!--   model_options = model_opts, -->
<!--   training_options = train_opts -->
<!-- ) -->

<!-- outfile <- file.path(getwd(),"model_us.hdf5") -->
<!-- multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE) -->

<!-- ``` -->

<!-- ## Downstream Analysis -->

<!-- ```{r} -->
<!-- # file_path <- "./model.hdf5" -->
<!-- # model <- load_model(file_path) -->
<!-- colData(multi_omics)$sample <- colData(multi_omics)$samplename -->

<!-- samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample") -->

<!-- plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90)) -->

<!-- p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5,  -->
<!--                  color_by = "group.y",  -->
<!--                  add_violin = TRUE) -->

<!-- p <- p +  -->
<!--   scale_color_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) + -->
<!--   scale_fill_manual(values=c("CNT"="black", "HFD"="red", "INT"="blue", "HFE"="green")) -->
<!-- print(p) -->
<!-- ``` -->

<!-- ```{r, fig.height=12, fig.width=12} -->
<!-- plot_factors(multi_omics_mofa,  -->
<!--   factors = 1:6, -->
<!--   color_by = "group.y" -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pca_dt <- MOFA2::get_factors(object = multi_omics_mofa, factors = 1:3) -->

<!-- MOFA2::get_variance_explained(object = multi_omics_mofa, factors = 1:3) -->

<!-- library(plotly) -->

<!-- d3plot <- plot_ly(x = pca_dt$group1[,1], y = pca_dt$group1[,2], z = pca_dt$group1[,3], color = multi_omics_mofa@samples_metadata$group.y) -->
<!-- d3plot <- d3plot %>% add_markers() -->
<!-- # d3plot <- d3plot %>% layout(scene = list(xaxis = list(title = paste0('PC1', "(", pcvar[1], "%)")), -->
<!-- #                      yaxis = list(title = paste0('PC2', "(", pcvar[2], "%)")), -->
<!-- #                      zaxis = list(title = paste0('PC4', "(", pcvar[4], "%)")))) -->
<!-- d3plot -->
<!-- ``` -->

<!-- ```{r} -->
<!-- plot_weights(multi_omics_mofa,  -->
<!--   view = 1, -->
<!--   factor = 2, -->
<!--   nfeatures = 10,     # Number of features to highlight -->
<!--   scale = T,          # Scale weights from -1 to 1 -->
<!--   abs = F            # Take the absolute value? -->
<!-- ) -->

<!-- weight_dt <- get_weights(object = multi_omics_mofa, factors = 2, as.data.frame = TRUE) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(view_id in seq_len(length(multi_omics_selected))){ -->
<!--     plot_data_heatmap(multi_omics_mofa, -->
<!--       view = view_id,         # view of interest -->
<!--       factor = 2,             # factor of interest -->
<!--       features = 40,          # number of features to plot (they are selected by weight) -->

<!--       # extra arguments that are passed to the `pheatmap` function -->
<!--       cluster_rows = TRUE, cluster_cols = TRUE, -->
<!--       show_rownames = TRUE, show_colnames = FALSE, annotation_samples = "group.y", -->
<!--     ) -->
<!-- } -->
<!-- ``` -->


## GFCV

## Loaded matrices

```{r}
library(MOFA2)
multi_omics_raw <- readRDS("./Merged_Multi_Omics_gfcv.rds")
multi_omics <- readRDS("./Multivariate_Merged_Multi_Omics_gfcv.rds")
names(experiments(multi_omics))

filter_list <- c("16S")

selected_omics <- lapply(filter_list, function(dt_name){
    omic_dt_names <- grep(dt_name, names(experiments(multi_omics)))
    omic_exp_list <- experiments(multi_omics[,,omic_dt_names])
    
    sig_vars_num <- sapply(omic_exp_list, function(omic_exp){
        var_dt <- rowData(omic_exp)
        p_val_dt <- var_dt[,grep("adj.p.value", colnames(var_dt))]
        sig_vars <- apply(p_val_dt, 1, function(p_vals){
            any(p_vals <= 0.05, na.rm = TRUE)
        })
        sum(unlist(sig_vars))
    })
    sig_vars_num
})
names(selected_omics) <- filter_list

a <- do.call(cbind, selected_omics)

selected_omics <- lapply(filter_list, function(dt_name){
    omic_dt_names <- grep(dt_name, names(experiments(multi_omics)))
    omic_exp_list <- experiments(multi_omics[,,omic_dt_names])
    
    sig_vars_num <- sapply(omic_exp_list, function(omic_exp){
        nrow(omic_exp)
    })
    sig_vars_num
})
names(selected_omics) <- filter_list

b <- do.call(cbind, selected_omics)
```

## Preparing the omics matrices 

```{r}
multi_omics_names <- names(experiments(multi_omics))

multi_omics_selected <- multi_omics_names[!grepl("Shotgun|16S", multi_omics_names)]
# multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("genus", multi_omics_names)])
multi_omics_selected <- c(multi_omics_selected, multi_omics_names[grepl("specie", multi_omics_names)])
multi_omics_selected
```

```{r}
filtered_exps <- lapply(multi_omics_selected, function(multi_omics_names){
    multi_omics_exp <- experiments(multi_omics)[[multi_omics_names]]

    if(grepl("RNASeq", multi_omics_names)){
        var_cd <- apply(assay(multi_omics_exp), 1, function(var_dt){
            sd(var_dt, na.rm = TRUE)/mean(var_dt, na.rm = TRUE)
        })
        multi_omics_exp <- multi_omics_exp[names(var_cd)[var_cd >= quantile(var_cd, 0)],]
    }
    multi_omics_exp
})
names(filtered_exps) <- multi_omics_selected
experiments(multi_omics) <- ExperimentList(filtered_exps)
```
```{r}
multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics)
plot_data_overview(multi_omics_mofa)
```


```{r}
multi_omics_mofa <- create_mofa_from_MultiAssayExperiment(mae = multi_omics[,,multi_omics_selected])
plot_data_overview(multi_omics_mofa)
```

## Parameters

```{r}
data_opts <- get_default_data_options(multi_omics_mofa)
head(data_opts)

model_opts <- get_default_model_options(multi_omics_mofa)
head(model_opts)
model_opts$num_factors <- 8

train_opts <- get_default_training_options(multi_omics_mofa)
head(train_opts)
```

## Model and training

```{r}
multi_omics_mofa <- prepare_mofa(
  object = multi_omics_mofa,
  data_options = data_opts,
  model_options = model_opts,
  training_options = train_opts
)

outfile <- file.path(getwd(),"model.hdf5")
multi_omics_mofa <- run_mofa(multi_omics_mofa, outfile, use_basilisk = TRUE)

```

## Downstream Analysis

```{r}
# file_path <- "./model.hdf5"
# model <- load_model(file_path)
colData(multi_omics)$sample <- colData(multi_omics)$samplename

samples_metadata(multi_omics_mofa) <- left_join(samples_metadata(multi_omics_mofa), as.data.frame(colData(multi_omics)), by = "sample")

plot_variance_explained(multi_omics_mofa) + theme(axis.text.x = element_text(angle = 90))

p <- plot_factor(multi_omics_mofa, group_by = "group.y", factors = 1:5, 
                 color_by = "group.y", 
                 add_violin = TRUE)

p <- p + 
  scale_color_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green")) +
  scale_fill_manual(values=c("CVM"="black", "CVF"="red", "GFM"="blue", "GFF"="green"))
print(p)
```

## Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```