---
title: "DMR Analysis"
output:
  html_document:
    toc: true
    toc_float: true
date: "2023-10-17"
params:
  methrix_path: "/home/rstudio/bioinf-d/lima_results_methylseq/Results/merged_methrix.rds"
  chromosome_filter: !r NULL
  outlier_samples: !r  c("X670E", "X594C")
  
  
  granges_path: "/home/rstudio/bioinf-d/lima_results_methylseq/Results/dmrseq.group.non_smoothed_global/dmrseq_results_non_smoothed_global.group.rds"
  
  output_dir: "/home/rstudio/bioinf-d/lima_results_methylseq/Results/"
  threads: 1
  
  random_seed: 1234
  group_var: "group"
  
  adj.p.value_thr: 0.1
  merge_granges: FALSE
  filter_significant_before_merging: TRUE

  merge_method: "disjoint"
  
  unique_gene_region_annotation_filter: FALSE 
  results_name: "non_smoothed_global"
  
  genome_id: "mm39"
  annotation_genome: "TxDb.Mmusculus.UCSC.mm39.knownGene"
  organism_db: "org.Mm.eg.db"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, echo = FALSE)
options(knitr.duplicate.label = "allow")
```

```{r}
library(methrix)
library(annotatr)
library(AnnotationHub)
library(ChIPseeker)
library(plotly)
library(ggplot2)
library(bsseq)
library(dplyr)
library(matrixStats)
library(params$annotation_genome, character.only = TRUE)
library(params$organism_db, character.only = TRUE)
source("../WGB_DNA_Methylation/plot_functions.R")
```

```{r}
merged_methrix <- readRDS(params$methrix_path)
results_name <- paste0(params$results_name, ".",  params$group_var)

if(!is.null(params$outlier_samples)){
  merged_methrix <- merged_methrix[,!colnames(merged_methrix) %in% params$outlier_samples]
}

if(!is.null(params$chromosome_filter)){
  all_chrom <- metadata(merged_methrix)$ref_CpG$chr
  merged_methrix <- subset_methrix(m = merged_methrix, 
                                  contigs = all_chrom[!(all_chrom %in% params$chromosome_filter)], 
                                  overlap_type = "any")
}
```

### Merge Results {.tabset}

```{r}
granges_results <- readRDS(params$granges_path)


if(params$merge_granges){
  unmerged_granges <- granges_results
  
  if(params$filter_significant_before_merging){
    granges_results <- lapply(granges_results, function(ctrs_res){
      p_val_cols <- mcols(ctrs_res)[grepl("\\.adj\\.p\\.value|qval", 
                                    colnames(mcols(ctrs_res)))]
      ctrs_res[rowSums(as.matrix(p_val_cols) <= params$adj.p.value_thr, na.rm = TRUE) >= 1]
    })
  }
  
  merge_method <- match.arg(params$merge_method, 
                            c("disjoint", "reduced", "union")) # Reduced and union give the exact same results
  granges_results <- switch(merge_method, "disjoint" = {
    ## Merge Option A - disjoint regions
    # Most Accuracy but includes duplicated and might cut DMR related to the same gene
    local_res_disjoin <- disjoin(Reduce(c, granges_results))
    
    for(ctrs_ranges_name in names(granges_results)){
      ctrs_ranges <- granges_results[[ctrs_ranges_name]]
      
      colnames(mcols(ctrs_ranges)) <- paste(ctrs_ranges_name, 
                                            colnames(mcols(ctrs_ranges)), 
                                            sep = ".") 
      
      found_overlaps <- findOverlaps(query = ctrs_ranges, local_res_disjoin)
  
      for(col_name in colnames(mcols(ctrs_ranges))){
        if(grepl("index", col_name)){next}
        mcols(local_res_disjoin)[[col_name]] <- NA  
        mcols(local_res_disjoin)[found_overlaps@to, col_name] <- mcols(ctrs_ranges)[found_overlaps@from, col_name]
      }
      mcols(local_res_disjoin)[[paste(ctrs_ranges_name, "id", sep = ".")]] <- NA
      mcols(local_res_disjoin)[found_overlaps@to, 
                              paste(ctrs_ranges_name, 
                                    "id", sep = ".")] <- found_overlaps@from
    }
    local_res_disjoin
  }, "reduced" = {
    # Merge Option B - reduce
    local_res_reduced <- reduce(Reduce(c, granges_results))
    
    for(ctrs_ranges_name in names(granges_results)){
      ctrs_ranges <- granges_results[[ctrs_ranges_name]]
    
      colnames(mcols(ctrs_ranges)) <- paste(ctrs_ranges_name,
                                            colnames(mcols(ctrs_ranges)), 
                                            sep = ".")
    
      found_overlaps <- findOverlaps(query = ctrs_ranges, local_res_reduced)
    
      for(col_name in colnames(mcols(ctrs_ranges))){
        if(grepl("index", col_name)){next}
        mcols(local_res_reduced)[col_name] <- NA
        # If multiple regions match to the same one, then the last one is kept. We might want to change this to keep the most significant from both or to try and combine the pvalues?
        mcols(local_res_reduced)[found_overlaps@to,col_name] <- mcols(ctrs_ranges)[found_overlaps@from, col_name]
      }
    }
    local_res_reduced
  }, "union" = {
    local_res_union <- Reduce(GenomicRanges::union, granges_results)
    for(ctrs_ranges_name in names(granges_results)){
      ctrs_ranges <- granges_results[[ctrs_ranges_name]]
    
      colnames(mcols(ctrs_ranges)) <- paste(ctrs_ranges_name,
                                            colnames(mcols(ctrs_ranges)), 
                                            sep = ".")
    
      found_overlaps <- findOverlaps(query = ctrs_ranges, local_res_union)
    
      for(col_name in colnames(mcols(ctrs_ranges))){
        if(grepl("index", col_name)){next}
        mcols(local_res_union)[col_name] <- NA
        mcols(local_res_union)[found_overlaps@to, col_name] <- mcols(ctrs_ranges)[found_overlaps@from, col_name]
      }
    }
    local_res_union
  })
  results_list <- c(list("All_DMRs" = granges_results), unmerged_granges)
  
  merge_method <- paste0( ".", merge_method)
  
  # peakAnnoList <- c(peakAnnoList, "Merged" = rowRanges(dna_met_exp))
  # lapply(unmerged_granges, function(ctrs_dt){
  #   length(unique(findOverlaps(ctrs_dt, granges_results)@from))/length(ctrs_dt)
  # })

} else{
  merge_method <- ""
  if(is(granges_results, "list")){
    granges_results <- granges_results[[1]]  
  }
  results_list <- c(list("All_DMRs" = granges_results))
}
```


```{r}
# test_granges <- lapply(granges_results, function(x){x[seqnames(x) == "chr2"]})
# test_granges_disjoin <- disjoin(Reduce(c, test_granges))
# 
# range_view <- GRanges(seqnames = "chr2", ranges = IRanges(start = min(start(test_granges_disjoin)),
#         end = max(end(test_granges_disjoin))))
# 
# trackViewer::plotGRanges(test_granges$`HFD-CNT`, test_granges_disjoin,
#                          range = range_view)
# trackViewer::plotGRanges(test_granges$`HFD-CNT`, test_granges_disjoin,
#                          range = test_granges$`INT-HFD`[246])
# 
# 
# test_granges$`INT-HFD`[246]
# 
# test_found <- lapply(test_granges, function(ctrs){
#   ctrs[findOverlaps(ctrs, test_granges$`INT-HFD`[246])@from]
# })
# 
# # This one works
# test_methrix <- subset_methrix(merged_methrix, regions = test_granges$`INT-HFD`[246])
# test <- get_region_summary(merged_methrix, regions = test_granges$`INT-HFD`[246], overlap_type = "within")
# 
# # This is diferent
# test2 <- get_region_summary(merged_methrix, regions = test_granges$`INT-HFD`[246], overlap_type = "any")
# test_methrix2 <- subset_methrix(merged_methrix, regions = test_granges$`INT-HFD`[246], overlap_type = "any")
# 
# # This works
# test_methrix_d <- subset_methrix(merged_methrix, regions = test_granges_disjoin[23229], overlap_type = "any")
# test_d <- get_region_summary(merged_methrix, regions = test_granges_disjoin[23229], overlap_type = "any")
# 
# test_methrix@elementMetadata
# 
# overlapped_regions <- test_granges_disjoin[findOverlaps(test_granges$`INT-HFD`[246], test_granges_disjoin)@to]
```

```{r}
## Mean methylation quantification
merged_reduced_res <- methrix::get_region_summary(m = merged_methrix, 
                                                  regions = granges_results, 
                                                  overlap_type = "any")
```

```{r}
## Annotation
txdb <- base::get(params$annotation_genome)

peakAnnoList <- lapply(results_list, annotatePeak, TxDb=txdb, 
                        tssRegion=c(-3000, 3000), verbose=FALSE)

# pdf("Any_Sig_Disjoint_All_Sig_ctrs_annotation_.pdf")
# ChIPseeker::plotAnnoBar(peakAnnoList)
# dev.off()
peakAnnoList <- lapply(peakAnnoList, as.GRanges)

annotated_dt <- peakAnnoList$All_DMRs

colnames(mcols(annotated_dt)) <- gsub("pval", "p.value", 
                                      colnames(mcols(annotated_dt)))
colnames(mcols(annotated_dt)) <- gsub("qval", "adj.p.value", 
                                      colnames(mcols(annotated_dt)))

loaded_org_db <- base::get(params$organism_db)

mcols(annotated_dt)$geneSymbol <- mapIds(x = loaded_org_db, 
                                        keys = mcols(annotated_dt)$geneId, 
                                        column = "SYMBOL", 
                                        keytype = "ENTREZID")
mcols(annotated_dt)$geneName <- mapIds(x = loaded_org_db, 
                                        keys = mcols(annotated_dt)$geneId, 
                                        column = "GENENAME", 
                                        keytype = "ENTREZID")
mcols(annotated_dt)$n_overlap_CpGs <- merged_reduced_res$n_overlap_CpGs

dna_met_exp <- SummarizedExperiment(
  assays = as.data.frame(merged_reduced_res)[,colnames(merged_reduced_res) %in% colnames(merged_methrix)],
  rowRanges = annotated_dt,
  colData = colData(merged_methrix))

rownames(dna_met_exp) <- paste0("DMR", rownames(dna_met_exp))

p_vals_cols <- grep("adj\\.p\\.value", colnames(rowData(dna_met_exp)), value = TRUE)

for(id in p_vals_cols){
  rowData(dna_met_exp)[[id]][is.na(rowData(dna_met_exp)[[id]])] <- 1
}
```

```{r, eval=params$unique_gene_region_annotation_filter}
exp_row_data <- rowData(dna_met_exp)

annotation_regions <- paste0(exp_row_data[["annotation"]], exp_row_data[["geneId"]])

unique_annotations <- unique(annotation_regions)
id_columns <- grepl("\\.id$", colnames(rowData(dna_met_exp)))

p_val_dt <- as.matrix(rowData(dna_met_exp)[,grepl("\\.adj\\.p\\.value$", colnames(rowData(dna_met_exp)))])
p_val_dt <- p_val_dt <= params$adj.p.value_thr

exp_row_data <- as.matrix(exp_row_data[,id_columns])
exp_row_data_test <- data.frame("no_na_counts" = ncol(exp_row_data) - rowCounts(exp_row_data, value = NA),
                                "num_sig" = rowSums(p_val_dt, na.rm = TRUE),
                                "dmr_width" = width(rowRanges(dna_met_exp)),
                                "annotation" = annotation_regions)

rownames(exp_row_data_test) <- rownames(exp_row_data)

exp_row_data_test <- exp_row_data_test[order(exp_row_data_test$num_sig, 
                                             exp_row_data_test$no_na_counts, 
                                             exp_row_data_test$dmr_width,
                                             decreasing = TRUE),]

keep_dmrs <- tapply(X = exp_row_data_test, 
                    INDEX = exp_row_data_test$annotation, 
                    FUN = function(x){
  rownames(x)[1]
})

dna_met_exp <- dna_met_exp[keep_dmrs,]

peakAnnoList <- c(peakAnnoList, "Merged and Filtered" = rowRanges(dna_met_exp))
```

```{r}
# saveRDS(dna_met_exp, paste0(params$output_dir, "dna_met_exp_",
#                               results_name, merge_method, ".rds"))

dna_met_exp <- dna_met_exp[rowSums(as.matrix(rowData(dna_met_exp)[,p_vals_cols]) <= params$adj.p.value_thr, 
                                   na.rm = TRUE) >= 1,]

saveRDS(dna_met_exp, paste0(params$output_dir, "dna_met_exp_", 
                              results_name, merge_method, ".rds"))
```

```{r}
methrix_sub <- subset_methrix(merged_methrix, 
                              regions = rowRanges(dna_met_exp), 
                              overlap_type = "any")

set.seed(params$random_seed)
random_cpgs <- sample(seq_len(nrow(methrix_sub)), 
                      size = min(5000, nrow(methrix_sub)))

plot_dt <- assay(methrix_sub[random_cpgs,])
plot_dt <- impute::impute.knn(data = plot_dt)$data
```

### Results Analysis {.tabset}

#### Number of contrasts in which a DMR is significant table

```{r}
num_significant <- rowSums(as.matrix(rowData(dna_met_exp)[,p_vals_cols]) <= params$adj.p.value_thr, 
                                   na.rm = TRUE)
table(num_significant)
```


#### Region annotations non significant

```{r}
dmr_annotation_plot(omic_dt = peakAnnoList)
```

#### Region annotations significant

```{r}
results_list_sig <- lapply(peakAnnoList, function(x){
  p_val_cols <- as.matrix(mcols(x)[,grepl("adj\\.p\\.value|qval", 
                                          colnames(mcols(x))), drop = FALSE])
  p_val_cols[is.na(p_val_cols)] <- 1
  x[rowSums(p_val_cols <= params$adj.p.value_thr, na.rm = TRUE) >= 1]
})
dmr_annotation_plot(omic_dt = results_list_sig)
```

#### PCA any significant CpGs

```{r}
set.seed(params$random_seed)
top_cpgs_pca <- methrix::methrix_pca(m = methrix_sub, top_var = 5000, var = "rand")


pc_plot <- methrix::plot_pca(top_cpgs_pca, m = methrix_sub, col_anno = params$group_var, 
                  show_labels = TRUE, pc_x = "PC1", pc_y = "PC2") + ggtitle("Random 5000 CpGs")
pc_plot <- pc_plot + stat_ellipse(type = "t")

print(pc_plot)

# png(paste0(params$output_dir, "top_5000_filtered_cpgs_beta_values_pca", results_name, ".png"))
# print(pc_plot)
# dev.off()
```

#### Heatmaps any significant {.tabset}

##### Beta values

```{r}
set.seed(params$random_seed)

quantile_breaks <- quantile(plot_dt, seq(0, 1, length.out = 101), na.rm = TRUE)
quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
heatmap_dt <- pheatmap::pheatmap(mat = t(plot_dt), main = "Random 5000 CpGs",
                  annotation_row = as.data.frame(colData(methrix_sub)[,params$group_var, drop = FALSE]),  
                  clustering_method = "average",
                  color = color_values, clustering_distance_rows = dist(scale(t(assay(methrix_sub)))),
                  breaks = quantile_breaks, show_colnames = FALSE)

# png(paste0(params$output_dir, "random_5000_filtered_cpgs_beta_values_heatmap", results_name, ".png"))
# print(plot(heatmap_dt$gtable))
# dev.off()
```

##### Scaled Beta values

```{r}
quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]

color_values <- viridis::viridis(n = length(quantile_breaks) - 1)

set.seed(params$random_seed)
heatmap_dt <- pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                  annotation_row = as.data.frame(colData(methrix_sub)[,params$group_var, drop = FALSE]),
                  clustering_method = "average", main = "Random 5000 CpGs",
                  color = color_values, clustering_distance_rows = dist(scale(t(assay(methrix_sub)))), 
                  breaks = quantile_breaks, show_colnames = FALSE)

# png(paste0(params$output_dir, "random_5000_filtered_cpgs_scaled_values_heatmap", results_name, ".png"))
# plot(heatmap_dt$gtable)
# dev.off()
```

#### Heatmap significant DMRs {.tabset}

##### Beta values

```{r}
plot_dt <- assay(dna_met_exp)
plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

quantile_breaks <- quantile(plot_dt, seq(0, 1, length.out = 101), na.rm = TRUE)
quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]

color_values <- viridis::viridis(n = length(quantile_breaks) - 1)

set.seed(params$random_seed)
heatmap_dt <- pheatmap::pheatmap(mat = t(plot_dt), 
                  annotation_row = as.data.frame(colData(dna_met_exp)[,params$group_var, drop = FALSE]),
                  clustering_method = "average", 
                  main = paste("DMRs Any Significant (Mean methylation value & adj pvalue <=",
                              params$adj.p.value_thr),
                  color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))), 
                  breaks = quantile_breaks, show_colnames = FALSE)
```

##### Scaled Beta values

```{r}
quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]

color_values <- viridis::viridis(n = length(quantile_breaks) - 1)

set.seed(params$random_seed)
heatmap_dt <- pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                  annotation_row = as.data.frame(colData(dna_met_exp)[,params$group_var, drop = FALSE]),
                  clustering_method = "average", 
                  main = paste("DMRs Any Significant (Mean methylation value & adj pvalue <=",
                              params$adj.p.value_thr),
                  color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))), 
                  breaks = quantile_breaks, show_colnames = FALSE)
```

#### CpG distribution any significant {.tabset}

##### Beta values {.tabset}

###### By sample 

```{r}
plot_list <- group_distribution_plots(methrix_sub, group_var = params$group_var, scale_dt = FALSE)
plot_list[[1]]
```

```{r}
plot_list[[2]]
```

###### By group

```{r, fig.height=10, fig.width=10}
sample_distribution_plot(methrix_sub, group_var = params$group_var, scale_dt = FALSE)
```

##### Scaled Beta values {.tabset}

###### By sample 

```{r}
plot_list <- group_distribution_plots(methrix_sub, group_var = params$group_var)
plot_list[[1]]
```

```{r}
plot_list[[2]]
```

###### By group

```{r, fig.height=10, fig.width=10}
sample_distribution_plot(methrix_sub, group_var = params$group_var)
```

```{r}
# ranges_dt <- rowRanges(dna_met_exp)
# 
# mcols(ranges_dt) <- mcols(ranges_dt)[,!colnames(mcols(ranges_dt)) %in% c("seqnames", "ranges", "strand", "seqlevels", "seqlengths", "isCircular", "start",
#     "end", "width", "element")]
# mcols(ranges_dt)$qval <- 1
# mcols(ranges_dt)$pval <- 1  
# mcols(ranges_dt)$stat <- 1
# dmrseq::plotDMRs(BSseq = bsseq_obj, regions = ranges_dt[1:100], testCovariate = "Condition")
```



```{r}


################# Comparing merge methodologies
# trackViewer::plotGRanges(granges_results, granges_results_r, range = GRanges(seqnames = "chr1", 
#                                             ranges = IRanges(start = 4877830, 
#                                                              end = 4877913)))
# 
# merged_disjoint_res <- methrix::get_region_summary(m = merged_methrix, regions = granges_results)
# 
# merged_reduced_res <- methrix::get_region_summary(m = merged_methrix, regions = granges_results_r)
# 
# 
# granges_results
# granges_results_r


# pca_disj <- methrix_pca(m = merged_methrix, ranges = granges_results)
# plot_pca(pca_disj, m = merged_methrix, col_anno = "Condition")
# pca_disj_f <- methrix_pca(m = merged_methrix, ranges = granges_results_filt)
# plot_pca(pca_disj_f, m = merged_methrix, col_anno = "Condition", pc_x = "PC1", pc_y = "PC2")
# 
# pca_red <- methrix_pca(m = merged_methrix, ranges = granges_results_r)
# plot_pca(pca_red, m = merged_methrix, col_anno = "Condition")
##############
```

## Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```

<!-- # ```{r} -->
<!-- # lapply(parwise_results, function(x){ -->
<!-- #   sig_pw <- x[x$qval <= 0.05] -->
<!-- #   res_overlaps <- findOverlaps(sig_pw, local_res_ctrs$`Global Test`)@from -->
<!-- #   c(length(unique(res_overlaps))/length(sig_pw), length(sig_pw)) -->
<!-- # }) -->
<!-- #  -->
<!-- # methrix_subset <- subset_methrix(merged_methrix,  -->
<!-- #                                 local_res_ctrs$`Global Test`) -->
<!-- # ``` -->

