---
title: "DMR Enrichment Analyses"
output: html_document
date: "2024-01-16"
params: 
    dmr_exp_list: !r dna_met_list[[1]]
    ora_filter: !r expression(Factor_1_weights_sig)

    txdb_id: "TxDb.Mmusculus.UCSC.mm39.refGene"
    msigdbr_specie: "Mus musculus"
    random_seed: 1234
    output_dir: ""
    results_name: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(knitr.duplicate.label = "allow")

```

## Required Packages

```{r}
library(MultiAssayExperiment)
library(rGREAT)
library(dplyr)
library(msigdbr)
library(knitr)
library(DT)
library(GOSemSim)
library(org.Mm.eg.db)
library(SummarizedExperiment)
library(params$txdb_id, character.only = TRUE)
source("../../scripts/Utility_Functions/utilities.R")

```

## rGreat {.tabset}

```{r}
enrich_results_list <- list()
```

```{r}
# Get filtered granges
granges_list <- lapply(params$dmr_exp_list , function(ctrs_res){
    sub_granges <- rowRanges(ctrs_res[eval(params$ora_filter, envir = rowData(ctrs_res)),])
    if(length(sub_granges) == 0){return(NA)}

    mcols(sub_granges) <- mcols(sub_granges)[!colnames(mcols(sub_granges)) %in% 
                                               c("seqnames", "ranges", "strand", 
                                                "seqlevels", "seqlengths",
                                                "isCircular", "start", "end", 
                                                "width", "element")]
    return(sub_granges)
})
```

### GO {.tabset}

```{r}
enrich_results_list$ORA <- list()
```

#### BP {.tabset}

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = "GO:BP", 
                tss_source = params$txdb_id))
})
```

```{r}
bp_data <- godata(OrgDb = org.Mm.eg.db, ont = "BP")

contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)

    enrich_res <- rename(enrich_res, c("ID" = "id"))
    enrich_res <- clusterProfiler:::simplify_internal(
        res = enrich_res, cutoff = 0.7, by = "fold_enrichment", select_fun = max, 
        ontology = "BP", semData = bp_data)
    # enrich_res <- test_f(res = enrich_res, cutoff = 0.6, ontology = "BP", by = "p_adjust", semData = bp_data)
    
    return(enrich_res)
})
contrs_names <- names(contrs_result_sig_list)

enrich_results_list$ORA <- c(enrich_results_list$ORA, contrs_result_sig_list)

template <- c(
    "##### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

#### MF {.tabset}

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = "GO:MF", 
                tss_source = params$txdb_id))
})
```

```{r}
mf_data <- godata(OrgDb = org.Mm.eg.db, ont = "MF")

contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)

    enrich_res <- rename(enrich_res, c("ID" = "id"))
    enrich_res <- clusterProfiler:::simplify_internal(
        res = enrich_res, cutoff = 0.7, by = "fold_enrichment", 
        select_fun = max, ontology = "MF", semData = mf_data)
    return(enrich_res)
})
contrs_names <- names(contrs_result_sig_list)

enrich_results_list$ORA <- base::Map(c, enrich_results_list$ORA, contrs_result_sig_list)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

#### CC {.tabset}

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = "GO:CC", 
                tss_source = params$txdb_id))
})
```

```{r}
cc_data <- godata(OrgDb = org.Mm.eg.db, ont = "CC")

contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)

    enrich_res <- rename(enrich_res, c("ID" = "id"))
    enrich_res <- clusterProfiler:::simplify_internal(
        res = enrich_res, cutoff = 0.7, by = "fold_enrichment", 
        select_fun = max, ontology = "CC", semData = cc_data)
    return(enrich_res)
})
contrs_names <- names(contrs_result_sig_list)

enrich_results_list$ORA <- base::Map(c, enrich_results_list$ORA, contrs_result_sig_list)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

### KEGG {.tabset}

```{r}
kegg_pw <- rGREAT::getKEGGPathways("mmu")
```

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = kegg_pw, 
                tss_source = params$txdb_id))
})
```

```{r}
contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)

    return(enrich_res)
})
contrs_names <- names(contrs_result_sig_list)

enrich_results_list$ORA <- base::Map(c, enrich_results_list$ORA, contrs_result_sig_list)

template <- c(
    "#### `r contrs_names[{{i}}]` \n",
    "```{r, echo = FALSE}\n",
    "datatable(contrs_result_sig_list[[{{i}}]], options = list(scrollX = TRUE))",
    "```\n",
    "\n"
)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`



### WP {.tabset}

```{r, results='hide'}
wiki_pw <- GSA::GSA.read.gmt("../../inputs/wikipathways-20240101-gmt-Mus_musculus.gmt")
wiki_genesets <- wiki_pw$genesets
names(wiki_genesets) <- wiki_pw$geneset.names
```

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = wiki_genesets, 
                tss_source = params$txdb_id))
})
```

```{r}
contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)
    return(enrich_res)
})
contrs_names <- names(contrs_result_sig_list)

enrich_results_list$ORA <- base::Map(c, enrich_results_list$ORA, contrs_result_sig_list)

```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

### MSIGdbr H {.tabset}

```{r}
library(msigdbr)
h_gene_sets <- msigdbr(species = params$msigdbr_specie, 
                        category = "H")

h_gene_sets <- split(h_gene_sets$entrez_gene, h_gene_sets$gs_name)
h_gene_sets <- lapply(h_gene_sets, as.character)  # just to make sure gene IDs are all in character.
```

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = h_gene_sets, 
                tss_source = params$txdb_id))
})
```

```{r}
contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)
    return(enrich_res)
})

enrich_results_list$ORA <- base::Map(c, enrich_results_list$ORA, contrs_result_sig_list)

contrs_names <- names(contrs_result_sig_list)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

### MSIGdbr C2 {.tabset}

```{r}
library(msigdbr)
h_gene_sets <- msigdbr(species = params$msigdbr_specie, 
                        category = "C2")

h_gene_sets <- split(h_gene_sets$entrez_gene, h_gene_sets$gs_name)
h_gene_sets <- lapply(h_gene_sets, as.character)  # just to make sure gene IDs are all in character.
```

```{r, cache=TRUE}
set.seed(params$random_seed)
great_results <- lapply(granges_list, function(granges_obj){
    if(length(granges_obj) == 0){return(NA)}
    
    return(great(gr = granges_obj, 
                gene_sets = h_gene_sets, 
                tss_source = params$txdb_id))
})
```

```{r}
contrs_result_sig_list <- lapply(great_results, function(gr_res){
    if(!is(gr_res, "GreatObject")){
        return(data.frame())
    }
    enrich_res <- getEnrichmentTable(gr_res)
    enrich_res <- subset(enrich_res, p_adjust <= 0.05)
    return(enrich_res)
})

enrich_results_list$ORA <- base::Map(c, enrich_results_list$ORA, contrs_result_sig_list)

contrs_names <- names(contrs_result_sig_list)
```

```{r}
tables_chunk <- significant_tables_chunks(contrs_result_sig_list = contrs_result_sig_list, 
                                          contrs_names = contrs_names, template = template)
```

`r knit(text=unlist(tables_chunk))`

```{r}
saveRDS(enrich_results_list,  paste0(params$output_dir, params$results_name, 
                                    "_dmr_enrich_results.rds"))
```


## Appendix

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```