---
title: "dmrseq_analysis"
output: html_document
date: "2023-12-06"
params:
  methrix_path: !r methrix_path
  output_dir: !r dmrseq_output
  threads: 1
  
  chromosome_filter: !r chromosome_filter
  outlier_samples: !r  NULL
  
  max_fraction_missing: 0.2
  
  random_seed: 1234
  group_var: "group"
  
  global_test: FALSE
  covariates: "batch"
  cutoff: 0.05
  bpSpan: 500
  minNumRegion: 3
  minInSpan: 11 
  smooth: FALSE
  maxGapSmooth: 2000
  maxGap: 200 
  maxPerms: 3
  
  results_name: "non_smoothed"
  
  dmrseq_thr: 0.05

  pairwiseDSS: TRUE
  dss_cpg_adj.p.value: 0.05
  dss_cpg_percentage_sig: 0.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, echo = FALSE)
options(knitr.duplicate.label = "allow")
```

```{r}
library(methrix)
library(plotly)
library(ggplot2)
library(annotatr)
library(AnnotationHub)
library(ChIPseeker)
library(impute)
library(bsseq)
source("./plot_functions.R")
```

```{r}
merged_methrix <- readRDS(params$methrix_path)

if(!is.null(params$outlier_samples)){
  merged_methrix <- merged_methrix[,!colnames(merged_methrix) %in% params$outlier_samples]
}

if(!is.null(params$chromosome_filter)){
  all_chrom <- metadata(merged_methrix)$ref_CpG$chr
  merged_methrix <- subset_methrix(m = merged_methrix, 
                                  contigs = all_chrom[!(all_chrom %in% params$chromosome_filter)])
}
```

## DMRSeq

```{r dmrseq}
results_name <- paste0(params$results_name, ".",  params$group_var)
dmrseq_path <- paste0(params$output_dir, "dmrseq_results_", results_name, ".rds")
# bsseq_path <- paste0(params$output_dir, "bsseq_obj.rds")

if(!file.exists(dmrseq_path)){
    
    # merged_methrix <- methrix::subset_methrix(merged_methrix, contigs = c("chr1"))
    
    bsseq_obj <- methrix::methrix2bsseq(merged_methrix)
    
    delayed_cov <- getCoverage(bsseq_obj, type="Cov") == 0
    
    # loci.idx <- which((DelayedMatrixStats::rowSums2(getCoverage(bsseq_obj, type="Cov")==0)/ncol(bsseq_obj)) <= params$max_fraction_missing)
  
    loci.idx <- lapply(levels(bsseq_obj[[params$group_var]]), function(cond_level){
        which((DelayedMatrixStats::rowSums2(delayed_cov, 
                                            cols = bsseq_obj[[params$group_var]] == cond_level)/sum(bsseq_obj[[params$group_var]] == cond_level)) < params$max_fraction_missing)
    })

    # loci.idx <- Reduce(union, loci.idx)
    loci.idx <- Reduce(intersect, loci.idx)
    # 
    
    bsseq_obj <- bsseq_obj[sort(loci.idx),]
    
    if(params$global_test){
        set.seed(params$random_seed)
        local_res_ctrs <- list("Global Test" = dmrseq::dmrseq(
            bs = bsseq_obj, cutoff = params$cutoff, 
            testCovariate = params$group_var, minNumRegion = params$minNumRegion,
            bpSpan = params$bpSpan, minInSpan = params$minInSpan, 
            smooth = params$smooth, maxGapSmooth = params$maxGapSmooth, 
            maxGap = params$maxGap, maxPerms = params$maxPerms, 
            BPPARAM = BiocParallel::SerialParam()))
    } else{
        ctrs_list <- combn(levels(merged_methrix[[params$group_var]]), 2)
        rm(loci.idx)
        rm(delayed_cov)
        
        local_res_ctrs <- list()
        
        set.seed(params$random_seed)
        for(row_ids in seq_len(ncol(ctrs_list))){
          ctrs_ids <- ctrs_list[,row_ids]
          set.seed(params$random_seed)
          local_res_ctrs[[paste0(ctrs_ids[2:1], collapse = "-")]] <- dmrseq::dmrseq(
              bs = bsseq_obj[,bsseq_obj[[params$group_var]] %in% ctrs_ids], cutoff = params$cutoff, 
              testCovariate = params$group_var, minNumRegion = params$minNumRegion,
              bpSpan = params$bpSpan, minInSpan = params$minInSpan, smooth = params$smooth,
              maxGapSmooth = params$maxGapSmooth, maxGap = params$maxGap, 
              maxPerms = params$maxPerms, 
              BPPARAM = BiocParallel::SerialParam())
        }
    }
    # saveRDS(bsseq_obj, bsseq_path)
    saveRDS(local_res_ctrs, dmrseq_path)
} else{
  local_res_ctrs <- readRDS(dmrseq_path)  
  bsseq_obj <- methrix2bsseq(merged_methrix)
}

library(GenomicRanges)
library(IRanges)

lapply(local_res_ctrs, function(x){
  sum(x$qval <= params$dmrseq_thr)
})
```

## Results Plots {.tabset}

### Mean methylation diff

```{r, eval = !params$global_test}
mean_methylation_diff <- lapply(names(local_res_ctrs), function(ctrs_dt_name){
  ctrs_dt <- local_res_ctrs[[ctrs_dt_name]]
  ctrs_dt <- ctrs_dt[ctrs_dt$qval <= 0.05]
  ctrs_ids <- strsplit(ctrs_dt_name, split = "-")[[1]]
  print(ctrs_ids)
  if(length(ctrs_dt) == 0){return(NULL)}
  mean_methylation_dt <- data.frame("mean_methylation_diff" = dmrseq::meanDiff(bsseq_obj[,bsseq_obj[[params$group_var]] %in% ctrs_ids],
                                                                                dmrs = ctrs_dt, testCovariate = params$group_var), 
                                    "ctrs_name" = ctrs_dt_name)
  mean_methylation_dt
})
plot_dt <- do.call(rbind, mean_methylation_diff)
hist_mean_meth <- ggplot(plot_dt, aes(x = mean_methylation_diff, fill = ctrs_name)) +
  geom_histogram(alpha = 0.6, position = "identity") + theme_minimal() + 
  scale_fill_brewer(palette = "Set2")
# ggplotly(hist_mean_meth)
hist_mean_meth
```

### DMR Widths

```{r}
plot_dt <- lapply(names(local_res_ctrs), function(ctrs_dt_name){
  ctrs_dt <- local_res_ctrs[[ctrs_dt_name]]
  beta_cols <- grepl("beta", colnames(mcols(ctrs_dt)))
  
  
  found_dt <- data.frame("dmr_widths" = width(ctrs_dt), 
                         "pvalues" = ctrs_dt$pval, 
                         "qvalues" = ctrs_dt$qval,
                         "ctrs_name" = ctrs_dt_name)
  cbind(found_dt, mcols(ctrs_dt)[,beta_cols, drop = FALSE])
})
plot_dt <- do.call(rbind, plot_dt)

plot_res <- ggplot(plot_dt, aes(x = dmr_widths, fill = ctrs_name)) + 
  geom_histogram(alpha = 0.6, position = "identity", aes(y = ..density..)) + 
  theme_minimal() + scale_fill_brewer(palette = "Set2")
plot_res
# ggplotly(plot_res)
```

### pvalue distributions

```{r}
plot_res <- ggplot(plot_dt, aes(x = pvalues, fill = ctrs_name)) + 
  geom_histogram(alpha = 0.6, position = "identity") + theme_minimal() + 
  scale_fill_brewer(palette = "Set2") + geom_vline(xintercept = 0.05)
plot_res
#ggplotly(plot_res)
```

### qvalue distributions

```{r}
plot_res <- ggplot(plot_dt, aes(x = qvalues, fill = ctrs_name)) + 
  geom_histogram(alpha = 0.6, position = "identity") + theme_minimal() + 
  scale_fill_brewer(palette = "Set2") + geom_vline(xintercept = 0.05)
plot_res
#ggplotly(plot_res)
```

### Beta values distribution

```{r}
plot_dt_beta <- reshape2::melt(plot_dt[,grepl("beta", colnames(plot_dt)) | colnames(plot_dt) == "ctrs_name", drop = FALSE], id.vars = "ctrs_name")
colnames(plot_dt_beta) <- c("Group", "Group_Beta", "Beta_Value")

beta_plot_var <- ifelse(params$global_test, "Group_Beta", "Group")

plot_res <- ggplot(plot_dt_beta, aes(x = Beta_Value, 
                                     fill = .data[[beta_plot_var]])) +
  geom_histogram(alpha = 0.6, position = "identity") + theme_minimal() +
  scale_fill_brewer(palette = "Set2")
plot_res
#ggplotly(plot_res)
```

### Granges Percentage found DMRs Overlaps

List of the percentage of DMRs of a contrast that can be found in each of the other contrasts DMRs.

```{r}
overlap_list <- lapply(names(local_res_ctrs), function(ctrs_name){
  ctrs_to_overlap <- local_res_ctrs[[ctrs_name]]
  # ctrs_to_overlap <- ctrs_to_overlap[ctrs_to_overlap$qval <= 0.05]
  
  other_ctrs_name_list <- names(local_res_ctrs)[!names(local_res_ctrs) %in% ctrs_name]
  
  ctrs_overlap <- sapply(other_ctrs_name_list, function(other_ctrs_name){
    other_ctrs <- local_res_ctrs[[other_ctrs_name]]
    # other_ctrs <- other_ctrs[other_ctrs$qval <= 0.05]
    
    length(unique(findOverlaps(ctrs_to_overlap, other_ctrs)@from))/length(ctrs_to_overlap)
  })
  names(ctrs_overlap) <- other_ctrs_name_list
  # ctrs_overlap[[ctrs_name]] <- NA
  ctrs_overlap
})

names(overlap_list) <- names(local_res_ctrs)
overlap_list
```

### Granges Percentage Significant DMRs Overlaps

List of the percentage of significant DMRs of a contrast that can be found in each of the other contrasts significant DMRs.

```{r}
overlap_list <- lapply(names(local_res_ctrs), function(ctrs_name){
  ctrs_to_overlap <- local_res_ctrs[[ctrs_name]]
  ctrs_to_overlap <- ctrs_to_overlap[ctrs_to_overlap$qval <= 0.05]
  
  other_ctrs_name_list <- names(local_res_ctrs)[!names(local_res_ctrs) %in% ctrs_name]
  
  ctrs_overlap <- sapply(other_ctrs_name_list, function(other_ctrs_name){
    other_ctrs <- local_res_ctrs[[other_ctrs_name]]
    other_ctrs <- other_ctrs[other_ctrs$qval <= 0.05]
    
    length(unique(findOverlaps(ctrs_to_overlap, other_ctrs)@from))/length(ctrs_to_overlap)
  })
  names(ctrs_overlap) <- other_ctrs_name_list
  # ctrs_overlap[[ctrs_name]] <- NA
  ctrs_overlap
})

names(overlap_list) <- names(local_res_ctrs)
overlap_list
```

### Heatmaps subset

```{r results_heatmaps, eval = !params$global_test}
heat_map_list <- lapply(names(local_res_ctrs), function(ctrs_name){
  ctrs <- local_res_ctrs[[ctrs_name]]
  sig_dmrs <- ctrs[ctrs$qval <= params$dmrseq_thr,]
  if(length(sig_dmrs) <= 2){return(NA)}
  ctrs_ids <- strsplit(ctrs_name, split = "-")[[1]]

  methrix_sub <- merged_methrix[,merged_methrix[[params$group_var]] %in% ctrs_ids]
  plot_dt <- methrix::get_region_summary(m = methrix_sub, 
                                          regions = sig_dmrs, 
                                          overlap_type = "any")
  plot_dt <- plot_dt[,6:ncol(plot_dt)]
  plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

  quantile_breaks <- quantile(scale(t(plot_dt)), 
                              seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  set.seed(params$random_seed)
  # selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
  # browser()
  pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                    main = paste("Heatmap All Significant DMRs (Scaled mean methylation)", ctrs_name),
                    annotation_row = as.data.frame(colData(methrix_sub)[,params$group_var, 
                                                                        drop = FALSE]),  
                    color = color_values, 
                    clustering_distance_rows = dist(scale(t(plot_dt))), 
                    breaks = quantile_breaks, show_colnames = FALSE)
})
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })

# pdf(paste0(params$output_dir, "ctrs_heatmap_", params$group_var, "_", params$results_name,".pdf"))
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })
# dev.off()
```

### Heatmaps full

```{r results_heatmaps_full}
heat_map_list <- lapply(names(local_res_ctrs), function(ctrs_name){
  ctrs <- local_res_ctrs[[ctrs_name]]
  sig_dmrs <- ctrs[ctrs$qval <= params$dmrseq_thr,]
  if(length(sig_dmrs) <= 2){return(NA)}
  ctrs_ids <- strsplit(ctrs_name, split = "-")[[1]]

  plot_dt <- methrix::get_region_summary(m = merged_methrix, regions = sig_dmrs, 
                                          overlap_type = "any")
  plot_dt <- plot_dt[,6:ncol(plot_dt)]
  plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

  quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  set.seed(params$random_seed)
  # selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
  # browser()
  pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                    main = paste("Heatmap All Significant DMRs (Scaled mean methylation)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var,
                                                                           drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))), 
                    breaks = quantile_breaks, show_colnames = FALSE, clustering_method = "average")
})
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })

# pdf(paste0(params$output_dir, "ctrs_heatmap_full", params$group_var, "_", params$results_name,".pdf"))
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })
# dev.off()
```

### Heatmap full CpGs

```{r results_heatmaps_full_cpgs}
heat_map_list <- lapply(names(local_res_ctrs), function(ctrs_name){
  ctrs <- local_res_ctrs[[ctrs_name]]
  sig_dmrs <- ctrs[ctrs$qval <= params$dmrseq_thr,]
  if(length(sig_dmrs) <= 2){return(NA)}
  ctrs_ids <- strsplit(ctrs_name, split = "-")[[1]]
  
  methrix_subset <- subset_methrix(merged_methrix, 
                                   regions = sig_dmrs, 
                                   overlap_type = "any")
  
  set.seed(params$random_seed)
  subset_cpgs <- sample(seq_len(nrow(methrix_subset)), 
                        size = min(5000, nrow(methrix_subset)))
  
  plot_dt <- assay(methrix_subset)[subset_cpgs,]
  plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

  quantile_breaks <- quantile(scale(t(plot_dt)), 
                              seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  set.seed(params$random_seed)
  # selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
  # browser()
  pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                    main = paste("Heatmap", nrow(plot_dt), 
                                "random CpGs significant DMRs (Scaled Beta)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var, drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))), 
                    breaks = quantile_breaks, show_colnames = FALSE, clustering_method = "average")
})
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })

# pdf(paste0(params$output_dir, "ctrs_heatmap_full_cpgs", params$group_var, "_", params$results_name,".pdf"))
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })
# dev.off()
```

### Pairwise DSS from DMRSeq significant DMRs

```{r, pairwiseDSS, eval = params$pairwiseDSS & params$global_test}
global_test_dt <- local_res_ctrs$`Global Test`[local_res_ctrs$`Global Test`$qval <= params$dmrseq_thr]

methrix_dss_subset <- subset_methrix(m = merged_methrix, 
                                     regions = global_test_dt, 
                                     overlap_type = "any")

filtered_dmrseq_path <- paste0(params$output_dir, "dmrseq_results_filtered_", 
                                results_name, ".rds")

if(!file.exists(filtered_dmrseq_path)){
  dss_path <- paste0(params$output_dir, "dss_results_list_", 
                     results_name, ".", params$group_var, ".rds")
  if(!file.exists(dss_path)){
    xfun::Rscript_call(rmarkdown::render, args = list(
          input = "./DSS_analysis.Rmd", 
          output_file = paste0(params$output_dir, "DSS_analysis_", 
                              results_name, ".", params$group_var, ".html"), 
          params = list(
            output_dir = params$output_dir,
            results_name = results_name,
      
            methrix_path = methrix_dss_subset,
            group_var = params$group_var,
            
            random_seed = params$random_seed,
      
            smoothing = params$smooth,
            dmr_pvalue = 0.001
        ), envir = new.env()
      )
    )
    dss_results_list <- readRDS(dss_path)
  
  } else{
    dss_results_list <- readRDS(dss_path)
  }
  methrix_cpg_granges <- GRanges(seqnames = methrix_dss_subset@elementMetadata$chr, 
                                 ranges = IRanges(start = methrix_dss_subset@elementMetadata$start, 
                                                  end = methrix_dss_subset@elementMetadata$start), 
                                 strand = methrix_dss_subset@elementMetadata$strand)
  
  overlap_cpg_region <- findOverlaps(query = global_test_dt, 
                                    subject = methrix_cpg_granges)
  overlap_cpg_region <- as.data.frame(overlap_cpg_region)
  
  cpgs_in_dmr <- lapply(unique(overlap_cpg_region$queryHits), function(dmr_id){
    overlap_cpg_region$subjectHits[dmr_id == overlap_cpg_region$queryHits]
  })
  
  all(unique(overlap_cpg_region$queryHits) == seq_along(global_test_dt))
  
  sig_res <- lapply(names(dss_results_list), function(ctrs_name){
    is_sig <- sapply(cpgs_in_dmr, function(dmr_id){
      cpgs_res <- dss_results_list[[ctrs_name]][dmr_id,]
      
      sig_cpgs <- cpgs_res[,"fdrs"] <= params$dss_cpg_adj.p.value
      if(sum(sig_cpgs, na.rm = TRUE)/length(sig_cpgs) >= params$dss_cpg_percentage_sig){
        return(0.05)
      } else{
        return(1)
      }
    })
    res_dt <- data.frame(is_sig)
    colnames(res_dt) <- paste0(ctrs_name, ".adj.p.value")
    res_dt
  })
  sig_res <- do.call(cbind, sig_res)
  
  mcols(global_test_dt) <- cbind(mcols(global_test_dt), sig_res)
  
  saveRDS(list("Global Test" = global_test_dt), filtered_dmrseq_path)  
} else{
  global_test_dt <- readRDS(filtered_dmrseq_path)[[1]]
}

any_sig <- rowAnys(as.matrix(mcols(global_test_dt)[,grep("adj.p.value", 
                                                        colnames(mcols(global_test_dt)))]) <= 0.05, 
                    value = TRUE)
sum(any_sig)

methrix_subset_2 <- subset_methrix(merged_methrix, 
                                   global_test_dt[any_sig], 
                                   overlap_type = "any")
```

```{r}
# sig_num <- lapply(names(dss_results_list), function(ctrs_name){
#   sapply(cpgs_in_dmr, function(dmr_id){
#     cpgs_res <- dss_results_list[[ctrs_name]][dmr_id,]
#     
#     sig_cpgs <- cpgs_res[,"fdrs"] <= params$dss_cpg_adj.p.value
#     sum(sig_cpgs, na.rm = TRUE)/length(sig_cpgs)*100
#   })
# })
# names(sig_num) <- names(dss_results_list)
# plot_dt <- reshape2::melt(as.data.frame(sig_num))
# colnames(plot_dt) <- c("Contrast", "Percentage_Significant")
# 
# ggplotly(ggplot(plot_dt, aes(x = Percentage_Significant, fill = Contrast)) + theme_minimal() + geom_histogram(position = "identity")) + 
#   xlab("Percentage Significant CpGs in the global DMRs")
```

```{r, pairwiseDSS_heatmap, eval = params$pairwiseDSS & params$global_test}
set.seed(params$random_seed)
subset_cpgs <- sample(seq_len(nrow(methrix_subset_2)), 
                        size = min(5000, nrow(methrix_subset_2)))
  
plot_dt <- assay(methrix_subset_2)[subset_cpgs,]
plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

quantile_breaks <- quantile(scale(t(plot_dt)), 
                            seq(0, 1, length.out = 101), na.rm = TRUE)
quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
color_values <- viridis::viridis(n = length(quantile_breaks) - 1)

set.seed(params$random_seed)
# selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
# browser()
pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                  main = paste("Heatmap", nrow(plot_dt), 
                              "random CpGs in any DMR with atleast",
                              params$dss_cpg_percentage_sig*100, 
                              "CpGs with pvalue lower than", 
                              params$dss_cpg_adj.p.value),
                  annotation_row = as.data.frame(colData(methrix_subset_2)[,params$group_var, 
                                                                          drop = FALSE]),
                  color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))),
                  breaks = quantile_breaks, show_colnames = FALSE, clustering_method = "average")
```

```{r}

####################
# methrix_subset_2 <- subset_methrix(merged_methrix, test_1[any_sig2], overlap_type = "any")
# 
# set.seed(params$random_seed)
# plot_dt <- assay(methrix_subset_2)
# plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data
# 
# quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
# quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
# color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
# 
# set.seed(params$random_seed)
# # selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
# # browser()
# pheatmap::pheatmap(mat = scale(t(plot_dt)), main = paste("Heatmap All Significant CpG any significant"),
#                   annotation_row = as.data.frame(colData(methrix_subset_2)[,params$group_var, drop = FALSE]),
#                   color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))),
#                   breaks = quantile_breaks, show_colnames = FALSE, clustering_method = "average")

```

## Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
