---
title: "DSS_analysis"
output: html_document
date: "2023-12-06"
params: 
  methrix_path: "/home/rstudio/bioinf-d/lima_results_methylseq/Results/merged_methrix.rds"
  output_dir: "/home/rstudio/bioinf-d/lima_results_methylseq/Results/DSS.group.non_smoothed/"
  
  dmrseq_results_path: "/home/rstudio/bioinf-d/lima_results_methylseq/Results/dmrseq_results.group.non_smoothed/dmrseq_results_non_smoothed.group.rds"
  
  chromosome_filter: !r NULL
  outlier_samples: !r  c("X670E", "X594C")
  random_seed: 1234
  group_var: "group"
  
  adj.p.value_thr: 0.05
  
  results_name: "non_smoothed"

  smoothing: FALSE
  dmr_pvalue: 0.001
  percentage_sig: 0.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, echo = FALSE)
options(knitr.duplicate.label = "allow")
```

```{r}
library(methrix)
library(plotly)
library(ggplot2)
library(annotatr)
library(AnnotationHub)
library(ChIPseeker)
library(impute)
library(limma)
source("./plot_functions.R")
```

```{r}
if(is(params$methrix_path, "methrix")){
  merged_methrix <- params$methrix_path
} else{
  merged_methrix <- readRDS(params$methrix_path)
  
  if(!is.null(params$outlier_samples)){
    merged_methrix <- merged_methrix[,!colnames(merged_methrix) %in% params$outlier_samples]
  }
  
  if(!is.null(params$chromosome_filter)){
    all_chrom <- metadata(merged_methrix)$ref_CpG$chr
    merged_methrix <- subset_methrix(m = merged_methrix, 
                                    contigs = all_chrom[!(all_chrom %in% params$chromosome_filter)])
  }  
}

```

### DSS analysis {.tabset}

```{r}
dss_results_path <- paste0(params$output_dir, "dss_results_", params$results_name, ".", params$group_var, ".rds")
dss_results_list_path <- paste0(params$output_dir, "dss_results_list_", params$results_name, ".", params$group_var, ".rds")

if(!file.exists(dss_results_list_path)){
  print("Calculating DSS models")
  dss_formula <- as.formula(paste("~", params$group_var, "+ 0"))
  
  dss_results <- DSS::DMLfit.multiFactor(BSobj = methrix2bsseq(merged_methrix), colData(merged_methrix),
                                         formula = dss_formula, smoothing = params$smoothing)

  ctrs_vec <- combn(levels(merged_methrix[[params$group_var]]), 2, FUN = function(x){paste0(x[2:1], collapse = "-")})
  mod_mat <- model.matrix(dss_formula, colData(merged_methrix))
  colnames(mod_mat) <- gsub(params$group_var, "", colnames(mod_mat))

  contrst_mat <- limma::makeContrasts(levels = mod_mat, contrasts = ctrs_vec)
  dss_results_list <- list()

  for(col_id in seq_len(ncol(contrst_mat))){
    ctrs_ids <- contrst_mat[,col_id, drop = FALSE]
    dss_results_list[[colnames(contrst_mat)[col_id]]] <- DSS::DMLtest.multiFactor(DMLfit = dss_results,
                                                                                Contrast = ctrs_ids)

  }
  saveRDS(dss_results, dss_results_path)
  saveRDS(dss_results_list, dss_results_list_path)
} else{
  dss_results_list <- readRDS(dss_results_list_path)
}
```

#### Histograms {.tabset}

##### pvalues

```{r}
lapply(dss_results_list, function(x){sum(x[,4] <= params$adj.p.value_thr, na.rm = TRUE)})

hist_list <-lapply(names(dss_results_list), function(ctrs_name){
  # data.frame("pvalues" = dss_results_list[[ctrs_name]][,4], "Contrast" = ctrs_name)
  hist(dss_results_list[[ctrs_name]][,4], main = paste(ctrs_name, "pvalues"))
})
# plot_dt <- data.table::rbindlist(plot_dt)
# ggplot(plot_dt) + geom_histogram(aes(x = pvalues, fill = Contrast), position = "identity") + ggtitle("DSS p-values histogram non smoothed")
```

##### qvalues

```{r}
lapply(dss_results_list, function(x){sum(x[,5] <= params$adj.p.value_thr, na.rm = TRUE)})

hist_list <- lapply(names(dss_results_list), function(ctrs_name){
  # data.frame("pvalues" = dss_results_list[[ctrs_name]][,4], "Contrast" = ctrs_name)
  hist(dss_results_list[[ctrs_name]][,5], main = paste(ctrs_name, "qvalues"))
})
```

#### Heatmaps {.tabset}

##### pvalues top 5000 {.tabset}

###### Beta values

```{r}
heatmal_list <- lapply(names(dss_results_list), function(ctrs_name){
  max_cpgs <- min(5000, nrow(dss_results_list[[ctrs_name]]))
  
  plot_dt <- assay(merged_methrix[order(dss_results_list[[ctrs_name]][,4], 
                                        decreasing = FALSE)[seq_len(max_cpgs)],])
  plot_dt <- plot_dt[rowSums(plot_dt, na.rm = TRUE) != 0,]
  
  plot_dt <- impute::impute.knn(data = plot_dt)$data

  quantile_breaks <- quantile(plot_dt, seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  pheatmap::pheatmap(mat = t(plot_dt), main = paste("Top", nrow(plot_dt), 
                                                    "most significant CpGs DSS (pvalue and clustering not all CpG)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var, drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))),
                    clustering_method = "average",
                    breaks = quantile_breaks, show_colnames = FALSE)

  # png(paste0(params$output_dir, "Top_5000_most_significant_cpgs_beta_values_heatmap", ctrs_name, ".png"))
  # plot(heatmap_dt$gtable)
  # dev.off()
})
```

###### Scaled Beta values

```{r}
heatmap_list <- lapply(names(dss_results_list), function(ctrs_name){
  max_cpgs <- min(5000, nrow(dss_results_list[[ctrs_name]]))

  plot_dt <- assay(merged_methrix[order(dss_results_list[[ctrs_name]][,4], 
                                        decreasing = FALSE)[seq_len(max_cpgs)],])
  plot_dt <- plot_dt[rowSums(plot_dt, na.rm = TRUE) != 0,]
  
  plot_dt <- impute::impute.knn(data = plot_dt)$data
  
  quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                    main = paste("Top", max_cpgs, "most significant CpGs DSS (by pvalue and clustering not all CpG)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var, drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))),
                    clustering_method = "average",
                    breaks = quantile_breaks, show_colnames = FALSE)

  # png(paste0(params$output_dir, "Top_5000_most_significant_cpgs_scaled_beta_values_heatmap", ctrs_name, ".png"))
  # plot(heatmap_dt$gtable)
  # dev.off()
})
```

##### qvalues (all sig contrasts with more than 2 significant CpGs) {.tabset}

###### Beta values

```{r}
heatmap_list <- lapply(names(dss_results_list), function(ctrs_name){
  sig_cpgs <- dss_results_list[[ctrs_name]][,5] <= params$adj.p.value_thr
  
  if(sum(sig_cpgs, na.rm = TRUE) < 5){return(NULL)}
  
  plot_dt <- assay(merged_methrix[which(sig_cpgs),])
  plot_dt <- plot_dt[rowSums(plot_dt, na.rm = TRUE) != 0,]

  plot_dt <- impute::impute.knn(data = plot_dt)$data
  
  quantile_breaks <- quantile(plot_dt, seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  pheatmap::pheatmap(mat = t(plot_dt), main = paste(nrow(plot_dt), "Significant CpGs DSS (by pvalue and clustering not all CpG)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var, drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))),
                    clustering_method = "average",
                    breaks = quantile_breaks, show_colnames = FALSE)

  # png(paste0(params$output_dir, "Top_5000_most_significant_cpgs_beta_values_heatmap", ctrs_name, ".png"))
  # plot(heatmap_dt$gtable)
  # dev.off()
})
```

###### Scaled Beta values

```{r}
heatmap_list <-lapply(names(dss_results_list), function(ctrs_name){
  sig_cpgs <- dss_results_list[[ctrs_name]][,5] <= params$adj.p.value_thr
  
  if(sum(sig_cpgs, na.rm = TRUE) < 5){return(NULL)}
  
  plot_dt <- assay(merged_methrix[which(sig_cpgs),])
  plot_dt <- plot_dt[rowSums(plot_dt, na.rm = TRUE) != 0,]
  plot_dt <- impute::impute.knn(data = plot_dt)$data

  quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  pheatmap::pheatmap(mat = scale(t(plot_dt)), main = paste(nrow(plot_dt), "Significant CpGs DSS (by qvalue and clustering not all CpG)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var, drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))),
                    clustering_method = "average",
                    breaks = quantile_breaks, show_colnames = FALSE)

  # png(paste0(params$output_dir, "Top_5000_most_significant_cpgs_scaled_beta_values_heatmap", ctrs_name, ".png"))
  # plot(heatmap_dt$gtable)
  # dev.off()
})
```

#### DSS DMR Calling {.tabset}

```{r}
dss_dmr_path <- paste0(params$output_dir, "dss_dmr_results_", params$results_name, ".", params$group_var, ".rds")

if(!file.exists(dss_dmr_path)){
  called_dmrs <- lapply(names(dss_results_list), function(ctrs_res_name){
    ctrs_res <- dss_results_list[[ctrs_res_name]]
    print(ctrs_res_name)
    colnames(ctrs_res)[3:4] <- c("stat", "pval")
    ctrs_res$pval[is.na(ctrs_res$pval)] <- 1
    return(DSS::callDMR(DMLresult = ctrs_res, 
                        p.threshold = params$dmr_pvalue, 
                        pct.sig = params$percentage_sig))
  })
  names(called_dmrs) <- names(dss_results_list)
  called_dmrs_granges <- lapply(called_dmrs, function(res){
    GenomicRanges::makeGRangesFromDataFrame(res, keep.extra.columns = TRUE)
  })
  called_dmrs_granges <- lapply(names(called_dmrs_granges), function(res_id){
    res <- called_dmrs_granges[[res_id]]
    mcols(res)[[paste0(res_id, ".adj.p.value")]] <- 0.05 
    res
  })
  names(called_dmrs_granges) <- names(called_dmrs)
  saveRDS(called_dmrs_granges, dss_dmr_path)
} else{
  called_dmrs_granges <- readRDS(dss_dmr_path)
}
```

```{r}
lengths(called_dmrs_granges)
```

```{r}
if(file.exists(params$dmrseq_results_path)){
  local_res_ctrs <- readRDS(params$dmrseq_results_path)
  
  mapply(function(dmrseq_res, dss_res){
    overlapping_found <- findOverlaps(dmrseq_res[dmrseq_res$qval <= 0.05], dss_res)
    
    length(unique(overlapping_found@to))/length(dss_res)
    
    
  }, dmrseq_res = local_res_ctrs, dss_res = called_dmrs_granges)
  
  mapply(function(dmrseq_res, dss_res){
    overlapping_found <- findOverlaps(dmrseq_res[dmrseq_res$qval <= 0.05], dss_res)
    length(unique(overlapping_found@from))/length(dmrseq_res[dmrseq_res$qval <= params$adj.p.value_thr])
    
    
  }, dmrseq_res = local_res_ctrs, dss_res = called_dmrs_granges)
}

```

Significant DMRs found with DSS (more than `params$percentage_sig` with pvalue higher than `params$dmr_pvalue`) that were also found by DMRseq (qval <= 0.05)

##### Heatmaps full

```{r results_heatmaps_full}
heat_map_list <- lapply(names(called_dmrs_granges), function(ctrs_name){
  ctrs <- called_dmrs_granges[[ctrs_name]]
  sig_dmrs <- ctrs[mcols(ctrs)[,grepl("adj.p.value", colnames(mcols(ctrs)))] <= params$adj.p.value_thr]
  if(length(sig_dmrs) <= 2){return(NA)}
  ctrs_ids <- strsplit(ctrs_name, split = "-")[[1]]

  plot_dt <- methrix::get_region_summary(m = merged_methrix, regions = sig_dmrs, 
                                          overlap_type = "any")
  plot_dt <- plot_dt[,6:ncol(plot_dt)]
  plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

  quantile_breaks <- quantile(scale(t(plot_dt)), seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  set.seed(params$random_seed)
  # selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
  # browser()
  pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                    main = paste("Heatmap All Significant DMRs (Scaled mean methylation)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var,
                                                                           drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))), 
                    breaks = quantile_breaks, show_colnames = FALSE, clustering_method = "average")
})
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })

# pdf(paste0(params$output_dir, "ctrs_heatmap_full", params$group_var, "_", params$results_name,".pdf"))
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })
# dev.off()
```

##### Heatmap full CpGs

```{r results_heatmaps_full_cpgs}
heat_map_list <- lapply(names(called_dmrs_granges), function(ctrs_name){
  ctrs <- called_dmrs_granges[[ctrs_name]]
  sig_dmrs <- ctrs[mcols(ctrs)[,grepl("adj.p.value", colnames(mcols(ctrs)))] <= params$adj.p.value_thr]
  if(length(sig_dmrs) <= 2){return(NA)}
  ctrs_ids <- strsplit(ctrs_name, split = "-")[[1]]
  
  methrix_subset <- subset_methrix(merged_methrix, 
                                   regions = sig_dmrs, 
                                   overlap_type = "any")
  
  set.seed(params$random_seed)
  subset_cpgs <- sample(seq_len(nrow(methrix_subset)), 
                        size = min(5000, nrow(methrix_subset)))
  
  plot_dt <- assay(methrix_subset)[subset_cpgs,]
  plot_dt <- impute::impute.knn(data = as.matrix(plot_dt))$data

  quantile_breaks <- quantile(scale(t(plot_dt)), 
                              seq(0, 1, length.out = 101), na.rm = TRUE)
  quantile_breaks <- quantile_breaks[!duplicated(quantile_breaks)]
  color_values <- viridis::viridis(n = length(quantile_breaks) - 1)
  
  set.seed(params$random_seed)
  # selected_random_dmrs <- sample(seq_len(nrow(plot_dt)), size = ceiling(nrow(plot_dt)*0.05))
  # browser()
  pheatmap::pheatmap(mat = scale(t(plot_dt)), 
                    main = paste("Heatmap", nrow(plot_dt), 
                                "random CpGs significant DMRs (Scaled Beta)", ctrs_name),
                    annotation_row = as.data.frame(colData(merged_methrix)[,params$group_var, drop = FALSE]),  
                    color = color_values, clustering_distance_rows = dist(scale(t(plot_dt))), 
                    breaks = quantile_breaks, show_colnames = FALSE, clustering_method = "average")
})
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })

# pdf(paste0(params$output_dir, "ctrs_heatmap_full_cpgs", params$group_var, "_", params$results_name,".pdf"))
# lapply(heat_map_list, function(x){
#   if(length(x) == 1){return(NULL)}
#   plot(x$gtable)
# })
# dev.off()
```

```{r}
rm(merged_methrix)
```


## Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}

```
